### 2021-07-15 plos_latex_template.tex
Initial Overleaf Import

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules towards an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Name1 Surname\textsuperscript{1,2\Yinyang},
Name2 Surname\textsuperscript{2\Yinyang},
Name3 Surname\textsuperscript{2,3\textcurrency},
Name4 Surname\textsuperscript{2},
Name5 Surname\textsuperscript{2\ddag},
Name6 Surname\textsuperscript{2\ddag},
Name7 Surname\textsuperscript{1,2,3*},
with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
\dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* correspondingauthor@institute.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract (optional from what I've seen)}

In July 2021, the authors of this article, along with a larger group of people, organized a global and virtual conference for users and developers of the R software, useR!2021. useR! conferences are non-profit conferences organized by volunteers for the community, supported by the R Foundation. Attendees are from the academia and the industry. We aimed at building a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone. We would like to share the lessons learned within a year of work organizing this conference, as 10 simple rules towards an inclusive conference. UseR! These rules apply to academic, industry, or mixed conferences; the rules are inspired by a global experience but can also be applied to regional or local conferences.

% % Please keep the Author Summary between 150 and 200 words
% % Use first person. PLOS ONE authors please skip this step. 
% % Author Summary not valid for PLOS ONE submissions.   
% \section*{Author summary (I don't think we need it)}
% Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}

Conferences are spaces to meet and reconnect with a specific community, learn from the new work in the field and share our own work. The larger the conference, the larger the network and opportunities to meet and learn. However, just like in science and academia in general, conferences can become an exclusive space for a privileged group (e.g. white, male, from a rich country, English-native speaker, with no physical disabilities).

This article aims at suggesting rules for opening conferences towards inclusiveness and diversity, and to truly welcome underrepresented minorities. They are based on the experience of the authors in the organization of a global and virtual conference for users and developers of the R software, useR!2021. We aimed at building a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone. [Need a ending sentence to connect to the rules.]

\section*{Rule 1: Embrace all dimensions of diversity}
The first step to be inclusive is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event. Make efforts to welcome --and even better, celebrate --all dimensions of human diversity, such as age, body ability, career stage, gender, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. Write a diversity statement (e.g. \url{https://user2021.r-project.org/about/diversity_statement/}) and express this welcoming spirit in social media to let people know that they are seen, respected and welcome. That this is their space and their community too. 
%ast be aware that participants may come from different backgrounds and that freedom of speech can be used to harm others, so don't let one dimension of diversity (i.e., political opinion) prevail over others. See rule 10.

\section*{Rule 2: Have an inclusive organizing team}
An inclusive conference can only be organized by an inclusive and diverse organizing team. Invite people from different regions, genders, socioeconomic situations, and other dimensions of diversity. Pay special attention to the usually underrepresented groups. If the organization is not used to working with a diverse group, it may be challenging. But the challenges will help your team grow, learn and have more inclusive practices. If the organizing team is global and remote, you may struggle with time zones, and meet at odd times. You may also have to review carefully the choice of words for emails,
%ast ¿algo más enfático? [do the effort to] teach yourself the correct choice of words...
social media publications and your website, to be respectful of every culture and situation;
%ast yo cortaría esto en dos separadas. 
team members could be quick to spot text that is not right, or that does not resonate with a particular group of people. An inclusive team will let you know if the organization is failing in any aspect of inclusiveness or accessibility.

\section*{Rule 3: Go beyond your limited networks to find the best people for everything}

When inviting people to the organizing team, keynote speakers, program committee, session chairs and others, always look for the best in their subject. If there is not much diversity in your first set of names, it does not mean that minorities are not good enough. It probably means that the system has always privileged some groups of people, and that you need to look further to find great people that are not on the spotlight. [At least in one book I've read a discussion about H-index that we could cite here]. Look beyond your narrow and often limited networks. And if you start by having an inclusive organizing team (rule 1 or 2), they can help find the right people.

\section*{Rule 4: Have registration rates according to the cost of living of participants}

Conferences, even virtual ones, should have a registration fee mainly for two reasons. The first one is that preparing the conference takes a lot of effort and costs money (e.g. commercial registration tools, captioning, or conference venue if in-person) and it should be reflected on a price. The second reason is that people tend to value more the things they pay for, and there is a higher risk of not showing up to a free even [any publication to back this up?]. On the other hand, if you are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when fixing the registration fees. Usually, there is a higher fee for people from the industry than for academia. Consider also a lower fee for non-profit organizations, government employees, or free lancers. 
% ast: don't forget people in vulnerable job positions, such as post-docs
You can also use conversion factors for country of residence using data from International Comparison Program report of the World Bank.

\section*{Rule 5: Do not give scholarships or grants to attend. Give fee waivers or discounts without tedious application forms.}

Offering scholarships or grants to attend your conference may be counterproductive, as they can be seen more of a competition and attract people that might not need the money (e.g. their PI could pay for it), but could apply for it to enrich their CV. If the goal is to help the people in need, make that clear. Do not give it a fancy name. It is a fee waiver or a discount for the people in need. And simplify the process for applicants. People with low resources already have a hard time applying for loans, grants and scholarships in their lives. Do not make life harder on them by asking them to write long paragraphs to convince you that they deserve your support. Many people that cannot afford conferences do not even try to apply for waivers/scholarships because they think it would be hard to get them. Trust that when they say they need it, it is because they do.

If your resources allow it, you could even take a step forwards and offer financial support: child care support, transportation fees (if in-person) or internet connection services (if virtual).

\section*{Rule 6: Make the conference accessible to all}

Conferences are usually inaccessible to people with hearing and visual impairments. Take them into account in every stage of the process. Welcome them to your organizing team so that they can be part of the decisions since the beginning, and remind you of inaccessible practices that need to improve. Communicate on social media with alternative text and videos with captions. Use screen-reader friendly platforms for conference registration, abstract submission and, if the conference is virtual or hybrid, look for a screen-reader friendly chat platform too. Provide captioning for presentations. If the conference is in-person, make sure that presenters speak to the microphone to make it easier for captioners or interpreters to listen to them. Provide accessibility guidelines for slides and presentations, and encourage the presenters to follow them. Accessibility means, for example, that the speakers should also provide raw and accessible material to their talks, such as RMarkdown, html or tex files, and that, if they pre-record their presentation, to make sure that their video is shown and their face is visible so that people with hearing impairment can have the possibility to read their lips. You can also give accessibility awards at the end of the conference to acknowledge the speakers who were mindful of inclusiveness when preparing their talks. Please find an example of accessibility guidelines at \url{https://user2021.r-project.org/participation/accessibility/}.
%ast Bear in mind that most accessibility practices are beneficial to everyone. Captions, for instance, are used by non-native speakers and people that access the conference in loud environments. The prior availability of material benefits people with low bandwidth and non-native speakers as well.

\section*{Rule 7: Put more effort in promoting your conference among groups of people who are usually underrepresented}

If some minorities are under represented in your conference, they may not feel part of the community or may not have the financial resources to attend. Promote the conference in both ways: as welcoming for everyone and supportive for those who do not have the money %ast: or "the credentials" to be there.
For the former, if part of the community has been historically discriminated, emphasize that they are particularly welcome in this event, and that the organization will make sure to make it a safe and inclusive environment for them. For the latter, advertise the fee waivers, financial support, and the easiness of the process. People who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be rather supportive in this case.

\section*{Rule 8: Have an important online component of the conference}

In-person interaction is priceless; however, it is more expensive for some, even unattainable for others. This is particularly true for global conferences. Global conferences usually take place in high-income countries, making it too expensive and for many people impossible to attend due to immigration requirements. [Share links to news of conference participants denied entry to countries, and maybe some narrating the whole absurd process for visas] Online conferences are more inclusive because there is not need for a visa or a big budget.

Alternatively, a conference could have an in-person and an online component. This could allow for a group of people to physically interact, while providing many others the opportunity to participate. The challenge in this kind of setting would be to make the online component equally relevant as the in-person component, and not just a consolation price to the less privileged in the community.

\section*{Rule 9: Don't let English be a barrier for good quality participation}

In international conferences, English is often the official language. Submissions, presentations, tutorials and workshops are in English. While one official language is extremely helpful to communicate and English is the primary language in scientific communication, opening the conference to other languages could make it less intimidating to people who are not so fluent in English. Excluding them may also mean missing great innovative pieces of work, because of a language barrier. So advertise your conference in several languages, and consider having workshops and presentations in other languages (with or without captions in English).

\section*{Rule 10: Have a code of conduct team}

To maintain and inclusive and safe space in your conference, there should be a code of conduct and a team to enforce it. `The purpose of a code of conduct is to protect members of a community from harm in that community's spaces. The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' (coc book)

The code of conduct team should receive training to have a clear understanding on what to enforce and what not to enforce in a code of conduct, how to receive reports, respond to incidents, and communicate the responses. A diverse code of conduct team could be more understanding of intersectionality issues in discrimination and harassment practices. 
%the code of condcut should be displayed proeminently in several spaces of the conference and it should be effective if it deters people from unaccepted behavior in the first place

\section*{Concluding remarks}

The ten rules stated here can be adapted depending on the conference settings. We are aware, from our own experience, that enforcing these rules require a lot of work that is not compensated financially or in promotions. The rewards from this practice can be a healthier, stronger and more inclusive community, empowering people that are historically marginalized, and learning from them, thus increasing the quality of the conference.

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR!2021 for their valuable contribution to an inclusive conference experience. And the R Foundation for their support to the conference.

% Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

% \nolinenumbers

% % Either type in your references using
% % \begin{thebibliography}{}
% % \bibitem{}
% % Text
% % \end{thebibliography}
% %
% % or
% %
% % Compile your BiBTeX database using our plos2015.bst
% % style file and paste the contents of your .bbl file
% % here. See http://journals.plos.org/plosone/s/latex for 
% % step-by-step instructions.
% % % 
% \begin{thebibliography}{10}

% \bibitem{bib1}
% Conant GC, Wolfe KH.
% \newblock {{T}urning a hobby into a job: how duplicated genes find new
%   functions}.
% \newblock Nat Rev Genet. 2008 Dec;9(12):938--950.

% \bibitem{bib2}
% Ohno S.
% \newblock Evolution by gene duplication.
% \newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
%   Springer-Verlag.; 1970.

% \bibitem{bib3}
% Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
% \newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
%   infection through a transposon insertion followed by a {D}uplication}.
% \newblock PLoS Genet. 2011 Oct;7(10):e1002337.

% \end{thebibliography}

\end{document}

### 2021-07-15 plos_latex_template.tex
adding some of Andrea's comments to the text and changing the style

@@
 
-% Use "Eq" instead of "Equation" for equation citations.
 \section*{Introduction}
@@
 The first step to be inclusive is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event.
-Make efforts to welcome --and even better, celebrate --all dimensions of human diversity, such as age, body ability, career stage, gender, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
-Write a diversity statement (e.g.
-\url{https://user2021.r-project.org/about/diversity_statement/}) and express this welcoming spirit in social media to let people know that they are seen, respected and welcome.
+We can make efforts to welcome --and even better, celebrate --all dimensions of human diversity, such as age, body ability, career stage, gender, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
+Publishing a diversity statement (e.g.
+\url{https://user2021.r-project.org/about/diversity_statement/}) and expressing this welcoming spirit in social media will let people know that they are seen, respected and welcome.
 That this is their space and their community too.
@@
 An inclusive conference can only be organized by an inclusive and diverse organizing team.
-Invite people from different regions, genders, socioeconomic situations, and other dimensions of diversity.
-Pay special attention to the usually underrepresented groups.
+We can invite people from different regions, genders, socioeconomic situations, and other dimensions of diversity.
+Special attention should be payed to the usually underrepresented groups.
 If the organization is not used to working with a diverse group, it may be challenging.
-But the challenges will help your team grow, learn and have more inclusive practices.
-If the organizing team is global and remote, you may struggle with time zones, and meet at odd times.
-You may also have to review carefully the choice of words for emails,
-%ast ¿algo más enfático? [do the effort to] teach yourself the correct choice of words...
-social media publications and your website, to be respectful of every culture and situation;
-%ast yo cortaría esto en dos separadas. 
-team members could be quick to spot text that is not right, or that does not resonate with a particular group of people.
-An inclusive team will let you know if the organization is failing in any aspect of inclusiveness or accessibility.
+But the challenges will help the team grow, learn and have more inclusive practices.
+If the organizing team is global and remote, time zones can be hard to manage, and meetings will end up happening at odd times.
+The team may also have to review carefully and learn to choose the right words for emails,
+social media publications and your website, to be respectful of every culture and situation. 
+Team members could be quick to spot text that is not right, or that does not resonate with a particular group of people.
+An inclusive team will let everyone know if the organization is failing in any aspect of inclusiveness or accessibility, and help improve it.
 
@@
 
-When inviting people to the organizing team, keynote speakers, program committee, session chairs and others, always look for the best in their subject.
-If there is not much diversity in your first set of names, it does not mean that minorities are not good enough.
-It probably means that the system has always privileged some groups of people, and that you need to look further to find great people that are not on the spotlight.
+When inviting people to the organizing team, keynote speakers, program committee, session chairs and others, we should always look for the best in their subject.
+If there is not much diversity in the first set of names, it does not mean that minorities are not good enough.
+It probably means that the system has always privileged some groups of people, and that we need to look further to find great people that are not on the spotlight.
 [At least in one book I've read a discussion about H-index that we could cite here].
-Look beyond your narrow and often limited networks.
-And if you start by having an inclusive organizing team (rule 1 or 2), they can help find the right people.
+Let's look beyond our narrow and often limited networks.
+And if we start by having an inclusive organizing team (rule 1 or 2), they can help find the right people.
 
@@
 The second reason is that people tend to value more the things they pay for, and there is a higher risk of not showing up to a free even [any publication to back this up?].
-On the other hand, if you are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when fixing the registration fees.
+On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when fixing the registration fees.
 Usually, there is a higher fee for people from the industry than for academia.
-Consider also a lower fee for non-profit organizations, government employees, or free lancers.
-% ast: don't forget people in vulnerable job positions, such as post-docs
-You can also use conversion factors for country of residence using data from International Comparison Program report of the World Bank.
+We should also consider a lower fee for non-profit organizations, government employees, or free lancers.
+And use conversion factors for country of residence using data from International Comparison Program report of the World Bank.
+It is important to include discounts for students as well as for postdocs, as these are often precarious positions.
+Postdoc status is not always well-defined in academia and can vary for each country, so their payment category should be explicitly defined.
 
@@
 
-Offering scholarships or grants to attend your conference may be counterproductive, as they can be seen more of a competition and attract people that might not need the money (e.g.
+Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen more of a competition and attract people that might not need the money (e.g.
 their PI could pay for it), but could apply for it to enrich their CV.
-If the goal is to help the people in need, make that clear.
+If the goal is to help the people in need, let's make that clear.
 Do not give it a fancy name.
 It is a fee waiver or a discount for the people in need.
-And simplify the process for applicants.
+And the process for applicants should be simplified.
 People with low resources already have a hard time applying for loans, grants and scholarships in their lives.
-Do not make life harder on them by asking them to write long paragraphs to convince you that they deserve your support.
+We should not make their lives harder by asking them to write long paragraphs to convince us that they deserve our support.
 Many people that cannot afford conferences do not even try to apply for waivers/scholarships because they think it would be hard to get them.
-Trust that when they say they need it, it is because they do.
+We should trust that when they say they need it, it is because they do.
 
-If your resources allow it, you could even take a step forwards and offer financial support: child care support, transportation fees (if in-person) or internet connection services (if virtual).
+If the conference resources allow it, one could even take a step forwards and offer financial support: child care support, transportation fees (if in-person) or internet connection services (if virtual).
 
@@
 Conferences are usually inaccessible to people with hearing and visual impairments.
-Take them into account in every stage of the process.
-Welcome them to your organizing team so that they can be part of the decisions since the beginning, and remind you of inaccessible practices that need to improve.
-Communicate on social media with alternative text and videos with captions.
-Use screen-reader friendly platforms for conference registration, abstract submission and, if the conference is virtual or hybrid, look for a screen-reader friendly chat platform too.
-Provide captioning for presentations.
-If the conference is in-person, make sure that presenters speak to the microphone to make it easier for captioners or interpreters to listen to them.
-Provide accessibility guidelines for slides and presentations, and encourage the presenters to follow them.
+They should be taken into account in every stage of the process.
+Welcoming them into the organizing team would allow them to take part in the decisions since the beginning, and remind us of inaccessible practices that need to improve.
+In social media communication, images should be accompanied with alternative text and videos with captions.
+Platforms for conference registration, abstract submission --and, if the conference is virtual or hybrid, chat platforms --should be screen-reader friendly (tested beforehand).
+Captioning for presentations should be available.
+If the conference is in-person, presenters should speak to the microphone to make it easier for captioners or interpreters to listen to them.
+Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged.
 Accessibility means, for example, that the speakers should also provide raw and accessible material to their talks, such as RMarkdown, html or tex files, and that, if they pre-record their presentation, to make sure that their video is shown and their face is visible so that people with hearing impairment can have the possibility to read their lips.
-You can also give accessibility awards at the end of the conference to acknowledge the speakers who were mindful of inclusiveness when preparing their talks.
+Accessibility awards at the end of the conference are a good way to acknowledge the speakers who were mindful of inclusiveness when preparing their talks.
 Please find an example of accessibility guidelines at \url{https://user2021.r-project.org/participation/accessibility/}.
-%ast Bear in mind that most accessibility practices are beneficial to everyone.
+
+Accessibility practices should not be left for the last minute.
+They require time and early decision-making.
+Conversely, inaccessible decisions are hard to undo; e.g.
+when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference.
+Bear in mind that most accessibility practices are beneficial to everyone.
 Captions, for instance, are used by non-native speakers and people that access the conference in loud environments.
@@
 
-\section*{Rule 7: Put more effort in promoting your conference among groups of people who are usually underrepresented}
+\section*{Rule 7: Put more effort in promoting the conference among groups of people who are usually underrepresented}
 
-If some minorities are under represented in your conference, they may not feel part of the community or may not have the financial resources to attend.
-Promote the conference in both ways: as welcoming for everyone and supportive for those who do not have the money %ast: or "the credentials" to be there.
-For the former, if part of the community has been historically discriminated, emphasize that they are particularly welcome in this event, and that the organization will make sure to make it a safe and inclusive environment for them.
-For the latter, advertise the fee waivers, financial support, and the easiness of the process.
+If some minorities are under represented in the conference, they may not feel part of the community or may not have the financial resources to attend.
+And effort must be made to promote the conference as welcoming for everyone and supportive for those who do not have the money %ast: or "the credentials" to be there.
+For the former, if part of the community has been historically discriminated, one should emphasize that they are particularly welcome in this event, and that the organization will make sure to make it a safe and inclusive environment for them.
+For the latter, fee waivers, financial support, and the easiness of the process should be advertised.
 People who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be rather supportive in this case.
@@
 Excluding them may also mean missing great innovative pieces of work, because of a language barrier.
-So advertise your conference in several languages, and consider having workshops and presentations in other languages (with or without captions in English).
+Advertising the conference in several languages, and considering having workshops and presentations in other languages (with or without captions in English) could help overcome this barrier.
 
@@
 
-To maintain and inclusive and safe space in your conference, there should be a code of conduct and a team to enforce it.
+To maintain and inclusive and safe space in the conference, there should be a code of conduct and a team to enforce it.
 `The purpose of a code of conduct is to protect members of a community from harm in that community's spaces.
 The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' (coc book)
+The code of conduct should be displayed prominently in several spaces of the conference to deter people from incurring in unaccepted behavior.
 
@@
 A diverse code of conduct team could be more understanding of intersectionality issues in discrimination and harassment practices.
-%the code of condcut should be displayed proeminently in several spaces of the conference and it should be effective if it deters people from unaccepted behavior in the first place
 

### 2021-08-19 plos_latex_template.tex
few edits to MS (JR)

@@
 \usepackage{array}
+
+%strikethrough
+\usepackage{soul}
 
@@
 
-In July 2021, the authors of this article, along with a larger group of people, organized a global and virtual conference for users and developers of the R software, useR!2021.
+In July 2021, the authors of this article, along with a larger group of people, organized a global and virtual conference for users and developers of the R software, useR!
+2021.
 useR!
-conferences are non-profit conferences organized by volunteers for the community, supported by the R Foundation.
-Attendees are from the academia and the industry.
-We aimed at building a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-We would like to share the lessons learned within a year of work organizing this conference, as 10 simple rules towards an inclusive conference.
-UseR!
-These rules apply to academic, industry, or mixed conferences; the rules are inspired by a global experience but can also be applied to regional or local conferences.
+conferences are strictly non-profit and are organized by volunteers from the R community and supported by the R Foundation.
+These global meetings include attendees from both academia and the industry.
+This year, we aimed at building a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone.
+In this article, we share a few critical lessons learned during the past year's work organizing this conference.
+We highlight these as the 10 simple rules to host an inclusive conference.
+These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but can also be easily adapted and applied to regional or local conferences.
 
@@
 
-Conferences are spaces to meet and reconnect with a specific community, learn from the new work in the field and share our own work.
-The larger the conference, the larger the network and opportunities to meet and learn.
-However, just like in science and academia in general, conferences can become an exclusive space for a privileged group (e.g.
-white, male, from a rich country, English-native speaker, with no physical disabilities).
+Conferences are spaces to meet and reconnect with members from a specific community, learn about new advances in the field and share our recent contributions.
+The larger the conference, the larger the opportunities to network and learn from your cohort.
+However, just like in science and academia in general, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arend_disparity_2019, timperley_he_2020, gewin_what_2019, brown_ableism_2018}.
 
-This article aims at suggesting rules for opening conferences towards inclusiveness and diversity, and to truly welcome underrepresented minorities.
-They are based on the experience of the authors in the organization of a global and virtual conference for users and developers of the R software, useR!2021.
-We aimed at building a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-[Need a ending sentence to connect to the rules.]
+Calls to improve diversity in many academic spaces, including conferences, are becoming frequent as communities and scientific societies are called out for overrepresenting specific groups of people.
+While these systematic issues are beginning to be addressed, there is scope for much improvement all around.
 
-\section*{Rule 1: Embrace all dimensions of diversity}
-The first step to be inclusive is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event.
-We can make efforts to welcome --and even better, celebrate --all dimensions of human diversity, such as age, body ability, career stage, gender, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
-Publishing a diversity statement (e.g.
-\url{https://user2021.r-project.org/about/diversity_statement/}) and expressing this welcoming spirit in social media will let people know that they are seen, respected and welcome.
-That this is their space and their community too.
-%ast be aware that participants may come from different backgrounds and that freedom of speech can be used to harm others, so don't let one dimension of diversity (i.e., political opinion) prevail over others.
-See rule 10.
+Diversity encompasses multiple dimensions: age, body ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
+There are implicit hierarchies along these axes, and some statuses (e.g., being cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
+Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
 
-\section*{Rule 2: Have an inclusive organizing team}
-An inclusive conference can only be organized by an inclusive and diverse organizing team.
-We can invite people from different regions, genders, socioeconomic situations, and other dimensions of diversity.
-Special attention should be payed to the usually underrepresented groups.
-If the organization is not used to working with a diverse group, it may be challenging.
-But the challenges will help the team grow, learn and have more inclusive practices.
-If the organizing team is global and remote, time zones can be hard to manage, and meetings will end up happening at odd times.
-The team may also have to review carefully and learn to choose the right words for emails,
-social media publications and your website, to be respectful of every culture and situation. 
-Team members could be quick to spot text that is not right, or that does not resonate with a particular group of people.
-An inclusive team will let everyone know if the organization is failing in any aspect of inclusiveness or accessibility, and help improve it.
+%some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}
 
-\section*{Rule 3: Go beyond your limited networks to find the best people for everything}
+This article suggests rules for pivoting traditional conferences towards inclusiveness and diversity and welcoming minoritized groups.
+These tips stem from the authors' experience in organizing a global and virtual conference for users and developers of the R programming language, useR!
+2021.
+We aimed at building a high-quality conference not just in terms of technical content but embracing everyone in a kind, inclusive, accessible, and welcoming environment, critical in the context of the COVID-19 pandemic.
+Most of the authors also have experience organizing other regional and national academic conferences and events in communities of practice such as R-Ladies and The Carpentries.
+Here, we share the lessons learned within the past year of organizing this global useR conference, summarized as 10 simple rules towards an inclusive conference.
 
-When inviting people to the organizing team, keynote speakers, program committee, session chairs and others, we should always look for the best in their subject.
-If there is not much diversity in the first set of names, it does not mean that minorities are not good enough.
-It probably means that the system has always privileged some groups of people, and that we need to look further to find great people that are not on the spotlight.
-[At least in one book I've read a discussion about H-index that we could cite here].
-Let's look beyond our narrow and often limited networks.
-And if we start by having an inclusive organizing team (rule 1 or 2), they can help find the right people.
+\section*{Rule 1: Set realistic but bold diversity indicators}
+.
+As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors.
+Next, set measurable indicators of success when organizing a diverse conference.
+The numbers resulting from these indicators are not the primary goal, and this is not about performative allyship.
+Instead, these numbers should guide and help hold the organizing team accountable towards these diverse and inclusive goals.
 
-\section*{Rule 4: Have registration rates according to the cost of living of participants}
+We found it greatly helpful to publish a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcomed, and that this is their space and community too.
+However, it is important to not stop with a statement alone, but identify specific goals and tasks that ensure compliance with these larger goals, for e.g., identifying accessible platforms, developing a detailed code of conduct, and accessibility guidelines (see other Rules below).
 
-Conferences, even virtual ones, should have a registration fee mainly for two reasons.
-The first one is that preparing the conference takes a lot of effort and costs money (e.g.
-commercial registration tools, captioning, or conference venue if in-person) and it should be reflected on a price.
-The second reason is that people tend to value more the things they pay for, and there is a higher risk of not showing up to a free even [any publication to back this up?].
-On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when fixing the registration fees.
-Usually, there is a higher fee for people from the industry than for academia.
-We should also consider a lower fee for non-profit organizations, government employees, or free lancers.
-And use conversion factors for country of residence using data from International Comparison Program report of the World Bank.
-It is important to include discounts for students as well as for postdocs, as these are often precarious positions.
-Postdoc status is not always well-defined in academia and can vary for each country, so their payment category should be explicitly defined.
+\section*{Rule 2: Have an inclusive and diverse organizing team}
+A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
+Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity.
+Particular attention should be paid to the usually underrepresented groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event).
+To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
+Even if creating and maintaining such an interdisciplinary team is challenging, the positive outcomes far outweigh any minor inconveniences, and the team will grow, learn, and embrace more inclusive practices.
+For instance, if the organizing team is global and remote, time zones can be hard to manage, with meetings scheduled at odd hours.
+The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best platforms to mechanisms to account for every culture and situation.
+A successful well-oiled inclusive team will provide a safe space for everyone not just to learn and grow but identify critical aspects that are working or failing to improve inclusiveness and accessibility together \cite{hong_groups_2004}.
+
+%rj: Yani's refs to have just in case:
+%https://sites.lsa.umich.edu/scottepage/wp-content/uploads/sites/344/2015/11/pnas.pdf; https://www.pnas.org/content/early/2014/11/13/1407301111; https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
+
+\section*{Rule 3: Finding the best people for each task}
+
+While this rule may seem intuitive, one would be surprised by how often conference organizers tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in.
+For e.g., a cisgender or non-neurodiverse person estimating what might work best for the LGBTQ+ or people with hearing and seeing difficulty.
+Therefore, to build your team of organizers for the conference, you need to actively seek out and recruit the best people for each task, akin to filling your team with the best, diversified portfolio.
+However, the intention to put together a diverse team alone is not enough.
+One needs to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions [\textbf{Fig.
+1}].
+To achieve this, it will be necessary to go beyond one's limited networks when inviting people to the organizing team, keynote speakers, program committee, session chairs, reviewers, and others.
+Ensuring diversity in each of these teams and panels is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups.
+If these groups aren't showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to turning every work and looking in unfamiliar places to bring in the talent.
+Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}, thus encouraging us to look further to find great people that are not routinely in the spotlight.
+We need to go beyond our narrow and often limited networks.
+So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for various aspects of the conference organization.
+The regional and local communities are also good sources to tap into.
+For example, for useR!
+2021, groups like AfricaR, R-Ladies, MiR, R Forwards, and LatinR were fundamental to reach people for the organizing team, potential attendees, and sponsors.
+
+\section*{Rule 4: Set registration rates commensurate with the registrant's cost of living}
+
+Conferences, even virtual ones, should have registration fees for two primary reasons.
+The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee.
+The second reason has to do with the human psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with registration /cite{eventbrite_ultimate_2017}[any publication to back this up?].
+On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates.
+Usually, there is a higher fee for people from the industry than for academia (~ proportionate to income).
+We should also consider a lower fee for non-profit organizations, government employees, or freelancers, and use conversion factors (for country of residence using data from International Comparison Program report of the World Bank \cite{arend_disparity_2019}).
+It is important to include discounts for students as well as for postdocs (early career trainees) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}.
+Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined.
+Disabilities may also increase the overall cost of conferences, especially in-person \cite{de_picker_rethinking_2020, irish_increasing_2020}; discounts would help reduce the financial burden for these factions too.
+
+Representation is an important aspect of diversity and inclusivity: see people like you, from your country, that speak your language is one of the best way to feel you belong, this is way the rates can not only being different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.
 
@@
 
-Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen more of a competition and attract people that might not need the money (e.g.
-their PI could pay for it), but could apply for it to enrich their CV.
-If the goal is to help the people in need, let's make that clear.
-Do not give it a fancy name.
-It is a fee waiver or a discount for the people in need.
-And the process for applicants should be simplified.
-People with low resources already have a hard time applying for loans, grants and scholarships in their lives.
-We should not make their lives harder by asking them to write long paragraphs to convince us that they deserve our support.
-Many people that cannot afford conferences do not even try to apply for waivers/scholarships because they think it would be hard to get them.
-We should trust that when they say they need it, it is because they do.
+Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
+If the goal is to help the people in need, the language needs to make that clear without shrouding it with fancy-sounding phrases.
+Ultimately, it is a fee waiver or a discount for the people who need additional financial support.
+Also, the process for applicants should be simplified.
+People with limited resources already have a hard time applying for loans, grants, and scholarships all their lives.
+We should not make the attendees' lives more challenging by asking for long essays to convince us that they deserve our support.
+\st{Many people that cannot afford international conferences do not even try to apply for waivers/scholarships because they are discouraged by the requirements, the process, and partly because they believe it would be hard for them to win financial/travel aid.} For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
+Also, the participant metadata (e.g., demographics, gender, career stage) is already a strong indicator of their potential need for aid.
+If the conference resources allow it, the organizers could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
 
-If the conference resources allow it, one could even take a step forwards and offer financial support: child care support, transportation fees (if in-person) or internet connection services (if virtual).
+\section*{Rule 6: Make the conference accessible to all, from start to end}
 
-\section*{Rule 6: Make the conference accessible to all}
+Conferences are usually inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
+Each of these groups should be considered at every stage of the planning process.
+Welcoming representatives from these groups into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule 2}).
+If the conference is in person, the venue must be accessible for people with mobility limitations.
+A quiet space for neurodivergent people and a space for breastfeeding should also be provided.
+Images used in communications about the conference should be accompanied by alternative text and videos with captions and sound.
+Platforms for conference registration, abstract submission -- and chat platforms, if the conference is virtual or hybrid -- should be screen-reader friendly and with low technology requirements (that do not have compatible hardware, software, and internet connection requirements to function, all this tested beforehand).
+Captioning for presentations should be available in more than one language if possible.
+If the conference is in-person, presenters should speak into a microphone to make it easier for captioners or interpreters to listen to them.
+Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
+Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., such as RMarkdown, HTML, or TeX files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed.
+Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.
 
-Conferences are usually inaccessible to people with hearing and visual impairments.
-They should be taken into account in every stage of the process.
-Welcoming them into the organizing team would allow them to take part in the decisions since the beginning, and remind us of inaccessible practices that need to improve.
-In social media communication, images should be accompanied with alternative text and videos with captions.
-Platforms for conference registration, abstract submission --and, if the conference is virtual or hybrid, chat platforms --should be screen-reader friendly (tested beforehand).
-Captioning for presentations should be available.
-If the conference is in-person, presenters should speak to the microphone to make it easier for captioners or interpreters to listen to them.
-Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged.
-Accessibility means, for example, that the speakers should also provide raw and accessible material to their talks, such as RMarkdown, html or tex files, and that, if they pre-record their presentation, to make sure that their video is shown and their face is visible so that people with hearing impairment can have the possibility to read their lips.
-Accessibility awards at the end of the conference are a good way to acknowledge the speakers who were mindful of inclusiveness when preparing their talks.
-Please find an example of accessibility guidelines at \url{https://user2021.r-project.org/participation/accessibility/}.
-
-Accessibility practices should not be left for the last minute.
-They require time and early decision-making.
-Conversely, inaccessible decisions are hard to undo; e.g.
-when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference.
+Accessibility practices are not afterthoughts that can be dealt with at the last minute.
+They require time and early decision-making \cite{irish_increasing_2020}.
+Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference.
 Bear in mind that most accessibility practices are beneficial to everyone.
@@
 
-\section*{Rule 7: Put more effort in promoting the conference among groups of people who are usually underrepresented}
+\section*{Rule 7: Take deliberate efforts to promote the conference among underrepresented groups}
 
-If some minorities are under represented in the conference, they may not feel part of the community or may not have the financial resources to attend.
-And effort must be made to promote the conference as welcoming for everyone and supportive for those who do not have the money %ast: or "the credentials" to be there.
-For the former, if part of the community has been historically discriminated, one should emphasize that they are particularly welcome in this event, and that the organization will make sure to make it a safe and inclusive environment for them.
-For the latter, fee waivers, financial support, and the easiness of the process should be advertised.
-People who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be rather supportive in this case.
+If some minoritized groups are underrepresented in the conference, they may not feel like they belong. 
+If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
+An effort must be made to promote the conference as welcoming and supportive for those without the necessary 'credentials' or resources.
+For the latter, fee waivers, financial support, and ease of application/registration should be advertised.
+People who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory against underrepresented groups.
+Providing financial and other support (e.g., childcare, flexible hours) will help minoritized groups partake more willingly in the organization and the actual conference.
+%ast likewise, some key people in your organizing team might not be able to collaborate without remuneration. Resources permitting, it may be a good idea to secure funds to offer stipends to them.
 
@@
 
-In-person interaction is priceless; however, it is more expensive for some, even unattainable for others.
-This is particularly true for global conferences.
-Global conferences usually take place in high-income countries, making it too expensive and for many people impossible to attend due to immigration requirements.
-[Share links to news of conference participants denied entry to countries, and maybe some narrating the whole absurd process for visas] Online conferences are more inclusive because there is not need for a visa or a big budget.
+In-person interaction is priceless; however, it is more expensive for some, even unattainable for others \cite{arend_disparity_2019, timperley_he_2020, niner_better_2021}.
+This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{sarabipour_evaluating_2020, gewin_what_2019}.
+Online conferences are more inclusive and do not need a visa or a big budget.
 
-Alternatively, a conference could have an in-person and an online component.
-This could allow for a group of people to physically interact, while providing many others the opportunity to participate.
-The challenge in this kind of setting would be to make the online component equally relevant as the in-person component, and not just a consolation price to the less privileged in the community.
+Alternatively, a conference could have a hybrid format with an in-person and online component.
+This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
+The challenge in this kind of setting would be to make the online component equally relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.
 
-\section*{Rule 9: Don't let English be a barrier for good quality participation}
+\section*{Rule 9: Don't let language restrict high-quality participation}
 
 In international conferences, English is often the official language.
-Submissions, presentations, tutorials and workshops are in English.
-While one official language is extremely helpful to communicate and English is the primary language in scientific communication, opening the conference to other languages could make it less intimidating to people who are not so fluent in English.
-Excluding them may also mean missing great innovative pieces of work, because of a language barrier.
-Advertising the conference in several languages, and considering having workshops and presentations in other languages (with or without captions in English) could help overcome this barrier.
+Submissions, presentations, tutorials, and workshops are in English.
+The platforms, the webpage, and official communications are also in English.
+While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{niner_better_2021}.
+Excluding them may potentially lead to missing innovative contributions brought about by a language barrier.
+Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier.
+For instance, hosting one international day/session per conference might be a great place to start!
 
-\section*{Rule 10: Have a code of conduct team}
+\section*{Rule 10: Have a code of conduct and a team to reinforce it}
 
-To maintain and inclusive and safe space in the conference, there should be a code of conduct and a team to enforce it.
+To ensure an inclusive, friendly, and safe space in the conference, the organizers need to a clearly define the code of conduct and a team to enforce it \cite{favaro_your_2016}.
 `The purpose of a code of conduct is to protect members of a community from harm in that community's spaces.
-The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' (coc book)
+The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' \cite{aurora_how_2018}
 The code of conduct should be displayed prominently in several spaces of the conference to deter people from incurring in unaccepted behavior.
 
-The code of conduct team should receive training to have a clear understanding on what to enforce and what not to enforce in a code of conduct, how to receive reports, respond to incidents, and communicate the responses.
-A diverse code of conduct team could be more understanding of intersectionality issues in discrimination and harassment practices.
+The code of conduct team should receive training to understand unambiguously what to and not to enforce in a code of conduct, how to receive reports, respond to incidents, and communicate their responses.
+A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.
 
@@
 
-The ten rules stated here can be adapted depending on the conference settings.
-We are aware, from our own experience, that enforcing these rules require a lot of work that is not compensated financially or in promotions.
-The rewards from this practice can be a healthier, stronger and more inclusive community, empowering people that are historically marginalized, and learning from them, thus increasing the quality of the conference.
+The ten rules stated here can be adapted depending on the conference format and settings.
+From our own experience, we are aware that enforcing these rules requires a tremendous amount of work and is not compensated financially or otherwise.
+However, the rewards from these suggested best practices will provide a healthier, stronger, and more inclusive community, empowering historically marginalized people and learning from them, thereby increasing the overall quality of the conference.
+Also, since these are early days and accessibility is not as widespread as it should be, the process might seem exacting.
+Hopefully, if more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
+After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.
 
 \section*{Acknowledgments}
-The authors of this piece would like to thank every single member of the organizing team of useR!2021 for their valuable contribution to an inclusive conference experience.
-And the R Foundation for their support to the conference.
+The authors of this piece would like to thank every single member of the organizing team of useR!
+2021 [insert full list of organizers <LINK>] for their valuable contribution to an inclusive conference experience, and the R Foundation for their financial support for useR!
+2021.
 
@@
 % % Compile your BiBTeX database using our plos2015.bst
+\bibliography{references}
 % % style file and paste the contents of your .bbl file

### 2021-08-21 plos_latex_template.tex
This was mostly reading Janani's, Heather's, and Matt's edits :)

@@
 
-In July 2021, the authors of this article, along with a larger group of people, organized a global and virtual conference for users and developers of the R software, useR!
-2021.
+In July 2021, the authors of this article, along with a larger group of people, organized the annual user conference of the R Project for Statistical Computing.
+This instance of the useR!
+conference was particularly well received as a virtual and explicitly global event, reaching users and developers of the R language from more than 120 countries.
 useR!
-conferences are strictly non-profit and are organized by volunteers from the R community and supported by the R Foundation.
-These global meetings include attendees from both academia and the industry.
-This year, we aimed at building a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-In this article, we share a few critical lessons learned during the past year's work organizing this conference.
-We highlight these as the 10 simple rules to host an inclusive conference.
-These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but can also be easily adapted and applied to regional or local conferences.
+conferences are non-profit events, organized by volunteers from the R community and supported by the R Foundation.
+The conference attracts a broad range of participants from academia, industry, government and the non-profit sector.
+For 2021, we aimed to build a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone.
+In this article, we share a few critical lessons learned in the process.
+We streamline our most important learnings in 10 simple rules to host an inclusive conference.
+These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
 
@@
 
-Conferences are spaces to meet and reconnect with members from a specific community, learn about new advances in the field and share our recent contributions.
+Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field and share our recent contributions.
 The larger the conference, the larger the opportunities to network and learn from your cohort.
-However, just like in science and academia in general, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arend_disparity_2019, timperley_he_2020, gewin_what_2019, brown_ableism_2018}.
+However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arend_disparity_2019, timperley_he_2020, gewin_what_2019, brown_ableism_2018}.
 
-Calls to improve diversity in many academic spaces, including conferences, are becoming frequent as communities and scientific societies are called out for overrepresenting specific groups of people.
-While these systematic issues are beginning to be addressed, there is scope for much improvement all around.
+Calls to improve diversity in many spaces, including conferences, are becoming more frequent as communities and scientific societies are getting called out for over-representing specific groups.
+While there has been some action to address systematic inequalities, there is still a lot of room for improvement.
 
-Diversity encompasses multiple dimensions: age, body ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
-There are implicit hierarchies along these axes, and some statuses (e.g., being cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
+Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
+There are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
 Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
@@
 
-This article suggests rules for pivoting traditional conferences towards inclusiveness and diversity and welcoming minoritized groups.
-These tips stem from the authors' experience in organizing a global and virtual conference for users and developers of the R programming language, useR!
+This article suggests rules to pivot traditional conferences towards inclusiveness and diversity and to welcome minoritized groups.
+These tips stem from the authors' experience of organizing a virtual and global conference for users and developers of the R programming language, useR!
 2021.
-We aimed at building a high-quality conference not just in terms of technical content but embracing everyone in a kind, inclusive, accessible, and welcoming environment, critical in the context of the COVID-19 pandemic.
-Most of the authors also have experience organizing other regional and national academic conferences and events in communities of practice such as R-Ladies and The Carpentries.
-Here, we share the lessons learned within the past year of organizing this global useR conference, summarized as 10 simple rules towards an inclusive conference.
+We embraced the challenge of organizing a high-quality conference in the context of the COVID-19 pandemic using the chance of a virtual conference environment to create a kind, truly inclusive, accessible, and welcoming experience for everyone.
+Most of the authors also have experience of  organizing other regional and national academic conferences and events in communities of practice such as R-Ladies and The Carpentries.
+Here, we share the lessons learned within the past year of organizing this global useR!
+conference, summarized as 10 simple rules towards an inclusive conference.
 
 \section*{Rule 1: Set realistic but bold diversity indicators}
-.
+
 As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors.
@@
 
-We found it greatly helpful to publish a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcomed, and that this is their space and community too.
-However, it is important to not stop with a statement alone, but identify specific goals and tasks that ensure compliance with these larger goals, for e.g., identifying accessible platforms, developing a detailed code of conduct, and accessibility guidelines (see other Rules below).
+Publishing a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too.
+However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct, and preparing accessibility guidelines (see other Rules below).
 
@@
 The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best platforms to mechanisms to account for every culture and situation.
-A successful well-oiled inclusive team will provide a safe space for everyone not just to learn and grow but identify critical aspects that are working or failing to improve inclusiveness and accessibility together \cite{hong_groups_2004}.
+A successful, inclusive team will not only be a safe space for people to learn and grow but also may help identify critical aspects that are working or failing to improve inclusiveness and accessibility together \cite{hong_groups_2004}.
 
@@
 
-\section*{Rule 3: Finding the best people for each task}
+\section*{Rule 3: Search proactively the best people for each task}
 
@@
 
+Disabled people often repeat "Nothing about us without us".
+This means the actual life experiences of people from URM are not replaceable with good intentions from people outside these groups.
+The same holds for other dimensions of diversity.
+No space will be truly inclusive and welcoming to URM without people from these groups -and their experiences-.
+This means people in privileged groups may need to step down from positions of power.
+Of course, no-one has the same experiences and experiences the same oppressions in life.
+When building a diverse team, bear in mind that
+
 \section*{Rule 4: Set registration rates commensurate with the registrant's cost of living}
@@
 The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee.
-The second reason has to do with the human psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with registration /cite{eventbrite_ultimate_2017}[any publication to back this up?].
+The second reason has to do with the human psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}.
 On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates.
@@
 
-\section*{Rule 5: Do not give scholarships or grants to attend. Give fee waivers or discounts without tedious application forms.}
+\section*{Rule 5: Abandon diversity scholarships}
 
+%rationale: 1.
+they might select for the already privileged or people who dont need those to be nice on the cvs 2.
+they select for outgoing extroverted privileged ("brave") people 3.
+they are time and emotionally consuming which selects out some people 4.
+they single out people.
+even if you controlled for 1 to 3 you would "badge" people - you are creating a minority while trying to avoid it.
 Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
-If the goal is to help the people in need, the language needs to make that clear without shrouding it with fancy-sounding phrases.
+If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases.
 Ultimately, it is a fee waiver or a discount for the people who need additional financial support.
 Also, the process for applicants should be simplified.
-People with limited resources already have a hard time applying for loans, grants, and scholarships all their lives.
+People with limited resources already have a hard time applying for loans, grants, and scholarships.
 We should not make the attendees' lives more challenging by asking for long essays to convince us that they deserve our support.
@@
 The authors of this piece would like to thank every single member of the organizing team of useR!
-2021 [insert full list of organizers <LINK>] for their valuable contribution to an inclusive conference experience, and the R Foundation for their financial support for useR!
+2021 [insert full list of organizers \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for their financial support for useR!
 2021.

### 2021-08-21 plos_latex_template.tex
accepted most changes from Matt, Heather, Janani, Andrea

@@
 % % Please keep the Author Summary between 150 and 200 words
-% % Use first person. PLOS ONE authors please skip this step. 
-% % Author Summary not valid for PLOS ONE submissions.   
-% \section*{Author summary (I don't think we need it)}
-% Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
@@
 
-Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
-There are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
-Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
-
-%some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}
-
 This article suggests rules to pivot traditional conferences towards inclusiveness and diversity and to welcome minoritized groups.
@@
 
-\section*{Rule 1: Set realistic but bold diversity indicators}
+\section*{Rule 1: Embrace all dimensions of diversity}
+
+The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.
+
+Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
+There are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
+Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
 
 As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors.
-Next, set measurable indicators of success when organizing a diverse conference.
-The numbers resulting from these indicators are not the primary goal, and this is not about performative allyship.
-Instead, these numbers should guide and help hold the organizing team accountable towards these diverse and inclusive goals.
+Next, think about set measurable indicators of success when organizing a diverse conference.
+The numbers resulting from these indicators are not the primary goal.
+Instead, this vision should guide and help hold the organizing team accountable towards these diverse and inclusive goals.
 
@@
 For instance, if the organizing team is global and remote, time zones can be hard to manage, with meetings scheduled at odd hours.
-The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best platforms to mechanisms to account for every culture and situation.
-A successful, inclusive team will not only be a safe space for people to learn and grow but also may help identify critical aspects that are working or failing to improve inclusiveness and accessibility together \cite{hong_groups_2004}.
+The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation.
+An inclusive team will let everyone know if the organization is failing in any aspect of inclusiveness or accessibility, and help improve it \cite{hong_groups_2004}.
 
@@
 While this rule may seem intuitive, one would be surprised by how often conference organizers tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in.
-For e.g., a cisgender or non-neurodiverse person estimating what might work best for the LGBTQ+ or people with hearing and seeing difficulty.
-Therefore, to build your team of organizers for the conference, you need to actively seek out and recruit the best people for each task, akin to filling your team with the best, diversified portfolio.
-However, the intention to put together a diverse team alone is not enough.
+For example, a cisgender or abled person estimating what might work best for transgender people or people with sensory disabilities.
+Therefore, to build your team of organizers for the conference, you need to actively seek out and recruit the best people for each task.
+The intention to put together a diverse team alone is not enough.
 One needs to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions [\textbf{Fig.
 1}].
-To achieve this, it will be necessary to go beyond one's limited networks when inviting people to the organizing team, keynote speakers, program committee, session chairs, reviewers, and others.
+To achieve this, it will be necessary to go beyond one's limited networks when inviting people to the organizing team, keynote speakers, program committee, session chairs, reviewers, and other roles.
 Ensuring diversity in each of these teams and panels is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups.
@@
 For example, for useR!
-2021, groups like AfricaR, R-Ladies, MiR, R Forwards, and LatinR were fundamental to reach people for the organizing team, potential attendees, and sponsors.
+2021, groups like AfricaR, R-Ladies, MiR, Forwards, and LatinR were fundamental to reach people for the organizing team, potential attendees, and sponsors.
 
@@
 The same holds for other dimensions of diversity.
-No space will be truly inclusive and welcoming to URM without people from these groups -and their experiences-.
+No space will be truly inclusive and welcoming to URM without people from these groups bringing their experience to bear.
 This means people in privileged groups may need to step down from positions of power.
 Of course, no-one has the same experiences and experiences the same oppressions in life.
-When building a diverse team, bear in mind that
+%When building a diverse team, bear in mind that
 
@@
 The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee.
-The second reason has to do with the human psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}.
+The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}.
 On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates.
-Usually, there is a higher fee for people from the industry than for academia (~ proportionate to income).
+Usually, there is a higher fee for people from the industry than for academia ($\sim$ proportionate to income).
 We should also consider a lower fee for non-profit organizations, government employees, or freelancers, and use conversion factors (for country of residence using data from International Comparison Program report of the World Bank \cite{arend_disparity_2019}).
-It is important to include discounts for students as well as for postdocs (early career trainees) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}.
+It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}.
 Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined.
@@
 
-Representation is an important aspect of diversity and inclusivity: see people like you, from your country, that speak your language is one of the best way to feel you belong, this is way the rates can not only being different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.
+Representation is an important aspect of diversity and inclusivity: seeing people like you, from your country, that speak your language is one of the best way to feel you belong, this is why the rates can not only be different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.
 
@@
 
-Conferences are usually inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
+Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
+People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
 Each of these groups should be considered at every stage of the planning process.
@@
 If the conference is in person, the venue must be accessible for people with mobility limitations.
-A quiet space for neurodivergent people and a space for breastfeeding should also be provided.
-Images used in communications about the conference should be accompanied by alternative text and videos with captions and sound.
-Platforms for conference registration, abstract submission -- and chat platforms, if the conference is virtual or hybrid -- should be screen-reader friendly and with low technology requirements (that do not have compatible hardware, software, and internet connection requirements to function, all this tested beforehand).
+Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided.
+Images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
+Platforms for conference registration and abstract submission, websites and chat platforms should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
+These aspects should be tested well in advance of going live.
 Captioning for presentations should be available in more than one language if possible.
-If the conference is in-person, presenters should speak into a microphone to make it easier for captioners or interpreters to listen to them.
+If the conference is in-person, presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them.
 Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
-Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., such as RMarkdown, HTML, or TeX files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed.
+Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed.
+Adding captions to recorded videos is relatively easy these days and is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
 Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.
@@
 
-\section*{Rule 8: Have an important online component of the conference}
+\section*{Rule 8: Have a strong online component of the conference}
 
@@
 While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{niner_better_2021}.
-Excluding them may potentially lead to missing innovative contributions brought about by a language barrier.
+Excluding them may potentially lead to missing innovative contributions due to a language barrier.
 Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier.
@@
 
-To ensure an inclusive, friendly, and safe space in the conference, the organizers need to a clearly define the code of conduct and a team to enforce it \cite{favaro_your_2016}.
-`The purpose of a code of conduct is to protect members of a community from harm in that community's spaces.
-The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' \cite{aurora_how_2018}
-The code of conduct should be displayed prominently in several spaces of the conference to deter people from incurring in unaccepted behavior.
+To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}.
+The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.'
+The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
 
-The code of conduct team should receive training to understand unambiguously what to and not to enforce in a code of conduct, how to receive reports, respond to incidents, and communicate their responses.
+The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses.
 A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.
@@
 The ten rules stated here can be adapted depending on the conference format and settings.
-From our own experience, we are aware that enforcing these rules requires a tremendous amount of work and is not compensated financially or otherwise.
+From our own experience, we are aware that enforcing these rules requires a tremendous amount of work that is generally not compensated financially or otherwise.
 However, the rewards from these suggested best practices will provide a healthier, stronger, and more inclusive community, empowering historically marginalized people and learning from them, thereby increasing the overall quality of the conference.
-Also, since these are early days and accessibility is not as widespread as it should be, the process might seem exacting.
+Since these are early days and accessibility is not as widespread as it should be, the process might seem exacting.
 Hopefully, if more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
@@
 The authors of this piece would like to thank every single member of the organizing team of useR!
-2021 [insert full list of organizers \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for their financial support for useR!
-2021.
+2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR!
+2021 and supporting us through the process.
 

### 2021-08-22 plos_latex_template.tex
Edited online rule

@@
 
-\section*{Rule 1: Embrace all dimensions of diversity}
+\section{Rule X: Define what diversity and inclusion mean for your conference}
 
@@
 
-As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors.
+As a starting point, identify which groups are usually unsupported or discriminated against in your community or event, and determine the contributing factors.
 Next, think about set measurable indicators of success when organizing a diverse conference.
@@
 
-\section*{Rule 2: Have an inclusive and diverse organizing team}
+\section{Rule X: Have a strong online component of the conference}
+
+In-person interaction is priceless; however, it is more expensive for some, even unattainable for others.
+This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arend_disparity_2019, gewin_what_2019}.
+Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
+Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipour_evaluating_2020, niner_better_2021}; in a study case, conference attendance accounted for $35\%$ of a researchers footprint .
+
+Alternatively, a conference could have a hybrid format with an in-person and online component.
+This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
+The challenge in this kind of setting would be to make the online component equally relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.
+
+\section{Rule X: Have an inclusive and diverse organizing team}
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
@@
 
-\section*{Rule 3: Search proactively the best people for each task}
+\section{Rule X: Search proactively the best people for each task}
 
@@
 
-\section*{Rule 4: Set registration rates commensurate with the registrant's cost of living}
+\section{Rule X: Set registration rates commensurate with the registrant's cost of living}
 
@@
 
-\section*{Rule 5: Abandon diversity scholarships}
+\section{Rule X: Abandon diversity scholarships}
 
@@
 
-\section*{Rule 6: Make the conference accessible to all, from start to end}
+\section{Rule X: Make the conference accessible to all, from start to end}
 
@@
 
-\section*{Rule 7: Take deliberate efforts to promote the conference among underrepresented groups}
+\section{Rule X: Take deliberate efforts to promote the conference among underrepresented groups}
 
@@
 
-\section*{Rule 8: Have a strong online component of the conference}
-
-In-person interaction is priceless; however, it is more expensive for some, even unattainable for others \cite{arend_disparity_2019, timperley_he_2020, niner_better_2021}.
-This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{sarabipour_evaluating_2020, gewin_what_2019}.
-Online conferences are more inclusive and do not need a visa or a big budget.
-
-Alternatively, a conference could have a hybrid format with an in-person and online component.
-This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-The challenge in this kind of setting would be to make the online component equally relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.
-
-\section*{Rule 9: Don't let language restrict high-quality participation}
+\section{Rule X: Don't let language restrict high-quality participation}
 
@@
 
-\section*{Rule 10: Have a code of conduct and a team to reinforce it}
+\section{Rule X: Have a code of conduct and a team to reinforce it}
 

### 2021-08-22 plos_latex_template.tex
RJ edited organizing team and search for people rules

@@
 This article suggests rules to pivot traditional conferences towards inclusiveness and diversity and to welcome minoritized groups.
-These tips stem from the authors' experience of organizing a virtual and global conference for users and developers of the R programming language, useR!
-2021.
+These tips stem from the authors' experience of organizing useR!
+2021, a virtual and global conference for users and developers of the R programming language \cite{r_core_team_2021}.
 We embraced the challenge of organizing a high-quality conference in the context of the COVID-19 pandemic using the chance of a virtual conference environment to create a kind, truly inclusive, accessible, and welcoming experience for everyone.
@@
 As a starting point, identify which groups are usually unsupported or discriminated against in your community or event, and determine the contributing factors.
+
 Next, think about set measurable indicators of success when organizing a diverse conference.
@@
 Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
+This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinson_r_medicine_2021}.
 Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipour_evaluating_2020, niner_better_2021}; in a study case, conference attendance accounted for $35\%$ of a researchers footprint .
@@
 To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
-Even if creating and maintaining such an interdisciplinary team is challenging, the positive outcomes far outweigh any minor inconveniences, and the team will grow, learn, and embrace more inclusive practices.
-For instance, if the organizing team is global and remote, time zones can be hard to manage, with meetings scheduled at odd hours.
-The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation.
-An inclusive team will let everyone know if the organization is failing in any aspect of inclusiveness or accessibility, and help improve it \cite{hong_groups_2004}.
+But this will only work if there is real inclusion.
+One would be surprised by how often conference leads tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in.
+For example, a cisgender or abled person estimating what might work best for transgender people or people with sensory disabilities.
+The same holds for other dimensions of diversity.
+A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.
+Even if creating and maintaining such a team and space is challenging, the positive outcomes far outweigh any minor inconveniences (e.g.
+time zones), and the team will grow, learn, and embrace more inclusive practices \cite{hong_groups_2004}.
+
+%The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation.
+% rj: left this out, Heather said it doesn't make much sense here
 
@@
 
-\section{Rule X: Search proactively the best people for each task}
+\section{Rule X: Make the conference accessible to all, from start to end}
 
-While this rule may seem intuitive, one would be surprised by how often conference organizers tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in.
-For example, a cisgender or abled person estimating what might work best for transgender people or people with sensory disabilities.
-Therefore, to build your team of organizers for the conference, you need to actively seek out and recruit the best people for each task.
-The intention to put together a diverse team alone is not enough.
-One needs to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions [\textbf{Fig.
-1}].
-To achieve this, it will be necessary to go beyond one's limited networks when inviting people to the organizing team, keynote speakers, program committee, session chairs, reviewers, and other roles.
-Ensuring diversity in each of these teams and panels is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups.
-If these groups aren't showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to turning every work and looking in unfamiliar places to bring in the talent.
-Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}, thus encouraging us to look further to find great people that are not routinely in the spotlight.
-We need to go beyond our narrow and often limited networks.
-So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for various aspects of the conference organization.
+Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
+People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
+Each of these groups should be considered at every stage of the planning process.
+Welcoming people from these groups into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule 2}).
+If the conference is in person, the venue must be accessible for people with mobility limitations.
+Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided.
+Images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
+Platforms for conference registration and abstract submission, websites and chat platforms should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
+These aspects should be tested well in advance of going live.
+Captioning for presentations should be available in more than one language if possible.
+If the conference is in-person, presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them.
+Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
+Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed.
+Adding captions to recorded videos is relatively easy these days and is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
+Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.
+
+Accessibility practices are not afterthoughts that can be dealt with at the last minute.
+They require time and early decision-making \cite{irish_increasing_2020}.
+Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference.
+Bear in mind that most accessibility practices are beneficial to everyone.
+Captions, for instance, are used by non-native speakers and people that access the conference in loud environments.
+The prior availability of material benefits people with low bandwidth and non-native speakers as well.
+
+\section{Rule X: Search proactively the best people for invited roles}
+
+When inviting people as keynote speakers, program committee, session chairs and others, we should always look for the best in their subject.
+Ensuring diversity in each of these roles is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups.
+If these groups are not showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to looking in unfamiliar places to bring in the talent.
+Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}.
+Rather than deter us, this should thus encourage us to look further to find great people that are not routinely in the spotlight.
+We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
+
+% So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for various aspects of the conference organization.
 The regional and local communities are also good sources to tap into.
 For example, for useR!
-2021, groups like AfricaR, R-Ladies, MiR, Forwards, and LatinR were fundamental to reach people for the organizing team, potential attendees, and sponsors.
+2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential attendees, and sponsors.
 
-Disabled people often repeat "Nothing about us without us".
+%Disabled people often repeat "Nothing about us without us".
 This means the actual life experiences of people from URM are not replaceable with good intentions from people outside these groups.
@@
 If the conference resources allow it, the organizers could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
-
-\section{Rule X: Make the conference accessible to all, from start to end}
-
-Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
-People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
-Each of these groups should be considered at every stage of the planning process.
-Welcoming representatives from these groups into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule 2}).
-If the conference is in person, the venue must be accessible for people with mobility limitations.
-Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided.
-Images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
-Platforms for conference registration and abstract submission, websites and chat platforms should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
-These aspects should be tested well in advance of going live.
-Captioning for presentations should be available in more than one language if possible.
-If the conference is in-person, presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them.
-Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
-Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed.
-Adding captions to recorded videos is relatively easy these days and is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
-Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.
-
-Accessibility practices are not afterthoughts that can be dealt with at the last minute.
-They require time and early decision-making \cite{irish_increasing_2020}.
-Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference.
-Bear in mind that most accessibility practices are beneficial to everyone.
-Captions, for instance, are used by non-native speakers and people that access the conference in loud environments.
-The prior availability of material benefits people with low bandwidth and non-native speakers as well.
 

### 2021-08-22 plos_latex_template.tex
RJ: edited accessibility rule

@@
 
-Next, think about set measurable indicators of success when organizing a diverse conference.
+Next, set measurable indicators of success when organizing a diverse conference.
 The numbers resulting from these indicators are not the primary goal.
@@
 \section{Rule X: Have an inclusive and diverse organizing team}
+\label{organizing_team}
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
@@
 People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
-Each of these groups should be considered at every stage of the planning process.
-Welcoming people from these groups into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule 2}).
+Inaccessible practices only exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people.
+For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
+
 If the conference is in person, the venue must be accessible for people with mobility limitations. 
 Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided.
-Images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
-Platforms for conference registration and abstract submission, websites and chat platforms should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
+Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
+Regardless of the conference format --online, in person or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
+Platforms for conference registration and abstract submission, websites and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
+Captioning for presentations should be available in more than one language if possible.
 These aspects should be tested well in advance of going live.
-Captioning for presentations should be available in more than one language if possible.
-If the conference is in-person, presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them.
-Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
-Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed.
-Adding captions to recorded videos is relatively easy these days and is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
+
+Some practices, such as making accessible slides and presentations, will require great efforts from the presenters if they are not used to them. 
+For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have.  
+Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
 Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.
 
-Accessibility practices are not afterthoughts that can be dealt with at the last minute.
+Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
 They require time and early decision-making \cite{irish_increasing_2020}. 
 Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
-Bear in mind that most accessibility practices are beneficial to everyone.
-Captions, for instance, are used by non-native speakers and people that access the conference in loud environments.
-The prior availability of material benefits people with low bandwidth and non-native speakers as well.
+Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{organizing_team}}).
 

### 2021-08-22 plos_latex_template.tex
RJ: minor edits to CoC rule
RJ: uniting fair registration rates with scholarship rules

@@
 
+\section{Rule X: Have a code of conduct and a team to reinforce it}
+
+To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}.
+The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.'
+The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
+
+The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses.
+A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.
+
 \section{Rule X: Search proactively the best people for invited roles}
@@
 
-\section{Rule X: Set registration rates commensurate with the registrant's cost of living}
+\section{Rule X: Set fair registration rates and provide financial support when needed}
 
@@
 The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}.
+
 On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates.
 Usually, there is a higher fee for people from the industry than for academia ($\sim$ proportionate to income).
-We should also consider a lower fee for non-profit organizations, government employees, or freelancers, and use conversion factors (for country of residence using data from International Comparison Program report of the World Bank \cite{arend_disparity_2019}).
+A lower fee for non-profit organizations, government employees, or freelancers should also be considered.
 It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}.
 Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined.
-Disabilities may also increase the overall cost of conferences, especially in-person \cite{de_picker_rethinking_2020, irish_increasing_2020}; discounts would help reduce the financial burden for these factions too.
+You can take into account the cost of living in their countries using conversion factors (e.g.
+from the International Comparison Program report of the World Bank \cite{arend_disparity_2019}).
 
-Representation is an important aspect of diversity and inclusivity: seeing people like you, from your country, that speak your language is one of the best way to feel you belong, this is why the rates can not only be different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.
-
-\section{Rule X: Abandon diversity scholarships}
+%Representation is an important aspect of diversity and inclusivity: seeing people like you, from your country, that speak your language is one of the best way to feel you belong, this is why the rates can not only be different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.
 
@@
 Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
-If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases.
+If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or tokenizing phrases (e.g.
+`grant recipient' or `diversity scholar', respectively).
+Do not call it a scholarship or a grant.
 Ultimately, it is a fee waiver or a discount for the people who need additional financial support.
@@
 People with limited resources already have a hard time applying for loans, grants, and scholarships.
-We should not make the attendees' lives more challenging by asking for long essays to convince us that they deserve our support.
-\st{Many people that cannot afford international conferences do not even try to apply for waivers/scholarships because they are discouraged by the requirements, the process, and partly because they believe it would be hard for them to win financial/travel aid.} For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
-Also, the participant metadata (e.g., demographics, gender, career stage) is already a strong indicator of their potential need for aid.
-If the conference resources allow it, the organizers could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
+Do not make the attendees' lives more challenging by asking for long essays to convince a committee that they deserve your support.
+For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
+Also, the participant metadata (e.g., demographics, gender, career stage, disabilities, if they disclose them) is already a strong indicator of their potential need for aid.
+If the conference resources allow it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
 
@@
 For instance, hosting one international day/session per conference might be a great place to start!
-
-\section{Rule X: Have a code of conduct and a team to reinforce it}
-
-To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}.
-The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.'
-The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
-
-The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses.
-A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.
 

### 2021-08-22 plos_latex_template.tex
RJ: Edits to communication rule

@@
 \section{Rule X: Have an inclusive and diverse organizing team}
-\label{organizing_team}
+\label{rule_organizing_team}
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
@@
 Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
-Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{organizing_team}}).
+Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}).
 
@@
 
-\section{Rule X: Take deliberate efforts to promote the conference among underrepresented groups}
-
-If some minoritized groups are underrepresented in the conference, they may not feel like they belong. 
-If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
-An effort must be made to promote the conference as welcoming and supportive for those without the necessary 'credentials' or resources.
-For the latter, fee waivers, financial support, and ease of application/registration should be advertised.
-People who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory against underrepresented groups.
-Providing financial and other support (e.g., childcare, flexible hours) will help minoritized groups partake more willingly in the organization and the actual conference.
-%ast likewise, some key people in your organizing team might not be able to collaborate without remuneration. Resources permitting, it may be a good idea to secure funds to offer stipends to them.
-
 \section{Rule X: Don't let language restrict high-quality participation}
+\label{rule_language}
 
@@
 For instance, hosting one international day/session per conference might be a great place to start!
+
+\section{Rule X: Take deliberate efforts to promote the conference among underrepresented groups}
+
+If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
+Publish a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines.
+Fee waivers, financial support, and ease of application/registration should also be advertised; people who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory against underrepresented groups.
+Advertise the conference in multiple languages (rule \ref{rule_language}).
+Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
+
+%Providing financial and other support (e.g., childcare, flexible hours) will help minoritized groups partake more willingly in the organization and the actual conference.
+%ast likewise, some key people in your organizing team might not be able to collaborate without remuneration. Resources permitting, it may be a good idea to secure funds to offer stipends to them.
 

### 2021-08-23 plos_latex_template.tex
RJ: added money rule

@@
 
+\section{Rule X: Do not expect the organizing team to work full time and for free}
+
+Having a diverse team and executing inclusive and accessible practices throughout the organizing period --it might be a year or two --may require a lot of effort.
+The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone.
+However, having a strong community as the only reward may be enough if you are in a privileged position.
+Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free.
+In addition, tasks such as taking and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
+Be mindful of each particular context, be flexible with hours or commitments, and revise your budget; if possible, secure funds to offer stipends to them early on.
+
 \section*{Concluding remarks}

### 2021-08-23 plos_latex_template.tex
RJ: edits to the conclusion

@@
 The ten rules stated here can be adapted depending on the conference format and settings.
-From our own experience, we are aware that enforcing these rules requires a tremendous amount of work that is generally not compensated financially or otherwise.
-However, the rewards from these suggested best practices will provide a healthier, stronger, and more inclusive community, empowering historically marginalized people and learning from them, thereby increasing the overall quality of the conference.
-Since these are early days and accessibility is not as widespread as it should be, the process might seem exacting.
-Hopefully, if more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
-After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.
+From our own experience, we are aware that enforcing these rules requires a tremendous amount of work.
+We engaged in most of these practices, and others, like having an explicit payment category for postdocs or paying the whole organizing team, we learned them too late, so we share them here as part of our learning.
+Since accessibility and inclusiveness are not as widespread as it should be, the process might seem exacting.
+But this is where the power of communities and individuals reside.
+If more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
+And we will make a change towards healthier, stronger, and more inclusive communities, empowering historically marginalized people and learning from them, thereby increasing the overall quality of our conferences.%After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.
 

### 2021-08-23 plos_latex_template.tex
ast reading and commenting and starting to rewrite rule1 :cookie:

@@
 
-Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field and share our recent contributions.
+Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
 The larger the conference, the larger the opportunities to network and learn from your cohort.
@@
 \section{Rule X: Define what diversity and inclusion mean for your conference}
-
+%
 The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.
@@
 
-As a starting point, identify which groups are usually unsupported or discriminated against in your community or event, and determine the contributing factors.
+Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context.
+Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your definition.
+Does this mean an even gender distribution in your speakers?
+Does this relate to people of all races -especially black people- being present as coordinators, presenters, and in the audience?
+Does this mean having LGBTQ+ friendly-spaces or community participation from key geographic regions?
+This vision should guide and help hold the organizing team accountable along the way.
 
-Next, set measurable indicators of success when organizing a diverse conference.
-The numbers resulting from these indicators are not the primary goal.
-Instead, this vision should guide and help hold the organizing team accountable towards these diverse and inclusive goals.
+Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}).
+%and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too.
+However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct, and preparing accessibility guidelines (see other Rules below).
 
-Publishing a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too.
-However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct, and preparing accessibility guidelines (see other Rules below).
+%Lack of representation, unwelcoming -or overtly aggressive- environments hinder participation (or future participation) of people who could otherwise become active community members.
+In extreme cases, they can divert career paths, affect lives, and exclude people from some fields.
+So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn.
+%Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator.
+%ast: some version of this might come in a box.
 
@@
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-The challenge in this kind of setting would be to make the online component equally relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.
+The challenge in this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.
 
@@
 %The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation.
-% rj: left this out, Heather said it doesn't make much sense here
+% rj: left this out, Heather said it doesn't make much sense here %ast: we can collect part of this in rule 1, since some self-teaching, and not expecting to be taught, is necessary
 
+%%Disabled people often repeat "Nothing about us without us", and this is true for every marginalized group.
+This means the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups.
+No space will be truly inclusive and welcoming without people from these groups bringing their experience to bear.
+This means people in privileged groups may need to step down from positions of power.
+%this might
 %rj: Yani's refs to have just in case:
@@
 Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
-People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
-Inaccessible practices only exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people.
+%People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
+Inaccessible practices exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people.
 For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
@@
 Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
-Regardless of the conference format --online, in person or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
-Platforms for conference registration and abstract submission, websites and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
+Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
+Platforms for conference registration and abstract submission, websites, and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
 Captioning for presentations should be available in more than one language if possible.
@@
 
-Some practices, such as making accessible slides and presentations, will require great efforts from the presenters if they are not used to them. 
+Some practices, such as making accessible slides and presentations, are not yet common practice and will require great efforts from the presenters if they are not used to them. 
 For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have.  
@@
 To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}.
-The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.'
+The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.
 The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
@@
 A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.
+%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.'
 
@@
 Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}.
-Rather than deter us, this should thus encourage us to look further to find great people that are not routinely in the spotlight.
+Rather than deter us, this should encourage us to look further to find great people that are not routinely in the spotlight.
 We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
@@
 
-%Disabled people often repeat "Nothing about us without us".
-This means the actual life experiences of people from URM are not replaceable with good intentions from people outside these groups.
-The same holds for other dimensions of diversity.
-No space will be truly inclusive and welcoming to URM without people from these groups bringing their experience to bear.
-This means people in privileged groups may need to step down from positions of power.
-Of course, no-one has the same experiences and experiences the same oppressions in life.
-%When building a diverse team, bear in mind that
-
 \section{Rule X: Set fair registration rates and provide financial support when needed}
@@
 The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee.
-The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}.
+The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events than for events with registration costs \cite{eventbrite_ultimate_2017}.
 
 On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates.
-Usually, there is a higher fee for people from the industry than for academia ($\sim$ proportionate to income).
+Usually, there is a higher fee for people from the industry than for academia.
 A lower fee for non-profit organizations, government employees, or freelancers should also be considered.
@@
 Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined.
-You can take into account the cost of living in their countries using conversion factors (e.g.
+You can take into account the cost of living in each country using conversion factors (e.g.
 from the International Comparison Program report of the World Bank \cite{arend_disparity_2019}).
+Consider that, while it is traditional for employers to provide --at least some --conference support in the Global North, this is not the case everywhere.
 
@@
 Also, the process for applicants should be simplified.
-People with limited resources already have a hard time applying for loans, grants, and scholarships.
-Do not make the attendees' lives more challenging by asking for long essays to convince a committee that they deserve your support.
+Applying for loans, grants, and scholarships is already an emotionally demanding task.
+Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support.
 For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
-Also, the participant metadata (e.g., demographics, gender, career stage, disabilities, if they disclose them) is already a strong indicator of their potential need for aid.
-If the conference resources allow it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
+\st{Also, the participant metadata (e.g., demographics, gender, career stage, disabilities, if they disclose them) is already a strong indicator of their potential need for aid.} Other If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
 
@@
 From our own experience, we are aware that enforcing these rules requires a tremendous amount of work.
-We engaged in most of these practices, and others, like having an explicit payment category for postdocs or paying the whole organizing team, we learned them too late, so we share them here as part of our learning.
-Since accessibility and inclusiveness are not as widespread as it should be, the process might seem exacting.
+We engaged in most of these practices, and learned others too late, like having an explicit payment category for postdocs or paying the whole organizing team, so we share them here as part of our learning process.
+Since accessibility and inclusiveness are not as widespread as they should be, the process might seem exacting.
 But this is where the power of communities and individuals reside.
@@
 2021 and supporting us through the process.
-
-% Cras egestas velit mauris, eu mollis turpis pellentesque sit amet.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Nam id pretium nisi.
-Sed ac quam id nisi malesuada congue.
-Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.
 

### 2021-08-25 plos_latex_template.tex
RJ: Added who the paper is for in the intro

@@
 
-In July 2021, the authors of this article, along with a larger group of people, organized the annual user conference of the R Project for Statistical Computing.
-This instance of the useR!
-conference was particularly well received as a virtual and explicitly global event, reaching users and developers of the R language from more than 120 countries.
+In July 2021, the authors of this article participated in the organization team of the annual user conference of the R Project for Statistical Computing.
 useR!
-conferences are non-profit events, organized by volunteers from the R community and supported by the R Foundation.
-The conference attracts a broad range of participants from academia, industry, government and the non-profit sector.
-For 2021, we aimed to build a high-quality conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-In this article, we share a few critical lessons learned in the process.
+conferences are non-profit events, organized by volunteers from the R community and arranged by the R Foundation.
+The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector.
+For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone.
+In this article, we share a few lessons learned in the process.
 We streamline our most important learnings in 10 simple rules to host an inclusive conference.
 These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
+
+%ast: matt's original sentence: reaching users and developers of the R language from more than 120 countries.
 
@@
 
-Calls to improve diversity in many spaces, including conferences, are becoming more frequent as communities and scientific societies are getting called out for over-representing specific groups.
+%ast will write a paragraph about why diversity and the systemic inequalities
 While there has been some action to address systematic inequalities, there is still a lot of room for improvement.
@@
 This article suggests rules to pivot traditional conferences towards inclusiveness and diversity and to welcome minoritized groups.
+It is directed as people who are part of a stable meetings commitee that oversees the site/location selection process or that coordinates with the local organizers (e.g.
+R Foundation Conference Committee, Ecological Society of America Meetings Committee).
+It is also directed at the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
+
 These tips stem from the authors' experience of organizing useR!
 2021, a virtual and global conference for users and developers of the R programming language \cite{r_core_team_2021}.
-We embraced the challenge of organizing a high-quality conference in the context of the COVID-19 pandemic using the chance of a virtual conference environment to create a kind, truly inclusive, accessible, and welcoming experience for everyone.
+We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive, accessible, and welcoming experience for everyone.
 Most of the authors also have experience of organizing other regional and national academic conferences and events in communities of practice such as R-Ladies and The Carpentries.
@@
 \section{Rule X: Define what diversity and inclusion mean for your conference}
-%
+\label{rule_diversity}
+
 The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.
@@
 
+To accomplish this diversity vision, you will have to take some conscious decisions toward removing bias from your conference's spaces.
+Having diverse people in decision-making positions will affect positively all the other aspects of your conference, not only internally because all the decision-making process will benefit from their input \ref{rule_organizing_team}, but also externally because representation matters \ref{rule_communication}.
+
+While you may want to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects as well.
+Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, are some examples of decisions that can make inclusion real.
+%Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
 Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}).
 %and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too.
-However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct, and preparing accessibility guidelines (see other Rules below).
+However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
 
-%Lack of representation, unwelcoming -or overtly aggressive- environments hinder participation (or future participation) of people who could otherwise become active community members.
+Lack of representation, unwelcoming -or overtly aggressive- environments hinder participation (or future participation) of people who could otherwise become active community members.
 In extreme cases, they can divert career paths, affect lives, and exclude people from some fields.
 So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn.
-%Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator.
+Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator.
 %ast: some version of this might come in a box.
@@
 \section{Rule X: Have a strong online component of the conference} 
-
+\label{rule_online}
 In-person interaction is priceless; however, it is more expensive for some, even unattainable for others.
@@
 This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinson_r_medicine_2021}.
-Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipour_evaluating_2020, niner_better_2021}; in a study case, conference attendance accounted for $35\%$ of a researchers footprint .
+Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipour_evaluating_2020, niner_better_2021}; in a study case, conference attendance accounted for $35\%$ of a researcher's footprint \cite{achten_carbon_2013}.
 
@@
 Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity.
-Particular attention should be paid to the usually underrepresented groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event).
+Particular attention should be paid to the usually marginalized groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event).
 To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
@@
 This means the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups.
-No space will be truly inclusive and welcoming without people from these groups bringing their experience to bear.
-This means people in privileged groups may need to step down from positions of power.
-%this might
 %rj: Yani's refs to have just in case:
@@
 \section{Rule X: Make the conference accessible to all, from start to end}
-
+\label{rule_accessibility}
 Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
@@
 If the conference is in person, the venue must be accessible for people with mobility limitations. 
-Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided.
+Accommodations such as a quiet space for neurodivergent people should also be provided.
 Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
@@
 
-\section{Rule X: Have a code of conduct and a team to reinforce it}
+\section{Rule X: Adopt a code of conduct and create a team to reinforce it}
+\label{rule_CoC}
 
@@
 
-\section{Rule X: Search proactively the best people for invited roles}
-
+\section{Rule X: Search proactively for the best people the conference's vision}
+\label{rule_presenters}
 When inviting people as keynote speakers, program committee, session chairs and others, we should always look for the best in their subject.
@@
 
-% So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for various aspects of the conference organization.
+% So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for *these invited roles*.
 The regional and local communities are also good sources to tap into.
@@
 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential attendees, and sponsors.
+%ast: i think this could come back, removing the specific part about "conference organization" - i edited that.
 
@@
 
-%Representation is an important aspect of diversity and inclusivity: seeing people like you, from your country, that speak your language is one of the best way to feel you belong, this is why the rates can not only be different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.
-
 %rationale: 1.
@@
 Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
-If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or tokenizing phrases (e.g.
+If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g.
 `grant recipient' or `diversity scholar', respectively).
@@
 Also, the process for applicants should be simplified.
-Applying for loans, grants, and scholarships is already an emotionally demanding task.
+Applying for loans, grants, and scholarships may be an emotionally demanding task.
 Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support.
 For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
-\st{Also, the participant metadata (e.g., demographics, gender, career stage, disabilities, if they disclose them) is already a strong indicator of their potential need for aid.} Other If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
+If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
 
@@
 
-\section{Rule X: Take deliberate efforts to promote the conference among underrepresented groups}
-
+\section{Rule X: Actively reach out and involve people who have been systematically excluded}
+\label{rule_communication}
+%this rule name should be more explicit towards communication, "promoting"
 If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
@@
 Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
-
-%Providing financial and other support (e.g., childcare, flexible hours) will help minoritized groups partake more willingly in the organization and the actual conference.
-%ast likewise, some key people in your organizing team might not be able to collaborate without remuneration. Resources permitting, it may be a good idea to secure funds to offer stipends to them.
 
@@
 
+\section{Rule X: Don't let perfect be the enemy of good}
+ While conference teams should have a bold vision and clear targets consensus about where to shine and where to compromise is equally important.
+As a team, this is a challenging journey.
+It means letting go of some ideals, or committing time and money into something that one does not find particularly useful or important.
+
+Remember, aspirations can be local and narrow and as a consequence shy away valuable contributors.
+A late bus is almost certainly better than no bus.
+A broader pool of contributors does not only allow organizers to add more and work on multiple aspects at a different pace, it also helps to split workload and responsibilities to avoid burning out over the marathon that organizing a conference is.
+
+Without allowing for compromises, workload will naturally channel to those who are committed the most and already carry the largest share of the load.
+Hence compromises are no sign of giving in but are crucial to delegation, building self-responsible teams and keep a good spirit in the long run.
+
 \section*{Concluding remarks}
 
-The ten rules stated here can be adapted depending on the conference format and settings.
-From our own experience, we are aware that enforcing these rules requires a tremendous amount of work.
-We engaged in most of these practices, and learned others too late, like having an explicit payment category for postdocs or paying the whole organizing team, so we share them here as part of our learning process.
+The ten rules stated here can be adapted depending on the conference format and settings. From our own experience, %it's worth it :P
+we are aware that enforcing these rules requires a tremendous amount of work.
+We engaged in most of these practices, and learned others too late, so we share them here as part of our learning process.
 Since accessibility and inclusiveness are not as widespread as they should be, the process might seem exacting.

### 2021-08-25 plos_latex_template.tex
RJ: new bib file and some of Andrea's changes to rule 1

@@
 The larger the conference, the larger the opportunities to network and learn from your cohort.
-However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arend_disparity_2019, timperley_he_2020, gewin_what_2019, brown_ableism_2018}.
+However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
 
@@
 
-The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.
+The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020}.
 
+When thinking about diversity, each one of us can have different points of view about the meaning of the term, depending on where we come from.
+If you are reading this, you may be here because you already have
 Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
-There are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
+While human diversity should be celebrated and respected in every space, the truth is that there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
 Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
@@
 To accomplish this diversity vision, you will have to take some conscious decisions toward removing bias from your conference's spaces.
-Having diverse people in decision-making positions will affect positively all the other aspects of your conference, not only internally because all the decision-making process will benefit from their input \ref{rule_organizing_team}, but also externally because representation matters \ref{rule_communication}.
+Having diverse people in decision-making positions will affect positively all the other aspects of your conference, not only internally because all the decision-making process will benefit from their input (see Rule \ref{rule_organizing_team}), but also externally because representation matters[see Rule \ref{rule_communication}].
 
 While you may want to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects as well.
-Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, are some examples of decisions that can make inclusion real.
+Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, and accomodating are some examples of decisions that can make inclusion real.
 %Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
-Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}).
-%and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too.
+Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the communication strategy see Rule \ref{rule_communication} and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too.
 However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
@@
 In-person interaction is priceless; however, it is more expensive for some, even unattainable for others.
-This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arend_disparity_2019, gewin_what_2019}.
+This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
 Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
-This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinson_r_medicine_2021}.
-Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipour_evaluating_2020, niner_better_2021}; in a study case, conference attendance accounted for $35\%$ of a researcher's footprint \cite{achten_carbon_2013}.
+This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}.
+Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts}.
 
@@
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-The challenge in this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.
+The challenge in this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
 
@@
 Even if creating and maintaining such a team and space is challenging, the positive outcomes far outweigh any minor inconveniences (e.g.
-time zones), and the team will grow, learn, and embrace more inclusive practices \cite{hong_groups_2004}.
+time zones), and the team will grow, learn, and embrace more inclusive practices \cite{hongGroupsDiverseProblem2004}.
 
@@
 \label{rule_accessibility}
-Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}.
+Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{priceAccessImaginedConstruction2009}.
 %People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
@@
 Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
-They require time and early decision-making \cite{irish_increasing_2020}. 
+They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}. 
 Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
@@
 
-To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}.
-The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.
+To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaroYourScienceConference2016}.
+The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2018}.
 The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
@@
 If these groups are not showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to looking in unfamiliar places to bring in the talent.
-Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}.
+Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,sarabipourChangingScientificMeetings2021}.
 Rather than deter us, this should encourage us to look further to find great people that are not routinely in the spotlight.
@@
 A lower fee for non-profit organizations, government employees, or freelancers should also be considered.
-It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}.
+It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
 Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined.
 You can take into account the cost of living in each country using conversion factors (e.g.
-from the International Comparison Program report of the World Bank \cite{arend_disparity_2019}).
+from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}).
 Consider that, while it is traditional for employers to provide --at least some --conference support in the Global North, this is not the case everywhere.
@@
 The platforms, the webpage, and official communications are also in English.
-While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{niner_better_2021}.
+While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}.
 Excluding them may potentially lead to missing innovative contributions due to a language barrier.
@@
 % % Compile your BiBTeX database using our plos2015.bst
-\bibliography{references}
+\bibliography{community-science}
 % % style file and paste the contents of your .bbl file

### 2021-08-25 plos_latex_template.tex
RJ: changes in accessibility rule

@@
 
-\section{Rule X: Make the conference accessible to all, from start to end}
+\section{Rule X: Make the conference accessible to people with disabilities}
 \label{rule_accessibility}
-Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{priceAccessImaginedConstruction2009}.
+
+Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
 %People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
-Inaccessible practices exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people.
+Ironically, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people.
 For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

### 2021-08-25 plos_latex_template.tex
RJ: spotlight rule

@@
 
+When thinking about diversity, each one of us can have different points of view about the meaning of the term, depending on where we come from.
+If you are reading this, you may be here because you already have some experience in Diversity, Inclusion, and Accessibility practices.
+Or you may want to improve diversity in your conference but do not know how or where to start.
+This article does not intend to be a comprehensive review of this topic --this would be impossible-- but will give some ideas pointing in key directions.
+
+Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
+While human diversity should be celebrated and respected in every way, the truth is that there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems --including conferences-- are consciously and unconsciously built.
+As a result, most people in the world fall into some category where they are part of a marginalized group --as opposed to a privileged one.
 The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020}.
 
-When thinking about diversity, each one of us can have different points of view about the meaning of the term, depending on where we come from.
-If you are reading this, you may be here because you already have
-Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
-While human diversity should be celebrated and respected in every space, the truth is that there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built.
-Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
+%Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
 
 Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context.
-Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your definition.
+Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your vision.
 Does this mean an even gender distribution in your speakers?
@@
 
+\section{Create a safe and welcoming environment}
+
 While you may want to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects as well.
-Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, and accomodating are some examples of decisions that can make inclusion real.
+Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, are just some examples of decisions that can make inclusion real.
+
 %Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
-Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the communication strategy see Rule \ref{rule_communication} and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too.
+
+Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}.
+
 However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
@@
 This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}.
-Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts}.
+Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
 
@@
 
-\section{Rule X: Search proactively for the best people the conference's vision}
-\label{rule_presenters}
-When inviting people as keynote speakers, program committee, session chairs and others, we should always look for the best in their subject.
-Ensuring diversity in each of these roles is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups.
-If these groups are not showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to looking in unfamiliar places to bring in the talent.
-Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,sarabipourChangingScientificMeetings2021}.
-Rather than deter us, this should encourage us to look further to find great people that are not routinely in the spotlight.
+\section{Rule X: Consciously unbias your spotlight roles}
+\label{rule_spotlight}
+
+When choosing or inviting people as keynote speakers, program committee, session chairs and other spotlight roles, it is likely that there will not be much diversity in the first set of names.
+Many of our biased lists are products of the existing systems that have always privileged some groups of people  \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
+Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight.
+Ensuring diversity in each of these roles needs to be a deliberate process.
 We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
 
-% So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for *these invited roles*.
-The regional and local communities are also good sources to tap into.
+Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them.
+An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles.
+The regional and local communities in your field are also good sources to tap into.
 For example, for useR!
-2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential attendees, and sponsors.
-%ast: i think this could come back, removing the specific part about "conference organization" - i edited that.
+2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
+
+Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative  ways \cite{cheng2020x+}.
+Give awards (see Rule \ref{rule_accessibility}) to those who prepared accessible slides and presentations, for instance, to acknowledge being mindful of inclusiveness.
+Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve.
+Defy the stereotypical criteria for success by acknowledging these community practices.
+However, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility.
+Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
 
@@
 If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
-Publish a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines.
-Fee waivers, financial support, and ease of application/registration should also be advertised; people who are used to applying and getting rejected for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory against underrepresented groups.
+Publish the diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines.
+Fee waivers, financial support, and ease of application/registration should also be advertised; people who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory against underrepresented groups.
 Advertise the conference in multiple languages (rule \ref{rule_language}).

### 2021-08-25 plos_latex_template.tex
rj + ast working on diversity, welcoming, accessibility, and spotlight rules

@@
 
-When thinking about diversity, each one of us can have different points of view about the meaning of the term, depending on where we come from.
+Each one of us can have different points of view about the meaning of diversity, depending on where we come from.
 If you are reading this, you may be here because you already have some experience in Diversity, Inclusion, and Accessibility practices.
@@
 This article does not intend to be a comprehensive review of this topic --this would be impossible-- but will give some ideas pointing in key directions.
+% RJ: Most of the first paragraph could pivot to the intro, but we can see later.
 
 Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
-While human diversity should be celebrated and respected in every way, the truth is that there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems --including conferences-- are consciously and unconsciously built.
+Human diversity should be celebrated and respected in every way.
+Nonetheless, there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems --including conferences-- are consciously and unconsciously built.
 As a result, most people in the world fall into some category where they are part of a marginalized group --as opposed to a privileged one.
-The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020}.
+The first step towards a diverse and inclusive conference is to recognize that people are diverse, and that some groups face discrimination (some more than others) and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020}.
 
@@
 Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your vision.
-Does this mean an even gender distribution in your speakers?
-Does this relate to people of all races -especially black people- being present as coordinators, presenters, and in the audience?
-Does this mean having LGBTQ+ friendly-spaces or community participation from key geographic regions?
+Would this translate into an even gender distribution in your speakers?
+Would it be representation of all races -especially Black people-- among the head organizers, speakers, and attendees?
+Would it be having LGBTQ+ friendly-spaces or community participation from key geographic regions?
 This vision should guide and help hold the organizing team accountable along the way.
 
-To accomplish this diversity vision, you will have to take some conscious decisions toward removing bias from your conference's spaces.
-Having diverse people in decision-making positions will affect positively all the other aspects of your conference, not only internally because all the decision-making process will benefit from their input (see Rule \ref{rule_organizing_team}), but also externally because representation matters[see Rule \ref{rule_communication}].
+To accomplish this diversity vision, you will have to take conscious decisions toward removing bias from your conference's spaces.
+Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the decision-making processes will benefit from their input (see Rule \ref{rule_organizing_team}), and because representation matters (see Rule \ref{rule_communication}).
+% RJ: we could improve the wording of the last sentence.
 
@@
 
-While you may want to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects as well.
-Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, are just some examples of decisions that can make inclusion real.
+While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
+Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
 
@@
 
-Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}.
+%Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}.
+However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
 
-However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
+Paying attention to some dimensions of diversity while neglecting others may have unintended consequences.
+Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members.
+In severe cases, they can divert career paths, affect lives, and exclude people from some fields.
 
-Lack of representation, unwelcoming -or overtly aggressive- environments hinder participation (or future participation) of people who could otherwise become active community members.
-In extreme cases, they can divert career paths, affect lives, and exclude people from some fields.
-So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn.
+%So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn.
 Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator.
 %ast: some version of this might come in a box.
+
+A key component in creating a safe environment during a conference is adopting a code of conduct and creating and training a team to enforce it \cite{favaroYourScienceConference2016}.
+The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2018}.
+The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
+An efficient code of conduct acts as a protection for the community because the people who may be target of unacceptable behavior tend to be people with less power or privilege.
+
+The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses.
+A diverse code of conduct team will be more understanding of power dynamics, and sensitive to discrimination and harassment.
+We strongly recommend reading How to enforce a Code of Conduct by Valerie Aurora (\url{https://frameshiftconsulting.com/resources/code-of-conduct-book/}) as an excellent starting point for the Code of Conduct team work.
 
@@
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-The challenge in this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
+The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
 
@@
 Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity.
-Particular attention should be paid to the usually marginalized groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event).
-To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
+Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}} discussing the dimensions of diversity lacking in your community and event).
+To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
 But this will only work if there is real inclusion.
@@
 %The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation.
-% rj: left this out, Heather said it doesn't make much sense here %ast: we can collect part of this in rule 1, since some self-teaching, and not expecting to be taught, is necessary
+% rj: left this out, Heather said it doesn't make much sense here %ast: we can collect part of this in rule 1, since some self-teaching, and not expecting to be taught, is necessary.
+nope.
+now i think this is here.
 
@@
 This means the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups.
+
 %rj: Yani's refs to have just in case:
@@
 Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
-Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.
 
@@
 Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}).
-
-\section{Rule X: Adopt a code of conduct and create a team to reinforce it}
-\label{rule_CoC}
-
-To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaroYourScienceConference2016}.
-The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2018}.
-The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
-
-The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses.
-A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.
-%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.'
 
@@
 
-%rationale: 1.
-they might select for the already privileged or people who dont need those to be nice on the cvs 2.
-they select for outgoing extroverted privileged ("brave") people 3.
-they are time and emotionally consuming which selects out some people 4.
-they single out people.
-even if you controlled for 1 to 3 you would "badge" people - you are creating a minority while trying to avoid it.
 Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
@@
 Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
+
+% RJ: could add some ideas from here 
+% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
 

### 2021-08-26 plos_latex_template.tex
RJ: version 2 of the conclusion

@@
 
-%ast: matt's original sentence: reaching users and developers of the R language from more than 120 countries.
+%andrea: matt's original sentence: reaching users and developers of the R language from more than 120 countries.
 
@@
 
-%ast will write a paragraph about why diversity and the systemic inequalities
+%andrea will write a paragraph about why diversity and the systemic inequalities
 While there has been some action to address systematic inequalities, there is still a lot of room for improvement.
+% Rocío: Somehow this part below should be linked to the systemic inequalities part
+Each one of us can have different points of view and experiences regarding inequality, depending on where we come from.
+If you are reading this, you may be here because you already have some experience in Diversity, Inclusion, and Accessibility practices.
+Or you may want to improve diversity in your conference but do not know how or where to start.
+This article does not intend to be a comprehensive review of this topic --this would be impossible-- but will give some ideas pointing in key directions.
+%Rocío: And finish with something more specific to connect with the next paragraph.
 
@@
 
-\section{Rule X: Define what diversity and inclusion mean for your conference}
+\section{Define what diversity and inclusion mean for your conference}
 \label{rule_diversity}
-
-Each one of us can have different points of view about the meaning of diversity, depending on where we come from.
-If you are reading this, you may be here because you already have some experience in Diversity, Inclusion, and Accessibility practices.
-Or you may want to improve diversity in your conference but do not know how or where to start.
-This article does not intend to be a comprehensive review of this topic --this would be impossible-- but will give some ideas pointing in key directions.
-% RJ: Most of the first paragraph could pivot to the intro, but we can see later.
 
@@
 The first step towards a diverse and inclusive conference is to recognize that people are diverse, and that some groups face discrimination (some more than others) and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020}.
-
-%Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
 
@@
 
-To accomplish this diversity vision, you will have to take conscious decisions toward removing bias from your conference's spaces.
-Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the decision-making processes will benefit from their input (see Rule \ref{rule_organizing_team}), and because representation matters (see Rule \ref{rule_communication}).
-% RJ: we could improve the wording of the last sentence.
+%Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
 
 \section{Create a safe and welcoming environment}
+\label{rule_inclusion}
 
@@
 
-%Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
-
-%Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}.
-However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
-
 Paying attention to some dimensions of diversity while neglecting others may have unintended consequences.
@@
 In severe cases, they can divert career paths, affect lives, and exclude people from some fields.
+%Rocío: it would be nice to have reference here.
 
-%So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn.
-Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator.
-%ast: some version of this might come in a box.
-
-A key component in creating a safe environment during a conference is adopting a code of conduct and creating and training a team to enforce it \cite{favaroYourScienceConference2016}.
+Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
 The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2018}.
 The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
-An efficient code of conduct acts as a protection for the community because the people who may be target of unacceptable behavior tend to be people with less power or privilege.
+An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.
 
@@
 
-\section{Rule X: Have a strong online component of the conference} 
+%So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn.
+Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator.
+%andrea: some version of this might come in a box.
+
+%Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}.
+However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
+
+\section{Have a strong online component of the conference} 
 \label{rule_online}
-In-person interaction is priceless; however, it is more expensive for some, even unattainable for others.
+In-person interaction is priceless; however, it is more expensive for some, and even unattainable for others.
 This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
@@
 
-\section{Rule X: Have an inclusive and diverse organizing team}
+\section{Have an inclusive and diverse organizing team}
 \label{rule_organizing_team}
+
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
@@
 To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
-But this will only work if there is real inclusion.
+If you already have an organizing team, check for biases in its composition.
+
+%liz: please check if this paragraph connects well with the whole rule.
+i am using nothing about us without us but also talking about all marginalized groups, in the context of building a team.
+att: andrea
+
+Gathering a diverse team will only work if there is real inclusion.
+Disabled people often say: "Nothing about us without us"; the same holds for other dimensions of diversity.
+This means that the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups.
+A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.
+So, if you plan to organize something related to any community, region, or specific dimension of diversity, involve people from these groups.
+Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cit{hong_groups_2004}.
+On the other hand, seeing diverse teams is important to create a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}).
+%Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
+
+Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences and the team will grow, learn, and embrace more inclusive practices \cite{hongGroupsDiverseProblem2004}.
+
+%How to work: 
+Most advice when working with a diverse team should apply to every team, but is especially true when working with people from marginalized groups.
+First, delegate.
 One would be surprised by how often conference leads tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in.
-For example, a cisgender or abled person estimating what might work best for transgender people or people with sensory disabilities.
-The same holds for other dimensions of diversity.
-A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.
-Even if creating and maintaining such a team and space is challenging, the positive outcomes far outweigh any minor inconveniences (e.g.
-time zones), and the team will grow, learn, and embrace more inclusive practices \cite{hongGroupsDiverseProblem2004}.
 
-%The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation.
-% rj: left this out, Heather said it doesn't make much sense here %ast: we can collect part of this in rule 1, since some self-teaching, and not expecting to be taught, is necessary.
-nope.
-now i think this is here.
+Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities.
+Nominate directly and offer the positions of power, even if this means stepping down.
+Offer mentorship and guidance if you encounter doubts, cases of impostor syndrome, doubts about the use of English in communication \ref{rule_language}.
+Break the expectations about lonley leadership and create smaller, co-led groups.
 
-%%Disabled people often repeat "Nothing about us without us", and this is true for every marginalized group.
-This means the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups.
+In addition to exclude derogatory or discriminatory language (which would constitute a violation of the code of conduct \ref{rule_inclusion} but also, should never happen), make the effort to teach yourself the best ways to communicate to account for every culture and situation.
+
+Finally, splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups .
+
+Take care of your team!
 
@@
 
-\section{Rule X: Make the conference accessible to people with disabilities}
+\section{Make the conference accessible to people with disabilities}
 \label{rule_accessibility}
@@
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
-%People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
 Ironically, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people.
@@
 Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
-Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}).
+On the other hand, it is likely that financial constraints may not allow you to make your conference completely accessible and there will be choices to make (e.g.
+choosing between captions and interpreters).
+Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
 
-\section{Rule X: Consciously unbias your spotlight roles}
+\section{Consciously unbias your spotlight roles}
 \label{rule_spotlight}
@@
 
-\section{Rule X: Set fair registration rates and provide financial support when needed}
+\section{Set fair registration rates and provide financial support when needed}
 
@@
 
-\section{Rule X: Don't let language restrict high-quality participation}
+\section{Don't let language restrict high-quality participation}
 \label{rule_language}
@@
 
-\section{Rule X: Actively reach out and involve people who have been systematically excluded}
+\section{Actively reach out to people who have been systematically excluded}
 \label{rule_communication}
-%this rule name should be more explicit towards communication, "promoting"
+%Andrea: this rule name should be more explicit towards communication, "promoting"
 If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
@@
 Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
-
-% RJ: could add some ideas from here 
+% Yani: We need to work a little more here, perhaps with some examples.
+% Rocío: could add some ideas from here 
 % https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
 
-\section{Rule X: Do not expect the organizing team to work full time and for free}
+\section{Do not expect the organizing team to work full time and for free}
 
@@
 
-\section{Rule X: Don't let perfect be the enemy of good}
+\section{Don't let perfect be the enemy of good}
  While conference teams should have a bold vision and clear targets consensus about where to shine and where to compromise is equally important.
@@
 It means letting go of some ideals, or committing time and money into something that one does not find particularly useful or important.
+%Yani: ughhhh...
+this doesn't sound good.
+I agree more with let go some ideal, for example, when I compromise in not have caption for tutorials, because we don't have money.
+Have caption in the other part of the conf was better because reach a more wide audience of the conference.
+I will try to show the compromise from the point of view of all the things we know we still have to do about inclusion, but we cannot do because we don't have money or the knowledge or the capacity at this time.
+This is also mention in some of the other rules.
+I don't remember if it is the one with the goals and indicators.
 
-Remember, aspirations can be local and narrow and as a consequence shy away valuable contributors.
-A late bus is almost certainly better than no bus.
+Remember, aspirations can be local and narrow and as a consequence shy away valuable contributors. A late bus is almost certainly better than no bus. 
+% Yani: I don't like this phrase.
+It is like we have something of a lower quality than the original idea, when I truly think we have a high quality conference because was inclusive and diverse.
+and science say that a diverse team get better results :)
 A broader pool of contributors does not only allow organizers to add more and work on multiple aspects at a different pace, it also helps to split workload and responsibilities to avoid burning out over the marathon that organizing a conference is.
+%Yani: We have a broader team because we want a more diverse pool of view and experience, this is a good phrase but sound like we people were an object.
+Perhaps we can rephrase and add to the rule that talk about the composition of the team.%andrea: i am taking split workload and responsibilities up, but they were already there.
 
@@
 
+%\section{Rule X: Have a bold vision but be ready to compromise when necessary}
+This rule is a complement to all the other rules.
+It is important to have a bold vision and clear targets.
+Some parts might be defined as baseline assumptions that are not up for discussion.
+However, there will always be compromises.
+Ideal solutions that meet all our expectations are often not possible.
+Don't let the perfect be the enemy of good.
+As a team, this is a challenging journey.
+It means letting go of some ideals, or committing time and money into something that one does not find particularly useful or important.
+We organized useR!
+during a global pandemic.
+Very early on, we tried to recognize that our resources were limited (both in time of the organizing team but also in money).
+We had to let go of certain aspects that were important to many of us.
+For an inclusive conference, we need to avoid giving in because considering all aspects seems impossible and burning ourselves and the team out because considering all aspects is actually impossible.
+
 \section*{Concluding remarks}
 
-The ten rules stated here can be adapted depending on the conference format and settings. From our own experience, %it's worth it :P
-we are aware that enforcing these rules requires a tremendous amount of work.
-We engaged in most of these practices, and learned others too late, so we share them here as part of our learning process.
 Since accessibility and inclusiveness are not as widespread as they should be, the process might seem exacting.
 But this is where the power of communities and individuals reside. 
-If more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
-And we will make a change towards healthier, stronger, and more inclusive communities, empowering historically marginalized people and learning from them, thereby increasing the overall quality of our conferences.%After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.
+If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead. 
+And you will make a change towards healthier, stronger, and more inclusive communities, empowering historically marginalized people and learning together, thereby increasing the overall quality of your conferences.
+
+We engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
+We organized useR! during a global pandemic  
+and, as a team, this was a challenging journey. 
+Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
+We had to let go of certain aspects that were important to many of us. From that experience, this is our last message: 
+Your conference may not become perfectly inclusive and accessible. 
+Don't let perfect be the enemy of good.
+
+% Try not to burn you and the team out trying to make it perfect.
 

### 2021-08-26 plos_latex_template.tex
RJ: Andrea's edits to organizing team

@@
 The larger the conference, the larger the opportunities to network and learn from your cohort.
-However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
+However, conferences can become discriminating spaces, in which members of some specific privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities) reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
 
-%andrea will write a paragraph about why diversity and the systemic inequalities
-While there has been some action to address systematic inequalities, there is still a lot of room for improvement.
+While there has been some action to address systematic inequalities [where], there is still a lot of room for improvement.
 % Rocío: Somehow this part below should be linked to the systemic inequalities part
@@
 
-This article suggests rules to pivot traditional conferences towards inclusiveness and diversity and to welcome minoritized groups.
-It is directed as people who are part of a stable meetings commitee that oversees the site/location selection process or that coordinates with the local organizers (e.g.
+This article suggests rules to pivot traditional conferences towards inclusiveness and diversity.
+It is directed as people who are part of a stable meetings committee that oversees the site/location selection process or that coordinates with the local organizers (e.g.
 R Foundation Conference Committee, Ecological Society of America Meetings Committee).
@@
 2021, a virtual and global conference for users and developers of the R programming language \cite{r_core_team_2021}.
-We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive, accessible, and welcoming experience for everyone.
+We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive and accessible experience for everyone.
 Most of the authors also have experience of organizing other regional and national academic conferences and events in communities of practice such as R-Ladies and The Carpentries.
-Here, we share the lessons learned within the past year of organizing this global useR!
-conference, summarized as 10 simple rules towards an inclusive conference.
+Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.
 
@@
 As a result, most people in the world fall into some category where they are part of a marginalized group --as opposed to a privileged one.
-The first step towards a diverse and inclusive conference is to recognize that people are diverse, and that some groups face discrimination (some more than others) and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020}.
+A more diverse and inclusive conference  starts by recognizing that people are diverse, and that some groups face discrimination (some more than others) and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020} because of these implicit hierarchies.
+
+Think about the reasons why you want to create a more diverse conference.
+This may be important to you, whether you come from a privileged background or not.
+Along this paper, Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes, and do not act as a benefactor or savior but as a genuine collaborator.
+
+%Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
 
@@
 Would it be representation of all races -especially Black people-- among the head organizers, speakers, and attendees?
-Would it be having LGBTQ+ friendly-spaces or community participation from key geographic regions?
+Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
 This vision should guide and help hold the organizing team accountable along the way.
-
-%Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
 
@@
 
-%So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn.
-Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator.
-%andrea: some version of this might come in a box.
-
 %Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}.
 However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
-
-\section{Have a strong online component of the conference} 
-\label{rule_online}
-In-person interaction is priceless; however, it is more expensive for some, and even unattainable for others.
-This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
-Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
-This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}.
-Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
-
-Alternatively, a conference could have a hybrid format with an in-person and online component.
-This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
 
@@
 So, if you plan to organize something related to any community, region, or specific dimension of diversity, involve people from these groups.
-Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cit{hong_groups_2004}.
-On the other hand, seeing diverse teams is important to create a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}).
+
+Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences. 
+Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
+On the other hand, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}).
 %Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
 
-Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences and the team will grow, learn, and embrace more inclusive practices \cite{hongGroupsDiverseProblem2004}.
-
-%How to work: 
 Most advice when working with a diverse team should apply to every team, but is especially true when working with people from marginalized groups.
 First, delegate.
-One would be surprised by how often conference leads tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in.
+Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities.
+Nominate directly and offer coordinating positions, even if this means stepping down.
+Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication \ref{rule_language}, or others.
+Break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
+Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while men take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}).
 
-Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities.
-Nominate directly and offer the positions of power, even if this means stepping down.
-Offer mentorship and guidance if you encounter doubts, cases of impostor syndrome, doubts about the use of English in communication \ref{rule_language}.
-Break the expectations about lonley leadership and create smaller, co-led groups.
+Words matter.
+In addition to excluding derogatory or discriminatory language (remember that the code of conduct applies to you too, Rule \ref{rule_inclusion}), make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
+Do not expect people to teach you --it's not their role-- and accept any feedback without being defensive.
 
-In addition to exclude derogatory or discriminatory language (which would constitute a violation of the code of conduct \ref{rule_inclusion} but also, should never happen), make the effort to teach yourself the best ways to communicate to account for every culture and situation.
-
-Finally, splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups .
-
-Take care of your team!
+And most importantly, take care of your team!
+Check on them regularly and make sure everyone is comfortable.
+This represents a great amount of work and you'll need to support each other in the long run.
 
@@
 %https://sites.lsa.umich.edu/scottepage/wp-content/uploads/sites/344/2015/11/pnas.pdf; https://www.pnas.org/content/early/2014/11/13/1407301111; https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
+
+\section{Have a strong online component of the conference} 
+\label{rule_online}
+
+In-person interaction is priceless; however, it is more expensive for some, and even unattainable for others.
+This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
+Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
+This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}.
+Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
+
+Alternatively, a conference could have a hybrid format with an in-person and online component.
+This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
+The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
 
@@
 choosing between captions and interpreters).
-Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
+Including people that are usually disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
 
@@
 
-\section{Don't let perfect be the enemy of good}
- While conference teams should have a bold vision and clear targets consensus about where to shine and where to compromise is equally important.
-As a team, this is a challenging journey.
-It means letting go of some ideals, or committing time and money into something that one does not find particularly useful or important.
-%Yani: ughhhh...
-this doesn't sound good.
-I agree more with let go some ideal, for example, when I compromise in not have caption for tutorials, because we don't have money.
-Have caption in the other part of the conf was better because reach a more wide audience of the conference.
-I will try to show the compromise from the point of view of all the things we know we still have to do about inclusion, but we cannot do because we don't have money or the knowledge or the capacity at this time.
-This is also mention in some of the other rules.
-I don't remember if it is the one with the goals and indicators.
-
-Remember, aspirations can be local and narrow and as a consequence shy away valuable contributors. A late bus is almost certainly better than no bus. 
-% Yani: I don't like this phrase.
-It is like we have something of a lower quality than the original idea, when I truly think we have a high quality conference because was inclusive and diverse.
-and science say that a diverse team get better results :)
-A broader pool of contributors does not only allow organizers to add more and work on multiple aspects at a different pace, it also helps to split workload and responsibilities to avoid burning out over the marathon that organizing a conference is.
-%Yani: We have a broader team because we want a more diverse pool of view and experience, this is a good phrase but sound like we people were an object.
-Perhaps we can rephrase and add to the rule that talk about the composition of the team.%andrea: i am taking split workload and responsibilities up, but they were already there.
-
-Without allowing for compromises, workload will naturally channel to those who are committed the most and already carry the largest share of the load.
-Hence compromises are no sign of giving in but are crucial to delegation, building self-responsible teams and keep a good spirit in the long run.
-
-%\section{Rule X: Have a bold vision but be ready to compromise when necessary}
-This rule is a complement to all the other rules.
-It is important to have a bold vision and clear targets.
-Some parts might be defined as baseline assumptions that are not up for discussion.
-However, there will always be compromises.
-Ideal solutions that meet all our expectations are often not possible.
-Don't let the perfect be the enemy of good.
-As a team, this is a challenging journey.
-It means letting go of some ideals, or committing time and money into something that one does not find particularly useful or important.
-We organized useR!
-during a global pandemic.
-Very early on, we tried to recognize that our resources were limited (both in time of the organizing team but also in money).
-We had to let go of certain aspects that were important to many of us.
-For an inclusive conference, we need to avoid giving in because considering all aspects seems impossible and burning ourselves and the team out because considering all aspects is actually impossible.
-
 \section*{Concluding remarks}
 
-Since accessibility and inclusiveness are not as widespread as they should be, the process might seem exacting.
-But this is where the power of communities and individuals reside. 
+The ten rules stated here can be adapted depending on the conference format and settings.
+When organizing useR!2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
+We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
+Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
+We had to let go of certain aspects that were important to many of us. 
+From that experience, this is our last message: Your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference. 
 If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
-And you will make a change towards healthier, stronger, and more inclusive communities, empowering historically marginalized people and learning together, thereby increasing the overall quality of your conferences.
-
-We engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
-We organized useR! during a global pandemic  
-and, as a team, this was a challenging journey. 
-Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
-We had to let go of certain aspects that were important to many of us. From that experience, this is our last message: 
-Your conference may not become perfectly inclusive and accessible. 
-Don't let perfect be the enemy of good.
+And you will make a change towards healthier, stronger, and more inclusive communities.
 

### 2021-08-26 plos_latex_template.tex
ast - some changes in rule 1. the don'ts.

@@
 Nonetheless, there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems --including conferences-- are consciously and unconsciously built.
-As a result, most people in the world fall into some category where they are part of a marginalized group --as opposed to a privileged one.
-A more diverse and inclusive conference  starts by recognizing that people are diverse, and that some groups face discrimination (some more than others) and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020} because of these implicit hierarchies.
+While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have excluded systematically whole groups of people from academia and from scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
 
-Think about the reasons why you want to create a more diverse conference.
-This may be important to you, whether you come from a privileged background or not.
-Along this paper, Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes, and do not act as a benefactor or savior but as a genuine collaborator.
-
-%Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.
+Think about the reasons why you want to create a more diverse conference and try not to treat diversity as a checklist, avoid using diversity and inclusion work for image purposes, and do not act as a benefactor or savior but as a genuine collaborator.
+%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe).
+i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
 
@@
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
+Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
+Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.
 
@@
 The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses.
-A diverse code of conduct team will be more understanding of power dynamics, and sensitive to discrimination and harassment.
+A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues.
 We strongly recommend reading How to enforce a Code of Conduct by Valerie Aurora (\url{https://frameshiftconsulting.com/resources/code-of-conduct-book/}) as an excellent starting point for the Code of Conduct team work.
@@
 Nominate directly and offer coordinating positions, even if this means stepping down.
-Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication \ref{rule_language}, or others.
+Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See Rule \ref{rule_language}), or others.
 Break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.

### 2021-08-27 plos_latex_template.tex
added comments Adithi

@@
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
+Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTIQA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
 Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.
@@
 %Rocío: it would be nice to have reference here.
+%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
 
@@
 Gathering a diverse team will only work if there is real inclusion.
-Disabled people often say: "Nothing about us without us"; the same holds for other dimensions of diversity.
+People with disabilities often say: "Nothing about us without us"; the same holds for other dimensions of diversity.
 This means that the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups.
@@
 
-%rj: Yani's refs to have just in case:
-%https://sites.lsa.umich.edu/scottepage/wp-content/uploads/sites/344/2015/11/pnas.pdf; https://www.pnas.org/content/early/2014/11/13/1407301111; https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
+%Rocío: Yani's ref to have just in case:
+% https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
 
@@
 The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
+%Adithi: After the conference is over, the conference materials and videos can be uploaded on free platforms, so that people who were left out of the main event due to any reason, still feel they were cared for and can go to the resources.
+Also when the event is online, a more timezone friendly approach could be used, all the events if online can be global friendly, people with dependents need hours to take care and the night to rest, if the conference is overnight, they lose sleep and suffer, instead if the conference could be friendly for all regions on certain days just like useR!
+2021.
 
@@
 choosing between captions and interpreters).
-Including people that are usually disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
+Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
+As Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".
 
@@
 
-\section{Actively reach out to people who have been systematically excluded}
+\section{Actively reach out and promote the conference to people who have been systematically excluded}
 \label{rule_communication}

### 2021-09-01 plos_latex_template.tex
Some editions by @nmorandeira - Correction of CoC book reference - adding key terms: intersectionality; well-being; ally; definition of privilege - adding conference locations and timezones - expanding examples of R-Communities to other academic conferences - comment on documenting the work for next conferences

@@
 
-In July 2021, the authors of this article participated in the organization team of the annual user conference of the R Project for Statistical Computing.
+The authors of this article participated in the organization team of the annual user conference of the R Project for Statistical Computing, held in July 2021.
 useR!
@@
 %andrea: matt's original sentence: reaching users and developers of the R language from more than 120 countries.
+
+%natalia: I changed the 1st sentence because we participated in the organizing team not just during the conference, but for several months before it.
+That was needed to work towards an inclusive conference
 
@@
 i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
-
+Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference.
+%Natalia: I'll look for a reference for privilege definition.
 Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context.
@@
 Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTIQA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
+Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}.
 Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.
@@
 Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
-The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2018}.
+The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}.
 The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
@@
 A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues.
-We strongly recommend reading How to enforce a Code of Conduct by Valerie Aurora (\url{https://frameshiftconsulting.com/resources/code-of-conduct-book/}) as an excellent starting point for the Code of Conduct team work.
+We strongly recommend reading “How to Respond to Code of Conduct Reports” \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.
+A Code of Conduct response committee needs to be trained and prepared in advance on how to decide on the incident reports.
+
+%Natalia: the book is by Aurora and Gardiner (2019), I corrected the title and added the reference including the link
 
@@
 
-And most importantly, take care of your team!
+And most importantly, take care of your team and their well-being!
 Check on them regularly and make sure everyone is comfortable.
@@
 % https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
+%Natalia: I don't know if this is the best place, but I'd add the term "well-being".
 
@@
 2021.
+
+If the conference is online and synchronic (e.g.
+with live sessions), timezones should be considered: you may change the favored timezone year to year, or work on several timezones in each conference edition.
+If the conference is hybrid or in-person, be sure the location does not favor the same people over years, move from one continent to another.
 
@@
 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
+When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.
 
@@
 For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
-If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).
+If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees and visa financial and administrative support (if in-person), or internet connection services (if virtual).
+If the conference is virtual, you may offer technical support so that attendees can record their videos, include captions or design accessible presentations.
 
@@
 We had to let go of certain aspects that were important to many of us. 
-From that experience, this is our last message: Your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference. 
+From that experience, this is our last message: your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference.
+Also, keep working to improve inclusion in the next edition of the conference!
+to document and balance your work, design surveys, ask for feedback from attendees and organizers.
+Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team.
 If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.

### 2021-09-13 plos_latex_template.tex
comments on the 1st rule by Batool

@@
 
-Think about the reasons why you want to create a more diverse conference and try not to treat diversity as a checklist, avoid using diversity and inclusion work for image purposes, and do not act as a benefactor or savior but as a genuine collaborator.
+Depending on the field, scientific or professional communities can have very different compositions.
+Thinking about which subgroups have been historically underrepresented in your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event.
+Doing this exercise while you ...
+Think about the reasons why you want to create a more diverse conference can avoid falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes., and do not act as a benefactor or savior but as a genuine collaborator.
 %andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe).
@@
 Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference.
+Privilege is largely invisible to those who have it.
+Once we understand our own privileges, we can begin addressing it on an individual and institutional basis.
 %Natalia: I'll look for a reference for privilege definition.
+%marcela: I modified the paragraph a bit, but it needs more work to connect Natalia's and Andrea's ideas. It's just a suggestion, feel free to re-write it again.
+%Batool: I tried making some paraphrasing - It's just a simple suggestion, not sure if it deliver the message you mean it.
+I also think that we may need to define "Privilege" at the 1st mention in the 1st paragraph in this rule.
+Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."
+
 Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context.
@@
 This vision should guide and help hold the organizing team accountable along the way.
+
+%batool: There are a few references which define privilege and might be good to add here including: Friedman, S., O’Brien, D., & McDonald, I.
+(2021).
+Deflecting Privilege: Class Identity and the Intergenerational Self.
+Sociology, 55(4), 716–733.
+https://doi.org/10.1177/0038038520982225
+%batool: Another one is: Crevani, L.
+(2019).
+Privilege in place: How organisational practices contribute to meshing privilege in place.
+Scandinavian Journal of Management, 35(2).
+https://doi.org/10.1016/j.scaman.2018.09.002
+%batool: please note that I didn't read both entire papers I mentioned above.
 
@@
 %Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
+
+Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of  attendants.
+If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, this sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
+Some examples are the newbies sessions for first timers, mixers lead by Rladies or MiR, or a leisure activity that reunites subgroups with the same interests: arts, exercise, sports, movies, etc.
 
@@
 On the other hand, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}).
+This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees.
+This gave a sense of variety and an appeal of “at least one networking session for everyone”.
+
 %Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
@@
 
+Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles, and accessibility needs.
+Social events in presencial conferences are often the most exclusive, and they are a very common place where you have code of conduct violations.
+Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond.
+An online event is a great opportunity to create spaces for people who do not enjoy in person networking.
+Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
+Respect people’s limits, preferences, and remember that “the usual” does not necessarily work with your community, and that one single networking activity will never serve the whole community.
+It is worth trying new and varied activities that might work with subgroups of attendees.
+
 If the conference is online and synchronic (e.g.
@@
 Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
+Do not forget about art shows and movies when making sure that the materials presented are accessible for all.
 Platforms for conference registration and abstract submission, websites, and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
@@
 For example, for useR!
-2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
+2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
 When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.
@@
 \section{Set fair registration rates and provide financial support when needed}
+\label{rule_costs}
 
@@
 \section{Do not expect the organizing team to work full time and for free}
+\label{rule_pay_team}
 

### 2021-09-13 plos_latex_template.tex
simple comments in the 2nd role by Batool

@@
 If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, this sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
-Some examples are the newbies sessions for first timers, mixers lead by Rladies or MiR, or a leisure activity that reunites subgroups with the same interests: arts, exercise, sports, movies, etc.
+Some examples are the newbies sessions for first timers, mixers lead by R-Ladies or MiR, or a leisure activity that reunites subgroups with the same interests: arts, exercise, sports, movies, etc.
 
@@
 The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}.
-The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
+The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior, which includes microaggression.
 An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.
@@
 A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues.
+There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC).
 We strongly recommend reading “How to Respond to Code of Conduct Reports” \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.

### 2021-09-13 plos_latex_template.tex
finished 1st round of comments - sorry for the delay - Batool

@@
 
-And most importantly, take care of your team and their well-being!
+And most importantly, take care of your team and prioritize their well-being!
 Check on them regularly and make sure everyone is comfortable.
@@
 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
-When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.
+When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged.
+Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.
 
@@
 Excluding them may potentially lead to missing innovative contributions due to a language barrier.
+
+Meeting the need of linguistically diverse audience is an important aspect in inclusion that is often overlooked.
+Non-native English speakers may feel intimidated to ask questions or raise their views.
+Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2).
+If the panellist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
+Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.
+
 Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier.
@@
 Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team.
+Always remember inclusion is an ambition that organisers can and should continually pursue by allowing room for learning and improvement.
 If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.

### 2021-09-15 plos_latex_template.tex
sara's small comment on rule #10
Rocío: rewrote the introduction

@@
 
-%natalia: I changed the 1st sentence because we participated in the organizing team not just during the conference, but for several months before it.
-That was needed to work towards an inclusive conference
-
 % % Please keep the Author Summary between 150 and 200 words
@@
 Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
-The larger the conference, the larger the opportunities to network and learn from your cohort.
-However, conferences can become discriminating spaces, in which members of some specific privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities) reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
+However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. %RJ: (Biggs et al. 2018) https://link.springer.com/article/10.1007/s11199-017-0800-9
+Conversely, efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact in our communities and the lives and careers of their members.
 
-While there has been some action to address systematic inequalities [where], there is still a lot of room for improvement.
-% Rocío: Somehow this part below should be linked to the systemic inequalities part
-Each one of us can have different points of view and experiences regarding inequality, depending on where we come from.
-If you are reading this, you may be here because you already have some experience in Diversity, Inclusion, and Accessibility practices.
-Or you may want to improve diversity in your conference but do not know how or where to start.
-This article does not intend to be a comprehensive review of this topic --this would be impossible-- but will give some ideas pointing in key directions.
-%Rocío: And finish with something more specific to connect with the next paragraph.
+This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
+Some scholarly or technological communities are formally structured as learned societies (e.g.
+the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g.
+R User Groups or Python User Groups), and others may not have any established structure.
+For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.
 
-This article suggests rules to pivot traditional conferences towards inclusiveness and diversity.
-It is directed as people who are part of a stable meetings committee that oversees the site/location selection process or that coordinates with the local organizers (e.g.
-R Foundation Conference Committee, Ecological Society of America Meetings Committee).
-It is also directed at the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
-
+The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, that coordinates with the local organizers of conferences, or even the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
 These tips stem from the authors' experience of organizing useR!
-2021, a virtual and global conference for users and developers of the R programming language \cite{r_core_team_2021}.
+2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
 We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive and accessible experience for everyone. 
-Most of the authors also have experience of organizing other regional and national academic conferences and events in communities of practice such as R-Ladies and The Carpentries.
 Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.
 
-\section{Define what diversity and inclusion mean for your conference}
+%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
+
+\section{Embrace all dimensions of diversity}
 \label{rule_diversity}
-
+%ast at the end i think we should go back to Embrace all dimensions of diversity
 Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
@@
 Depending on the field, scientific or professional communities can have very different compositions.
-Thinking about which subgroups have been historically underrepresented in your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event.
-Doing this exercise while you ...
-Think about the reasons why you want to create a more diverse conference can avoid falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes., and do not act as a benefactor or savior but as a genuine collaborator.
+Thinking about which subgroups have been historically excluded from your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event.
+Doing this exercise while you think about the reasons why you want to create a more diverse conference can prevent falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes.
 %andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe).
 i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
+%Yanina: They are going to blame what they want to blame.
+You can't control what people are going to do and say and we shouldn't write defensively because we can't take up space in our message with that.
+People who do not want to change will not change because of our paper.
+I imagine this as a guide and help for those people who do want to do something and want to do their best and do not know what or how to do it.
+Or they have already done something and want to continue deepening those changes.
+The rest is not worth spending time and energy.
+But perhaps the goals of the paper change in some other discussion and I lost that.
 Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference.
 Privilege is largely invisible to those who have it.
-Once we understand our own privileges, we can begin addressing it on an individual and institutional basis.
+Once we understand our own privileges, we can begin addressing them on an individual and institutional basis.
 %Natalia: I'll look for a reference for privilege definition.
@@
 Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
-This vision should guide and help hold the organizing team accountable along the way.
+This is not an invitation to prioritize some dimensions of diversity above others This vision should guide and help hold the organizing team accountable along the way.
 
@@
 
+%Yanina: It's probably a language problem but how I read this rule It is like to give the organizers a carte blanche to define what diversity and inclusion is and with that they are covered.
+What if they decide that everything is fine if they are all white people, native English speakers and without disabilities?
+I don't think we want to encourage that.
+I mean, you need a bottom line and from them you build to improve.
+Plus, as it is well highligth in these rules, it is difficult to realize our own privileges, so, again, you cannot have a definition of diversity and inclusion for the conference if the team that defines it is not diverse.
+How are you going to have different ideas from the ones you generally have?
+
 \section{Create a safe and welcoming environment}
@@
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTIQA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
+Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
 Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}.
@@
 
-Paying attention to some dimensions of diversity while neglecting others may have unintended consequences.
+Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences.
 Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members.
-In severe cases, they can divert career paths, affect lives, and exclude people from some fields.
+A negative conference experience can divert career paths, affect lives, and exclude people from some fields.
 %Rocío: it would be nice to have reference here.
+%ast: maybe https://link.springer.com/article/10.1007/s11199-017-0800-9
 %Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
@@
 Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, this sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
-Some examples are the newbies sessions for first timers, mixers lead by R-Ladies or MiR, or a leisure activity that reunites subgroups with the same interests: arts, exercise, sports, movies, etc.
+If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
+Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
 
@@
 Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See Rule \ref{rule_language}), or others.
-Break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
-Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while men take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}).
+You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
+Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}).
 
-Words matter.
+Set the tone. %adding here from Liz's rule but maybe this could go in communication in general
 In addition to excluding derogatory or discriminatory language (remember that the code of conduct applies to you too, Rule \ref{rule_inclusion}), make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
-Do not expect people to teach you --it's not their role-- and accept any feedback without being defensive.
+Do not expect people to teach you --it's not their role-- and accept feedback without being defensive.
 
@@
 % https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
-%Natalia: I don't know if this is the best place, but I'd add the term "well-being".
+
+\section{Consciously unbias your spotlight roles}
+\label{rule_spotlight}
+
+When choosing or inviting people as keynote speakers, program committee, session chairs and other spotlight roles, it is likely that there will not be much diversity in the first set of names.
+Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
+Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight.
+Ensuring diversity in each of these roles needs to be a deliberate process.
+We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
+
+Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them.
+An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles.
+The regional and local communities in your field are also good sources to tap into.
+For example, for useR!
+2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
+When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.
+Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged.
+Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.
+
+Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative  ways \cite{cheng2020x+}.
+Give awards (see Rule \ref{rule_accessibility}) to those who prepared accessible slides and presentations, for instance, to acknowledge being mindful of inclusiveness.
+Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve.
+Defy the stereotypical criteria for success by acknowledging these community practices.
+However, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility.
+Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
 
@@
 
-In-person interaction is priceless; however, it is more expensive for some, and even unattainable for others.
-This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
-Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
+In-person interaction is priceless; however, going to conferences is expensive, and many people do not attend for lack of financial support.
+This inaccessibility %ast: maybe inaccessibility is not the word here because we want to focus on disabilities with this word
+is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
+Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
 This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}.
-Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
+\cite{https://link.springer.com/content/pdf/10.1007/s10869-011-9232-7.pdf} found that the order of communication medium influenced perceived inclusion, which in turn influenced individual participation.
+The study suggests that within diverse groups in which minority member is different from the rest of team and may feel excluded, initial computer-mediated communication may help the minority member feel more included.
+Furthermore, an online format is more environmentally-friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
+
+%Yanina: I add the reference about people feeling better using on-line and computer mediated communication when belong to a excluded group.
 
@@
 
-Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles, and accessibility needs.
-Social events in presencial conferences are often the most exclusive, and they are a very common place where you have code of conduct violations.
+Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles and accessibility needs.
+Social events in in-person conferences are often the most exclusive, and they are a very common place where you have code of conduct violations.
 Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond.
@@
 
-\section{Consciously unbias your spotlight roles}
-\label{rule_spotlight}
+\section{Don't let language restrict high-quality participation}
+\label{rule_language}
 
-When choosing or inviting people as keynote speakers, program committee, session chairs and other spotlight roles, it is likely that there will not be much diversity in the first set of names.
-Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
-Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight.
-Ensuring diversity in each of these roles needs to be a deliberate process.
-We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
+In international conferences, English is often the official language.
+Submissions, presentations, tutorials, and workshops are in English.
+The platforms, the webpage, and official communications are also in English.
+While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}.
+Excluding them may potentially lead to missing innovative contributions due to a language barrier.
 
-Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them.
-An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles.
-The regional and local communities in your field are also good sources to tap into.
-For example, for useR!
-2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
-When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged.
-Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.
+Meeting the need of linguistically diverse audience is an important aspect in inclusion that is often overlooked.
+Non-native English speakers may feel intimidated to ask questions or raise their views.
+Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2).
+If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
+Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.
 
-Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative  ways \cite{cheng2020x+}.
-Give awards (see Rule \ref{rule_accessibility}) to those who prepared accessible slides and presentations, for instance, to acknowledge being mindful of inclusiveness.
-Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve.
-Defy the stereotypical criteria for success by acknowledging these community practices.
-However, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility.
-Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
+Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier.
+For instance, hosting one international day/session per conference might be a great place to start!
+
+\section{Express this welcoming spirit in your communication strategy}
+\label{rule_communication}
+%Andrea: this rule name should be more explicit towards communication, "promoting"
+\begin{itemize}
+    \item Promotion: Actively reach out and promote the conference to people who have been systematically excluded/Outreach/promotion plan that reaches communities that are traditionally excluded in scientific meetings
+    \item Express this welcoming spirit in your communication strategy/Be public about your commitment to equity and inclusion: display the coc prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process
+    \item Languages
+    \item Words matter: Be careful how you talk about minoritized people, Assertively set tone of inclusion
+\end{itemize}
+
+If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
+Publish the diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines.
+Fee waivers, financial support, and ease of application/registration should also be advertised; people who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
+Advertise the conference in multiple languages (rule \ref{rule_language}).
+Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
+% Yani: We need to work a little more here, perhaps with some examples.
+% Rocío: could add some ideas from here 
+% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
 
@@
 If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees and visa financial and administrative support (if in-person), or internet connection services (if virtual).
-If the conference is virtual, you may offer technical support so that attendees can record their videos, include captions or design accessible presentations.
+If the conference is virtual, you may offer technical support so that attendees can record their videos, include captions, or design accessible presentations.
 
-\section{Don't let language restrict high-quality participation}
-\label{rule_language}
-
-In international conferences, English is often the official language.
-Submissions, presentations, tutorials, and workshops are in English.
-The platforms, the webpage, and official communications are also in English.
-While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}.
-Excluding them may potentially lead to missing innovative contributions due to a language barrier.
-
-Meeting the need of linguistically diverse audience is an important aspect in inclusion that is often overlooked.
-Non-native English speakers may feel intimidated to ask questions or raise their views.
-Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2).
-If the panellist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
-Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.
-
-Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier.
-For instance, hosting one international day/session per conference might be a great place to start!
-
-\section{Actively reach out and promote the conference to people who have been systematically excluded}
-\label{rule_communication}
-%Andrea: this rule name should be more explicit towards communication, "promoting"
-If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
-Publish the diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines.
-Fee waivers, financial support, and ease of application/registration should also be advertised; people who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory against underrepresented groups.
-Advertise the conference in multiple languages (rule \ref{rule_language}).
-Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
-% Yani: We need to work a little more here, perhaps with some examples.
-% Rocío: could add some ideas from here 
-% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
-
-\section{Do not expect the organizing team to work full time and for free}
+\subsection{Do not expect the organizing team to work full time and for free}
 \label{rule_pay_team}
-
 Having a diverse team and executing inclusive and accessible practices throughout the organizing period --it might be a year or two --may require a lot of effort.
@@
 Be mindful of each particular context, be flexible with hours or commitments, and revise your budget; if possible, secure funds to offer stipends to them early on.
+
+\section{Diversity and inclusion are processes}
+\label{rule_process}
+
+%Sara: I think the last rule should be about the process and reinforce to actually discuss and include diversity as a principle and not as a checklist (as it is on the text).
+Maybe also something about accountability and permanence.
+I think that this rule go be together with the rule_organizing_team given that it is relative to the team work and their dynamics.
+%Sara: copying Yani comment here -  I like the "accountable and permanent" name.
+I think that an idea about progressive improvement can be mention here, like improve something every edition and no go back on the conquers goals.
+Now, to be accountable, you need to be transparent, and you need a clear structure; you get that with the structure of your team, with your declaration and guidelines, and with your action to reinforce those declarations and guidelines (if you have only declarations then you have only empty words, that sound good but you no need to comply.
+We discuss a lot about this).
+For me, one more time, putting together a diverse team is what makes the other thing happened.
+We can cite The tyranny of structurelessness.
+The Second Wave, 2(1), 1972.
+on this rule.
+Perhaps we can talk about governance.
+%ast: i'll put the last version of the tenth rule just to have an idea of how we left this
+%past names: Have a bold vision but be ready to compromise when necessary/Aim for progress not perfection
+%more people rules that have a continuity/accountability theme
+%Aim for Progress, not Perfection
+%Plan, balance and keep working to improve inclusion in the next edition of the conference!
+%Assess whether equity and inclusion goals were met during the meeting
+%Be accountable and permanent
+%Use feedback from previous conferences or other conferences in your field to identify barriers to inclusion and plan around them
+
+This rule is a complement to all the other rules.
+It is important to have a bold vision and clear targets.
+Some parts might be defined as baseline assumptions that are not up for discussion.
+However, there will always be compromises.
+Ideal solutions that meet all our expectations are often not possible.
+Don't let the perfect be the enemy of good.
+As a team, this is a challenging journey.
+Very early on, we tried to recognize that our resources were limited (both in time of the organizing team but also in money).
+We had to let go of certain aspects that were important to many of us.
+For an inclusive conference, we need to avoid giving in because considering all aspects seems impossible and burning ourselves and the team out because considering all aspects is actually impossible.
 

### 2021-09-15 plos_latex_template.tex
Rocío: minor comments in the rules

@@
 Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."
+% Rocío: The paragraph above has quite some redundancy with the paragraph below.
 
@@
 Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
+% Rocío: I'm not sure if "variable" is the right term for non-binary since this is not a statistical analysis but real life.
 Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}.
@@
 
+% Rocío: there's a link missing between this networking part and the paragraph above. This can go in a networking box.
 Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
@@
 The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}.
-The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior, which includes microaggression.
+The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior, which includes microaggression .
+%Rocío: Microagression. I know it's not always evident, but that's a detail I would expect in a CoC, not necessarily here unless explicitly justified.
 An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.
@@
 Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-On the other hand, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}).
+In addition, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}).
 This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees.
 This gave a sense of variety and an appeal of “at least one networking session for everyone”.
+%Rocío: this networking part could go in a box.
 
@@
 2021.
+% Rocío: we need to add that just because there's an online component, it doesn't mean that the in-person part of hybrid conferences can be exclusionary.
 
@@
 Social events in in-person conferences are often the most exclusive, and they are a very common place where you have code of conduct violations.
+% Rocío: I think that we should justify why we say that "social events in in-person conferences are often the most exclusive"
 Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond.
@@
 
+% Rocío: Could start with an introduction (or put somewhere else) that online can easily make conferences more inclusive for many people, but not necessarily for people with disabilities, and that they can be easily left behind unless we actively work on their inclusion.
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
@@
 As Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".
+%Rocío: maybe slightly rephrase the last sentence since it seems to imply that everybody knows Sasha Costanza-Chock.
 
@@
 \label{rule_costs}
+
+% Rocío: Rewrite as a design rule.
 

### 2021-09-17 plos_latex_template.tex
Several changes in rules + reorganization + comments :see_no_evil: (~Rocío, Andrea, Yani, Marce, Sara)

@@
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
-% Rocío: I'm not sure if "variable" is the right term for non-binary since this is not a statistical analysis but real life.
+Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and acknowledging gender is not binary  are just some examples of decisions that can make inclusion real.
 Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}.

### 2021-09-17 plos_latex_template.tex
very very minors- Batool

@@
     \item Promotion: Actively reach out and promote the conference to people who have been systematically excluded/Outreach/promotion plan that reaches communities that are traditionally excluded in scientific meetings
-    \item Express this welcoming spirit in your communication strategy/Be public about your commitment to equity and inclusion: display the coc prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process
+    \item Express this welcoming spirit in your communication strategy: Be public about your commitment to equity and inclusion: display the CoC prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process
     \item Languages
@@
 % https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
+% Batool: I will work on this rule to include more details by 19th of Sept
 

### 2021-09-18 plos_latex_template.tex
RJ: Rewriting rule 1 diversity - haven't added refs yet

@@
 
+% keeping references:
 %   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
@@
 \label{rule_diversity}
-%ast at the end i think we should go back to Embrace all dimensions of diversity
+
 Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
 Human diversity should be celebrated and respected in every way.
-Nonetheless, there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems --including conferences-- are consciously and unconsciously built.
+Nonetheless, we live in a world with implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.
+%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).
 While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have excluded systematically whole groups of people from academia and from scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
 
-Depending on the field, scientific or professional communities can have very different compositions.
-Thinking about which subgroups have been historically excluded from your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event.
-Doing this exercise while you think about the reasons why you want to create a more diverse conference can prevent falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes.
-%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe).
-i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
-%Yanina: They are going to blame what they want to blame.
-You can't control what people are going to do and say and we shouldn't write defensively because we can't take up space in our message with that.
-People who do not want to change will not change because of our paper.
-I imagine this as a guide and help for those people who do want to do something and want to do their best and do not know what or how to do it.
-Or they have already done something and want to continue deepening those changes.
-The rest is not worth spending time and energy.
-But perhaps the goals of the paper change in some other discussion and I lost that.
-Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference.
+To be proactive allies in our conferences and communities, we should start by examining our own privileges (unearned advantages given by society to some people but not all).
 Privilege is largely invisible to those who have it.
-Once we understand our own privileges, we can begin addressing them on an individual and institutional basis.
-%Natalia: I'll look for a reference for privilege definition.
-%marcela: I modified the paragraph a bit, but it needs more work to connect Natalia's and Andrea's ideas. It's just a suggestion, feel free to re-write it again.
-%Batool: I tried making some paraphrasing - It's just a simple suggestion, not sure if it deliver the message you mean it.
-I also think that we may need to define "Privilege" at the 1st mention in the 1st paragraph in this rule.
-Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."
-% Rocío: The paragraph above has quite some redundancy with the paragraph below.
+And we will not be able to help end discrimination and oppression if we do not recognize our privileges first and the injustices that come with them.
+Once we understand our own privileges, we can begin addressing them on an individual, collective, and institutional basis.
+Recognizing our privileges--particularly in our field and in our scientific or professional community--will help identify which subgroups have been the most excluded or discriminated against. 
+These are the groups we need to make more effort to include. 
+Depending on the field, region or community, it could be Black, LGBTQIA+, Muslim, or others. 
+Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.
 
-Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context.
-Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your vision.
-Would this translate into an even gender distribution in your speakers?
-Would it be representation of all races -especially Black people-- among the head organizers, speakers, and attendees?
+Investing more in effort in the most excluded groups does not mean neglecting the others (see Rule 2 for more on that). 
+But it will prevent from falling into the mistake of treating diversity as a checklist.
+It will also guide the vision of diversity for your conference--and your strategies to achieve it:
+Will a more diverse conference would translate into an even gender distribution in your speakers? 
+Would it be representation of all races--especially Black people--among the head organizers, speakers, and attendees? 
 Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
-This is not an invitation to prioritize some dimensions of diversity above others This vision should guide and help hold the organizing team accountable along the way.
+This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way. 
+%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).
 
@@
 
-%Yanina: It's probably a language problem but how I read this rule It is like to give the organizers a carte blanche to define what diversity and inclusion is and with that they are covered.
-What if they decide that everything is fine if they are all white people, native English speakers and without disabilities?
-I don't think we want to encourage that.
-I mean, you need a bottom line and from them you build to improve.
-Plus, as it is well highligth in these rules, it is difficult to realize our own privileges, so, again, you cannot have a definition of diversity and inclusion for the conference if the team that defines it is not diverse.
-How are you going to have different ideas from the ones you generally have?
-
 \section{Create a safe and welcoming environment}
@@
 Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}. 
+%sara: i'd move the intersectionality phrase to the last paragraph of the first rule, right before 'This vision should guide and help hold the organizing team accountable along the way.' Here, I'd just write:
 Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.
@@
 
-% Rocío: there's a link missing between this networking part and the paragraph above. This can go in a networking box.
-Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
-Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
+%i took the networking part to expand communication att.ast we can switch back afterwards
 
@@
 The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}.
-The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior, which includes microaggression .
+The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
 %Rocío: Microagression. I know it's not always evident, but that's a detail I would expect in a CoC, not necessarily here unless explicitly justified.
@@
 Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity.
-Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}} discussing the dimensions of diversity lacking in your community and event).
+Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}}) discussing the dimensions of diversity lacking in your community and event.
 To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
@@
 Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences.
+%sara: inconveniences here doesn't sound good.
+I'd use: the positive outcomes are worth it
 Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}).
+In addition, a diverse team plays an important role at creating a welcoming space because representation -seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes- is one of the best ways to create a sense of belonging (see Rule \ref{rule_inclusion}).
 This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees.
 This gave a sense of variety and an appeal of “at least one networking session for everyone”.
-%Rocío: this networking part could go in a box.
-
-%Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
+%ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general.
+Sara: maybe you can use this as we spoke relating team work here.
+And I think the working together part is not here but the following paragraph
+%sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics
 
 Most advice when working with a diverse team should apply to every team, but is especially true when working with people from marginalized groups.
-First, delegate.
-Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities.
-Nominate directly and offer coordinating positions, even if this means stepping down.
+First, rethink power relationships inside the group and share power.
+Be committed to examine power relations inside your organizing team and challenge it during all the process.
+Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer coordinating positions, and let people from privileged groups step down. 
+Build an environment in which every person has their voice and enforce voices from systematically excluded groups.
 Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See Rule \ref{rule_language}), or others.
@@
 Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}).
+Make work visible, discuss collectively what should be done (and why), divide and account work.
+Being transparent and sharing the information is key to make work visible and collective.
+%sara: this relates to make labor visible from DF, maybe we can cite them here too.
 
+%sara: agree that this fits better in the communication rule
 Set the tone. %adding here from Liz's rule but maybe this could go in communication in general
-In addition to excluding derogatory or discriminatory language (remember that the code of conduct applies to you too, Rule \ref{rule_inclusion}), make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
+In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
 Do not expect people to teach you --it's not their role-- and accept feedback without being defensive.
@@
 \section{Consciously unbias your spotlight roles}
-\label{rule_spotlight}
+\label{rule_spotlight}%sara: rule_unbias?
 
@@
 
-Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them.
+Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them.
 An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles.
@@
 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
-When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.
+When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask for their requirements and barriers, and explicitly invite them to contribute to the conference.
 Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged.
@@
 Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative ways \cite{cheng2020x+}.
-Give awards (see Rule \ref{rule_accessibility}) to those who prepared accessible slides and presentations, for instance, to acknowledge being mindful of inclusiveness.
+You can, for instance, reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), to acknowledge being mindful of inclusiveness.
 Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve.
 Defy the stereotypical criteria for success by acknowledging these community practices.
-However, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility.
+Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility.
 Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
@@
 Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles and accessibility needs.
-Social events in in-person conferences are often the most exclusive, and they are a very common place where you have code of conduct violations.
+Social events in in-person conferences are often the most exclusive (i.e., galas, dinner nights at expensive venues), and they are a very common place where you have code of conduct violations.
 % Rocío: I think that we should justify why we say that "social events in in-person conferences are often the most exclusive"
+%ast: exclusionary? i added a couple of examples
+
 Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond.
@@
 \label{rule_communication}
-%Andrea: this rule name should be more explicit towards communication, "promoting"
+%ast: sara mentioned something that can be good spirit for the communication rule: don't do anything to the outside that you are not doing for your team.
+it's communication strategy but also there should be coherence between your actions and your words.
+
 \begin{itemize}
@@
     \item Words matter: Be careful how you talk about minoritized people, Assertively set tone of inclusion
+    \item coherence between actions and words
+    \item \textbf{networking} when you gather a diverse groups of attendants you also have to take care of the spaces where they will interact
 \end{itemize}
@@
 
-\section{Set fair registration rates and provide financial support when needed}
-\label{rule_costs}
+% Rocío: there's a link missing between this networking part and the paragraph above. This can go in a networking box.
+Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
+If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
+Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
 
+\section{Allocate financial resources to support the presence of diverse people in every space of the conference
+}
+\label{rule_financial}
+%ast: new overall rationale (feel free to comment and change): conferences are expensive to organize and some work on limited budgets.
+this repercutes in fee costs -but fees for attendance are exclusionary- there should be fees, less no-shows.
+fees should be set fairly.
+%scholarships or grants to attend are a common way to boost participation from diverse groups but they attract people who believe they'll win -already privileged- people who do not need them but could use the prestige (for lack of another word) to fit in their cv,
 % Rocío: Rewrite as a design rule.
-
+\subsection{Set fair registration rates and provide financial support when needed}
 Conferences, even virtual ones, should have registration fees for two primary reasons.
@@
 
-Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
+Offering scholarships or grants to attend the conference are a common way to boost participation from diverse groups, and some conferences ask for cover letters or applications  but they may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
 If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g.
@@
 For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
+
 If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees and visa financial and administrative support (if in-person), or internet connection services (if virtual).

### 2021-09-18 plos_latex_template.tex
Rocío: Minor edits to rule 2 inclusion

@@
 
-Investing more in effort in the most excluded groups does not mean neglecting the others (see Rule 2 for more on that). 
+Investing more in effort in the most excluded groups does not mean neglecting the others (see Rule 2 \ref{rule_inclusion} for more on that). 
 But it will prevent from falling into the mistake of treating diversity as a checklist.
@@
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and acknowledging gender is not binary are just some examples of decisions that can make inclusion real.
-Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}. 
-%sara: i'd move the intersectionality phrase to the last paragraph of the first rule, right before 'This vision should guide and help hold the organizing team accountable along the way.' Here, I'd just write:
+Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and acknowledging that gender is not binary are just some examples of decisions that can make inclusion real.
 Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.
@@
 %Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
-
+% Rocío: Andrea, could the newbies part go in the networking (box)? I don't see how to make it fit here. 
 %i took the networking part to expand communication att.ast we can switch back afterwards
@@
 The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
-%Rocío: Microagression. I know it's not always evident, but that's a detail I would expect in a CoC, not necessarily here unless explicitly justified.
 An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.
 
-The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses.
+The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly.
 A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues.
@@
 We strongly recommend reading “How to Respond to Code of Conduct Reports” \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.
-A Code of Conduct response committee needs to be trained and prepared in advance on how to decide on the incident reports.
 
@@
 However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
+% Rocío: This could be moved to communication (Batool, Andrea)
 

### 2021-09-18 plos_latex_template.tex
Rocío: major edits to rule 3 team including pay your team

@@
 
-Investing more in effort in the most excluded groups does not mean neglecting the others (see Rule 2 \ref{rule_inclusion} for more on that). 
+Investing more in effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that). 
 But it will prevent from falling into the mistake of treating diversity as a checklist.
@@
 Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences.
-Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members.
+Lack of representation, unwelcoming--or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members.
 A negative conference experience can divert career paths, affect lives, and exclude people from some fields.
@@
 There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC).
-We strongly recommend reading “How to Respond to Code of Conduct Reports” \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.
+We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.
 
@@
 Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity.
-Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}}) discussing the dimensions of diversity lacking in your community and event.
+Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}}).
 To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
-If you already have an organizing team, check for biases in its composition.
+If you already have an organizing team, check for gaps in its composition.
+%Rocío: I changed "biases" for "gaps".
+I'm sure other statisticians would've found the phrasing weird.
+To really understand what is meant by biases, the reader would have to go to rule unbias, and rules should be self-contained.
+% Rocío: Idea of this paragraph: build a diverse team
 
-%liz: please check if this paragraph connects well with the whole rule.
+Gathering a diverse team will only work if there is real inclusion.
+People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity.
+This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups.
+A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.
+%liz, please check if this paragraph connects well with the whole rule.
 i am using nothing about us without us but also talking about all marginalized groups, in the context of building a team.
 att: andrea
+% Rocío: Idea of this paragraph: real inclusion when having diverse members
 
-Gathering a diverse team will only work if there is real inclusion.
-People with disabilities often say: "Nothing about us without us"; the same holds for other dimensions of diversity.
-This means that the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups.
-A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.
-So, if you plan to organize something related to any community, region, or specific dimension of diversity, involve people from these groups.
+Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes are worthwhile. 
+Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
+In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging (see \textbf{Rule \ref{rule_inclusion}}).
+% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least for now
 
-Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences.
-%sara: inconveniences here doesn't sound good.
-I'd use: the positive outcomes are worth it
-Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an important role at creating a welcoming space because representation -seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes- is one of the best ways to create a sense of belonging (see Rule \ref{rule_inclusion}).
-This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees.
+%This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees.
 This gave a sense of variety and an appeal of “at least one networking session for everyone”.
@@
 
-Most advice when working with a diverse team should apply to every team, but is especially true when working with people from marginalized groups.
+The following advice should apply to any kind of team, but are specially relevant when working with people from marginalized groups.
 First, rethink power relationships inside the group and share power.
 Be committed to examine power relations inside your organizing team and challenge it during all the process.
-Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer coordinating positions, and let people from privileged groups step down. 
+% Rocío: Could we explain power relationship / share power ideas in more simple terms?
+Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
 Build an environment in which every person has their voice and enforce voices from systematically excluded groups.
-Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See Rule \ref{rule_language}), or others.
+Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others.
 You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
-Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}).
+Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).
 Make work visible, discuss collectively what should be done (and why), divide and account work.
@@
 %sara: this relates to make labor visible from DF, maybe we can cite them here too.
+% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
 
-%sara: agree that this fits better in the communication rule
-Set the tone. %adding here from Liz's rule but maybe this could go in communication in general
-In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
-Do not expect people to teach you --it's not their role-- and accept feedback without being defensive.
-
-And most importantly, take care of your team and prioritize their well-being!
-Check on them regularly and make sure everyone is comfortable.
-This represents a great amount of work and you'll need to support each other in the long run.
-
-%Rocío: Yani's ref to have just in case:
-% https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
+Most importantly, take care of your team. 
+Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort.
+The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone.
+However, having a strong community as the only reward may be enough for those who are in more privileged position. 
+Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free. 
+In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
+Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
+Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}). 
+% Rocío: Idea of this paragraph: take care of your team.
 
 \section{Consciously unbias your spotlight roles}
-\label{rule_spotlight}%sara: rule_unbias?
+\label{rule_unbias}
 
@@
 Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
+
+%sara: agree that this fits better in the communication rule
+Set the tone. %adding here from Liz's rule but maybe this could go in communication in general
+In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
+Do not expect people to teach you --it's not their role-- and accept feedback without being defensive.
 

### 2021-09-19 plos_latex_template.tex
Rocío: Rewriting rule online

@@
 Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.
+% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
 
@@
 This gave a sense of variety and an appeal of “at least one networking session for everyone”.
+(Marcela)
 %ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general.
@@
 %sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics
+% Rocío: the rule is getting too long and, since this is focused on networking, I'd prefer it to go to a networking section.
 
@@
 
-Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them.
+Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them.
 An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles.
 The regional and local communities in your field are also good sources to tap into.
-For example, for useR!
-2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
-When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask for their requirements and barriers, and explicitly invite them to contribute to the conference.
+For example, for useR!2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
+When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their requirements and barriers, and explicitly invite them to contribute to the conference.
+% Rocío: ask about their requirements and barriers sounds a bit confusing.
+% Rocío: Not everyone knows what a community of practice is.
+Could we add a definition or use other words?
+Definition: "Communities of practice are groups of people who share a concern, a set of problems, or a passion about a topic, and who deepen their knowledge and expertise in this area by interacting on an ongoing basis".
+(Wenger, E., McDermott, R.
+A., & Snyder, W.
+(2002).
+Cultivating Communities of Practice: A Guide to Managing Knowledge.
+Harvard Business Press.)
 Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged.
@@
 Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
+% Rocío: Since the system usually rewards the most competitive people, appealing cvs and so on, I think it's worth it to explain that rewarding collaborative work is also part of unbiasing; not only because it's mostly done by women, people of color, etc., but because this kind of amazing work is usually neglected in our conferences.
+It may not be obvious to everyone that the content of this paragraph is not only "bring these people to the spotlight", but that it also means unbiasing the conference.
+I don't think I'll have time to add lines about this in the beginning of the paragraph so if anybody could do that, it would be great.
 
@@
 
-In-person interaction is priceless; however, going to conferences is expensive, and many people do not attend for lack of financial support.
-This inaccessibility %ast: maybe inaccessibility is not the word here because we want to focus on disabilities with this word
-is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
+In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
+Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for global conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
+% Rocío: we can add our Nature Correspondence letter to the citations later on. 
 Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
+% Rocío: cite Saliba 2020
 This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}.
-\cite{https://link.springer.com/content/pdf/10.1007/s10869-011-9232-7.pdf} found that the order of communication medium influenced perceived inclusion, which in turn influenced individual participation.
-The study suggests that within diverse groups in which minority member is different from the rest of team and may feel excluded, initial computer-mediated communication may help the minority member feel more included.
-Furthermore, an online format is more environmentally-friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
+% Rocío: we can add Roosetal.
+and our Nature Correspondence letter to the citations here too.
+I would've liked one that talks about increase in participation of tutors and presenters but haven't found any.
+The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year, without depending on a conference central location.
+% Rocío: main idea of the paragraph: online conferences are more inclusive
 
-%Yanina: I add the reference about people feeling better using on-line and computer mediated communication when belong to a excluded group.
+Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings.
+%Rocío: cite Saliba 2020.
+However, social events in in-person conferences can be exclusive (e.g.
+galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur.
+% Rocío: any reference to back the last statement?
+The CoC guide?
+Organizers should invest time in creating opportunities to meet and bond virtually. 
+An online event is a great opportunity to create spaces for people who do not enjoy in person networking. 
+Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
+Respect people's limits, preferences, and remember that `the usual' does not necessarily work with everyone in your community, and that one single networking activity will never serve the whole community.
+It is worth trying new and varied activities that might work with subgroups of attendees.
+Moreover, virtual communication may make people from minoritized groups feel more included, thus participate more (e.g.
+CITATION).
+% RJ: https://link.springer.com/content/pdf/10.1007/s10869-011-9232-7.pdf
+% Rocío: Networking paragraph.
+
+In addition to being more inclusive, an online format is more environmentally-friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
+Besides, if conference materials are systematically stored and pre-recorded, they can be made available to people who missed them live or even after the conference is over. 
+% Rocío: Short paragraph of additional advantages of online conferences that are not about inclusion.
 
@@
 The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
-%Adithi: After the conference is over, the conference materials and videos can be uploaded on free platforms, so that people who were left out of the main event due to any reason, still feel they were cared for and can go to the resources.
-Also when the event is online, a more timezone friendly approach could be used, all the events if online can be global friendly, people with dependents need hours to take care and the night to rest, if the conference is overnight, they lose sleep and suffer, instead if the conference could be friendly for all regions on certain days just like useR!
-2021.
-% Rocío: we need to add that just because there's an online component, it doesn't mean that the in-person part of hybrid conferences can be exclusionary.
-
-Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles and accessibility needs.
-Social events in in-person conferences are often the most exclusive (i.e., galas, dinner nights at expensive venues), and they are a very common place where you have code of conduct violations.
-% Rocío: I think that we should justify why we say that "social events in in-person conferences are often the most exclusive"
-%ast: exclusionary? i added a couple of examples
-
-Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond.
-An online event is a great opportunity to create spaces for people who do not enjoy in person networking.
-Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
-Respect people’s limits, preferences, and remember that “the usual” does not necessarily work with your community, and that one single networking activity will never serve the whole community.
-It is worth trying new and varied activities that might work with subgroups of attendees.
-
-If the conference is online and synchronic (e.g.
-with live sessions), timezones should be considered: you may change the favored timezone year to year, or work on several timezones in each conference edition.
-If the conference is hybrid or in-person, be sure the location does not favor the same people over years, move from one continent to another.
+Needless to say, an inclusive online component of a hybrid conference is not a free pass to have an exclusionary in-person component.
+% Rocío: I know this is a bit aggressive so I wouldn't mind if anybody helps with the wording.
+For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
+% Rocío: Hybrid conferences paragraph.
 
@@
 \end{itemize}
+
+% Rocío: This ref (https://www.nature.com/articles/d41586-020-01489-0 already in the .bib) can be cited for networking
 

### 2021-09-19 plos_latex_template.tex
Rocío: Editing rule 9 financial

@@
 Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
-Do not forget about art shows and movies when making sure that the materials presented are accessible for all.
 Platforms for conference registration and abstract submission, websites, and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
 Captioning for presentations should be available in more than one language if possible.
+Do not forget about making your social events accessible (e.g. art shows and movies) and making sure that the materials presented are accessible for all. 
 These aspects should be tested well in advance of going live.
@@
 
-\section{Express this welcoming spirit in your communication strategy}
+% Rocío: Pasting Batool's text in her 10 simple rules because there are very valuable things we could use: "Rule: Meeting the need of linguistically diverse audience.
+Most international conferences are held in English.
+However, English is only spoken by 4.83 percent of the world population.
+If the panellist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
+Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.
+If possible, ensure that some sessions are delivered by languages other than English"
+
+\section{Express the welcoming spirit in your communication strategy}
 \label{rule_communication}
@@
 it's communication strategy but also there should be coherence between your actions and your words.
+%Rocío: I love that.
+Please add it.
 
@@
 \end{itemize}
-
-% Rocío: This ref (https://www.nature.com/articles/d41586-020-01489-0 already in the .bib) can be cited for networking
 
@@
 
-% Rocío: there's a link missing between this networking part and the paragraph above. This can go in a networking box.
-Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
+% Rocío: pasting this part of the text here because it may go under Andrea's idea for communication: "Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
 If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
-Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
+Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc."
 
-%sara: agree that this fits better in the communication rule
+%sara: this fits better in the communication rule
 Set the tone. %adding here from Liz's rule but maybe this could go in communication in general
@@
 \label{rule_financial}
+
 %ast: new overall rationale (feel free to comment and change): conferences are expensive to organize and some work on limited budgets.
@@
 %scholarships or grants to attend are a common way to boost participation from diverse groups but they attract people who believe they'll win -already privileged- people who do not need them but could use the prestige (for lack of another word) to fit in their cv,
-% Rocío: Rewrite as a design rule.
-\subsection{Set fair registration rates and provide financial support when needed}
+
+\subsection*{Inclusion comes with a cost}
+
+Many inclusive practices come with a cost and it is important to add them to the budget when planning the conference. 
+Good quality and professional captioning, interpreters or translators, accessible and screen-reader friendly registration tools--and chat platform if the conference is online--, paying your organizing team (see \textbf{Rule \ref{rule_team}}), and code of conduct training, are of the some expenses that you may have to include in your budget.
+Budgets are limited, so it is important to estimate costs in advance and define your priorities.
+
+\subsection*{Set fair registration rates and provide financial support when needed}
+
 Conferences, even virtual ones, should have registration fees for two primary reasons.
-The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee.
-The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events than for events with registration costs \cite{eventbrite_ultimate_2017}.
+The first one is that preparing the conference costs money and this is likely to translate to the participation fee.
+The second reason has to do with psychology: people value more the things they pay for, and there is a lower attendance rate for free events than for events with registration costs \cite{eventbrite_ultimate_2017}.
 
@@
 A lower fee for non-profit organizations, government employees, or freelancers should also be considered.
-It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
-Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined.
-You can take into account the cost of living in each country using conversion factors (e.g.
+It is important to include discounts for students to encourage their participation \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
+Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, they should be allowed to locate themselves in a category they consider affordable (i.e.
+student, researcher or `pay what you can').
+To fix the registration costs, you can take into account the cost of living in each country using conversion factors (e.g.
 from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}).
-Consider that, while it is traditional for employers to provide --at least some --conference support in the Global North, this is not the case everywhere.
+Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere.
 
 Offering scholarships or grants to attend the conference are a common way to boost participation from diverse groups, and some conferences ask for cover letters or applications  but they may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
-If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g.
-`grant recipient' or `diversity scholar', respectively).
-Do not call it a scholarship or a grant.
-Ultimately, it is a fee waiver or a discount for the people who need additional financial support.
 Also, the process for applicants should be simplified.
@@
 Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support.
-For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
+For this stage of the process, a certain degree of trust goes a long way-- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
+Only if the applicants for fee waivers are numerous and your budget cannot allow for that many fee waivers, you could ask for a short application.
+You could try to reduce this risk by finding enough sponsoring money so that you do not depend on registration fees to pay for your conference.
 
@@
 
-\subsection{Do not expect the organizing team to work full time and for free}
-\label{rule_pay_team}
-Having a diverse team and executing inclusive and accessible practices throughout the organizing period --it might be a year or two --may require a lot of effort.
-The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone.
-However, having a strong community as the only reward may be enough if you are in a privileged position.
-Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free.
-In addition, tasks such as taking and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
-Be mindful of each particular context, be flexible with hours or commitments, and revise your budget; if possible, secure funds to offer stipends to them early on.
+% If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g.
+`grant recipient' or `diversity scholar', respectively).
+Do not call it a scholarship or a grant.
+Ultimately, it is a fee waiver or a discount for the people who need additional financial support.
+% Rocío: I removed the part of naming the scholarship/grant/waiver, because if it goes somewhere it should be communication. Andrea, Batool, please see if it fits there.
 

### 2021-09-19 plos_latex_template.tex
Rocío: mostly editing and updating references

@@
 Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
-However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. %RJ: (Biggs et al. 2018) https://link.springer.com/article/10.1007/s11199-017-0800-9
+However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
 Conversely, efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact in our communities and the lives and careers of their members.
@@
 Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.
-% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
+% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure
 
@@
 Nonetheless, we live in a world with implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.
-%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).
 While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have excluded systematically whole groups of people from academia and from scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
@@
 This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way.
-%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).
 
@@
 %batool: please note that I didn't read both entire papers I mentioned above.
+%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).
+% Rocío: I feel like we don't need to cite any reference about privilege the way it is now. But feel free to disagree.
 
@@
 Lack of representation, unwelcoming--or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members.
-A negative conference experience can divert career paths, affect lives, and exclude people from some fields.
-%Rocío: it would be nice to have reference here.
-%ast: maybe https://link.springer.com/article/10.1007/s11199-017-0800-9
+A negative conference experience can divert career paths, affect lives, and exclude people from some fields \cite{biggsAcademicConferenceChilly2018}.
 %Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
@@
 Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
-The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}.
+The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
 The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
@@
 There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC).
-We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.
-
-%Natalia: the book is by Aurora and Gardiner (2019), I corrected the title and added the reference including the link
+We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct team work.
 
@@
 %sara: this relates to make labor visible from DF, maybe we can cite them here too.
+%Rocío: sure.
 % Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
@@
 % Rocío: we can add our Nature Correspondence letter to the citations later on. 
-Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities.
-% Rocío: cite Saliba 2020
-This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}.
-% Rocío: we can add Roosetal.
-and our Nature Correspondence letter to the citations here too.
-I would've liked one that talks about increase in participation of tutors and presenters but haven't found any.
+Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
+This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020}.
+% Rocío: we can add our Nature Correspondence letter to the citations here too. I would've liked one that talks about increase in participation of tutors and presenters but haven't found any. 
 The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year, without depending on a conference central location.
@@
 
-Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings.
-%Rocío: cite Saliba 2020.
+Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings \cite{salibaGettingGripsOnline2020}.
 However, social events in in-person conferences can be exclusive (e.g.
@@
 It is worth trying new and varied activities that might work with subgroups of attendees.
-Moreover, virtual communication may make people from minoritized groups feel more included, thus participate more (e.g.
-CITATION).
-% RJ: https://link.springer.com/content/pdf/10.1007/s10869-011-9232-7.pdf
+Moreover, virtual communication may make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012}).
 % Rocío: Networking paragraph.
@@
 Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
-Platforms for conference registration and abstract submission, websites, and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
+Platforms for conference registration and abstract submission, websites, and chat platforms--if used--should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
 Captioning for presentations should be available in more than one language if possible.
@@
 Many inclusive practices come with a cost and it is important to add them to the budget when planning the conference. 
-Good quality and professional captioning, interpreters or translators, accessible and screen-reader friendly registration tools--and chat platform if the conference is online--, paying your organizing team (see \textbf{Rule \ref{rule_team}}), and code of conduct training, are of the some expenses that you may have to include in your budget.
+Good quality and professional captioning, interpreters or translators, accessible and screen-reader friendly registration tools--and chat platform if the conference is online--, paying your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), and code of conduct training, are of the some expenses that you may have to include in your budget.
 Budgets are limited, so it is important to estimate costs in advance and define your priorities.
@@
 Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team.
-Always remember inclusion is an ambition that organisers can and should continually pursue by allowing room for learning and improvement.
+Always remember inclusion is an ambition that organizers can and should continually pursue by allowing room for learning and improvement.
 If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
@@
 \section*{Acknowledgments}
-The authors of this piece would like to thank every single member of the organizing team of useR!
-2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR!
-2021 and supporting us through the process.
+The authors of this piece would like to thank every single member of the organizing team of useR!2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR!2021 and supporting us through the process.
 

### 2021-09-23 plos_latex_template.tex
some changes by Batool and Liz's revision and comments

@@
 att: andrea
+% Liz: Yes!
+
 % Rocío: Idea of this paragraph: real inclusion when having diverse members
@@
 
+% Liz: I know this section is getting long, and maybe this is too blunt, but: 
+% People who have experience with exclusion have the social and technical 
+% expertise about removing barriers to inclusion.Actively ask where the barriers are and how to mitigate them.
+
 The following advice should apply to any kind of team, but are specially relevant when working with people from marginalized groups.
@@
 Build an environment in which every person has their voice and enforce voices from systematically excluded groups.
+% Liz: Use of "voice" in this way makes deaf people feel excluded.
+Maybe it could be something like "Build an environment in which every person has input..." Or anything as long as it doesn't suggest you have to have a voice to participate.
 Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others.
@@
 
+% Liz: I have a little concern about mentoring WRT English, but I'm a white English speaker, so take it with a grain of salt.
+I think people who look like me need to do more accepting and less mentoring around English.
+If someone explicitly wants my help, I'm happy to give it because I love words and language, and I might learn something cool about another language.
+But I'm not sure we should be telling the English-priveleged readers that they are language mentors.
+
 Most importantly, take care of your team. 
@@
 Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free. 
+% Liz: Yes!
+This is especially true when *some* people are paid to work on the conference.
+I was just asked to do accessibility work that should have been done by the contracted, paid organizer.
+This is hugely problematic and will be true for lots of conferences outside the open software community.
 In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
@@
 % Rocío: Idea of this paragraph: take care of your team.
+% Liz: also, include people from diverse backgrounds in all aspects of planning, not just the diversity-related ones, unless that is their only interest.
+In some organizations, I could be on the scientific committee and participate in many other ways, but I'm only called on for accessibility.
 
@@
 Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.
+% Liz: maybe consider a broader group than "early career researchers." One problem is that marginalized people have many fewer opportunities, and as they age that attenuation of opportunities compounds.
+Someone can be in the margins even years after finishing grad school or being a postdoc (not to mention that people without degrees could have contributions.
 
@@
 I don't think I'll have time to add lines about this in the beginning of the paragraph so if anybody could do that, it would be great.
+
+% Liz: I might be able to, but I'm not sure where you want the additional text; where the beginning of the paragraph is because of how this is formatted.
+I couldn't agree more that being truly inclusive means not restricting pepole's role.
+Every subteam should be diverse, in addition to the whole organizing team.
+It's not just about representation but needs to encompass integration across the whole event.
 
@@
 % Rocío: main idea of the paragraph: online conferences are more inclusive
+% Liz: but only if you make them that way. The R conference was a standout and innovative compared to others I've considred attending since Covid.
 
@@
 % Rocío: I know this is a bit aggressive so I wouldn't mind if anybody helps with the wording.
+% Liz: I like it, but you could say
+% an inclusive online component doesn't eliminate the need to organize an inclusive in-person component.
+
 For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
@@
 For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
+% Liz: I don't think it's actually ironic. maybe surprising or counterintuitive?
 
 If the conference is in person, the venue must be accessible for people with mobility limitations. 
+$ Liz: let's stay away from deficit-model language like "limitations." Possible aternatives:
+% people who use wheelchairs or other assistive mobility devices
+% ... must be physically accessible for disabled people
+
 Accommodations such as a quiet space for neurodivergent people should also be provided.
+% Liz: I hate "accommodateions because it sounds like something optional or for comfort or luxurious Consider something like "facilities"?
+
 Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
+% Liz: avoid using disabilities as nouns like -- "the blind", "the deaf" etc. I woul duse "people who are deaf or hard-of-hearing" above. There's controversy about whether the word person
+% should come first ("person who is blind") or the identity should come first ("blind person:) but the person should always be acknowledged (like "participants with disabilities", blind scientist", etc..
+It's kind of objectifying without a person-related word.
 Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
+% Liz: not just alt-text, the whole site should be screen-reader accessible so:
+% communication about the conference should be fully screen-reader accessible including alt-text, and any videos should have both captions and sound.
+
 Platforms for conference registration and abstract submission, websites, and chat platforms--if used--should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
@@
 
+% Liz: It is essential to provide an accessible version of the slides for people who won't be seeing them One of the most important assumptions about presentations are that they will be both seen and heard.
+Usually, neither of these channels is a complete source of information about the presentation.
+For example, when discussing computer programming, code is usually displayed on slides while the presenter talks about them.
+They seldom read the details of what appears on the slides in enough detail for a blind person to learn, and a deaf person will also miss some explanation.
+Presenters  should provide as much information as possible on both channels, but in order to compensate for the fact that this is not common practice, blind people need to be able to read an accessible version of the slides at the same time they listen to the presentation.
+It is critical to make these materials available before the conference starts..
+
 Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
@@
 Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
+% Liz: "for people who use whellchairs" rather than "for wheelchairs"
+
 On the other hand, it is likely that financial constraints may not allow you to make your conference completely accessible and there will be choices to make (e.g.
@@
 Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
-As Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".
+% Liz: maybe strengthen the last phrase to use their expertise to solve accessibility problems and collaborate on decisions about accessibility practices.
+
+% Liz: As design justice author Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".
 %Rocío: maybe slightly rephrase the last sentence since it seems to imply that everybody knows Sasha Costanza-Chock.
+% not sure if that's a good characterization-- I am just learning about her right now.
 
@@
 
-Meeting the need of linguistically diverse audience is an important aspect in inclusion that is often overlooked.
+Meeting the need 
+% Liz: People with disabilities don't want to be considered "needy".
+We reject the term "special needs" in part because it frames the person with the disability as being the burden causing the accommodation, where maybe it's the responsibility of society to be open to everyone.
+So I wonder if it's the same for people who speak other languages.
+Maybe they are not needy or inferior.
+Or maybe it's wrong for me to apply disability rights philosophy to language.
+% The linguistic diversity of conference participants is often overlooked,
+% resulting in missed opportunities for them, as well as for the conference
+% to benefit from their contributions.
+
+of linguistically diverse audience is an important aspect in inclusion that is often overlooked.
+
  Non-native English speakers may feel intimidated to ask questions or raise their views.
@@
     \item Promotion: Actively reach out and promote the conference to people who have been systematically excluded/Outreach/promotion plan that reaches communities that are traditionally excluded in scientific meetings
+% Liz: including explicit descriptions of accessibility and other inclusion practices.
+Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers.
     \item Express this welcoming spirit in your communication strategy: Be public about your commitment to equity and inclusion: display the CoC prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process
@@
 Do not expect people to teach you --it's not their role-- and accept feedback without being defensive.
-
+$ Liz: should be "without being offended"
 \section{Allocate financial resources to support the presence of diverse people in every space of the conference

### 2021-09-25 plos_latex_template.tex
update date
AST: addressing some of Liz's comments and tiny changes

@@
 It will also guide the vision of diversity for your conference--and your strategies to achieve it:
-Will a more diverse conference would translate into an even gender distribution in your speakers? 
-Would it be representation of all races--especially Black people--among the head organizers, speakers, and attendees? 
+Will a more diverse conference translate into an even gender distribution in your speakers? 
+Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
 Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
@@
 Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
-Build an environment in which every person has their voice and enforce voices from systematically excluded groups.
+Build an environment in which every person can express their position and give prevalence to the positions from systematically excluded groups.
 % Liz: Use of "voice" in this way makes deaf people feel excluded.
 Maybe it could be something like "Build an environment in which every person has input..." Or anything as long as it doesn't suggest you have to have a voice to participate.
-Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others.
+%i rephrased slightly, avoiding "support" it's not a question of privileged people giving support but of positions being given importance
+Offer support and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others.
 You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
@@
 But I'm not sure we should be telling the English-priveleged readers that they are language mentors.
-
+%ast. i agree with liz and wrote this from the point of view of a non-native speaker- in terms of support rather than mentorship. replaced with support but feel free to check
 Most importantly, take care of your team. 
@@
 For example, for useR!2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
-When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their requirements and barriers, and explicitly invite them to contribute to the conference.
+When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their barriers for participation, and explicitly invite them to contribute to the conference.
 % Rocío: ask about their requirements and barriers sounds a bit confusing.
@@
 Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged.
-Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.
+Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers and people from minoritized groups, fostering collaboration and building their skills.
 % Liz: maybe consider a broader group than "early career researchers." One problem is that marginalized people have many fewer opportunities, and as they age that attenuation of opportunities compounds.
@@
 However, social events in in-person conferences can be exclusive (e.g.
-galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur.
+galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}.
 % Rocío: any reference to back the last statement?
 The CoC guide?
+ast: yes
 Organizers should invest time in creating opportunities to meet and bond virtually. 
@@
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
-Ironically, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people.
+Importantly, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people.
 For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
-% Liz: I don't think it's actually ironic. maybe surprising or counterintuitive?
+% Liz: I don't think it's actually ironic. maybe surprising or counterintuitive? ast: absolutely! i put importantly at the end to highlight that without opposition
 
 If the conference is in person, the venue must be accessible for people with mobility limitations. 
-$ Liz: let's stay away from deficit-model language like "limitations." Possible aternatives:
+% Liz: let's stay away from deficit-model language like "limitations." Possible aternatives:
 % people who use wheelchairs or other assistive mobility devices
@@
 
-Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
+Presenters should always speak into a microphone to make it easier for people who are hard of hearing and for captioners or interpreters to listen to them. 
 % Liz: avoid using disabilities as nouns like -- "the blind", "the deaf" etc. I woul duse "people who are deaf or hard-of-hearing" above. There's controversy about whether the word person
@@
 They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}. 
-Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
+Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs in an in-person conference. 
 % Liz: "for people who use whellchairs" rather than "for wheelchairs"
@@
 However, English is only spoken by 4.83 percent of the world population.
-If the panellist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
+If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
 Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.
@@
     \item Promotion: Actively reach out and promote the conference to people who have been systematically excluded/Outreach/promotion plan that reaches communities that are traditionally excluded in scientific meetings
-% Liz: including explicit descriptions of accessibility and other inclusion practices.
+    \item Include explicit descriptions of accessibility and other inclusion practices in your communication strategy.
 Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers.
@@
 % https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
-% Batool: I will work on this rule to include more details by 19th of Sept
+% Batool: I will work on this rule to include more details by 25th of Sept
 
@@
 In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
-Do not expect people to teach you --it's not their role-- and accept feedback without being defensive.
-$ Liz: should be "without being offended"
+Do not expect people to teach you --it's not their role-- and accept feedback without being offended.
+
 \section{Allocate financial resources to support the presence of diverse people in every space of the conference

### 2021-10-05 plos_latex_template.tex
added marks2021 reference on conferences, diversity & disability to .bib and suggested places for it in .tex
Updates from Overleaf

@@
 Please add it.
+An inclusive communication is to give each individual within the conference a voice and sense of belonging.
+Embedding inclusion in your communication strategy can only come by design, it shouldn’t be a thoughtless default.
+In the essense of your communication strategy, you should pritorise and attend to the needs of your team, in other words, don't do anything to the outside that you are not doing for your team.
+As you plan the conference, actively reach out and promote the conference to people who have been systematically excluded.
+Design a promotion plan that reaches communities that are traditionally excluded in scientific meetings.
+This could be by including representatives from these communities in the organising team.Always  include explicit descriptions of accessibility and other inclusion practices.
+Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers.
 
 \begin{itemize}
-    \item Promotion: Actively reach out and promote the conference to people who have been systematically excluded/Outreach/promotion plan that reaches communities that are traditionally excluded in scientific meetings
-    \item Include explicit descriptions of accessibility and other inclusion practices in your communication strategy.
-Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers.
     \item Express this welcoming spirit in your communication strategy: Be public about your commitment to equity and inclusion: display the CoC prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process

### 2021-10-05 plos_latex_template.tex
refine rule 8 - Batool

@@
 %   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
+% Liz adding reference marks2021meeting
 
@@
 % Rocío: Could start with an introduction (or put somewhere else) that online can easily make conferences more inclusive for many people, but not necessarily for people with disabilities, and that they can be easily left behind unless we actively work on their inclusion.
+% Liz: let me know if you want me to take a shot at this.adding reference marks2021meeting
+
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
@@
 An inclusive communication is to give each individual within the conference a voice and sense of belonging.
-Embedding inclusion in your communication strategy can only come by design, it shouldn’t be a thoughtless default.
-In the essense of your communication strategy, you should pritorise and attend to the needs of your team, in other words, don't do anything to the outside that you are not doing for your team.
+Embedding inclusion in your communication strategy can only come by design, it should not be a thoughtless default.
+In the essence of your communication strategy, there should be coherence between actions and words.
+Always attend to the needs of your team, in other words, don't do anything to the outside that you are not doing for your team.
 As you plan the conference, actively reach out and promote the conference to people who have been systematically excluded.
 Design a promotion plan that reaches communities that are traditionally excluded in scientific meetings.
-This could be by including representatives from these communities in the organising team.Always  include explicit descriptions of accessibility and other inclusion practices.
+This could be by including representatives from these communities in the organizing team.
+Always include explicit descriptions of accessibility and other inclusion practices.
 Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers.
+Express this welcoming spirit in your communication strategy: Be public about your commitment to equity and inclusion: display the CoC prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process.
+Be careful with the language you use in promoting to the conference and how you talk about minoritized people, Assertively set tone of inclusion.
 

### 2021-10-05 plos_latex_template.tex
polish rule 8 - Batool

@@
 Always attend to the needs of your team, in other words, don't do anything to the outside that you are not doing for your team.
-As you plan the conference, actively reach out and promote the conference to people who have been systematically excluded.
-Design a promotion plan that reaches communities that are traditionally excluded in scientific meetings.
-This could be by including representatives from these communities in the organizing team.
-Always include explicit descriptions of accessibility and other inclusion practices.
-Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers.
-Express this welcoming spirit in your communication strategy: Be public about your commitment to equity and inclusion: display the CoC prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process.
-Be careful with the language you use in promoting to the conference and how you talk about minoritized people, Assertively set tone of inclusion.
 
-\begin{itemize}
-    \item Express this welcoming spirit in your communication strategy: Be public about your commitment to equity and inclusion: display the CoC prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process
-    \item Languages
-    \item Words matter: Be careful how you talk about minoritized people, Assertively set tone of inclusion
-    \item coherence between actions and words
-    \item \textbf{networking} when you gather a diverse groups of attendants you also have to take care of the spaces where they will interact
-\end{itemize}
+As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded.
+If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
+This could through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct and accessibility guidelines, and providing fee waivers and financial support.
+People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
+If possible, include representative from these communities within the organizing team (rule \ref{rule_diversity}).
+Also, try to advertise the conference in multiple languages (rule \ref{rule_language}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
 
-If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
-Publish the diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines.
-Fee waivers, financial support, and ease of application/registration should also be advertised; people who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
-Advertise the conference in multiple languages (rule \ref{rule_language}).
-Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
+Always include explicit descriptions of accessibility and other inclusion practices (rule \ref{rule_accessibility}).
+Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusions.
+Be careful with the language you use in promoting to the conference and how you talk about minoritized people, assertively set tone of inclusion.
+In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
+Do not expect people to teach you --it's not their role-- and accept feedback without being offended.
+
+Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
+If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
+Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
+
 % Yani: We need to work a little more here, perhaps with some examples.
@@
 %sara: this fits better in the communication rule
-Set the tone. %adding here from Liz's rule but maybe this could go in communication in general
-In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
+%adding here from Liz's rule but maybe this could go in communication in general
+%In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
 Do not expect people to teach you --it's not their role-- and accept feedback without being offended.

### 2021-10-06 plos_latex_template.tex
polish rule 8 - Batool

@@
 Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs in an in-person conference.
-% Liz: "for people who use whellchairs" rather than "for wheelchairs"
 
@@
 Please add it.
-An inclusive communication is to give each individual within the conference a voice and sense of belonging.
+An inclusive communication is to give each individual within the conference the opportunity to express themselves and a sense of belonging.
 Embedding inclusion in your communication strategy can only come by design, it should not be a thoughtless default.
@@
 People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
-If possible, include representative from these communities within the organizing team (rule \ref{rule_diversity}).
+If possible, include representative from these communities within the organizing team (rule \ref{rule_inclusion}).
 Also, try to advertise the conference in multiple languages (rule \ref{rule_language}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
@@
 
-Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
+Although conference attendees should have ample opportunities to network with each other, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
+If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community(Rule \ref{rule_unbias})).
 Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
@@
 % Batool: I will work on this rule to include more details by 25th of Sept
-
-% Rocío: pasting this part of the text here because it may go under Andrea's idea for communication: "Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community.
-Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc."
 
@@
 \label{rule_financial}
-
-%ast: new overall rationale (feel free to comment and change): conferences are expensive to organize and some work on limited budgets.
-this repercutes in fee costs -but fees for attendance are exclusionary- there should be fees, less no-shows.
-fees should be set fairly.
-%scholarships or grants to attend are a common way to boost participation from diverse groups but they attract people who believe they'll win -already privileged- people who do not need them but could use the prestige (for lack of another word) to fit in their cv,
 

### 2021-10-06 plos_latex_template.tex
minor changes - Batool

@@
 
-Investing more in effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that). 
+Investing more effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that). 
 But it will prevent from falling into the mistake of treating diversity as a checklist.
@@
 % Liz: I don't think it's actually ironic. maybe surprising or counterintuitive? ast: absolutely! i put importantly at the end to highlight that without opposition
-
 If the conference is in person, the venue must be accessible for people with mobility limitations. 
@@
 % ... must be physically accessible for disabled people
-
 Accommodations such as a quiet space for neurodivergent people should also be provided.
@@
 % to benefit from their contributions.
-
 of linguistically diverse audience is an important aspect in inclusion that is often overlooked.
-
 Non-native English speakers may feel intimidated to ask questions or raise their views.
@@
 People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
-If possible, include representative from these communities within the organizing team (rule \ref{rule_inclusion}).
-Also, try to advertise the conference in multiple languages (rule \ref{rule_language}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
+If possible, include representatives from these communities within the organizing team (\textbf{Rule \ref{rule_inclusion}}).
+Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
 
-Always include explicit descriptions of accessibility and other inclusion practices (rule \ref{rule_accessibility}).
+Always include explicit descriptions of accessibility and other inclusion practices (\textbf{Rule \ref{rule_accessibility}}).
 Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusions.
@@
 
-Although conference attendees should have ample opportunities to network with each other, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community(Rule \ref{rule_unbias})).
+As we mentioned previously, conference attendees should have ample opportunities to network with each other.
+However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
+If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
 Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.

### 2021-10-08 plos_latex_template.tex
ast: financial rule + general reading, all WIP.

@@
 useR!
-conferences are non-profit events, organized by volunteers from the R community and arranged by the R Foundation.
+conferences are non-profit events organized by volunteers from the R community and arranged by the R Foundation.
 The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector.
@@
 Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
-However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
-Conversely, efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact in our communities and the lives and careers of their members.
+However, they are likely to reproduce the systematic discrimination occurring in other spaces in our fields and become [bad experiences]. 
+Efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact on our communities and the lives and careers of their members.
 
@@
 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
-We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive and accessible experience for everyone. 
-Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.
-% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure
+We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive, and accessible experience for everyone. 
+Here, we share the lessons learned within the past year of organizing this global useR! 2021, summarized as ten simple rules towards an inclusive conference.
+We have organized these rules in three groups, because they refer to the foundations (Rule 1 and 2), the actions regarding people (Rules 3 to 5), the actions regarding conference design (Rules 6 to 9), and a general synthesis %ast this is an attempt, but i am not a fan of rule 10 alone at the end.
+actually the whole classification when written seems off to me (right now, this might change)
+% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure [x]
 
@@
 Human diversity should be celebrated and respected in every way.
-Nonetheless, we live in a world with implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.
-While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have excluded systematically whole groups of people from academia and from scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
+Nonetheless, we live in a world with implicit hierarchies along these axes.
+Some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.
+While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
 
@@
 These are the groups we need to make more effort to include. 
-Depending on the field, region or community, it could be Black, LGBTQIA+, Muslim, or others. 
+Depending on the field, region, or community, it could be Black People, LGBTQIA+, Muslim, or others. 
+%ast: ö check phrasing, add "people"
 Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.
@@
 A negative conference experience can divert career paths, affect lives, and exclude people from some fields \cite{biggsAcademicConferenceChilly2018}.
-%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
-% Rocío: Andrea, could the newbies part go in the networking (box)? I don't see how to make it fit here. 
-%i took the networking part to expand communication att.ast we can switch back afterwards
+%ast maybe this last sentence could go in the introduction because it's general i marked [bad experience] in the intro
 
@@
 There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC).
-We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct team work.
-
-%Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}.
-However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
-% Rocío: This could be moved to communication (Batool, Andrea)
+We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
 
@@
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
-Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity.
-Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}}).
-To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
+Build a team with people from different regions, genders, races, socioeconomic statuses, and other aspects of diversity.
+Go beyond balancing all genders in this effort, and pay attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}).
+To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large.
 If you already have an organizing team, check for gaps in its composition.
-%Rocío: I changed "biases" for "gaps".
-I'm sure other statisticians would've found the phrasing weird.
-To really understand what is meant by biases, the reader would have to go to rule unbias, and rules should be self-contained.
+
+%ast: I added race -even if race is a construct- because focus is given to gender and the spaces tend to be predominantly white and US/Europe-centric.
+In other parts of the text I was using racialized but here we can keep it simple, i think.
+another thing: i avoided "gender balancing" because idk if that comes across as too binary, even if what is done is adding cis women to the team
+
 % Rocío: Idea of this paragraph: build a diverse team
@@
 People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity.
-This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups.
+This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups [cite design justice].
 A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.
-%liz, please check if this paragraph connects well with the whole rule.
-i am using nothing about us without us but also talking about all marginalized groups, in the context of building a team.
-att: andrea
-% Liz: Yes!
 
@@
 %This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees.
-This gave a sense of variety and an appeal of “at least one networking session for everyone”.
+This gave a sense of variety and an appeal of "at least one networking session for everyone".
 (Marcela)
@@
 % People who have experience with exclusion have the social and technical 
-% expertise about removing barriers to inclusion.Actively ask where the barriers are and how to mitigate them.
+% expertise about removing barriers to inclusion. Actively ask where the barriers are and how to mitigate them.
 
-The following advice should apply to any kind of team, but are specially relevant when working with people from marginalized groups.
+The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups.
 First, rethink power relationships inside the group and share power.
@@
 Maybe it could be something like "Build an environment in which every person has input..." Or anything as long as it doesn't suggest you have to have a voice to participate.
-%i rephrased slightly, avoiding "support" it's not a question of privileged people giving support but of positions being given importance
+%ast: i rephrased slightly, avoiding "support" it's not a question of privileged people giving support but of positions being given importance
 Offer support and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others.
 You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
-Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).
+Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).
 Make work visible, discuss collectively what should be done (and why), divide and account work.
@@
 
-When choosing or inviting people as keynote speakers, program committee, session chairs and other spotlight roles, it is likely that there will not be much diversity in the first set of names.
+%idea1: our lists are biased
+
+When choosing or inviting people as keynote speakers, program committee, session chairs, and other spotlight roles, it is likely that there will not be much diversity in the first set of names.
 Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
@@
 When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their barriers for participation, and explicitly invite them to contribute to the conference.
-% Rocío: ask about their requirements and barriers sounds a bit confusing.
 % Rocío: Not everyone knows what a community of practice is.
@@
 Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative ways \cite{cheng2020x+}.
-You can, for instance, reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), to acknowledge being mindful of inclusiveness.
+You can, for instance, reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
 Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve.
@@
 % Liz: I might be able to, but I'm not sure where you want the additional text; where the beginning of the paragraph is because of how this is formatted.
-I couldn't agree more that being truly inclusive means not restricting pepole's role.
+I couldn't agree more that being truly inclusive means not restricting people's role.
 Every subteam should be diverse, in addition to the whole organizing team.
 It's not just about representation but needs to encompass integration across the whole event.
+
+% ast: i agree a lot with Rocío here, the change in what is given importance is a key part of the unbiasing.
+i had the impression that the part of not restricting people's role goes above, in team rule.
 
@@
 
-\section{Allocate financial resources to support the presence of diverse people in every space of the conference
-}
+\section{Allocate financial resources to support your conference goals}
 \label{rule_financial}
 
-\subsection*{Inclusion comes with a cost}
+%Inclusion comes with a cost
 
-Many inclusive practices come with a cost and it is important to add them to the budget when planning the conference. 
-Good quality and professional captioning, interpreters or translators, accessible and screen-reader friendly registration tools--and chat platform if the conference is online--, paying your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), and code of conduct training, are of the some expenses that you may have to include in your budget.
-Budgets are limited, so it is important to estimate costs in advance and define your priorities.
+Conference budgets are limited and rely mostly on sponsorship and attendance fees.
+Some expenses are more or less fixed but allocation of resources has to be intentional to support the goals of the conference.
+It is important to estimate the costs for these inclusive practices and define your priorities in advance.
+This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership training, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}).
+It also includes additional support for attendees: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
+While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team.
+Don't be afraid to innovate and resist the institutional inertia.
+"This has always been like that" will not take the conference towards structural change.
+%should we talk about the pandemic? something like: 
+With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
+Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.
 
-\subsection*{Set fair registration rates and provide financial support when needed}
-
-Conferences, even virtual ones, should have registration fees for two primary reasons.
-The first one is that preparing the conference costs money and this is likely to translate to the participation fee.
-The second reason has to do with psychology: people value more the things they pay for, and there is a lower attendance rate for free events than for events with registration costs \cite{eventbrite_ultimate_2017}.
-
-On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates.
+%Set fair registration rates
+Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates.
 Usually, there is a higher fee for people from the industry than for academia.
 A lower fee for non-profit organizations, government employees, or freelancers should also be considered.
-It is important to include discounts for students to encourage their participation \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
-Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, they should be allowed to locate themselves in a category they consider affordable (i.e.
-student, researcher or `pay what you can').
-To fix the registration costs, you can take into account the cost of living in each country using conversion factors (e.g.
-from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}).
+It is important to include discounts for students \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012} and other early career stages, such as postdoc or trainee.
+The definition of such positions and the associated work conditions vary for each country.
 Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere.
+In addition to this, low rates in strong currencies can represent significant amounts of money in other currencies.
+You can adjust the registration fee by the cost of living in each country using conversion factors (e.g.
+from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example).
+In general, people should be allowed to locate themselves in a category they consider affordable, even with the possibility of a "pay what you can" approach.
+Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}.
+Offering fee waivers to part of the participants is also a good option, but even then, other costs can be prohibitive.
 
-Offering scholarships or grants to attend the conference are a common way to boost participation from diverse groups, and some conferences ask for cover letters or applications  but they may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them.
-Also, the process for applicants should be simplified.
+Offering scholarships or grants to attend the conference is a common way to boost participation from different groups, and some conferences ask for cover letters or applications to them.
+These applications may be counterproductive, if they attract people that do not need the money but apply to enrich their CVs, and can filter out people who do not feel entitled to earn them.
 Applying for loans, grants, and scholarships may be an emotionally demanding task.
 Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support.
-For this stage of the process, a certain degree of trust goes a long way-- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility.
-Only if the applicants for fee waivers are numerous and your budget cannot allow for that many fee waivers, you could ask for a short application.
-You could try to reduce this risk by finding enough sponsoring money so that you do not depend on registration fees to pay for your conference.
-
-If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees and visa financial and administrative support (if in-person), or internet connection services (if virtual).
-If the conference is virtual, you may offer technical support so that attendees can record their videos, include captions, or design accessible presentations.
-
-% If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g.
-`grant recipient' or `diversity scholar', respectively).
-Do not call it a scholarship or a grant.
-Ultimately, it is a fee waiver or a discount for the people who need additional financial support.
-% Rocío: I removed the part of naming the scholarship/grant/waiver, because if it goes somewhere it should be communication. Andrea, Batool, please see if it fits there.
+For this stage of the process, a certain degree of trust goes a long way-- when someone says they need the support, it is best to give them the waiver rather than second guess their eligibility.
+If the applicants for fee waivers are numerous and your budget cannot allow for that many fee waivers, you could ask for a short application.
 

### 2021-10-11 plos_latex_template.tex
AST: i was checking the output of the PDF and deleted extra tex files that overleaf saves separately and @rociojoo had added (sorry!)  nothing changes technically. the output seems to be off when you look at it on overleaf but if you open the link to the actual PDF it's ok. you can add them againif you want, Rocío but overleaf does not do anything to them.

@@
 In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
-Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for global conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
+Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
 % Rocío: we can add our Nature Correspondence letter to the citations later on. 
@@
 % Rocío: we can add our Nature Correspondence letter to the citations here too. I would've liked one that talks about increase in participation of tutors and presenters but haven't found any. 
-The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year, without depending on a conference central location.
+The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.
 % Rocío: main idea of the paragraph: online conferences are more inclusive
@@
 galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}.
-% Rocío: any reference to back the last statement?
-The CoC guide?
-ast: yes
 Organizers should invest time in creating opportunities to meet and bond virtually.
-An online event is a great opportunity to create spaces for people who do not enjoy in person networking. 
-Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
-Respect people's limits, preferences, and remember that `the usual' does not necessarily work with everyone in your community, and that one single networking activity will never serve the whole community.
+An online event is a great opportunity to create spaces for people who do not enjoy in-person networking. 
+Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
+Respect people's limits, preferences, and remember that 'the usual' does not necessarily work for everyone in your community, and that one single networking activity will never serve the whole community.
 It is worth trying new and varied activities that might work with subgroups of attendees.
@@
 
-Alternatively, a conference could have a hybrid format with an in-person and online component.
+Alternatively, a conference could have a hybrid format with an in-person and an online component.
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
 The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
-Needless to say, an inclusive online component of a hybrid conference is not a free pass to have an exclusionary in-person component.
-% Rocío: I know this is a bit aggressive so I wouldn't mind if anybody helps with the wording.
-% Liz: I like it, but you could say
-% an inclusive online component doesn't eliminate the need to organize an inclusive in-person component.
-
+Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
+%ast: I added Liz's suggestion and rephrased slightly to convey that online is not automatically inclusive
 For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
@@
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
+Even when the organizing team implements other inclusive practices, [ö improve this] the concern about the participation of people with disabilities tends to be partial, and is one of the lesser known attended features of inclusiveness.
+
 Importantly, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people.
 For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
-% Liz: I don't think it's actually ironic. maybe surprising or counterintuitive? ast: absolutely! i put importantly at the end to highlight that without opposition
-If the conference is in person, the venue must be accessible for people with mobility limitations. 
-% Liz: let's stay away from deficit-model language like "limitations." Possible aternatives:
-% people who use wheelchairs or other assistive mobility devices
-% ... must be physically accessible for disabled people
-Accommodations such as a quiet space for neurodivergent people should also be provided.
-% Liz: I hate "accommodateions because it sounds like something optional or for comfort or luxurious Consider something like "facilities"?
 
-Presenters should always speak into a microphone to make it easier for people who are hard of hearing and for captioners or interpreters to listen to them. 
-% Liz: avoid using disabilities as nouns like -- "the blind", "the deaf" etc. I woul duse "people who are deaf or hard-of-hearing" above. There's controversy about whether the word person
-% should come first ("person who is blind") or the identity should come first ("blind person:) but the person should always be acknowledged (like "participants with disabilities", blind scientist", etc..
-It's kind of objectifying without a person-related word.
-Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound.
-% Liz: not just alt-text, the whole site should be screen-reader accessible so:
-% communication about the conference should be fully screen-reader accessible including alt-text, and any videos should have both captions and sound.
+If the conference is in person, the venue must be accessible for people who use wheelchairs, have important signs in Braille, and a sound system compatible with hearing devices, just to name a few important features.
+Presenters should always speak into a microphone.
+Take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, menus that include sugar-free, gluten-free options.
+%ast: I removed the captioners part because it goes slightly off-topic and the sound system is apart for them when in person. 
+Regardless of the conference format --online, in person, or hybrid-- all images used in the communication strategy of the conference should have alternative text, including the website.
+Likewise, any videos should have both captions and sound.
 
-Platforms for conference registration and abstract submission, websites, and chat platforms--if used--should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
-Captioning for presentations should be available in more than one language if possible.
-Do not forget about making your social events accessible (e.g. art shows and movies) and making sure that the materials presented are accessible for all. 
+Platforms for conference registration and abstract submission, websites, and chat platforms --if used-- should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
+Captioning or interpretation for presentations should be available in more than one language if possible.
+Sign languages are more complicated, because the vary from region to region, but if your conference has a clearly regional scope, a specific sign language (such as ASL, American Sign Language, or LIBRAS, the Brazilian Sign Language) can be adopted as official.
+
+Do not forget about making your social events accessible and making sure that the materials presented are accessible for all.
+This includes physical activities that should not restrict participation based on body type or ability.
+
 These aspects should be tested well in advance of going live.
 
+%teach about accessibility
 Some practices, such as making accessible slides and presentations, are not yet common practice and will require great efforts from the presenters if they are not used to them. 
 For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have.
-Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
+Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, in screen-reader-frendly formats (e.g., R Markdown, HTML, or \TeX{} files).
+Accessibility in slides includes alternative text, and written code instead of print screen.
 
-% Liz: It is essential to provide an accessible version of the slides for people who won't be seeing them One of the most important assumptions about presentations are that they will be both seen and heard.
-Usually, neither of these channels is a complete source of information about the presentation.
-For example, when discussing computer programming, code is usually displayed on slides while the presenter talks about them.
-They seldom read the details of what appears on the slides in enough detail for a blind person to learn, and a deaf person will also miss some explanation.
-Presenters  should provide as much information as possible on both channels, but in order to compensate for the fact that this is not common practice, blind people need to be able to read an accessible version of the slides at the same time they listen to the presentation.
-It is critical to make these materials available before the conference starts..
+ If presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
 
 Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
-They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}. 
-Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs in an in-person conference.
+They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}, (see Rule \ref{rule_financial} about allocating resources).
+Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
 
@@
 choosing between captions and interpreters).
-Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices.
-% Liz: maybe strengthen the last phrase to use their expertise to solve accessibility problems and collaborate on decisions about accessibility practices.
-
-% Liz: As design justice author Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".
-%Rocío: maybe slightly rephrase the last sentence since it seems to imply that everybody knows Sasha Costanza-Chock.
-% not sure if that's a good characterization-- I am just learning about her right now.
+Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot accessibility problems and solve them, and collaborate on decisions about accessibility practices (see \textbf{Rule \ref{rule_organizing_team}}).
+Simulating disability is not a substitute for actual work with the people directly affected by the decision outcomes\cite{costanzachockDesign2020}
 
@@
 
-Meeting the need 
 % Liz: People with disabilities don't want to be considered "needy".
@@
 Or maybe it's wrong for me to apply disability rights philosophy to language.
-% The linguistic diversity of conference participants is often overlooked,
-% resulting in missed opportunities for them, as well as for the conference
-% to benefit from their contributions.
-of linguistically diverse audience is an important aspect in inclusion that is often overlooked.
+The linguistic diversity of conference participants is often overlooked, resulting in missed opportunities for them, as well as for the conference to benefit from their contributions.
 Non-native English speakers may feel intimidated to ask questions or raise their views.
@@
 
-% Rocío: Pasting Batool's text in her 10 simple rules because there are very valuable things we could use: "Rule: Meeting the need of linguistically diverse audience.
-Most international conferences are held in English.
-However, English is only spoken by 4.83 percent of the world population.
-If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
-Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.
-If possible, ensure that some sessions are delivered by languages other than English"
+%ast: I think ome ideas of the first and second paragraphs can be joined because some parts are in both (intimidation, the importance of not missing opportunities) and maybe some practical tips can be made?
+Do we add an invitation to "bear" with subtitles and other languages to the English monolingual audience?
+an invitation to be respectful of "bad English", from submissions to presentations?
+%in person: live interpretation is doable. online: translated captions, pre-recorded talks, english trancripts
 
@@
 As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded.
-If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
-This could through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct and accessibility guidelines, and providing fee waivers and financial support.
+If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
+This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines [*here] and advertising support for attendees, such as fee waivers and financial support.
 People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
 If possible, include representatives from these communities within the organizing team (\textbf{Rule \ref{rule_inclusion}}).
+%ast: i think this sentence can go out, plus these communities
 Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
@@
 Always include explicit descriptions of accessibility and other inclusion practices (\textbf{Rule \ref{rule_accessibility}}).
-Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusions.
-Be careful with the language you use in promoting to the conference and how you talk about minoritized people, assertively set tone of inclusion.
+Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.%ast this explicit description seems to belong to the previous paragraph, maybe where I marked [*here]?
+Be careful with the language you use in promoting the conference and how you talk about minoritized people, assertively set tone of inclusion.
 In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
@@
 However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
+If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
 Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
+%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)
 
@@
 % Batool: I will work on this rule to include more details by 25th of Sept 
-
-%sara: this fits better in the communication rule
-%adding here from Liz's rule but maybe this could go in communication in general
-%In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
-Do not expect people to teach you --it's not their role-- and accept feedback without being offended.
+% ast: Thanks Batool! i added a couple of comments.
 
@@
 
-%Inclusion comes with a cost
-
 Conference budgets are limited and rely mostly on sponsorship and attendance fees.
@@
 It is important to estimate the costs for these inclusive practices and define your priorities in advance.
-This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership training, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}).
+This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}).
 It also includes additional support for attendees: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
@@
 Don't be afraid to innovate and resist the institutional inertia.
-"This has always been like that" will not take the conference towards structural change.
-%should we talk about the pandemic? something like: 
+"We have never payed for this, and this has always been like that" will not take the conference towards structural change.
+%ast should we talk about the pandemic? something like: 
 With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
@@
 
-%Set fair registration rates
 Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates.
@@
 In addition to this, low rates in strong currencies can represent significant amounts of money in other currencies.
-You can adjust the registration fee by the cost of living in each country using conversion factors (e.g.
-from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example).
+You can adjust the registration fee by the cost of living in each country using conversion factors (e.g., from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example).
 In general, people should be allowed to locate themselves in a category they consider affordable, even with the possibility of a "pay what you can" approach.
@@
 
-Offering scholarships or grants to attend the conference is a common way to boost participation from different groups, and some conferences ask for cover letters or applications to them.
-These applications may be counterproductive, if they attract people that do not need the money but apply to enrich their CVs, and can filter out people who do not feel entitled to earn them.
-Applying for loans, grants, and scholarships may be an emotionally demanding task.
-Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support.
-For this stage of the process, a certain degree of trust goes a long way-- when someone says they need the support, it is best to give them the waiver rather than second guess their eligibility.
-If the applicants for fee waivers are numerous and your budget cannot allow for that many fee waivers, you could ask for a short application.
+Offering scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and some conferences ask for cover letters or applications to them.
+These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses.
+For online conferences, granting fee waivers is easier, and diversity scholarships can be replaced with broader participation support.
+However, these applications may be counterproductive if they attract people that do not need the money but apply to enrich their CVs, and filter out people who do not feel entitled to earn them.
+Applying for loans, grants, and scholarships is an emotionally demanding task.
+Simplify applications as much as you can and be specific about the target audience of these programs.
+The broader the support, the simpler the application.
 
@@
 The ten rules stated here can be adapted depending on the conference format and settings.
-When organizing useR!2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
+When organizing useR! 2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
 We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
@@
 \section*{Acknowledgments}
-The authors of this piece would like to thank every single member of the organizing team of useR!2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR!2021 and supporting us through the process.
+The authors of this piece would like to thank every single member of the organizing team of useR!
+2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR!
+2021 and supporting us through the process.
 

### 2021-10-13 plos_latex_template.tex
Rocío: edits to last paragraph of the intro

@@
 For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-In this article, we share a few lessons learned in the process.
-We streamline our most important learnings in 10 simple rules to host an inclusive conference.
+In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference.
 These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
@@
 Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
-However, they are likely to reproduce the systematic discrimination occurring in other spaces in our fields and become [bad experiences]. 
+However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
+An exclusionary conference experience can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018}.
 Efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact on our communities and the lives and careers of their members.
@@
 
-The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, that coordinates with the local organizers of conferences, or even the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
+The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
+The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
 These tips stem from the authors' experience of organizing useR!
 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
-We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive, and accessible experience for everyone. 
+We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone. 
 Here, we share the lessons learned within the past year of organizing this global useR! 2021, summarized as ten simple rules towards an inclusive conference.
-We have organized these rules in three groups, because they refer to the foundations (Rule 1 and 2), the actions regarding people (Rules 3 to 5), the actions regarding conference design (Rules 6 to 9), and a general synthesis %ast this is an attempt, but i am not a fan of rule 10 alone at the end.
-actually the whole classification when written seems off to me (right now, this might change)
+We have organized these rules in three groups.
+Rules 1, 2 and 10 refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
+The next two rules are focused on the people that participate in the conference: 
+Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
+Rules 5 to 9 are rules about actions regarding components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
+These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
+% Rocío: the last sentence was copied from the abstract because we need a closing sentence.
 % Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure [x]
@@
 another thing: i avoided "gender balancing" because idk if that comes across as too binary, even if what is done is adding cis women to the team
+% Rocío: what about ethnicity instead of race? I agree with the gender balancing thing.
 
@@
 % expertise about removing barriers to inclusion. Actively ask where the barriers are and how to mitigate them.
+% Rocío: let's make sure this is included somewhere.
 
@@
 Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
-Build an environment in which every person can express their position and give prevalence to the positions from systematically excluded groups.
+Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
 % Liz: Use of "voice" in this way makes deaf people feel excluded.
@@
 Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings \cite{salibaGettingGripsOnline2020}.
-However, social events in in-person conferences can be exclusive (e.g.
+However, social events in in-person conferences can be exclusionary (e.g.
 galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}.
@@
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
-Even when the organizing team implements other inclusive practices, [ö improve this] the concern about the participation of people with disabilities tends to be partial, and is one of the lesser known attended features of inclusiveness.
+Even when the organizing team implements other inclusive practices, the concern about the participation of people with disabilities tends to be partial, and is one of the .
 
@@
 Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
-
-On the other hand, it is likely that financial constraints may not allow you to make your conference completely accessible and there will be choices to make (e.g.
-choosing between captions and interpreters).
-Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot accessibility problems and solve them, and collaborate on decisions about accessibility practices (see \textbf{Rule \ref{rule_organizing_team}}).
-Simulating disability is not a substitute for actual work with the people directly affected by the decision outcomes\cite{costanzachockDesign2020}
 
@@
 People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
-If possible, include representatives from these communities within the organizing team (\textbf{Rule \ref{rule_inclusion}}).
-%ast: i think this sentence can go out, plus these communities
 Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.

### 2021-10-13 plos_latex_template.tex
Rocío: shortening rule 1

@@
 
+% Rocío: Main message of the paragraph: Why we are writing this piece
 Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
@@
 
+% Rocío: Main message of the paragraph: We do this for communities and people. What do we mean by communities and why are they important? (Why are they important may need some text)
 This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
@@
 
+% Rocío: Main message of the paragraph: Roadmap of the paper
 The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
@@
 
+% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
 Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
@@
 
-To be proactive allies in our conferences and communities, we should start by examining our own privileges (unearned advantages given by society to some people but not all).
-Privilege is largely invisible to those who have it.
-And we will not be able to help end discrimination and oppression if we do not recognize our privileges first and the injustices that come with them.
-Once we understand our own privileges, we can begin addressing them on an individual, collective, and institutional basis.
-Recognizing our privileges--particularly in our field and in our scientific or professional community--will help identify which subgroups have been the most excluded or discriminated against. 
+% Rocío: more effort towards including the more excluded
+Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against.
 These are the groups we need to make more effort to include. 
-Depending on the field, region, or community, it could be Black People, LGBTQIA+, Muslim, or others. 
-%ast: ö check phrasing, add "people"
-Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.
-
-Investing more effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that). 
-But it will prevent from falling into the mistake of treating diversity as a checklist.
-It will also guide the vision of diversity for your conference--and your strategies to achieve it:
+Investing more effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that), but it will guide the vision of diversity for your conference--and your strategies to achieve it:
 Will a more diverse conference translate into an even gender distribution in your speakers? 
@@
 Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
-This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way.
+Depending on the field, region, or community, it could be Black people, LGBTQIA+, Muslim, or other people.
+Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account. 
+% This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way.
 

### 2021-10-13 plos_latex_template.tex
Rocío: Shortening rule 2

@@
 However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
+Lack of representation and unwelcoming--or overtly aggressive--environments hinder participation (or future participation) of people who could otherwise become active community members. 
 An exclusionary conference experience can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018}.
@@
 These are the groups we need to make more effort to include. 
-Investing more effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that), but it will guide the vision of diversity for your conference--and your strategies to achieve it:
+Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it:
 Will a more diverse conference translate into an even gender distribution in your speakers? 
@@
 
+% Rocío: main idea: Inclusion for everyone
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and acknowledging that gender is not binary are just some examples of decisions that can make inclusion real.
+Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, and acknowledging that gender is not binary are just some examples of decisions that can make inclusion real.
 Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.
 
-Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences.
-Lack of representation, unwelcoming--or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members.
-A negative conference experience can divert career paths, affect lives, and exclude people from some fields \cite{biggsAcademicConferenceChilly2018}.
-%ast maybe this last sentence could go in the introduction because it's general i marked [bad experience] in the intro
-
+% Rocío: main idea: Code of conduct for a safe place
 Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
-The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
-The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
-An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.
-
+The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
 The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
 A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
-There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC).
+There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
 We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
+% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.
+
+% Rocío: Could have a paragraph about active bystandership—Andrea
 

### 2021-10-13 plos_latex_template.tex
Rocío: edits to communication rule

@@
 
+% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
 In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
-Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}.
-% Rocío: we can add our Nature Correspondence letter to the citations later on. 
+Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}.
 Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
-This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020}.
-% Rocío: we can add our Nature Correspondence letter to the citations here too. I would've liked one that talks about increase in participation of tutors and presenters but haven't found any. 
+This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
 The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.
-% Rocío: main idea of the paragraph: online conferences are more inclusive
-% Liz: but only if you make them that way. The R conference was a standout and innovative compared to others I've considred attending since Covid.
 
-Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings \cite{salibaGettingGripsOnline2020}.
-However, social events in in-person conferences can be exclusionary (e.g.
-galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}.
-Organizers should invest time in creating opportunities to meet and bond virtually.
-An online event is a great opportunity to create spaces for people who do not enjoy in-person networking. 
-Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
-Respect people's limits, preferences, and remember that 'the usual' does not necessarily work for everyone in your community, and that one single networking activity will never serve the whole community.
-It is worth trying new and varied activities that might work with subgroups of attendees.
-Moreover, virtual communication may make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012}).
-% Rocío: Networking paragraph.
-
-In addition to being more inclusive, an online format is more environmentally-friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.
-Besides, if conference materials are systematically stored and pre-recorded, they can be made available to people who missed them live or even after the conference is over. 
-% Rocío: Short paragraph of additional advantages of online conferences that are not about inclusion.
-
+% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
+Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
 Alternatively, a conference could have a hybrid format with an in-person and an online component.
@@
 The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
-Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
-%ast: I added Liz's suggestion and rephrased slightly to convey that online is not automatically inclusive
-For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
-% Rocío: Hybrid conferences paragraph.
+%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
+%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
+% Rocío: Now I'm not sure if we need the last two sentences that I commented.
+
+% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
+% Rocío: I'm still doubting if this should be the 2nd paragraph, and I think that it needs some editing.
+Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
+However, in-person settings might not be comfortable or appealing to everyone for socializing; 
+some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues), and common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. 
+On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
+\cite{trianaDoesOrderFacetoFace2012}).
+Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that one single networking activity will never serve the whole community.
+It is worth trying new and varied activities that might work with subgroups of attendees.
+Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
 
@@
 \label{rule_communication}
-%ast: sara mentioned something that can be good spirit for the communication rule: don't do anything to the outside that you are not doing for your team.
-it's communication strategy but also there should be coherence between your actions and your words.
-%Rocío: I love that.
-Please add it.
-An inclusive communication is to give each individual within the conference the opportunity to express themselves and a sense of belonging.
-Embedding inclusion in your communication strategy can only come by design, it should not be a thoughtless default.
-In the essence of your communication strategy, there should be coherence between actions and words.
-Always attend to the needs of your team, in other words, don't do anything to the outside that you are not doing for your team.
 
+%Rocío: Main idea: Include people in your communication 
+Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
 As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
 If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
-This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines [*here] and advertising support for attendees, such as fee waivers and financial support.
-People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory.
-Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.
+This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines and advertising financial support, or other creative ways that show that everyone is seen, respected, and welcome (Fig.
+MARMOTS).
+Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, email lists)--
+part of your community may use a different platform than the one who regularly attend the conference. 
+With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too (e.g.
+having accessibility guidelines on your website sends a welcoming message to people with disabilities), without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.
 
-Always include explicit descriptions of accessibility and other inclusion practices (\textbf{Rule \ref{rule_accessibility}}).
-Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.%ast this explicit description seems to belong to the previous paragraph, maybe where I marked [*here]?
-Be careful with the language you use in promoting the conference and how you talk about minoritized people, assertively set tone of inclusion.
-In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation.
-Do not expect people to teach you --it's not their role-- and accept feedback without being offended.
+%Rocío: Main idea: Use inclusive language
+Inclusive language--language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups--should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
+Make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. 
+Do not expect minoritized people to teach you--it's not their role--and accept feedback without being offended.
 
+% Rocío: not sure if this belongs to the rule—Andrea make a convincing case :-P
 As we mentioned previously, conference attendees should have ample opportunities to network with each other.

### 2021-10-13 plos_latex_template.tex
Rocío: shortening financial rule

@@
 
+% Rocío: main idea: Budgets are limited and need to choose priorities.
 Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
-Some expenses are more or less fixed but allocation of resources has to be intentional to support the goals of the conference.
-It is important to estimate the costs for these inclusive practices and define your priorities in advance.
-This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}).
-It also includes additional support for attendees: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
-While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team.
+While some expenses might be more or less fixed, % Rocío: can we give examples?
+allocation of resources has to be intentional to support the goals of the conference towards inclusion. 
+It is important to estimate the costs for these inclusive practices and define your priorities in advance, like paying the organizing team, necessary training (e.g.
+code of conduct), and accessibility practices (e.g.
+captions).
+Aditional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
+Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
+Instead of depending on registration fees, ask sponsors to support specific items in your inclusion expenses.
+
+% Rocío: main idea: help take the burden out of participants for more inclusive participation
+Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
+(see https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example of conversion rates based on country of origin and career status). 
+In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
+Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}.
+Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
+Scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and conferences usually ask for cover letters or applications from them. 
+These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses. 
+For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper. 
+Lower costs for attendance and offering financial support instead of prestigious grants--that could look good in a CV--may reduce the number of people applying for them and allow you to help all applicants.
+Applying for loans, grants, and scholarships might be an emotionally demanding task, so if this scenario is likely, simplify the process of asking for financial support.
+
+% Rocío: Let's keep this in mind for the 10th rule or conclusion
+% While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team.
 Don't be afraid to innovate and resist the institutional inertia.
 "We have never payed for this, and this has always been like that" will not take the conference towards structural change.
-%ast should we talk about the pandemic? something like: 
-With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
-Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.
-
-Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates.
-Usually, there is a higher fee for people from the industry than for academia.
-A lower fee for non-profit organizations, government employees, or freelancers should also be considered.
-It is important to include discounts for students \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012} and other early career stages, such as postdoc or trainee.
-The definition of such positions and the associated work conditions vary for each country.
-Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere.
-In addition to this, low rates in strong currencies can represent significant amounts of money in other currencies.
-You can adjust the registration fee by the cost of living in each country using conversion factors (e.g., from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example).
-In general, people should be allowed to locate themselves in a category they consider affordable, even with the possibility of a "pay what you can" approach.
-Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}.
-Offering fee waivers to part of the participants is also a good option, but even then, other costs can be prohibitive.
-
-Offering scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and some conferences ask for cover letters or applications to them.
-These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses.
-For online conferences, granting fee waivers is easier, and diversity scholarships can be replaced with broader participation support.
-However, these applications may be counterproductive if they attract people that do not need the money but apply to enrich their CVs, and filter out people who do not feel entitled to earn them.
-Applying for loans, grants, and scholarships is an emotionally demanding task.
-Simplify applications as much as you can and be specific about the target audience of these programs.
-The broader the support, the simpler the application.
+% %ast should we talk about the pandemic? something like: 
+% With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
+% Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.
 

### 2021-10-13 plos_latex_template.tex
Rocío: Adding Sara's version of rule 10

@@
 
-This rule is a complement to all the other rules.
+% This rule is a complement to all the other rules.
 It is important to have a bold vision and clear targets.
@@
 For an inclusive conference, we need to avoid giving in because considering all aspects seems impossible and burning ourselves and the team out because considering all aspects is actually impossible.
+
+For many people working with diversity, one typical aim is institutionalizing diversity (Ahmed 2012).
+However, our aim is also a symptom that diversity and equity are not part of institutions, academic societies, and conferences.
+This rule is about the process of making diversity and inclusion possible and a complement to all the other rules.
+We believe we need to be accountable and we need to aim for permanence.
+Institutionality comes from a set of norms, values, and priorities and these determine who is allowed and how (Ahmed 2012).
+Our long-term goal is to set new norms, values, and priorities based on diversity.
+Organizing a diverse conference is part of the process of change.
+However, we cannot do everything at one conference or a few of them.
+It is important to have a bold vision and clear targets.
+Keep in mind that ideal actions are often not possible and resources are limited (both time of the organizing team and money.
+
+If from the beginning we set diversity goals and actions for the conference, we can use that to be transparent and, later, accountable.
+Assess whether equity and inclusion goals were met during the meeting.
+This can be done internally within the organizing team and often also through the conference’s participants.
+Plan, balance and keep working to improve inclusion in the next edition of the conference.
+Try to propose permanent data collection to assess whether diversity and inclusion are actually being seen at the conference.
+We know that diversity and equity are not institutionalized, in fact, the system works to stop those who are changing the system (Ahmed 2019).
+We will find resistance, but diversity is a school.
+We also learn more about how people are being excluded, we learn that there are actions we can take to include more, and then, we can set higher goals on how to make inclusion possible.
 

### 2021-10-13 plos_latex_template.tex
Rocío: Edits to rule 10

--- 2021-10-13 plos_latex_template.tex
@@
 
-%Sara: I think the last rule should be about the process and reinforce to actually discuss and include diversity as a principle and not as a checklist (as it is on the text).
-Maybe also something about accountability and permanence.
-I think that this rule go be together with the rule_organizing_team given that it is relative to the team work and their dynamics.
-%Sara: copying Yani comment here -  I like the "accountable and permanent" name.
-I think that an idea about progressive improvement can be mention here, like improve something every edition and no go back on the conquers goals.
-Now, to be accountable, you need to be transparent, and you need a clear structure; you get that with the structure of your team, with your declaration and guidelines, and with your action to reinforce those declarations and guidelines (if you have only declarations then you have only empty words, that sound good but you no need to comply.
-We discuss a lot about this).
-For me, one more time, putting together a diverse team is what makes the other thing happened.
-We can cite The tyranny of structurelessness.
-The Second Wave, 2(1), 1972.
-on this rule.
-Perhaps we can talk about governance.
-%ast: i'll put the last version of the tenth rule just to have an idea of how we left this
-%past names: Have a bold vision but be ready to compromise when necessary/Aim for progress not perfection
-%more people rules that have a continuity/accountability theme
-%Aim for Progress, not Perfection
-%Plan, balance and keep working to improve inclusion in the next edition of the conference!
-%Assess whether equity and inclusion goals were met during the meeting
-%Be accountable and permanent
-%Use feedback from previous conferences or other conferences in your field to identify barriers to inclusion and plan around them
+% Rocío: main idea: we're not changing everything at the beginning, the conference will not be perfectly inclusive
+Do not expect the conference to be perfect. 
+Resources, both in time and money, are always limited. 
+Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
+There will be resistance, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
 
-This rule is a complement to all the other rules.
+% Rocío: main idea: be part of the process
+Diversity and inclusion will be the norm in the long-term if we see conferences as part of the structural change. 
+After the conference is over, assess whether equity and inclusion goals were met during the meeting. 
+This can be done internally within the organizing team and taking into account the attendees' point of view. 
+Identify the things that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
+Make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next edition of the conference.
+
+% For many people working with diversity, one typical aim is institutionalizing diversity (Ahmed 2012).
+However, our aim is also a symptom that diversity and equity are not part of institutions, academic societies, and conferences.
+This rule is about the process of making diversity and inclusion possible and a complement to all the other rules.
+We believe we need to be accountable and we need to aim for permanence.
+Institutionality comes from a set of norms, values, and priorities and these determine who is allowed and how (Ahmed 2012).
+Our long-term goal is to set new norms, values, and priorities based on diversity.
+Organizing a diverse conference is part of the process of change.
+However, we cannot do everything at one conference or a few of them.
 It is important to have a bold vision and clear targets.
-Some parts might be defined as baseline assumptions that are not up for discussion.
-However, there will always be compromises.
-Ideal solutions that meet all our expectations are often not possible.
-Don't let the perfect be the enemy of good.
-As a team, this is a challenging journey.
-Very early on, we tried to recognize that our resources were limited (both in time of the organizing team but also in money).
-We had to let go of certain aspects that were important to many of us.
-For an inclusive conference, we need to avoid giving in because considering all aspects seems impossible and burning ourselves and the team out because considering all aspects is actually impossible.
+Keep in mind that ideal actions are often not possible and resources are limited (both time of the organizing team and money.
+
+% If from the beginning we set diversity goals and actions for the conference, we can use that to be transparent and, later, accountable.
+Assess whether equity and inclusion goals were met during the meeting.
+This can be done internally within the organizing team and often also through the conference’s participants.
+Plan, balance and keep working to improve inclusion in the next edition of the conference.
+Try to propose permanent data collection to assess whether diversity and inclusion are actually being seen at the conference.
+We know that diversity and equity are not institutionalized, in fact, the system works to stop those who are changing the system (Ahmed 2019).
+We will find resistance, but diversity is a school.
+We also learn more about how people are being excluded, we learn that there are actions we can take to include more, and then, we can set higher goals on how to make inclusion possible.
 

### 2021-10-13 plos_latex_template.tex
adding an suggestion of alt.tex for the figure
adding the figure's png, caption and alt text (unsuccessfully)

@@
 \usepackage{soul}
+
+%sara: for commenting the pdf to add alt text (if anyone knows a better solution for atl text please help)
+\usepackage{pdfcomment}
 
@@
 Here, we share the lessons learned within the past year of organizing this global useR! 2021, summarized as ten simple rules towards an inclusive conference.
-We have organized these rules in three groups.
+We have organized these rules in three groups (Figure \ref{fig:diagram}).
+%sara: i think here is a good place to add the diagram: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
+%sara: still need to better format figure size and get the alt text right :P
 Rules 1, 2 and 10 refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
@@
 % Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure [x]
+
+\begin{figure}[!h]
+\centering
+%sara: without success i'm trying to use pdfcomment package and pdftooltip to make the alt-text. any help?
+\pdftooltip{
+\includegraphics[width=\textwidth]{figs/10_rules_diversity.png}}{
+%sara: suggestion for alt text:
+Diagram of how the 10 rules are organized into three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9).
+The diagram has five rows.
+Each line contains rectangular boxes with the rule number and a short title for each rule.
+Each rule box is colored based on their group (grounding rules are grey, people-related rules are green and design rules are pink).
+First row is for the grounding rules 1.
+Multidimensional diversity and 2.
+Safe and inclusive environment in grey boxes.
+Second row is for people-related rules 3.
+Organizing team and 4.
+Unbias in green boxes.
+Third row is for the design rules 5.
+Online component, 6.
+Accessibility for disabilities, 7.
+Language inclusivity in pink boxes.
+The fourth row is for the design rules 8.
+Communication and 9.
+Financial support and budget in pink boxes.
+The last row is for the grounding rule 10.
+Diversity and inclusion as a process in grey boxes.
+}
+\caption{Diagram of the rules organized in three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9).}
+\label{fig:diagram}
+\end{figure}
 

### 2021-10-15 plos_latex_template.tex
Rocío: editing rule language

@@
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
-Build a team with people from different regions, genders, races, socioeconomic statuses, and other aspects of diversity.
+Build a team with people from different regions, genders, races and ethniticities, socioeconomic statuses, and other aspects of diversity.
 Go beyond balancing all genders in this effort, and pay attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}).
@@
 
-%ast: I added race -even if race is a construct- because focus is given to gender and the spaces tend to be predominantly white and US/Europe-centric.
-In other parts of the text I was using racialized but here we can keep it simple, i think.
-another thing: i avoided "gender balancing" because idk if that comes across as too binary, even if what is done is adding cis women to the team
-% Rocío: what about ethnicity instead of race? I agree with the gender balancing thing.
+% Rocío: what about ethnicity instead of race? %i am not sure
 
@@
 
-In international conferences, English is often the official language.
-Submissions, presentations, tutorials, and workshops are in English.
-The platforms, the webpage, and official communications are also in English.
-While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}.
-Excluding them may potentially lead to missing innovative contributions due to a language barrier.
+% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
+In international conferences, the linguistic diversity of conference participants is often overlooked. 
+English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
+While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker, a privilege.
+Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
+and conferences may miss potentially innovative contributions.
 
-% Liz: People with disabilities don't want to be considered "needy".
-We reject the term "special needs" in part because it frames the person with the disability as being the burden causing the accommodation, where maybe it's the responsibility of society to be open to everyone.
-So I wonder if it's the same for people who speak other languages.
-Maybe they are not needy or inferior.
-Or maybe it's wrong for me to apply disability rights philosophy to language.
-The linguistic diversity of conference participants is often overlooked, resulting in missed opportunities for them, as well as for the conference to benefit from their contributions.
-Non-native English speakers may feel intimidated to ask questions or raise their views.
-Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2).
-If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language.
-Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.
+% Rocío: main idea: make conferences more linguistically inclusive
+Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (Rule 2). 
+Advertising the conference in several languages and considering having non-English workshops and presentations to help overcome this barrier. 
+Allow the audience to ask questions in their native language and invite others to help translate it; conversely, if the presenter is not comfortable with questions in English, encourage people in the audience to help by rephrasing their questions or try to translate them; after all, the community should be supportive.
 
-Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier.
-For instance, hosting one international day/session per conference might be a great place to start!
+% Not citing ninerBetterWhomLeveling2021 anymore for now.
 
@@
 
-% Rocío: not sure if this belongs to the rule—Andrea make a convincing case :-P
-As we mentioned previously, conference attendees should have ample opportunities to network with each other.
+% % Rocío: not sure if this belongs to the rule—Andrea make a convincing case :-P
+% As we mentioned previously, conference attendees should have ample opportunities to network with each other.
 However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.

### 2021-10-15 plos_latex_template.tex
Rocío: edits to rule 4

@@
 
+% Rocío: Idea of this paragraph: build a diverse team
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
-Build a team with people from different regions, genders, races and ethniticities, socioeconomic statuses, and other aspects of diversity.
-Go beyond balancing all genders in this effort, and pay attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}).
+Build a team with people from different regions, genders, races and ethnicity, socioeconomic statuses, and other aspects of diversity. 
+Go beyond balancing all genders in this effort and pay special attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). 
 To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. 
@@
 
-% Rocío: what about ethnicity instead of race? %i am not sure
-
-% Rocío: Idea of this paragraph: build a diverse team
-
+% Rocío: Idea of this paragraph: real inclusion when having diverse members
 Gathering a diverse team will only work if there is real inclusion. 
 People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. 
-This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups [cite design justice].
-A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.
+This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups \cite{costanzachockDesign2020}.
+People who have experience with exclusion have the social and technical expertise about removing barriers to inclusion.
+A truly inclusive and welcoming space is one in which everyone in the team is invited to be part of the solution.
+However, do not expect people from minoritized groups to only work on these solutions, and respect their interests in other areas of the organization.
 
-% Rocío: Idea of this paragraph: real inclusion when having diverse members
-
-Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes are worthwhile. 
+% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
+Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
 Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
 In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging (see \textbf{Rule \ref{rule_inclusion}}).
-% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least for now
+
+% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
+The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. 
+Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly, offer leading positions, and let people from privileged groups step down. 
+Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
+Offer support and encouragement if you encounter cases of impostor syndrome. 
+You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
+
+% Rocío: Idea of this paragraph: take care of your team.
+Most importantly, take care of your team. 
+Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort. 
+The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. 
+However, having a strong community as the only reward may be enough for those who are in more privileged positions. 
+Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.
+In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
+Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
+Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}).
+
+% % Rocío: might go to the next rule
+% Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).
+Make work visible, discuss collectively what should be done (and why), divide and account work.
+Being transparent and sharing the information is key to make work visible and collective.
+%sara: this relates to make labor visible from DF, maybe we can cite them here too.
+%Rocío: sure.
 
@@
 
-% Liz: I know this section is getting long, and maybe this is too blunt, but: 
-% People who have experience with exclusion have the social and technical 
-% expertise about removing barriers to inclusion. Actively ask where the barriers are and how to mitigate them.
-% Rocío: let's make sure this is included somewhere.
-
-The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups.
-First, rethink power relationships inside the group and share power.
-Be committed to examine power relations inside your organizing team and challenge it during all the process.
-% Rocío: Could we explain power relationship / share power ideas in more simple terms?
-Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
-Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
-% Liz: Use of "voice" in this way makes deaf people feel excluded.
-Maybe it could be something like "Build an environment in which every person has input..." Or anything as long as it doesn't suggest you have to have a voice to participate.
-%ast: i rephrased slightly, avoiding "support" it's not a question of privileged people giving support but of positions being given importance
-Offer support and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others.
-You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
-Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).
-Make work visible, discuss collectively what should be done (and why), divide and account work.
-Being transparent and sharing the information is key to make work visible and collective.
-%sara: this relates to make labor visible from DF, maybe we can cite them here too.
-%Rocío: sure.
-% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
-
-% Liz: I have a little concern about mentoring WRT English, but I'm a white English speaker, so take it with a grain of salt.
-I think people who look like me need to do more accepting and less mentoring around English.
-If someone explicitly wants my help, I'm happy to give it because I love words and language, and I might learn something cool about another language.
-But I'm not sure we should be telling the English-priveleged readers that they are language mentors.
-%ast. i agree with liz and wrote this from the point of view of a non-native speaker- in terms of support rather than mentorship. replaced with support but feel free to check
-Most importantly, take care of your team. 
-Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort.
-The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone.
-However, having a strong community as the only reward may be enough for those who are in more privileged position. 
-Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free. 
-% Liz: Yes!
-This is especially true when *some* people are paid to work on the conference.
-I was just asked to do accessibility work that should have been done by the contracted, paid organizer.
-This is hugely problematic and will be true for lots of conferences outside the open software community.
-In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
-Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
-Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}). 
-% Rocío: Idea of this paragraph: take care of your team.
-% Liz: also, include people from diverse backgrounds in all aspects of planning, not just the diversity-related ones, unless that is their only interest.
-In some organizations, I could be on the scientific committee and participate in many other ways, but I'm only called on for accessibility.
-
 \section{Consciously unbias your spotlight roles}
@@
 
-%idea1: our lists are biased
-
+%idea of paragraph 1: our lists are biased
 When choosing or inviting people as keynote speakers, program committee, session chairs, and other spotlight roles, it is likely that there will not be much diversity in the first set of names. 
@@
 We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
-
 Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them.
@@
 The regional and local communities in your field are also good sources to tap into.
-For example, for useR!2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
-When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their barriers for participation, and explicitly invite them to contribute to the conference.
-% Rocío: Not everyone knows what a community of practice is.
-Could we add a definition or use other words?
-Definition: "Communities of practice are groups of people who share a concern, a set of problems, or a passion about a topic, and who deepen their knowledge and expertise in this area by interacting on an ongoing basis".
-(Wenger, E., McDermott, R.
-A., & Snyder, W.
-(2002).
-Cultivating Communities of Practice: A Guide to Managing Knowledge.
-Harvard Business Press.)
-Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged.
-Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers and people from minoritized groups, fostering collaboration and building their skills.
-% Liz: maybe consider a broader group than "early career researchers." One problem is that marginalized people have many fewer opportunities, and as they age that attenuation of opportunities compounds.
-Someone can be in the margins even years after finishing grad school or being a postdoc (not to mention that people without degrees could have contributions.
 
-Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative ways \cite{cheng2020x+}.
-You can, for instance, reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
-Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve.
-Defy the stereotypical criteria for success by acknowledging these community practices.
-Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility.
-Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
-% Rocío: Since the system usually rewards the most competitive people, appealing cvs and so on, I think it's worth it to explain that rewarding collaborative work is also part of unbiasing; not only because it's mostly done by women, people of color, etc., but because this kind of amazing work is usually neglected in our conferences.
-It may not be obvious to everyone that the content of this paragraph is not only "bring these people to the spotlight", but that it also means unbiasing the conference.
-I don't think I'll have time to add lines about this in the beginning of the paragraph so if anybody could do that, it would be great.
-
-% Liz: I might be able to, but I'm not sure where you want the additional text; where the beginning of the paragraph is because of how this is formatted.
-I couldn't agree more that being truly inclusive means not restricting people's role.
-Every subteam should be diverse, in addition to the whole organizing team.
-It's not just about representation but needs to encompass integration across the whole event.
-
-% ast: i agree a lot with Rocío here, the change in what is given importance is a key part of the unbiasing.
-i had the impression that the part of not restricting people's role goes above, in team rule.
+% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
+In academia and the industry, not only does the list of people considered in spotlight roles is usually biased, but the spotlight roles themselves are.
+Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+}.
+Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
+You can, for instance, give more space in your conference to talks about community building, and reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
+Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to diversity and inclusion; recognize their areas of expertise and interest.
 
@@
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
-Even when the organizing team implements other inclusive practices, the concern about the participation of people with disabilities tends to be partial, and is one of the .
+Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
+This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it is not a substitute for real-life experience \cite{costanzachockDesign2020}.
+Accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020} and should not be dealt with at the last minute.
+Conversely, inaccessible decisions are hard to correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
 
@@
 Take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, menus that include sugar-free, gluten-free options.
-%ast: I removed the captioners part because it goes slightly off-topic and the sound system is apart for them when in person. 
+
 Regardless of the conference format --online, in person, or hybrid-- all images used in the communication strategy of the conference should have alternative text, including the website.
@@
 Captioning or interpretation for presentations should be available in more than one language if possible. 
-Sign languages are more complicated, because the vary from region to region, but if your conference has a clearly regional scope, a specific sign language (such as ASL, American Sign Language, or LIBRAS, the Brazilian Sign Language) can be adopted as official.
-
 Do not forget about making your social events accessible and making sure that the materials presented are accessible for all.

### 2021-10-15 plos_latex_template.tex
Rocío: edits to intro and conclusions

@@
 % Rocío: Main message of the paragraph: Why we are writing this piece
-Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
+Conferences, from the Latin \textit{conferre} or `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
 However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
-Lack of representation and unwelcoming--or overtly aggressive--environments hinder participation (or future participation) of people who could otherwise become active community members. 
-An exclusionary conference experience can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018}.
-Efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact on our communities and the lives and careers of their members.
-
-% Rocío: Main message of the paragraph: We do this for communities and people. What do we mean by communities and why are they important? (Why are they important may need some text)
-This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
-Some scholarly or technological communities are formally structured as learned societies (e.g.
-the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g.
-R User Groups or Python User Groups), and others may not have any established structure.
-For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.
+Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, producing knowledge, and becoming active community members that conferences can provide to the non-marginalized \cite{hendersonThoughtfulGatheringsGendering2020}.
+Exclusionary conference experiences can even divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
+This message is not new. 
+Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
+and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus an the online format to open doors for inclusion.
 
 % Rocío: Main message of the paragraph: Roadmap of the paper
+This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for an online component. 
 The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
@@
 We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone. 
-Here, we share the lessons learned within the past year of organizing this global useR! 2021, summarized as ten simple rules towards an inclusive conference.
-We have organized these rules in three groups (Figure \ref{fig:diagram}).
-%sara: i think here is a good place to add the diagram: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
+Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards an inclusive conference.
+The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
 %sara: still need to better format figure size and get the alt text right :P
@@
 Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
-Rules 5 to 9 are rules about actions regarding components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
+Rules 5 to 9 are rules about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
 These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
-% Rocío: the last sentence was copied from the abstract because we need a closing sentence.
-% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure [x]
 
@@
 \end{figure}
+
+% Rocío: I'm leaving this out now and seeing if we don't need this description of communities
+%% Rocío: Main message of the paragraph: We do this for communities and people. What do we mean by communities and why are they important? (Why are they important may need some text)
+%This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
+%Some scholarly or technological communities are formally structured as learned societies (e.g.
+the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g.
+R User Groups or Python User Groups), and others may not have any established structure.
+%For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.
 
@@
 In academia and the industry, not only does the list of people considered in spotlight roles is usually biased, but the spotlight roles themselves are.
-Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+}.
+Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
 Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
-You can, for instance, give more space in your conference to talks about community building, and reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
-Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to diversity and inclusion; recognize their areas of expertise and interest.
+You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
+Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest.
 
@@
 On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
-\cite{trianaDoesOrderFacetoFace2012}).
+\cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}).
 Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that one single networking activity will never serve the whole community.
@@
 
-% Rocío: Could start with an introduction (or put somewhere else) that online can easily make conferences more inclusive for many people, but not necessarily for people with disabilities, and that they can be easily left behind unless we actively work on their inclusion.
-% Liz: let me know if you want me to take a shot at this.adding reference marks2021meeting
-
 Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
-This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it is not a substitute for real-life experience \cite{costanzachockDesign2020}.
-Accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020} and should not be dealt with at the last minute.
-Conversely, inaccessible decisions are hard to correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
+This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}.
+Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020} and should not be dealt with at the last minute.
+Key decisions in this respect are hard to correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
 
-Importantly, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people.
-For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.
+For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
+In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free options.
 
-If the conference is in person, the venue must be accessible for people who use wheelchairs, have important signs in Braille, and a sound system compatible with hearing devices, just to name a few important features.
-Presenters should always speak into a microphone.
-Take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, menus that include sugar-free, gluten-free options.
+Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly and keyboard accessible, and have alternative text.
+Any videos should have captions and a transcript.
+The organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
+Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.
 
-Regardless of the conference format --online, in person, or hybrid-- all images used in the communication strategy of the conference should have alternative text, including the website.
-Likewise, any videos should have both captions and sound.
-
-Platforms for conference registration and abstract submission, websites, and chat platforms --if used-- should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection).
-Captioning or interpretation for presentations should be available in more than one language if possible. 
-Do not forget about making your social events accessible and making sure that the materials presented are accessible for all.
-This includes physical activities that should not restrict participation based on body type or ability.
-
-These aspects should be tested well in advance of going live.
-
-%teach about accessibility
-Some practices, such as making accessible slides and presentations, are not yet common practice and will require great efforts from the presenters if they are not used to them. 
-For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have.
-Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, in screen-reader-frendly formats (e.g., R Markdown, HTML, or \TeX{} files).
-Accessibility in slides includes alternative text, and written code instead of print screen.
-
-If presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example).
-
-Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
-They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}, (see Rule \ref{rule_financial} about allocating resources).
-Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
+Accessibility practices must include social events and networking, and include activities that do not restrict participation based on body type or ability.
+Importantly, all these accessibility practices are inclusive not only for people with disabilities but to a broader spectrum of people.
+For instance, captions are helpful for non-native speakers, having the material available for download helps attendeed with low bandwidth connection, etc.
+Accessibility efforts should be explicitly displayed in the official communications of the conference (see \textbf{Rule \ref{rule_communication}}).
 
@@
 
+This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
 The ten rules stated here can be adapted depending on the conference format and settings.
-When organizing useR! 2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
 We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
-Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
-We had to let go of certain aspects that were important to many of us. 
-From that experience, this is our last message: your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference.
-Also, keep working to improve inclusion in the next edition of the conference!
-to document and balance your work, design surveys, ask for feedback from attendees and organizers.
-Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team.
-Always remember inclusion is an ambition that organizers can and should continually pursue by allowing room for learning and improvement.
-If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
+We engaged in most of the practices mentioned here and learned others along the way. 
+Organizing a conference and implementing inclusive practices are both learning experiences;
+do not expect to achieve everything, set priorities and remember rule 10: diversity and inclusion are processes. 
+If you are part of a stable meetings committee, you could encourage organizers to follow these rules. 
+If more conferences and domains apply them, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
 And you will make a change towards healthier, stronger, and more inclusive communities.
-
-% Try not to burn you and the team out trying to make it perfect.
 

### 2021-10-16 plos_latex_template.tex
revision by Dorothea

@@
 % color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
+\usepackage[table,dvipsnames]{xcolor}
 
@@
 \begin{document}
+
+\newcommand{\fede}[1]{\textcolor{ForestGreen}{Fede: #1}}
+\newcommand{\rocio}[1]{\textcolor{MidnightBlue}{Rocio: #1}}
+\newcommand{\doro}[1]{\textcolor{RedViolet}{Dorothea: #1}}
+\newcommand{\jani}[1]{\textcolor{YellowOrange}{Janina: #1}}
+\newcommand{\nextcoauthor}[1]{\textcolor{RawSienna}{helloauthors: #1}}
+
 \vspace*{0.2in}
@@
 
-The authors of this article participated in the organization team of the annual user conference of the R Project for Statistical Computing, held in July 2021.
+The authors of this article participated in the organization of the annual user conference of the R Project for Statistical Computing, held in July 2021.
 useR!
@@
 In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference.
-These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
+These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.\fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}
 
@@
 % Rocío: Main message of the paragraph: Why we are writing this piece
-Conferences, from the Latin \textit{conferre} or `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+Conferences, from the Latin \textit{conferre} \doro{remove or, replace with -} or `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
 However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
-Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, producing knowledge, and becoming active community members that conferences can provide to the non-marginalized \cite{hendersonThoughtfulGatheringsGendering2020}.
-Exclusionary conference experiences can even divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
+Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, \doro{add "sharing and"} producing knowledge, and becoming active community members that conferences can provide to the non-marginalized \cite{hendersonThoughtfulGatheringsGendering2020}.
+Exclusionary conference experiences can \doro{Suggestion: remove "even"} even divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}.
 This message is not new. 
 Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
-and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus an the online format to open doors for inclusion.
+and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion.
 
@@
 This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for an online component.
+\doro{stronger: "to have a strong online component for all conferences"}
 The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
@@
 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
-We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone. 
+We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone \doro{is everyone too bold?
+everyone is never happy.
+"many"?}.
 Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards an inclusive conference.
@@
 
+% Dorothea: I like the intro as it is now, and I don't think the following paragraph is necessary.
 % Rocío: I'm leaving this out now and seeing if we don't need this description of communities
@@
 Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
+\doro{I would leave out the following sentence.
+It doesn't add a new thought and only partially repeats the sentence before.
+The questions are a stronger argument.
+Or we could write: "The answer to these questions depend on your field, region or community."}
 Depending on the field, region, or community, it could be Black people, LGBTQIA+, Muslim, or other people.
@@
 % Rocío: I feel like we don't need to cite any reference about privilege the way it is now. But feel free to disagree.
+%Dorothea: I agree with Rocío.
+citations for priviledge feel like citation for gravity...the current wording is not saying "there is priviledge, as shown by xyz.
+it says there is priviledge and we need to see where it is particularly strong in our field.
 
@@
 Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
+%Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
 The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
@@
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
-Build a team with people from different regions, genders, races and ethnicity, socioeconomic statuses, and other aspects of diversity. 
+Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
 Go beyond balancing all genders in this effort and pay special attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). 
@@
 Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort. 
+%Dorothea: I would remove the following sentence, I think we have made the point that it is worthwhile
 The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. 
@@
 We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
+\doro{The last sentence sounds odd. Maybe just: "until there is ample representation."}
 Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them.
@@
 % Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
-In academia and the industry, not only does the list of people considered in spotlight roles is usually biased, but the spotlight roles themselves are.
+In academia and the industry, not only is the list of people considered in spotlight roles usually biased, but the spotlight roles themselves are.
 Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
 Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
-You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
+You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
+\doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
 Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest.
@@
 Key decisions in this respect are hard to correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
-
+\doro{repetition with the next sentence}
 For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
@@
 Importantly, all these accessibility practices are inclusive not only for people with disabilities but to a broader spectrum of people.
-For instance, captions are helpful for non-native speakers, having the material available for download helps attendeed with low bandwidth connection, etc.
+For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
 Accessibility efforts should be explicitly displayed in the official communications of the conference (see \textbf{Rule \ref{rule_communication}}).
@@
 Allow the audience to ask questions in their native language and invite others to help translate it; conversely, if the presenter is not comfortable with questions in English, encourage people in the audience to help by rephrasing their questions or try to translate them; after all, the community should be supportive.
+
+% Rocío: we could eventually cite https://conferenceinference.wordpress.com/2020/11/30/when-language-is-not-a-barrier-a-tale-fr[…]istically-inclusive-conference-toma-pustelnikovaite/
 
@@
 While some expenses might be more or less fixed, % Rocío: can we give examples?
+% Dorothea: this depends so much on the conference type. I wanted to write "venue"but this is not true of course, so I don't know what to add. 
 allocation of resources has to be intentional to support the goals of the conference towards inclusion. 
@@
 Instead of depending on registration fees, ask sponsors to support specific items in your inclusion expenses.
+%Dorothea: I slightly disagree with that one. Maybe just leave the first part out?
 
@@
 Applying for loans, grants, and scholarships might be an emotionally demanding task, so if this scenario is likely, simplify the process of asking for financial support. 
+%Dorothea: I understand what we want to say, but there is a slight contradiction between "reduce the prestige so that less people apply and we can help everyone and reduce the barriers so that more people apply"
 

### 2021-10-22 plos_latex_template.tex
Some edits by LA added by AST

@@
 \newcommand{\jani}[1]{\textcolor{YellowOrange}{Janina: #1}}
-\newcommand{\nextcoauthor}[1]{\textcolor{RawSienna}{helloauthors: #1}}
+\newcommand{\la}[1]{\textcolor{RawSienna}{Laura: #1}}
 
@@
 {\Large
-\textbf\newline{Ten simple rules towards an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Ten simple rules towards hosting an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 useR!
-conferences are non-profit events organized by volunteers from the R community and arranged by the R Foundation.
+conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation.
 The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector.
@@
 In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference.
-These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.\fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}
+These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
+The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events.
+\fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}
 
@@
 % Rocío: Main message of the paragraph: Why we are writing this piece
-Conferences, from the Latin \textit{conferre} \doro{remove or, replace with -} or `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
 However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
-Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, \doro{add "sharing and"} producing knowledge, and becoming active community members that conferences can provide to the non-marginalized \cite{hendersonThoughtfulGatheringsGendering2020}.
-Exclusionary conference experiences can \doro{Suggestion: remove "even"} even divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}.
+Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
+Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
 This message is not new. 
@@
 and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion.
+%https://www.cscce.org/resources/organizing-community-events/
 
 % Rocío: Main message of the paragraph: Roadmap of the paper
-This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for an online component.
-\doro{stronger: "to have a strong online component for all conferences"}
+This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.
 The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
@@
 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
-We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone \doro{is everyone too bold?
-everyone is never happy.
-"many"?}.
-Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards an inclusive conference.
+We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
+Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
 The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
 %sara: still need to better format figure size and get the alt text right :P
-Rules 1, 2 and 10 refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
-The next two rules are focused on the people that participate in the conference: 
+Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
+The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
 Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
-Rules 5 to 9 are rules about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
-These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
+The third group includes rules 5 to 9.
+These rules are about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
+These rules apply equally to academic, industry, or mixed conferences.
+While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
 
@@
 Nonetheless, we live in a world with implicit hierarchies along these axes.
-Some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.
+Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults centered around which all systems--including conferences--are consciously and unconsciously built.
 While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
@@
 These are the groups we need to make more effort to include. 
-Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it:
-Will a more diverse conference translate into an even gender distribution in your speakers? 
+Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
+Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
 Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
 Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
-\doro{I would leave out the following sentence.
-It doesn't add a new thought and only partially repeats the sentence before.
-The questions are a stronger argument.
-Or we could write: "The answer to these questions depend on your field, region or community."}
-Depending on the field, region, or community, it could be Black people, LGBTQIA+, Muslim, or other people.
+%"The answer to these questions depend on your field, region or community.
 Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.
-% This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way.
 
@@
 % Rocío: main idea: Code of conduct for a safe place
-Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
+Adopting a code of conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
 %Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
 The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
-The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
+The code of conduct team must receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
 A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
 There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
-We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
+We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the code of conduct teamwork.
 % Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.
@@
 % Rocío: Could have a paragraph about active bystandership—Andrea
+%ast: tiny: check CoC capitalization (LA)
 
@@
 % Rocío: Idea of this paragraph: build a diverse team
-A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
+A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
 Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
-Go beyond balancing all genders in this effort and pay special attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). 
-To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. 
+Go beyond balancing all genders in this effort and pay special attention to to intersectional diversity (see \textbf{Rule \ref{rule_diversity}}). 
+To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the general population.
 If you already have an organizing team, check for gaps in its composition.
@@
 Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging (see \textbf{Rule \ref{rule_inclusion}}).
+In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).
 
@@
 The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. 
-Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly, offer leading positions, and let people from privileged groups step down. 
+Do not expect self-nomination and voting to work as mechanisms to counteract systemic inequalities.
+Instead, nominate directly and offer leading positions, and let people from privileged groups step down.
 Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
@@
 
-\section{Consciously unbias your spotlight roles}
+\section{Consciously mitigate bias in your spotlight roles}
 \label{rule_unbias}
@@
 \doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
-Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest.
+Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest and offer them space to showcase them.
 
@@
 Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}.
-Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
+Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
 This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
@@
 Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
-Alternatively, a conference could have a hybrid format with an in-person and an online component.
+Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well into a hybrid format that maximizes the experience of everyone attending your conference.
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
@@
 This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}.
-Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020} and should not be dealt with at the last minute.
-Key decisions in this respect are hard to correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.
-\doro{repetition with the next sentence}
+Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}.
+As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference.
+If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute.
+Key decisions in this respect are hard to correct.
 For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free options.
+In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free, and alcohol-free options.
 
@@
 Accessibility practices must include social events and networking, and include activities that do not restrict participation based on body type or ability.
-Importantly, all these accessibility practices are inclusive not only for people with disabilities but to a broader spectrum of people.
+Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
 For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
@@
 Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
-and conferences may miss potentially innovative contributions.
+and conferences may miss innovative contributions.
 
@@
 Scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and conferences usually ask for cover letters or applications from them. 
-These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses. 
+These programs are important for in-person participation to support travel and lodging expenses. 
 For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper. 
@@
 The authors of this piece would like to thank every single member of the organizing team of useR!
-2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR!
+2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for trusting us with the organization of useR!
 2021 and supporting us through the process.

### 2021-10-24 plos_latex_template.tex
Federico's edits

@@
 For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference.
-These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
+In this article, we streamline\fede{outline} our most important learnings\fede{lessons learnt} in 10 simple rules to host an inclusive conference.\fede{general comment on this: we should clearly state that this whole-virtual setting was a clear consequence of the pandemics, AND it was planned like this from the beginning, and not changing in-between to this format} These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
 The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events.
@@
 % Rocío: Main message of the paragraph: Why we are writing this piece
-Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.\fede{between this and the next sentence, there is a strong break.
+need some ligation, something like "Despite the many efforts in this direction, ..."}
 However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
@@
 Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
-and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion.
-%https://www.cscce.org/resources/organizing-community-events/
+and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %https://www.cscce.org/resources/organizing-community-events/
 
 % Rocío: Main message of the paragraph: Roadmap of the paper
-This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.
+This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.\fede{instead of "whether they are" -> ", which can be applied to both commonly existing formats, ..."; the fact we suggest to have an online component is not so tight to this, I guess?}
 The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
@@
 While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
+\fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}
 
@@
 % Rocío: Main idea: what is diversity and why does it matter? (inequalities)
-Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
+Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few\fede{we could substantiate this with one or two refs?}.
 Human diversity should be celebrated and respected in every way.
@@
 These are the groups we need to make more effort to include.
+\fede{General thought: I like the way we say WE and YOU.
+I still think I saw most such 10SR papers speak to the readers addressing them with the roles.
+Which would be "Organizers", "Attendees".
+Just my 2 cents as avid reader of 10SR papers ;)}
 Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
@@
 The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
-The code of conduct team must receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
-A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
+The code of conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly. 
+Assembling a diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
 There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
@@
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
-Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
+Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
 Go beyond balancing all genders in this effort and pay special attention to to intersectional diversity (see \textbf{Rule \ref{rule_diversity}}). 
@@
 % Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
-Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
+Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--especially when you are not used to them--but the positive outcomes are worthwhile. 
 Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
@@
 Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
-You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
+You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?}\fede{or, "promote a format for..."} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
 \doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
@@
 % Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
-In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
+In-person interaction in conferences is priceless, but often this is true and possible only for the ones who can afford to attend. 
 Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}.
 Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
-This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
-The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.
+This means that online conferences will have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
+The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year, without depending on a conference central location.
 
@@
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
+The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a "consolation prize" to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
 %Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
@@
 It is worth trying new and varied activities that might work with subgroups of attendees.
-Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
+Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school}
 
@@
 For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free, and alcohol-free options.
+In addition to this, the organizers should take care proactively of "invisible disabilities", for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free, and alcohol-free options.
 
@@
 Any videos should have captions and a transcript.
+
 The organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
@@
 
-Accessibility practices must include social events and networking, and include activities that do not restrict participation based on body type or ability.
+Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability.
 Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
@@
 English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
-While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker, a privilege.
-Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
+While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.
+Non-native English speakers could miss opportunities to attend or actively participate in conferences (e.g. asking questions or participating in discussions),
 and conferences may miss innovative contributions.
@@
 MARMOTS).
-Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, email lists)--
-part of your community may use a different platform than the one who regularly attend the conference. 
+Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
+part of your community may use a different platform than the ones who regularly attend the conference. 
 With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too (e.g.
@@
 
-\section{Allocate financial resources to support your conference goals}
+\section{Allocate adequate financial resources to support your conference goals}
 \label{rule_financial}
@@
 % Rocío: main idea: Budgets are limited and need to choose priorities.
-Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
+Conference budgets are usually limited and rely mostly on sponsorship and attendance fees. 
 While some expenses might be more or less fixed, % Rocío: can we give examples?
@@
 % Rocío: main idea: be part of the process
-Diversity and inclusion will be the norm in the long-term if we see conferences as part of the structural change. 
+Diversity and inclusion will be the norm in the long-term, if we see conferences as part of the structural change. 
 After the conference is over, assess whether equity and inclusion goals were met during the meeting. 
@@
 Identify the things that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
-Make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next edition of the conference.
+Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next edition of the conference.
 
@@
 This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
-The ten rules stated here can be adapted depending on the conference format and settings.
+This set of ten rules stated here can be adapted depending on the conference format and settings.
 We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
 We engaged in most of the practices mentioned here and learned others along the way.
+
 Organizing a conference and implementing inclusive practices are both learning experiences;

### 2021-10-28 plos_latex_template.tex
federico's and laura's edits plus many replies and edits responding to them, by rocío and andrea

--- 2021-10-22 plos_latex_template.tex
@@
 \newcommand{\la}[1]{\textcolor{RawSienna}{Laura: #1}}
+\newcommand{\as}[1]{\textcolor{Violet}{Andrea: #1}}
 
@@
 
-The authors of this article participated in the organization of the annual user conference of the R Project for Statistical Computing, held in July 2021.
+The authors of this article participated in the organization of the annual conference of the R Project for Statistical Computing, held in July 2021.
 useR!
@@
 For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference.
-These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
+In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference.\fede{general comment on this: we should clearly state that this whole-virtual setting was a clear consequence of the pandemics, AND it was planned like this from the beginning, and not changing in-between to this format} These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
 The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events.
-\fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}
 
@@
 % Rocío: Main message of the paragraph: Why we are writing this piece
-Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.\fede{between this and the next sentence, there is a strong break.
+need some ligation, something like "Despite the many efforts in this direction, ..."}
 However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
-Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
+Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
 Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
 This message is not new. 
-Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
-and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion.
-%https://www.cscce.org/resources/organizing-community-events/
+Barriers such as unattainable registration costs, sexism, \as{prejudice against disable people ( ableism)}, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
+and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021}, with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/
 
 % Rocío: Main message of the paragraph: Roadmap of the paper
-This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.
+This article focuses on practices for hosting inclusive conferences.
 The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
-The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
+The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
 These tips stem from the authors' experience of organizing useR!
@@
 The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
-Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
+Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of counteracting implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
 The third group includes rules 5 to 9.
-These rules are about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
-These rules apply equally to academic, industry, or mixed conferences.
+These rules are about components of the conference that should be carefully planned for: an online component, accessibility for people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
+These rules apply equally to conferences where the expected audience is academic, from the industry, or a mixed group.
 While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
+\fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}
 
@@
 % keeping references:
-%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
-% Liz adding reference marks2021meeting
+%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018, marks2021meeting}.
 
@@
 Nonetheless, we live in a world with implicit hierarchies along these axes.
-Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults centered around which all systems--including conferences--are consciously and unconsciously built.
+Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
+\as{People outside these groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia.
+Furthermore, no diversity dimension operates in isolation from the others, they intersect in complex ways to determine people's experience in the world--a concept know as \textit{intersectionality}--so that anyone can suffer from some forms of oppression and be privileged along some other dimension, and some groups are particularly vulnerable and systematically left behind.}
 While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
@@
 These are the groups we need to make more effort to include.
-Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
+\fede{General thought: I like the way we say WE and YOU.
+I still think I saw most such 10SR papers speak to the readers addressing them with the roles.
+Which would be "Organizers", "Attendees".
+Just my 2 cents as avid reader of 10SR papers ;)}
+Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
 Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
@@
 Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
-%"The answer to these questions depend on your field, region or community.
-Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.
+The answer to these questions depend on your field, region, or community.
 
@@
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, and acknowledging that gender is not binary are just some examples of decisions that can make inclusion real.
-Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.
+Accommodating for religious practices, setting specific accommodations for breastfeeding and child care, scheduling events exclusive for diverse groups (e.g., LGBTQIA+-friendly spaces), promoting and respecting pronouns for all genders and having gender-neutral bathrooms, are just some examples of decisions that are relatively easy to implement and can have a large impact in making inclusion real (see \cite{noauthor_discover2021} for some great advice on `low-hanging fruits').
+Importantly, by being proactive in creating such a welcoming space, you can be respectful of everybody's privacy and avoid requiring anyone to disclose personal information–such as revealing a disability, the sexual orientation, gender identity, or mental health issues, just to a name a few.
 
 % Rocío: main idea: Code of conduct for a safe place
-Adopting a code of conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
-%Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
-The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
-The code of conduct team must receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
-A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
-There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
-We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the code of conduct teamwork.
+Adopting a Code of Conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
+The Code of Conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
+The Code of Conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly. 
+Since people from marginalized groups are more likely to be victims of Code of Conduct violations and to see their claims dismissed due to power dynamics, assembling a diverse Code of Conduct team will be essential to deal with discrimination and harassment issues and the power dynamics underneath them and deal with them in a sensitive, victim-centered manner.
+It may also be more understanding of different disciplinary cultures and geographical and cultural considerations.
+We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
 % Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.
 
-% Rocío: Could have a paragraph about active bystandership—Andrea
-%ast: tiny: check CoC capitalization (LA)
+% Rocío: Could have a paragraph about active bystandership—Andrea%ö
 
@@
 
-% Rocío: Idea of this paragraph: build a diverse team
+% Rocío: Idea of this paragraph: build a diverse team, and representation
+
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
-Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
-Go beyond balancing all genders in this effort and pay special attention to to intersectional diversity (see \textbf{Rule \ref{rule_diversity}}). 
-To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the general population.
-If you already have an organizing team, check for gaps in its composition.
+Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
+Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
+In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).
+If you already started assembling an organizing team, check for gaps in its composition.
+The regional and local communities, groups, or associations in your field are good sources to tap into.
 
-% Rocío: Idea of this paragraph: real inclusion when having diverse members
-Gathering a diverse team will only work if there is real inclusion. 
+% Rocío: Idea of this paragraph: non-transferability and do not restrict people
 People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity.
 This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups \cite{costanzachockDesign2020}.
-People who have experience with exclusion have the social and technical expertise about removing barriers to inclusion.
-A truly inclusive and welcoming space is one in which everyone in the team is invited to be part of the solution.
-However, do not expect people from minoritized groups to only work on these solutions, and respect their interests in other areas of the organization.
+People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.
+However, they must have the freedom to choose in which areas of the conference they want to work and not be restricted to diversity and inclusion aspects.
 
-% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
-Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
-Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).
-
-% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
-The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. 
-Do not expect self-nomination and voting to work as mechanisms to counteract systemic inequalities.
-Instead, nominate directly and offer leading positions, and let people from privileged groups step down.
+% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
+When working with people from marginalized groups do not assume that self-nomination and voting will work as mechanisms to choose leadership positions.
+Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
 Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
-Offer support and encouragement if you encounter cases of impostor syndrome. 
-You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.
-
-% Rocío: Idea of this paragraph: take care of your team.
-Most importantly, take care of your team. 
-Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort. 
-%Dorothea: I would remove the following sentence, I think we have made the point that it is worthwhile
-The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. 
-However, having a strong community as the only reward may be enough for those who are in more privileged positions. 
-Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.
-In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
-Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
-Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}).
+Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.
+In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.
 
@@
 % Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).
-Make work visible, discuss collectively what should be done (and why), divide and account work.
-Being transparent and sharing the information is key to make work visible and collective.
-%sara: this relates to make labor visible from DF, maybe we can cite them here too.
-%Rocío: sure.
 
-%This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees.
-This gave a sense of variety and an appeal of "at least one networking session for everyone".
-(Marcela)
-%ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general.
-Sara: maybe you can use this as we spoke relating team work here.
-And I think the working together part is not here but the following paragraph
-%sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics
-% Rocío: the rule is getting too long and, since this is focused on networking, I'd prefer it to go to a networking section.
-
-\section{Consciously mitigate bias in your spotlight roles}
+\section{Consciously counteract bias in your spotlight roles}
+% in the conference program
 \label{rule_unbias}
@@
 %idea of paragraph 1: our lists are biased
-When choosing or inviting people as keynote speakers, program committee, session chairs, and other spotlight roles, it is likely that there will not be much diversity in the first set of names. 
-Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
-Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight. 
-Ensuring diversity in each of these roles needs to be a deliberate process. 
-We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.
-\doro{The last sentence sounds odd. Maybe just: "until there is ample representation."}
-Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them.
-An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles.
-The regional and local communities in your field are also good sources to tap into.
+
+When choosing or inviting people for visible and valued roles in the conference––such as keynote speakers, program committee members, session chairs––, it is likely that there will not be much diversity in the first set of names.
+Our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
+Rather than deter us, this should encourage us to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not stay with people in privileged groups (e.g.
+avoid `manels' and all-white panels, for instance).
+
+Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. 
+Even diverse selection committees can be subjected to unconscious bias, by interpreting information such as the location of the institution or the names of the authors and attaching value to their work according to negative or positive stereotypes.
+Selection committees need to be reminded of such biases, and try to counteract them when evaluating submissions and presentations, which should have clear evaluation criteria.
+\cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}
 
 % Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
-In academia and the industry, not only is the list of people considered in spotlight roles usually biased, but the spotlight roles themselves are.
-Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
-Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
-You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness.
-\doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
-Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest and offer them space to showcase them.
 
-\section{Have a strong online component of the conference} 
+Activities such as scientific publication and development of software are often regarded as higher value than care-taking roles like community building or teaching. 
+These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
+Research about these topics is equally undervalued compared to other areas of knowledge production.
+Counteracting bias also means broadening the range of topics of the conference and giving spotlight/visibility to usually undervalued research and the people behind it.
+%Defy the stereotypical criteria for merit by acknowledging these community practices and the people behind them.
+You can also reframe the awards ceremony to acknowledge those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
+
+\section{Have a strong online component} 
 \label{rule_online}
@@
 % Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
-In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
+In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
 Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}.
@@
 This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
-The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.
+The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones or deciding to rotate the favored timezone every year, without depending on a conference central location.
 
 % Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
-Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
-Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well into a hybrid format that maximizes the experience of everyone attending your conference.
+Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
+Alternatively, a conference could have a hybrid format with an in-person and an online components \la{that articulate well into a hybrid format that maximizes the experience of everyone attending your conference}.
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
 The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
-%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
+
+%an online component is not inclusive \textit{per se} without taking many other precautions
+
 %For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
@@
 Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
-However, in-person settings might not be comfortable or appealing to everyone for socializing; 
-some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues), and common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. 
+However, in-person settings might not be comfortable or appealing to everyone for socializing and 
+some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
 On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
 \cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}).
-Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that one single networking activity will never serve the whole community.
-It is worth trying new and varied activities that might work with subgroups of attendees.
-Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.
+Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community.
+It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference %LA cita 
+.
+Some ideas in this line are: offering the option of written chat instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}
+
+% we need to check if something here (newbies!) is not missing in the above paragraphs
+% As we mentioned previously, conference attendees should have ample opportunities to network with each other.
+However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
+If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
+Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
+%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)
 
@@
 
-Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
+Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}.
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
@@
 For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free, and alcohol-free options.
+In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.
 
 Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly and keyboard accessible, and have alternative text.
-Any videos should have captions and a transcript.
+Any videos should have good-quality captions and a transcript.
+
 The organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
@@
 
-Accessibility practices must include social events and networking, and include activities that do not restrict participation based on body type or ability.
+Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability.
 Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
 For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
-Accessibility efforts should be explicitly displayed in the official communications of the conference (see \textbf{Rule \ref{rule_communication}}).
+Accessibility efforts should be explicitly displayed in the official communication channels managed by the conference organizers (see \textbf{Rule \ref{rule_communication}}).
 
@@
 % Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
-In international conferences, the linguistic diversity of conference participants is often overlooked. 
+In international meetings, the linguistic diversity of the participants is often overlooked. 
 English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
-While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker, a privilege.
-Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
-and conferences may miss innovative contributions.
+While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.
+Non-native English speakers could miss opportunities to attend or actively participate in conferences (e.g. asking questions or participating in discussions),
+and conferences may in turn miss innovative contributions.
 
 % Rocío: main idea: make conferences more linguistically inclusive
-Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (Rule 2). 
-Advertising the conference in several languages and considering having non-English workshops and presentations to help overcome this barrier. 
-Allow the audience to ask questions in their native language and invite others to help translate it; conversely, if the presenter is not comfortable with questions in English, encourage people in the audience to help by rephrasing their questions or try to translate them; after all, the community should be supportive.
+Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (Rule \ref{rule_inclusion}).
+This may be done at different levels.
+First, by having spaces in languages other than English, both regarding the technical content of the conference and the social and networking aspects.
+These exclusive spaces have the benefit of not requiring technical adjustments, but have the problem of not communicating outside their specific public.
+Second, for presentations spoken in English, identify some other key languages for the conference and provide at least English-to-English captions (see Rule \ref{rule_accessibility}), but also translated captions or live interpretation into these key languages.
+Allow for multilingual Q \& A sessions.
+Third, include workshops, presentations, and events in languages other than English in the schedule, and provide captions or live interpretation to English.
+
+When doing this, remember your native English speaking audience to be respectful of accents and English mistakes from non-native English speakers and don't let them be discouraged by having to read captions.
+Promote the attendance to presentations in languages other than English as key components of your schedule and advertise the conference in languages other than English .
 
@@
 
-%ast: I think ome ideas of the first and second paragraphs can be joined because some parts are in both (intimidation, the importance of not missing opportunities) and maybe some practical tips can be made?
-Do we add an invitation to "bear" with subtitles and other languages to the English monolingual audience?
-an invitation to be respectful of "bad English", from submissions to presentations?
-%in person: live interpretation is doable. online: translated captions, pre-recorded talks, english trancripts
-
 \section{Express the welcoming spirit in your communication strategy}
+% media
 \label{rule_communication}
@@
 As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
-If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment.
-This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines and advertising financial support, or other creative ways that show that everyone is seen, respected, and welcome (Fig.
-MARMOTS).
-Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, email lists)--
-part of your community may use a different platform than the one who regularly attend the conference. 
-With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too (e.g.
-having accessibility guidelines on your website sends a welcoming message to people with disabilities), without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.
+Include all the work to make the conference a safe and inclusive environment in your communication strategy, 
+to emphasize that they* are particularly welcome in this event.
+This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see Rule \ref{rule_financial}).
+Try to come up with creative ways to show that everyone is seen, respected, and welcome.
+\la{For example, useR!
+2021 created "Margot, the marmot" a mascot for the conference.
+Margot is a cartoon with different designs reflecting folks that the conference wanted to reach.
+In particular, Margot wears different scarfs.
+The scarfs include colors of flags from different communities in R such as MiR´s, Rainbow R, ..., etc.
+(Figure 2)."}.
+Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
+part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
+% allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
+With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too.
 
@@
 Inclusive language--language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups--should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
-Make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. 
-Do not expect minoritized people to teach you--it's not their role--and accept feedback without being offended.
+Make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation; do not expect minoritized people to teach you--it's not their role--and accept feedback without being offended.
 
-% % Rocío: not sure if this belongs to the rule—Andrea make a convincing case :-P
-% As we mentioned previously, conference attendees should have ample opportunities to network with each other.
-However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
-Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
-%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)
-
-% Yani: We need to work a little more here, perhaps with some examples.
 % Rocío: could add some ideas from here 
 % https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
-% Batool: I will work on this rule to include more details by 25th of Sept 
-% ast: Thanks Batool! i added a couple of comments.
 
-\section{Allocate financial resources to support your conference goals}
+\section{Allocate adequate financial resources to support your conference goals}
 \label{rule_financial}
@@
 Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
-While some expenses might be more or less fixed, % Rocío: can we give examples?
-% Dorothea: this depends so much on the conference type. I wanted to write "venue"but this is not true of course, so I don't know what to add. 
-allocation of resources has to be intentional to support the goals of the conference towards inclusion. 
-It is important to estimate the costs for these inclusive practices and define your priorities in advance, like paying the organizing team, necessary training (e.g.
-code of conduct), and accessibility practices (e.g.
-captions).
-Aditional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
+While some expenses might be more or less fixed or known, (e.g., venue, internet platforms, catering) % Rocío: can we give examples?
+% Dorothea: this depends so much on the conference type.
+I wanted to write "venue"but this is not true of course, so I don't know what to add.
+%ast i added a couple more, i meant the most "obvious" budget items  in contrast to the less obvious
+allocation of resources has to be intentional to support the goals of inclusion of the conference. 
+It is important to estimate the costs for these practices and define your priorities in advance (e.g. paying the organizing team, Code of Conduct training, captioning).
+Additional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
 Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
-Instead of depending on registration fees, ask sponsors to support specific items in your inclusion expenses.
+When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion, than making generic demands for funding.
 %Dorothea: I slightly disagree with that one. Maybe just leave the first part out?
@@
 % Rocío: main idea: help take the burden out of participants for more inclusive participation
-Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
-(see https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example of conversion rates based on country of origin and career status). 
+Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
+(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
 In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
-Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}.
-Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
-Scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and conferences usually ask for cover letters or applications from them. 
-These programs are important for in-person participation to support travel and lodging expenses. 
-For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper. 
-Lower costs for attendance and offering financial support instead of prestigious grants--that could look good in a CV--may reduce the number of people applying for them and allow you to help all applicants.
-Applying for loans, grants, and scholarships might be an emotionally demanding task, so if this scenario is likely, simplify the process of asking for financial support. 
+Resources permitting, you can aim to have a conference with no registration costs.
+This is especially true for online events, but bear in mind that free events have a lower attendance rate than non-free events \cite{eventbrite_ultimate_2017}.
+%[here anything that might be general support and may allow to help all attendants, fee waivers, ]
+%Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
+Scholarships or grants to attend the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses, in in-person settings.
+Design the diversity-related grants to be explicit about the groups you want to support, to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance, while privileged people apply even when they do not need the support).
+Conferences usually ask for cover letters or applications to assign these grants, and this can be a time-consuming, emotionally-demanding task.
+Simplify as much as you can the process of asking for financial support.
+People who do not have the time (e.g., because they have care-taking roles) might decide not to apply if the process is overtly complicated.
+\la{Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups.
+Whenever possible, facilitate ways around money transfer (e.g., book flights, hotel reservations, waive conference fees).}
 %Dorothea: I understand what we want to say, but there is a slight contradiction between "reduce the prestige so that less people apply and we can help everyone and reduce the barriers so that more people apply"
+
+%For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper.
 
@@
 \section{Diversity and inclusion are processes}
+Make the conference part of a long-term process for inclusion
 \label{rule_process}
-
+%continuity, accountability, learning and being open to share the info, the teahcings, transmit what you learnt and share the list of people, expand the networks, tell others what was difficult - 
 % Rocío: main idea: we're not changing everything at the beginning, the conference will not be perfectly inclusive
@@
 Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
-There will be resistance, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
+\la{Due to these systemic problems, even the minimum changes to how conferences have been historically organized will find resistance}, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
 
 % Rocío: main idea: be part of the process
-Diversity and inclusion will be the norm in the long-term if we see conferences as part of the structural change. 
+Diversity and inclusion will be the norm in the long-term, if we see conferences as part of the required structural change. 
 After the conference is over, assess whether equity and inclusion goals were met during the meeting. 

### 2021-11-05 plos_latex_template.tex
Laura Ación on hybrid conferences

@@
 Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.
+
+% Saranjeet: At times regional institutes and universities do not support nor recognize the time put in organization tasks.
+As such one needs to carry out their day work for the institute while working extra hours for organizing.
+It might be challenging to find a way to recognize organizers that works internationally.
+However, there should certainly be some mechanism put into place which recognizes the time of the organizers.
 In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.
@@
 
+%Saranjeet: About time zones, although we can rotate and have different days dedicated to different time zones, we could also try spreading the conference time to be longer.
+So that each day the conference runs live for a couple of hours or so and the recorded material is made available/uploaded on YouTube (say) soon (every day).
+This might help because, not everyone might have the luxury of attending 4-6 hours of the conference on a weekday, even though it is running in their local timezone.
+However, people might be able to attend a couple of hours even on weekdays and will not get the feeling that they have missed a huge chunk of the conference.
+
+% Saranjeet: Not to include in this paper, but a general suggestion.
+Why not select a platform which helps in live broadcasting of the conference on LinkedIn as well?
+For example, Streamyard (https://streamyard.com/) is one such multistreaming platform, that can simultaneously stream to YouTube, LinkedIn, Twitter, and other platforms
+
 % Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
@@
 
+In fact, experts in theorizing about and building communities of practice [Wenger, 1999] who have been holding hybrid events since 2011 [Wenger-Tayner, 2021] consider that inclusive hybrid events are the future.
+These authors emphasize the need that all participants of hybrid events must go through the same event quality to the point that in-person and online participants pay the same registration amount to participate in high-quality hybrid conferences.
+These authors recommend having a buddy system, where every person attending online has an in-person attendant that makes sure the online person can follow the event and participate in the event as if they were in the room [Wenger-Trayner, 2020].
+Wenger-Trayner (2020, 2021) also recommends having a chat function for back-channel conversations between all attendants, a virtual space for all online attendants to talk to each other, and allocating time to small group conversations, often mixing in-person and online groups.
+The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space.
 %an online component is not inclusive \textit{per se} without taking many other precautions

### 2021-11-06 plos_latex_template.tex
rj and ast resolving lots of comments

@@
 
-%andrea: matt's original sentence: reaching users and developers of the R language from more than 120 countries.
-
 % % Please keep the Author Summary between 150 and 200 words
@@
 However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
-Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
+Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from opportunities for learning, sharing, and producing knowledge, and becoming active community members \cite{hendersonThoughtfulGatheringsGendering2020}.
 Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
 This message is not new. 
-Barriers such as unattainable registration costs, sexism, \as{prejudice against disable people ( ableism)}, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
-and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021}, with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/
+Barriers such as unattainable registration costs, sexism, prejudice against disable people (ableism), among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
+and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021},
+with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/
 
@@
 We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
-Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
+Here, we share the lessons learned within the time organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
 The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
@@
 
-% Dorothea: I like the intro as it is now, and I don't think the following paragraph is necessary.
-% Rocío: I'm leaving this out now and seeing if we don't need this description of communities
-%% Rocío: Main message of the paragraph: We do this for communities and people. What do we mean by communities and why are they important? (Why are they important may need some text)
-%This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
-%Some scholarly or technological communities are formally structured as learned societies (e.g.
-the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g.
-R User Groups or Python User Groups), and others may not have any established structure.
-%For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.
-
-% keeping references:
-%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018, marks2021meeting}.
-
 \section{Embrace all dimensions of diversity}
@@
 These are the groups we need to make more effort to include.
-\fede{General thought: I like the way we say WE and YOU.
-I still think I saw most such 10SR papers speak to the readers addressing them with the roles.
-Which would be "Organizers", "Attendees".
-Just my 2 cents as avid reader of 10SR papers ;)}
 Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
-Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
-Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
-Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
-The answer to these questions depend on your field, region, or community.
-
-%batool: There are a few references which define privilege and might be good to add here including: Friedman, S., O’Brien, D., & McDonald, I.
-(2021).
-Deflecting Privilege: Class Identity and the Intergenerational Self.
-Sociology, 55(4), 716–733.
-https://doi.org/10.1177/0038038520982225
-%batool: Another one is: Crevani, L.
-(2019).
-Privilege in place: How organisational practices contribute to meshing privilege in place.
-Scandinavian Journal of Management, 35(2).
-https://doi.org/10.1016/j.scaman.2018.09.002
-%batool: please note that I didn't read both entire papers I mentioned above.
-%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).
-% Rocío: I feel like we don't need to cite any reference about privilege the way it is now. But feel free to disagree.
-%Dorothea: I agree with Rocío.
-citations for priviledge feel like citation for gravity...the current wording is not saying "there is priviledge, as shown by xyz.
-it says there is priviledge and we need to see where it is particularly strong in our field.
+This may translate into different actions depending on your field, region, or community, e.g.
+a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
 
@@
 The Code of Conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
-The Code of Conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly. 
-Since people from marginalized groups are more likely to be victims of Code of Conduct violations and to see their claims dismissed due to power dynamics, assembling a diverse Code of Conduct team will be essential to deal with discrimination and harassment issues and the power dynamics underneath them and deal with them in a sensitive, victim-centered manner.
+The Code of Conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly creating protocols on how to respond.
+This should be done during the early phases of the conference planning because Code of conduct violations can happen before the conference (e.g.
+social media, organization meetings).
+Since people from marginalized groups are more likely to be victims of Code of Conduct violations and to see their claims dismissed due to power dynamics, assembling a diverse Code of Conduct team will be essential to deal with discrimination and harassment issues and the power dynamics underneath them and deal with them in a sensitive manner.
 It may also be more understanding of different disciplinary cultures and geographical and cultural considerations.
 We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
-% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.
 
@@
 Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.
-
 % Saranjeet: At times regional institutes and universities do not support nor recognize the time put in organization tasks.
@@
 However, there should certainly be some mechanism put into place which recognizes the time of the organizers.
+% Rocío: Add other types of recognition besides money (website, public recognition on social media, during the conference, etc.)
 In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.
@@
 
-\section{Consciously counteract bias in your spotlight roles}
-% in the conference program
+\section{Consciously counteract bias in the conference program}
 \label{rule_unbias}
@@
 
+% our committees are biased and unconscious bias exists
 Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. 
@@
 % Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
-
 Activities such as scientific publication and development of software are often regarded as higher value than care-taking roles like community building or teaching. 
@@
 Counteracting bias also means broadening the range of topics of the conference and giving spotlight/visibility to usually undervalued research and the people behind it.
-%Defy the stereotypical criteria for merit by acknowledging these community practices and the people behind them.
+%Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
 You can also reframe the awards ceremony to acknowledge those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
@@
 Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
-This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
-The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones or deciding to rotate the favored timezone every year, without depending on a conference central location.
+This means that online conferences have a greater reach, not only in terms of attendees but in terms of speakers  \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
+The online format may also make it easier to be inclusive of geographic regions by encompassing several time zones or deciding to rotate the favored time zone every year, without depending on a conference central location.
 
@@
 
-% Saranjeet: Not to include in this paper, but a general suggestion.
-Why not select a platform which helps in live broadcasting of the conference on LinkedIn as well?
-For example, Streamyard (https://streamyard.com/) is one such multistreaming platform, that can simultaneously stream to YouTube, LinkedIn, Twitter, and other platforms
+% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
+Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
+However, in-person settings might not be comfortable or appealing to everyone for socializing and some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
+On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
+\cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}).
+Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community.
+It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference.
+Some ideas in this line are: offering the option of written chat instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}
+% we need to check if something here (newbies!) is not missing in the above paragraphs
+% As we mentioned previously, conference attendees should have ample opportunities to network with each other.
+However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
+If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
+Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
 
@@
 Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
-Alternatively, a conference could have a hybrid format with an in-person and an online components \la{that articulate well into a hybrid format that maximizes the experience of everyone attending your conference}.
+Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well and maximize the experience of everyone attending your conference.
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
+The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{jooKeepOnlineOption2021,ninerBetterWhomLeveling2021}.
 
@@
 The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space.
-%an online component is not inclusive \textit{per se} without taking many other precautions
 
-%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
-% Rocío: Now I'm not sure if we need the last two sentences that I commented.
-
-% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
-% Rocío: I'm still doubting if this should be the 2nd paragraph, and I think that it needs some editing.
-Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
-However, in-person settings might not be comfortable or appealing to everyone for socializing and 
-some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
-On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
-\cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}).
-Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community.
-It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference %LA cita 
-.
-Some ideas in this line are: offering the option of written chat instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}
-
-% we need to check if something here (newbies!) is not missing in the above paragraphs
-% As we mentioned previously, conference attendees should have ample opportunities to network with each other.
-However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
-Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
-%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)
+% [Wenger, 1999] https://doi.org/10.1017/CBO9780511803932
+% [Wenger-Trayner, 2020] https://wenger-trayner.com/all/in-person-and-online-events/
+% [Wenger-Trayner, 2021] https://wenger-trayner.com/reflections/in-person-online-hybrid-the-future/
 
@@
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
-This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}.
+This is one key aspect where participation in the organizing team allows disabled people to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}.
 Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}.
@@
 For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.
+In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, autism and tis spectrum).
+This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.
 
@@
 Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability.
-Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
+Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
 For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
-Accessibility efforts should be explicitly displayed in the official communication channels managed by the conference organizers (see \textbf{Rule \ref{rule_communication}}).
+Accessibility efforts should be explicitly displayed in the official communication channels  (see \textbf{Rule \ref{rule_communication}}).
 
@@
 
-% Not citing ninerBetterWhomLeveling2021 anymore for now.
-
 \section{Express the welcoming spirit in your communication strategy}
@@
 part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
-% allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
+% rj: allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
 With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too.
@@
 Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
-While some expenses might be more or less fixed or known, (e.g., venue, internet platforms, catering) % Rocío: can we give examples?
-% Dorothea: this depends so much on the conference type.
-I wanted to write "venue"but this is not true of course, so I don't know what to add.
-%ast i added a couple more, i meant the most "obvious" budget items  in contrast to the less obvious
-allocation of resources has to be intentional to support the goals of inclusion of the conference. 
+While some expenses might be more or less fixed or known, (e.g., venue, internet platforms, catering), allocation of resources has to be intentional to support the goals of inclusion of the conference.
 It is important to estimate the costs for these practices and define your priorities in advance (e.g. paying the organizing team, Code of Conduct training, captioning).
@@
 Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
-When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion, than making generic demands for funding.
-%Dorothea: I slightly disagree with that one. Maybe just leave the first part out?
+When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion than making generic demands for funding.
 
@@
 %Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
-Scholarships or grants to attend the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses, in in-person settings.
-Design the diversity-related grants to be explicit about the groups you want to support, to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance, while privileged people apply even when they do not need the support).
+Scholarships or grants to attend the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses in in-person settings.
+Design the diversity-related fellowship to be explicit about the groups you want to support, to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance, while privileged people apply even when they do not need the support).
 Conferences usually ask for cover letters or applications to assign these grants, and this can be a time-consuming, emotionally-demanding task.
@@
 
-% Rocío: Let's keep this in mind for the 10th rule or conclusion
-% While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team.
-Don't be afraid to innovate and resist the institutional inertia.
-"We have never payed for this, and this has always been like that" will not take the conference towards structural change.
-% %ast should we talk about the pandemic? something like: 
-% With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
-% Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.
-
-\section{Diversity and inclusion are processes}
-Make the conference part of a long-term process for inclusion
+\section{Make the conference part of a long-term process for inclusion}
 \label{rule_process}
@@
 This can be done internally within the organizing team and taking into account the attendees' point of view. 
-Identify the things that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
+Identify the aspects that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
 Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next edition of the conference.

### 2021-11-17 plos_latex_template.tex
Rocío: edits to the intro and fig. 1

@@
 \newcommand{\as}[1]{\textcolor{Violet}{Andrea: #1}}
+\newcommand{\rita}[1]{\textcolor{CarnationPink}{Rita: #1}}
 
@@
 
-% Rocío: Main message of the paragraph: Why we are writing this piece
-Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.\fede{between this and the next sentence, there is a strong break.
-need some ligation, something like "Despite the many efforts in this direction, ..."}
-However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
-Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from opportunities for learning, sharing, and producing knowledge, and becoming active community members \cite{hendersonThoughtfulGatheringsGendering2020}.
-Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
-This message is not new. 
-Barriers such as unattainable registration costs, sexism, prejudice against disable people (ableism), among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
-and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021},
-with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/
+% Rocío: Main message of the paragraph: Conferences are great, but for some
+Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+The larger the conference, the larger the opportunities to network and learn from the cohort.
+However, opportunities for participating in conferences are not equal for all. 
+Just like in academia and tech in general, conferences can become exclusive spaces for privileged groups (i.e.
+with unearned advantages given by society to some but not all, e.g.
+white, male, from a rich country, English-native speaker, with no physical disabilities) (\cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}).
+% However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
 
-% Rocío: Main message of the paragraph: Roadmap of the paper
-This article focuses on practices for hosting inclusive conferences.
-The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
-The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
-These tips stem from the authors' experience of organizing useR!
+This article proposes 10 rules to host inclusive conferences.
+They are organized in three sections: the pillars (embracing diversity, creating a welcoming environment, and having a diverse organizing team), design rules (counteracting bias in the conference program),  (having a strong online component, achieving accessibility for people with disabilities, embracing language diversity, having inclusive communication strategies, and allocating financial resources to support inclusive practices), and a process rule: make the conference part of a long-term process for inclusion.
+These practices stem from the experience of organizing useR!
 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
 We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
-Here, we share the lessons learned within the time organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
-The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
-%sara: still need to better format figure size and get the alt text right :P
-Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
-The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
-Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of counteracting implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
-The third group includes rules 5 to 9.
+The list of authors is composed of the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
+Therefore, the rules are a result of our lessons learned from diverse perspectives and expertise (i.e. our different backgrounds, profiles, and roles in the conference). 
+Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes.
+They are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
+The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
+
+% This message is not new. 
+% Barriers such as unattainable registration costs, sexism, prejudice against disable people (ableism), among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
+% %timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018, marks2021meeting
+% and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021},
+% with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/
+% %By inclusion, we mean creating conferences organized by and dedicated to a group of attendees that is representative of the whole population.
+This means creating a welcoming, respectful, and safe conference that tries to counteract systemic oppression.
+
+% % Rocío: Main message of the paragraph: Roadmap of the paper
+% This article focuses on practices for hosting inclusive conferences.
+% The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
+% The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
+% These tips \rita{tips looks to me too informal for an article, maybe used the word `suggestions` or `advice`?} stem from the authors' experience of organizing useR!
+2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
+% We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
+% Here, we share the lessons learned within the time organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
+% The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
+% %sara: still need to better format figure size and get the alt text right :P
+
+% Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
+% The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
+% Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of counteracting implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
+% The third group includes rules 5 to 9.
 These rules are about components of the conference that should be carefully planned for: an online component, accessibility for people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
-These rules apply equally to conferences where the expected audience is academic, from the industry, or a mixed group.
+% These rules apply equally to conferences where the expected audience is academic, from the industry, or a mixed group.
 While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
-\fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}
+% \fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}
 
@@
 %sara: without success i'm trying to use pdfcomment package and pdftooltip to make the alt-text. any help?
+%%JPNG am thinking of the metaphor of gears in a machine as a proxy for the process.
+Think also about the inclusiveness of the different rules and how some facilitate others.
+The most encompassing rule is the number 10.
+This rule recognises the community learning process of DEI instead of being solely a practical and handy way of doing things.
+This should be the more comprehensive and inclusive level (gear) that include all others.
+Rules 1 and 2 are crucial because they represent two of the most important concepts we need to grasp: diversity and the requirements to express it freely.
+Rules 3 to 9 are the necessary practical pre-conditions that make the community learning process about DEI possible in the framework of a conference.
+I think that rules 1 and 2 can be represented by bigger toothed wheels than rules 3 to 9, but all need to fit and function together.The teeth of the toothed wheels would be representations of human forms representing the diverse teams involved.
 \pdftooltip{
-\includegraphics[width=\textwidth]{figs/10_rules_diversity.png}}{
+\includegraphics[width=\textwidth]{figs/rules.png}}{
 %sara: suggestion for alt text:
-Diagram of how the 10 rules are organized into three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9).
+%Diagram of how the 10 rules are organized into three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9).
 The diagram has five rows.
@@
 }
-\caption{Diagram of the rules organized in three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9).}
+\caption{Diagram of the rules organized in three groups: the pillars (rules 1, 2, and 3), design rules (rules 4 to 9), and the process (rule 10).}
 \label{fig:diagram}
@@
 Nonetheless, we live in a world with implicit hierarchies along these axes.
-Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
-\as{People outside these groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia.
+Some statuses (e.g., cisgender, white, male, from the North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
+\as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia.
 Furthermore, no diversity dimension operates in isolation from the others, they intersect in complex ways to determine people's experience in the world--a concept know as \textit{intersectionality}--so that anyone can suffer from some forms of oppression and be privileged along some other dimension, and some groups are particularly vulnerable and systematically left behind.}
@@
 a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
+For academic conferences consider also to select for a talk master and PhD student and not to give voice only to senior scientist or principal investigator.
 
@@
 
-A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
+A genuinely inclusive conference can only be coordinated  by an inclusive and diverse organizing team, in charge of the conference decision-making.
 Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
@@
 However, they must have the freedom to choose in which areas of the conference they want to work and not be restricted to diversity and inclusion aspects. 
-
+%ast: i dont know if adding something about the other side of the spectrum would work, something in the lines of "On the other hand, the non-transferability of abilities and experiences does not mean that people from privileged groups cannot work together to fight all forms of oppression.
+on the contrary," using power, co-liberation logic.
 % Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
@@
 \label{rule_unbias}
-
+%la: LA: repeating parts of my comment in Figure 1:maybe "mitigate bias" - unbiasedness is a reasonable goal once systemic biases are gone, in the meantime, introducing biases favoring marginalized folks is required to reach equity.
+Thus, while unbiasedness is the ultimate goal, the current rules should point to mitigate biases and that may require to bias in favor of marginalized groups.
+Equity and Justice call for biases in favor of marginalized folks.
+Unbiasing aims for equality that, in the current status quo, will keep favoring those who are already in the mainstream and will keep increasing marginalization for those currently in the margins.
+This is a great reference regarding the framework "mainstream/margins" https://www.trainingforchange.org/wp-content/uploads/2017/11/Mainstream-Margin-in-Groups.pdf See the cartoon about equality, equity, and justice here: https://onlinepublichealth.gwu.edu/resources/equity-vs-equality/ and here: https://www.reddit.com/r/coolguides/comments/o3ivsg/equality_equity_and_justice_explained_better/
 %idea of paragraph 1: our lists are biased
@@
 % our committees are biased and unconscious bias exists
-Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. 
+Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. %ast is something missing in this transition?
+
 Even diverse selection committees can be subjected to unconscious bias, by interpreting information such as the location of the institution or the names of the authors and attaching value to their work according to negative or positive stereotypes.
@@
 %Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
-You can also reframe the awards ceremony to acknowledge those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
+You can also reframe the awards ceremony \as{to acknowledge work beyond the traditional definition of merit/quality and include} those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
 
@@
 However, people might be able to attend a couple of hours even on weekdays and will not get the feeling that they have missed a huge chunk of the conference.
-
+%ast add something about online giving flexibility in formats and how that depends on to the context of the conference
 % Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
@@
 \cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}).
+%JPNG: I think in-person networking during a conference has a disadvantage in contrast to online and virtual interactions.
+In-person networking suffers from being contingent; they are meaningful only to the people that happen to be in the same place and at the precise time.
+If you weren't there during the coffee break, you lost the networking opportunity.
+On the contrary, written media like chats are recorded and stored, allowing you to join the conversation at every time while catching up on what has been said.
+This creates the opportunity for long term interactions during the conference, which is good for everyone and not only for people with some disadvantages (language) or different personality traits (introversion, shyness, etc.).
+I saw this kind of interaction not only during useR!2021 but also in other virtual conferences I had the opportunity to assist with.
+If you think this idea is relevant, we could add something like: "On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
+[29, 30]).
+Written media like chats create opportunities for long term interactions and networking by allowing more people to join the conversation even when it has started some time ago."
 Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community.
 It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference.
-Some ideas in this line are: offering the option of written chat instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}
+%la: see https://www.trainingforchange.org/wp-content/uploads/2017/11/Mainstream-Margin-in-Groups.pdf
+Some ideas in this line are: offering the option of written chat in addition to voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}
 % we need to check if something here (newbies!) is not missing in the above paragraphs
@@
 The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{jooKeepOnlineOption2021,ninerBetterWhomLeveling2021}.
+\rita{Sessions of an hybrid conference could have both chairs online, as attendee or speaker.
+This mean that is important when organizing the conference to include a third person who is in the venue to help the in-person participants, take the questions that are both online and in-person and communicate with the remote session chairs.}
 
@@
 Wenger-Trayner (2020, 2021) also recommends having a chat function for back-channel conversations between all attendants, a virtual space for all online attendants to talk to each other, and allocating time to small group conversations, often mixing in-person and online groups.
-The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space.
+\rita{The chat system of the conference can have a list of all the participants, with their registration name and if they are on-line or in-person, to facilitate the connections.}  The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space.
+\rita{The venue selection must respond to some criteria such as give online support both bot in-person and online participants, to have stable and secure connection for the in-person participant to join the online event.
+Moreover, offer the option to create virtual room to connect in-person and online participant for networking events, such as scientific interest group, or LGBTQIA+-friendly spaces.
+Exhibition stands in the venue need to have facilities to connect with online participants (virtual stand).
+Furthermore, the poster session need to be carefully thought to give the best experience to both in-person and online participants.
+This should have both components to allow online participants to see the poster and have the possibility to ask questions.
+However, it also need allow who want to present their poster virtually to be seen by in-person participants and respond to questions about their works.
+Lastly to better plan your hybrid conference ask in advance who wish to travel to the venue and who wish to participate online.
+This will save you time and money for looking for a suitable venue.}
 
@@
 For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, autism and tis spectrum).
-This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.
+In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, ADHD, autism and its spectrum).
+This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.%ast: check the flow here
 
@@
 and conferences may in turn miss innovative contributions.
+%lA: in case it is not mentioned later: If your budget does not allow to accommodate everyone, at a bare minimum, you should inform potential attendant which accommodations you do provide and which you will not.
+This will help everyone to know up-front what to expect and save their time if your conference does not accomodate their needs adequately.%ast: responsible communication!
 
@@
 Second, for presentations spoken in English, identify some other key languages for the conference and provide at least English-to-English captions (see Rule \ref{rule_accessibility}), but also translated captions or live interpretation into these key languages.
-Allow for multilingual Q \& A sessions.
+Allow for multilingual Q \& A sessions with an interpreter.
 Third, include workshops, presentations, and events in languages other than English in the schedule, and provide captions or live interpretation to English.
@@
 Include all the work to make the conference a safe and inclusive environment in your communication strategy, 
-to emphasize that they* are particularly welcome in this event.
+to emphasize that everybody is welcome in this event.
 This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see Rule \ref{rule_financial}).
@@
 Inclusive language--language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups--should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
-Make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation; do not expect minoritized people to teach you--it's not their role--and accept feedback without being offended.
+Make the effort to teach yourself the vocabulary %rita A reference to this vocabulary could be add on the conference main page?
+or in the code of conduct?
+%ast- good idea, we can point to resources here like the reference i added before: https://www2.le.ac.uk/projects/oer/oers/ssds/oers/study-skills/studyskills.pdf the main aspect of the "teach yourself" part is that you don't ask people to teach you but go for it, but giving some links is nice.
+and the best ways to communicate to account for every culture and situation; do not expect minoritized people to teach you--it's not their role--and accept feedback without being offended.
 
@@
 Conferences usually ask for cover letters or applications to assign these grants, and this can be a time-consuming, emotionally-demanding task.
-Simplify as much as you can the process of asking for financial support.
+Simplify as much as you can the process of asking for financial support %rita Could you give a concrete example?
+ast - yes!
+ö
 People who do not have the time (e.g., because they have care-taking roles) might decide not to apply if the process is overtly complicated.
@@
 \label{rule_process}
-%continuity, accountability, learning and being open to share the info, the teahcings, transmit what you learnt and share the list of people, expand the networks, tell others what was difficult - 
-% Rocío: main idea: we're not changing everything at the beginning, the conference will not be perfectly inclusive
-Do not expect the conference to be perfect. 
-Resources, both in time and money, are always limited. 
+
+Make the conference part of a long-term process for inclusion.
+Do not expect the conference to encompass all the dimensions of diversity at once.
+Resources, both in time and money, are always limited, so you need to have a clear diversity vision and understand what goals are feasible.
+Be explicit about the biases you are tackling and acknowledge which audiences your conference is not doing enough to include.
+Conceiving diversity and inclusion as a step-by-step process through successive conference versions can increase success and reduce unintended discrimination in the long run.
 Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
@@
 % Rocío: main idea: be part of the process
-Diversity and inclusion will be the norm in the long-term, if we see conferences as part of the required structural change. 
+Diversity and inclusion will be the norm in the long-term if we see conferences as part of the required structural change. 
 After the conference is over, assess whether equity and inclusion goals were met during the meeting. 
 This can be done internally within the organizing team and taking into account the attendees' point of view. 
+%jpng PNG: We are witnessing a cultural change in people attitudes to diversity and inclusion that will make these social goals essential components of conferences and meetings organization.
+In this process of change, feedback and evaluation of the successes and failures are crucial.
+Assessing whether equity and inclusion goals were met during the meeting can be done internally within the organizing team and pondering the attendees’ point of view.
 Identify the aspects that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
 Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next edition of the conference.
-
-% For many people working with diversity, one typical aim is institutionalizing diversity (Ahmed 2012).
-However, our aim is also a symptom that diversity and equity are not part of institutions, academic societies, and conferences.
-This rule is about the process of making diversity and inclusion possible and a complement to all the other rules.
-We believe we need to be accountable and we need to aim for permanence.
-Institutionality comes from a set of norms, values, and priorities and these determine who is allowed and how (Ahmed 2012).
-Our long-term goal is to set new norms, values, and priorities based on diversity.
-Organizing a diverse conference is part of the process of change.
-However, we cannot do everything at one conference or a few of them.
-It is important to have a bold vision and clear targets.
-Keep in mind that ideal actions are often not possible and resources are limited (both time of the organizing team and money.
-
-% If from the beginning we set diversity goals and actions for the conference, we can use that to be transparent and, later, accountable.
-Assess whether equity and inclusion goals were met during the meeting.
-This can be done internally within the organizing team and often also through the conference’s participants.
-Plan, balance and keep working to improve inclusion in the next edition of the conference.
-Try to propose permanent data collection to assess whether diversity and inclusion are actually being seen at the conference.
-We know that diversity and equity are not institutionalized, in fact, the system works to stop those who are changing the system (Ahmed 2019).
-We will find resistance, but diversity is a school.
-We also learn more about how people are being excluded, we learn that there are actions we can take to include more, and then, we can set higher goals on how to make inclusion possible.
 

### 2021-11-17 plos_latex_template.tex
Changes to section structure

@@
 
-\section{Embrace all dimensions of diversity}
+\section*{The pillars}
+
+To make a more inclusive conference, it is necessary to identify which groups of people are being excluded, the barriers they may face to be part of the conference, and strategies to reduce them. 
+The first rule is about embracing all dimensions of diversity and prioritizing the
+ participation of groups that are often marginalized.
+However, increasing diversity is not enough. Rule 2 is about creating a safe and welcoming environment to make everyone feel part of the conference and the community. 
+Inclusion should start from the organizing team. Rule 3 refers to the importance of building an inclusive and diverse organizing team, and provides advice on work dynamics.
+
+\subsection*{Rule 1: Embrace all dimensions of diversity}
 \label{rule_diversity}
@@
 
-\section{Create a safe and welcoming environment}
+\subsection*{Rule 2: Create a safe and welcoming environment}
 \label{rule_inclusion}
@@
 
-\section{Have an inclusive and diverse organizing team}
+\subsection*{Rule 3: Have an inclusive and diverse organizing team}
 \label{rule_organizing_team}
@@
 Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).
+In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see Rule 2 about a safe and welcoming environment).
 If you already started assembling an organizing team, check for gaps in its composition.
@@
 
-\section{Consciously counteract bias in the conference program}
+\section*{Design Rules}
+
+These rules concern diverse elements in the design of the conference. They are all important pieces for inclusion.
+In rule 4, we introduce ways to counteract bias in the conference program, in keynotes, program committee, abstract selection, and other spaces. 
+Rule 5 describes the advantages of online conferences reducing barriers for participation, and provides advice for organizing online and hybrid conferences.
+Rule 6 focuses on practices to make all conference platforms and spaces accessible to people with disabilities. 
+In rule 7, we advocate for more language-inclusive conferences, and provide suggestions to encourage and ensure full participation of non-native English speakers. 
+Rule 8 provides tips for an inclusive media strategy. 
+In rule 9, we address budgeting for inclusive practices, and helping participants with affordable registration costs, scholarships, and other forms of financial support.
+
+\subsection*{Rule 4: Consciously counteract bias in the conference program}
 \label{rule_unbias}
@@
 
-\section{Have a strong online component} 
+\subsection*{Rule 5: Have a strong online component} 
 \label{rule_online}
@@
 
-\section{Make the conference accessible to people with disabilities}
+\subsection*{Rule 6: Make the conference accessible to people with disabilities}
 \label{rule_accessibility}
@@
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
-This is one key aspect where participation in the organizing team allows disabled people to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}.
+This is one key aspect where participation in the organizing team allows disabled people to take part in the decisions from the beginning (see rule 3 about the organizing team), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}.
 Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}.
@@
 For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
-Accessibility efforts should be explicitly displayed in the official communication channels  (see \textbf{Rule \ref{rule_communication}}).
+Accessibility efforts should be explicitly displayed in the official communication channels (see rule 8 about media strategy).
 
-\section{Don't let language restrict high-quality participation}
+\subsection*{Rule 7: Don't let language restrict high-quality participation}
 \label{rule_language}
@@
 % Rocío: main idea: make conferences more linguistically inclusive
-Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (Rule \ref{rule_inclusion}).
+Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (rule 2 about a welcoming environment for all).
 This may be done at different levels.
@@
 These exclusive spaces have the benefit of not requiring technical adjustments, but have the problem of not communicating outside their specific public.
-Second, for presentations spoken in English, identify some other key languages for the conference and provide at least English-to-English captions (see Rule \ref{rule_accessibility}), but also translated captions or live interpretation into these key languages.
+Second, for presentations spoken in English, identify some other key languages for the conference and provide at least English-to-English captions (see rule 6 about accessibility for disabilities), but also translated captions or live interpretation into these key languages.
 Allow for multilingual Q \& A sessions with an interpreter.
@@
 
-\section{Express the welcoming spirit in your communication strategy}
+\subsection*{Rule 8: Express the welcoming spirit in your communication strategy}
 % media
@@
 to emphasize that everybody is welcome in this event.
-This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see Rule \ref{rule_financial}).
+This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see rule 9 about financial resources).
 Try to come up with creative ways to show that everyone is seen, respected, and welcome.
@@
 (Figure 2)."}.
-Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
+Also, try to advertise the conference in multiple languages (see rule 7 about language) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
 part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
@@
 
-\section{Allocate adequate financial resources to support your conference goals}
+\subsection*{Rule 9: Allocate adequate financial resources to support your conference goals}
 \label{rule_financial}
@@
 
-\section{Make the conference part of a long-term process for inclusion}
+\section*{The Process}
+
+One conference can make a difference in the path towards inclusion.
+The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encounters to the next organizers.
+
+\subsection*{Rule 10: Make the conference part of a long-term process for inclusion}
 \label{rule_process}

### 2021-11-17 plos_latex_template.tex
Rocío: adding authors

@@
 \\
-Name1 Surname\textsuperscript{1,2\Yinyang},
-Name2 Surname\textsuperscript{2\Yinyang},
-Name3 Surname\textsuperscript{2,3\textcurrency},
-Name4 Surname\textsuperscript{2},
-Name5 Surname\textsuperscript{2\ddag},
-Name6 Surname\textsuperscript{2\ddag},
-Name7 Surname\textsuperscript{1,2,3*},
-with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
+Rocío Joo\textsuperscript{1*}, %,2\Yinyang},
+Andrea Sánchez-Tapia\textsuperscript{2},
+Sara Mortara\textsuperscript{3},
+Heather Turner\textsuperscript{4}, %\ddag},
+Dorothea Hug Peter\textsuperscript{5}, %\ddag},
+Matt Bannert\textsuperscript{6},
+Batool Almazrouq\textsuperscript{7},
+Liz Hare\textsuperscript{8},
+Janani Ravi\textsuperscript{9},
+Laura Ación\textsuperscript{10},
+Juan Pablo Narváez\textsuperscript{11},
+Marcela Alfaro Córdoba\textsuperscript{12},
+Federico Marini\textsuperscript{13},
+Rita Giordano\textsuperscript{14},
+Natalia Morandeira\textsuperscript{15},
+Adithi Upadhya\textsuperscript{16},
+Joselyn Chávez\textsuperscript{17\Yinyang},
+Anicet Ebou\textsuperscript{18\Yinyang},
+Silvia Canelón\textsuperscript{19\Yinyang},
+Yanina Bellini Saibene\textsuperscript{20}%,3\textcurrency}
 \\
 \bigskip
-\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
 \\
@@
 \textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{4} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{5} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{6} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{7} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{8} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{9} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{11} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{12} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{14} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{15} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{17} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{18} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{19} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
 \\
@@
 % Also use this double-dagger symbol for special authorship notes, such as senior authorship.
-\ddag These authors also contributed equally to this work.
+% \ddag These authors also contributed equally to this work.
 
 % Current address notes
-\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
 % \textcurrency b Insert second current address 
@@
 % Deceased author note
-\dag Deceased
-
-% Group/Consortium Author Note
-\textpilcrow Membership list can be found in the Acknowledgments section.
+% \dag Deceased
 
 % Use the asterisk to denote corresponding authorship and provide email address in note below.
-* correspondingauthor@institute.edu
+* rocio.joo@globalfishingwatch.org
 

### 2021-11-20 plos_latex_template.tex
Rocío: editing conclusion

@@
 We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
+
 The list of authors is composed of the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
@@
 
-One conference can make a difference in the path towards inclusion.
-The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encounters to the next organizers.
+A conference can make a difference in the path towards inclusion.
+The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encountered to the next organizers.
 
@@
 
-Make the conference part of a long-term process for inclusion.
-Do not expect the conference to encompass all the dimensions of diversity at once.
-Resources, both in time and money, are always limited, so you need to have a clear diversity vision and understand what goals are feasible.
-Be explicit about the biases you are tackling and acknowledge which audiences your conference is not doing enough to include.
+% Rocío: main idea: no first attempt is guaranteed to succeed
+Make the conference part of a long-term process for inclusion. Do not expect the conference to encompass all the dimensions of diversity at once.
 Conceiving diversity and inclusion as a step-by-step process through successive conference versions can increase success and reduce unintended discrimination in the long run.
-Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
-\la{Due to these systemic problems, even the minimum changes to how conferences have been historically organized will find resistance}, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
+Diversity and inclusion will be the norm in the long-term if we see conferences as part of the required structural change.
 
-% Rocío: main idea: be part of the process
-Diversity and inclusion will be the norm in the long-term if we see conferences as part of the required structural change. 
+% Ask for feedback to know the extent of your success
 After the conference is over, assess whether equity and inclusion goals were met during the meeting.
-This can be done internally within the organizing team and taking into account the attendees' point of view. 
-%jpng PNG: We are witnessing a cultural change in people attitudes to diversity and inclusion that will make these social goals essential components of conferences and meetings organization.
-In this process of change, feedback and evaluation of the successes and failures are crucial.
-Assessing whether equity and inclusion goals were met during the meeting can be done internally within the organizing team and pondering the attendees’ point of view.
+Ask for feedback about the conference by posting a survey during and after the conference to let attendees evaluate whether the inclusion goals of the conference were met.
+When designing feedback surveys, ask for different information that would let you know about the demographic composition of the attendees to the conference. 
+Remember that some of this information varies greatly from region to region and adapt your options to the local or regional context.
+When choosing categories, avoid letting people in minoritized categories fall in the "other" option.
+You can also try to make open answer options to let people self-identify and express their opinions freely.
+Let all responses be optional and anonymous, but encourage your attendance to answer and commit to data privacy.
 Identify the aspects that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
-Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next edition of the conference.
+
+% Pass
+Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next editions of the conference.
+If you are part of a stable meetings committee, encourage organizers to follow these rules and help set up new standards for inclusion.
+Share also your new and extended network and nominate people for future leadership.
+A good way to do this is documenting your processes, lessons, and guides in blog posts, reports and presentations.
+They may be useful not only for your conference series but for other communities that are having the same questions that you are.
+
+%sale? Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
+%\la{Due to these systemic problems, even the minimum changes to how conferences have been historically organized will find resistance}, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
+%responbile communication (not here, see laura's comment) Be explicit about the biases you are tackling and acknowledge which audiences your conference is not doing enough to include.
 
@@
 
-This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
-This set of ten rules stated here can be adapted depending on the conference format and settings.
-We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
-We engaged in most of the practices mentioned here and learned others along the way.
-
+This article suggests changes in conference planning towards diversity and inclusion to make them networking and learning opportunities for all. 
 Organizing a conference and implementing inclusive practices are both learning experiences;
-do not expect to achieve everything, set priorities and remember rule 10: diversity and inclusion are processes. 
-If you are part of a stable meetings committee, you could encourage organizers to follow these rules. 
-If more conferences and domains apply them, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
-And you will make a change towards healthier, stronger, and more inclusive communities.
+you may not achieve everything you aimed for, set priorities and remember rule 10: diversity and inclusion are processes. 
+Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that a single conference cannot change. 
+Due to these systemic problems, even the minimum changes to how conferences have been historically organized may find resistance, but your courage to make these changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of people who are often excluded.
+Moreover, if more conferences and domains apply these rules, there will be less resistance, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
+% Rocío: I don't like the ending.
 

### 2021-11-24 plos_latex_template.tex
ast, rj addressing the structure of all rules and comments by laura, juan pablo, rita, federico

@@
 However, opportunities for participating in conferences are not equal for all. 
-Just like in academia and tech in general, conferences can become exclusive spaces for privileged groups (i.e.
-with unearned advantages given by society to some but not all, e.g.
-white, male, from a rich country, English-native speaker, with no physical disabilities) (\cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}).
+Just like in academia and tech in general, conferences can become exclusive spaces for privileged groups, i.e.
+with unearned advantages given by society to some but not all (e.g.
+white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
 % However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
@@
 This article proposes 10 rules to host inclusive conferences.
-They are organized in three sections: the pillars (embracing diversity, creating a welcoming environment, and having a diverse organizing team), design rules (counteracting bias in the conference program),  (having a strong online component, achieving accessibility for people with disabilities, embracing language diversity, having inclusive communication strategies, and allocating financial resources to support inclusive practices), and a process rule: make the conference part of a long-term process for inclusion.
+They are organized in three sections: the pillars (embracing diversity, creating a welcoming environment, and having a diverse organizing team), design rules (counteracting bias in the conference program, having a strong online component, achieving accessibility for people with disabilities, embracing language diversity, having inclusive communication strategies, and allocating financial resources to support inclusive practices), and a process rule: make the conference part of a long-term process for inclusion.
 These practices stem from the experience of organizing useR!
@@
 %sara: without success i'm trying to use pdfcomment package and pdftooltip to make the alt-text. any help?
-%%JPNG am thinking of the metaphor of gears in a machine as a proxy for the process.
-Think also about the inclusiveness of the different rules and how some facilitate others.
-The most encompassing rule is the number 10.
-This rule recognises the community learning process of DEI instead of being solely a practical and handy way of doing things.
-This should be the more comprehensive and inclusive level (gear) that include all others.
-Rules 1 and 2 are crucial because they represent two of the most important concepts we need to grasp: diversity and the requirements to express it freely.
-Rules 3 to 9 are the necessary practical pre-conditions that make the community learning process about DEI possible in the framework of a conference.
-I think that rules 1 and 2 can be represented by bigger toothed wheels than rules 3 to 9, but all need to fit and function together.The teeth of the toothed wheels would be representations of human forms representing the diverse teams involved.
+
 \pdftooltip{
@@
 
-To make a more inclusive conference, it is necessary to identify which groups of people are being excluded, the barriers they may face to be part of the conference, and strategies to reduce them. 
+To make a more inclusive conference, it is necessary to identify which groups of people are being excluded, the barriers they may face to be part of the conference, and strategies to overcome them. 
 The first rule is about embracing all dimensions of diversity and prioritizing the
@@
 However, increasing diversity is not enough. Rule 2 is about creating a safe and welcoming environment to make everyone feel part of the conference and the community. 
-Inclusion should start from the organizing team. Rule 3 refers to the importance of building an inclusive and diverse organizing team, and provides advice on work dynamics.
+Inclusion should start from the organizing team. Rule 3 refers to the importance of starting with an inclusive and diverse organizing team, and provides advice on work dynamics.
 
@@
 Nonetheless, we live in a world with implicit hierarchies along these axes.
-Some statuses (e.g., cisgender, white, male, from the North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
+Some statuses (e.g., cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
 \as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia.
-Furthermore, no diversity dimension operates in isolation from the others, they intersect in complex ways to determine people's experience in the world--a concept know as \textit{intersectionality}--so that anyone can suffer from some forms of oppression and be privileged along some other dimension, and some groups are particularly vulnerable and systematically left behind.}
+No form of oppression operates in isolation from the others, which means some groups are particularly vulnerable and systematically left behind.}
 While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
+Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against.
+These are the groups we need to make more effort to include.
+Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
+% vision
+% back to the indicators? 
+% realistic but bold (back?)
+% be aware and promote awareness
+%¡LAS BARRERAS!
+%Express the diversity and inclusion vision for your conference by writing a diversity statement.
 
 % Rocío: more effort towards including the more excluded
-Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against.
-These are the groups we need to make more effort to include.
-Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
 This may translate into different actions depending on your field, region, or community, e.g.
 a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
-For academic conferences consider also to select for a talk master and PhD student and not to give voice only to senior scientist or principal investigator.
 
@@
 While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
-Accommodating for religious practices, setting specific accommodations for breastfeeding and child care, scheduling events exclusive for diverse groups (e.g., LGBTQIA+-friendly spaces), promoting and respecting pronouns for all genders and having gender-neutral bathrooms, are just some examples of decisions that are relatively easy to implement and can have a large impact in making inclusion real (see \cite{noauthor_discover2021} for some great advice on `low-hanging fruits').
+% Edit and make more specific (below)
+Promote respect of pronouns, schedule events exclusive for diverse groups (e.g., LGBTQIA+-friendly spaces), and if the conference has an in-person component, set spaces for religious practices, breastfeeding and child care, and include menu options that account for diverse dietary and religious requirements.
+These are just some examples of decisions that are relatively easy to implement and can have a large impact in making inclusion real (see \cite{noauthor_discover2021} for some great advice on `low-hanging fruits').
 Importantly, by being proactive in creating such a welcoming space, you can be respectful of everybody's privacy and avoid requiring anyone to disclose personal information–such as revealing a disability, the sexual orientation, gender identity, or mental health issues, just to a name a few.
-
+% Rocío: Could have a paragraph about active bystandership—Andrea%ö
+%promote a supportive and respectful environment where attendants act as active bystanders. y definir '¬¬ 
 % Rocío: main idea: Code of conduct for a safe place
@@
 
-% Rocío: Could have a paragraph about active bystandership—Andrea%ö
-
 \subsection*{Rule 3: Have an inclusive and diverse organizing team}
@@
 
-A genuinely inclusive conference can only be coordinated  by an inclusive and diverse organizing team, in charge of the conference decision-making.
-Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
-Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see Rule 2 about a safe and welcoming environment).
+A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team in charge of the conference decision-making.
+Organizers should be a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
+Having diverse people in decision-making positions will positively affect all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
+In addition, a diverse team plays an important role in creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see Rule 2 about a safe and welcoming environment).
 If you already started assembling an organizing team, check for gaps in its composition.
@@
 This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups \cite{costanzachockDesign2020}.
-People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.
+People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
 However, they must have the freedom to choose in which areas of the conference they want to work and not be restricted to diversity and inclusion aspects. 
-%ast: i dont know if adding something about the other side of the spectrum would work, something in the lines of "On the other hand, the non-transferability of abilities and experiences does not mean that people from privileged groups cannot work together to fight all forms of oppression.
-on the contrary," using power, co-liberation logic.
 % Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
@@
 Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
-Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
+Build an environment in which every person can express their opinion and pay special attention to people from systematically excluded groups.
 Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.
-% Saranjeet: At times regional institutes and universities do not support nor recognize the time put in organization tasks.
-As such one needs to carry out their day work for the institute while working extra hours for organizing.
-It might be challenging to find a way to recognize organizers that works internationally.
-However, there should certainly be some mechanism put into place which recognizes the time of the organizers.
 % Rocío: Add other types of recognition besides money (website, public recognition on social media, during the conference, etc.)
@@
 
-These rules concern diverse elements in the design of the conference. They are all important pieces for inclusion.
+These rules concern diverse elements in the design of the conference and operational tasks. They are all important pieces for inclusion.
 In rule 4, we introduce ways to counteract bias in the conference program, in keynotes, program committee, abstract selection, and other spaces. 
@@
 \label{rule_unbias}
-%la: LA: repeating parts of my comment in Figure 1:maybe "mitigate bias" - unbiasedness is a reasonable goal once systemic biases are gone, in the meantime, introducing biases favoring marginalized folks is required to reach equity.
-Thus, while unbiasedness is the ultimate goal, the current rules should point to mitigate biases and that may require to bias in favor of marginalized groups.
-Equity and Justice call for biases in favor of marginalized folks.
-Unbiasing aims for equality that, in the current status quo, will keep favoring those who are already in the mainstream and will keep increasing marginalization for those currently in the margins.
-This is a great reference regarding the framework "mainstream/margins" https://www.trainingforchange.org/wp-content/uploads/2017/11/Mainstream-Margin-in-Groups.pdf See the cartoon about equality, equity, and justice here: https://onlinepublichealth.gwu.edu/resources/equity-vs-equality/ and here: https://www.reddit.com/r/coolguides/comments/o3ivsg/equality_equity_and_justice_explained_better/
 %idea of paragraph 1: our lists are biased
@@
 avoid `manels' and all-white panels, for instance).
+Make sure the composition of all these groups is balanced across topics, areas, and representation across regions, gender, and race.
+%rita: For academic conferences consider also to select for a talk master and PhD students and not to give voice only to senior scientist or principal investigator.
+lo pongo acá porque sí faltan cosas prácticas.
 
 % our committees are biased and unconscious bias exists
-Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. %ast is something missing in this transition?
+%Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field.
 
-Even diverse selection committees can be subjected to unconscious bias, by interpreting information such as the location of the institution or the names of the authors and attaching value to their work according to negative or positive stereotypes.
-Selection committees need to be reminded of such biases, and try to counteract them when evaluating submissions and presentations, which should have clear evaluation criteria.
+The final list of accepted talks can also be biased. 
+On the one hand, because people do not feel confident to submit their work. 
+To counteract bias in submission, promote your calls beyond the usual communities and keep in touch with local groups and communities of practice to encourage submissions by their members.
+Some of these communities have pre-submission mechanisms to help their members prepare and improve their submissions.%sin el ejemplo no se entiende
+% 1. Clear rules on reviews: double blind, double open, kind of open, etc.
+% Be transparent on the evaluation criteria
+% 2. Diverse program committee (above)
+% 3. Unconscious bias (test)
+% 4. Curate the final list against 
+%see: https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
+On the other hand, proposals may be biased because of even diverse selection committees can be subjected to unconscious or implicit bias %explain!.
+Unconscious bias happens when we interpret information, such as names of the authors, we assume their origin or ethnicity, their institution of origin, and we attach a positive or negative value to their work according to negative or positive stereotypes.
+To counteract this...
+The review process should guarantee the fair evaluation of the submitted work, and this may  will be double blind, open, or partially open. 
+In some fields, it's not possible to guarantee anonymity of the presenters, and in some and external cues such as country of origin and links to repositories. 
+Create clear evaluation criteria for the submitted proposals, and share them with the authors and reviewers.
+Remind the reviewers that unconscious bias exists and encourage them to take unconscious 
+Selection committees need to be reminded of such biases, and be aware of them when evaluating submissions and presentations.
 \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}
@@
 % Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
-Activities such as scientific publication and development of software are often regarded as higher value than care-taking roles like community building or teaching. 
+Activities such as scientific publication and development of software are often regarded as higher value than other roles like community building or teaching. 
 These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
 Research about these topics is equally undervalued compared to other areas of knowledge production.
-Counteracting bias also means broadening the range of topics of the conference and giving spotlight/visibility to usually undervalued research and the people behind it.
+Counteracting bias also means changing the standard program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, // and giving visibility to the whole range of activities and practitioners that contribute to the overall topics of the conference.
+
 %Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
-You can also reframe the awards ceremony \as{to acknowledge work beyond the traditional definition of merit/quality and include} those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
+You can also reframe the awards ceremony to acknowledge work beyond the traditional definition of merit/quality and include those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
 
@@
 % Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
+%in 2021 an inclusive conference has to be an online or hybrid conference.
+this means a radical change in the functioning of all spaces, to include people who are at home for any reason.
+acá contexto de covid - within ...
+y ser más firmes con lo de hybrid.
+joo - wenger.
+%primer parrafo como está cite joo, cosas prácticas de online - están en las citas, poner acá.
+2.
+networking hybrid qué es, tipos, active and distributed serían nuestras opciones - omitir las otras por no ser inclusivas.
+todo lo inclusivo de online con todo lo inclusivo de presencial.
+un par de ejemplos de lo presencial.
+o adaptaciones que requiere la parte presencial cuando se vuelve híbrido (chat, q and a, streaming) opportunities of networking - que ser chair o keynote o presenter no cambie si estás online o presencial, no requiera que estés presente.
+%practical: choose the platforms and be aware that reach varies between countries
+As stated in [the nature correspondence]: The COVID pandemic forced us to embrace online conferences (see C.
+Woolston Nature 582, 135–136; 2020), which remove some of the barriers that disproportionately affect marginalized groups.
+These include the cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications (H.
+J.
+Niner \& S.
+N.
+Wassermann Front.
+Mar.
+Sci.
+8, 638025; 2021; Gewin, V.
+Nature 569, 297–299; 2019).
+Inclusive conferences must have an online component [cite all the other references].
 In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
@@
 Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
-However, in-person settings might not be comfortable or appealing to everyone for socializing and some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
-On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
-\cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}).
-%JPNG: I think in-person networking during a conference has a disadvantage in contrast to online and virtual interactions.
-In-person networking suffers from being contingent; they are meaningful only to the people that happen to be in the same place and at the precise time.
-If you weren't there during the coffee break, you lost the networking opportunity.
-On the contrary, written media like chats are recorded and stored, allowing you to join the conversation at every time while catching up on what has been said.
-This creates the opportunity for long term interactions during the conference, which is good for everyone and not only for people with some disadvantages (language) or different personality traits (introversion, shyness, etc.).
-I saw this kind of interaction not only during useR!2021 but also in other virtual conferences I had the opportunity to assist with.
-If you think this idea is relevant, we could add something like: "On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
+%However, in-person settings might not be comfortable or appealing to everyone for socializing and some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
+However, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
+\cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}.
+% JPNG: "On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
 [29, 30]).
@@
 % Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
-Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
 Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well and maximize the experience of everyone attending your conference.
 This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
+Conference organizers who have been holding hybrid events since 2011 [Wenger-Tayner, 2021] consider that inclusive hybrid events are the future.
+These authors emphasize the need that all participants of hybrid events must go through the same event quality to the point that in-person and online participants pay the same registration amount to participate in high-quality hybrid conferences.
+These authors recommend having a buddy system, where every person attending online has an in-person attendant that makes sure the online person can follow the event and participate in the event as if they were in the room [Wenger-Trayner, 2020].
+Wenger-Trayner (2020, 2021) also recommends having a chat function for back-channel conversations between all attendants, a virtual space for all online attendants to talk to each other, and allocating time to small group conversations, often mixing in-person and online groups.
+% Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
 The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{jooKeepOnlineOption2021,ninerBetterWhomLeveling2021}.
@@
 
-In fact, experts in theorizing about and building communities of practice [Wenger, 1999] who have been holding hybrid events since 2011 [Wenger-Tayner, 2021] consider that inclusive hybrid events are the future.
-These authors emphasize the need that all participants of hybrid events must go through the same event quality to the point that in-person and online participants pay the same registration amount to participate in high-quality hybrid conferences.
-These authors recommend having a buddy system, where every person attending online has an in-person attendant that makes sure the online person can follow the event and participate in the event as if they were in the room [Wenger-Trayner, 2020].
-Wenger-Trayner (2020, 2021) also recommends having a chat function for back-channel conversations between all attendants, a virtual space for all online attendants to talk to each other, and allocating time to small group conversations, often mixing in-person and online groups.
 \rita{The chat system of the conference can have a list of all the participants, with their registration name and if they are on-line or in-person, to facilitate the connections.}  The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space.
@@
 
+% Be more direct
 Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}.
@@
 Key decisions in this respect are hard to correct.
-For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
+
+For in-person components of conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
 In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, ADHD, autism and its spectrum).
-This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.%ast: check the flow here
+This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.
 
@@
 Any videos should have good-quality captions and a transcript.
-
 The organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
@@
 For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
-Accessibility efforts should be explicitly displayed in the official communication channels (see rule 8 about media strategy).
+Accessibility efforts should be explicitly displayed in the official communication channels (see rule 8 about communication strategy).
+
+%Live captioning or interpretation!
+and surtitles (Rita) LA: provide live, high-quality (automated captions are often referred to as "craptions"), captioning or simultaneous translation in English for persons who only speak English.
+After all, captions and simultaneous translations are common place in the audiovisual industry and your conference audience will likely be already familiar with this if they ever watched a movie from a non-English speaking country.
 
@@
 and conferences may in turn miss innovative contributions.
-%lA: in case it is not mentioned later: If your budget does not allow to accommodate everyone, at a bare minimum, you should inform potential attendant which accommodations you do provide and which you will not.
-This will help everyone to know up-front what to expect and save their time if your conference does not accomodate their needs adequately.%ast: responsible communication!
 
@@
 This may be done at different levels.
-First, by having spaces in languages other than English, both regarding the technical content of the conference and the social and networking aspects.
-These exclusive spaces have the benefit of not requiring technical adjustments, but have the problem of not communicating outside their specific public.
-Second, for presentations spoken in English, identify some other key languages for the conference and provide at least English-to-English captions (see rule 6 about accessibility for disabilities), but also translated captions or live interpretation into these key languages.
-Allow for multilingual Q \& A sessions with an interpreter.
-Third, include workshops, presentations, and events in languages other than English in the schedule, and provide captions or live interpretation to English.
+For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities) follow the presentations.
+Whenever possible, identify other key languages for the conference and provide translated captions or live interpretation into these key languages, including multilingual Q \& A sessions.
+The following step would be having sessions and events in languages other than English, both regarding the technical content of the conference and the social and networking aspects.
+Networking sessions in languages other than English can include captions or interpretation or not, which has the benefit of not requiring major technical adjustments.
+Keynotes and regular talks should always be accompanied by captions or live interpretation to English.
 
-When doing this, remember your native English speaking audience to be respectful of accents and English mistakes from non-native English speakers and don't let them be discouraged by having to read captions.
-Promote the attendance to presentations in languages other than English as key components of your schedule and advertise the conference in languages other than English .
+Promote the attendance to presentations in languages other than English as key components of your schedule and announce when captions are available. 
+Remind your native English speaking audience to be respectful of accents and mistakes in English from non-native English speakers.
 
@@
 %Rocío: Main idea: Include people in your communication 
-Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
-As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
-Include all the work to make the conference a safe and inclusive environment in your communication strategy, 
-to emphasize that everybody is welcome in this event.
-This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see rule 9 about financial resources).
+The communication strategy of your conference should aim to increase the number and diversity of participants, and be inclusive by design.
+Actively reach out and promote the conference to people who have been systematically excluded. 
+Publish the diversity statement, the Code of Conduct, the accessibility guidelines, and  options for financial support (see rule 9 about financial resources), to emphasize that everybody is welcome in the event.
 Try to come up with creative ways to show that everyone is seen, respected, and welcome.
-\la{For example, useR!
-2021 created "Margot, the marmot" a mascot for the conference.
-Margot is a cartoon with different designs reflecting folks that the conference wanted to reach.
-In particular, Margot wears different scarfs.
-The scarfs include colors of flags from different communities in R such as MiR´s, Rainbow R, ..., etc.
-(Figure 2)."}.
-Also, try to advertise the conference in multiple languages (see rule 7 about language) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
-part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
-% rj: allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
-With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too.
+For example, useR!
+2021 created a mascot for the conference wearing different scarfs reflecting folks from different groups that the conference wanted to reach (Fig.
+2).
+The diverse organizing team should be able to help deciding which languages to emphasize and which social media platforms to utilize in the promotion of the conference (e.g.
+Twitter, Facebook, LinkedIn, conference website, mailing lists).
 
+%lA: in case it is not mentioned later: If your budget does not allow to accommodate everyone, at a bare minimum, you should inform potential attendant which accommodations you do provide and which you will not.
+This will help everyone to know up-front what to expect and save their time if your conference does not accomodate their needs adequately.%ast: responsible communication!
 %Rocío: Main idea: Use inclusive language
 Inclusive language--language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups--should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
-Make the effort to teach yourself the vocabulary %rita A reference to this vocabulary could be add on the conference main page?
-or in the code of conduct?
-%ast- good idea, we can point to resources here like the reference i added before: https://www2.le.ac.uk/projects/oer/oers/ssds/oers/study-skills/studyskills.pdf the main aspect of the "teach yourself" part is that you don't ask people to teach you but go for it, but giving some links is nice.
-and the best ways to communicate to account for every culture and situation; do not expect minoritized people to teach you--it's not their role--and accept feedback without being offended.
+Make the effort to teach yourself the vocabulary about disabilities, racialized groups, and gender and sexual orientations. %
+Do not expect minoritized people to teach you--it's not their role--, and accept feedback without being offended.
+Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms. 
+%
+% links, examples, this changes with the context
+%practice: create a standards in inclusive language for your community, to build collective knowledge than can be passed to
 
@@
 % Rocío: main idea: Budgets are limited and need to choose priorities.
-Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
-While some expenses might be more or less fixed or known, (e.g., venue, internet platforms, catering), allocation of resources has to be intentional to support the goals of inclusion of the conference.
-It is important to estimate the costs for these practices and define your priorities in advance (e.g. paying the organizing team, Code of Conduct training, captioning).
-Additional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
+Allocation of resources to support the goals of inclusion of the conference has to be intentional. 
+Estimate the costs for these practices and define your priorities in advance (e.g. paying the organizing team, Code of Conduct training, captioning).
+Additional support for attendees could also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
 Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
@@
 
+% Conference budgets are limited and rely mostly on sponsorship and attendance fees.
+
 % Rocío: main idea: help take the burden out of participants for more inclusive participation
-Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
+When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
 (see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
-In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
-Resources permitting, you can aim to have a conference with no registration costs.
-This is especially true for online events, but bear in mind that free events have a lower attendance rate than non-free events \cite{eventbrite_ultimate_2017}.
-%[here anything that might be general support and may allow to help all attendants, fee waivers, ]
-%Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
-Scholarships or grants to attend the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses in in-person settings.
-Design the diversity-related fellowship to be explicit about the groups you want to support, to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance, while privileged people apply even when they do not need the support).
-Conferences usually ask for cover letters or applications to assign these grants, and this can be a time-consuming, emotionally-demanding task.
-Simplify as much as you can the process of asking for financial support %rita Could you give a concrete example?
-ast - yes!
-ö
-People who do not have the time (e.g., because they have care-taking roles) might decide not to apply if the process is overtly complicated.
-\la{Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups.
-Whenever possible, facilitate ways around money transfer (e.g., book flights, hotel reservations, waive conference fees).}
-%Dorothea: I understand what we want to say, but there is a slight contradiction between "reduce the prestige so that less people apply and we can help everyone and reduce the barriers so that more people apply"
-
-%For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper.
+%In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
+Resources permitting, allow the possibility of a `pay what you can' option.
+You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than non-free events \cite{eventbrite_ultimate_2017}.
+%Other costs can be prohibitive-> scholarships
+Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
+Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance).
+Conferences usually ask for cover letters or applications to assign these scholarships, and this can be a time-consuming, emotionally-demanding task. 
+Simplify as much as you can the process of asking for financial support; 
+people who do not have the time (e.g., because they have family responsibilities) might decide not to apply if the process is overtly complicated. 
+Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups.
+Whenever possible, facilitate ways around money transfer (e.g., book flights, hotel reservations, waive conference fees).
 
@@
 
-A conference can make a difference in the path towards inclusion.
-The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encountered to the next organizers.
+A conference can make a difference in the path towards inclusion/Conferences are powerful community-building spaces, that can foster change in future events and the community itself.
+The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encountered to future organizers.
 
@@
 % Rocío: main idea: no first attempt is guaranteed to succeed
-Make the conference part of a long-term process for inclusion. Do not expect the conference to encompass all the dimensions of diversity at once.
-Conceiving diversity and inclusion as a step-by-step process through successive conference versions can increase success and reduce unintended discrimination in the long run.
-Diversity and inclusion will be the norm in the long-term if we see conferences as part of the required structural change.
+Assess whether equity and inclusion goals were met during the conference. 
+Part of it can be done internally by evaluating the diversity indicators you defined at the beginning of the planning process. % examples
 
 % Ask for feedback to know the extent of your success
-After the conference is over, assess whether equity and inclusion goals were met during the meeting.
-Ask for feedback about the conference by posting a survey during and after the conference to let attendees evaluate whether the inclusion goals of the conference were met.
-When designing feedback surveys, ask for different information that would let you know about the demographic composition of the attendees to the conference. 
-Remember that some of this information varies greatly from region to region and adapt your options to the local or regional context.
-When choosing categories, avoid letting people in minoritized categories fall in the "other" option.
-You can also try to make open answer options to let people self-identify and express their opinions freely.
-Let all responses be optional and anonymous, but encourage your attendance to answer and commit to data privacy.
-Identify the aspects that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
+Ask for feedback about the conference experience of the participants by posting a survey during and after the conference. 
+Make open answer options to let people self-identify and express their opinions freely. Let all responses be optional and anonymous and commit to confidentiality and data privacy. 
+The survey results will allow you to identify the aspects that worked towards your inclusion goals, and understand what went wrong and how people were excluded.
+Do not expect everything to be perfectly inclusive, or to encompass all the dimensions of diversity at once. %You're part of a process and contributing to structural change.
 
-% Pass
-Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next editions of the conference.
-If you are part of a stable meetings committee, encourage organizers to follow these rules and help set up new standards for inclusion.
-Share also your new and extended network and nominate people for future leadership.
-A good way to do this is documenting your processes, lessons, and guides in blog posts, reports and presentations.
-They may be useful not only for your conference series but for other communities that are having the same questions that you are.
-
+% Pass the information
+Document your processes, lessons, and guides in blog posts, reports, and presentations. 
+Facilitate contacts with the people you worked with along the process. 
+Make sure to share this information with future organizers,
+they will have a starting point and their work to improve inclusion in the following editions of the conference will be easier.
+If you are part of a stable meetings committee, encourage organizers to follow these rules and advocate for setting up new standards for inclusion. 
+%hablar del pasado - shoulder of people before, federico
 %sale? Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
 %\la{Due to these systemic problems, even the minimum changes to how conferences have been historically organized will find resistance}, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
-%responbile communication (not here, see laura's comment) Be explicit about the biases you are tackling and acknowledge which audiences your conference is not doing enough to include.
 
@@
 % Rocío: I don't like the ending.
+(Guidelines/documentation) may be useful not only for your conference series but also for other communities with the same questions that you have. % Give examples of how others have helped us.
 

### 2021-11-26 plos_latex_template.tex
Few abstract/pillars changes ~ @jananiravi

@@
 
-The authors of this article participated in the organization of the annual conference of the R Project for Statistical Computing, held in July 2021.
+The authors of this article recently organized useR!
+2021, the annual conference of the R Project for Statistical Computing.
 useR!
@@
 The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector.
-For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone.
-In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference.\fede{general comment on this: we should clearly state that this whole-virtual setting was a clear consequence of the pandemics, AND it was planned like this from the beginning, and not changing in-between to this format} These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.
-The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events.
+For 2021, we aimed to build a high-quality virtual and explicitly global conference providing a kind, inclusive, accessible, and welcoming environment for everyone.
+In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference.
+These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules also apply to regional events.
+These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events as we exit the pandemic.
 
@@
 
-To make a more inclusive conference, it is necessary to identify which groups of people are being excluded, the barriers they may face to be part of the conference, and strategies to overcome them. 
-The first rule is about embracing all dimensions of diversity and prioritizing the
- participation of groups that are often marginalized.
-However, increasing diversity is not enough. Rule 2 is about creating a safe and welcoming environment to make everyone feel part of the conference and the community. 
-Inclusion should start from the organizing team. Rule 3 refers to the importance of starting with an inclusive and diverse organizing team, and provides advice on work dynamics.
+To organize a more inclusive conference, it is necessary to identify upfront which groups of people are being excluded, the barriers that keep them from active partaking, and targeted strategies to overcome them.
+\textbf{Rule 1} is about embracing all dimensions of diversity and prioritizing the participation of groups that are often marginalized.
+However, increasing diversity alone is not enough. \textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the conference attendees and the larger community. 
+Inclusion should start from the organizing team. \textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team, and provides tips on work dynamics.
 
@@
 Nonetheless, we live in a world with implicit hierarchies along these axes.
-Some statuses (e.g., cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
+Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
 \as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia.
@@
 % Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
-%in 2021 an inclusive conference has to be an online or hybrid conference.
-this means a radical change in the functioning of all spaces, to include people who are at home for any reason.
-acá contexto de covid - within ...
-y ser más firmes con lo de hybrid.
-joo - wenger.
-%primer parrafo como está cite joo, cosas prácticas de online - están en las citas, poner acá.
-2.
-networking hybrid qué es, tipos, active and distributed serían nuestras opciones - omitir las otras por no ser inclusivas.
-todo lo inclusivo de online con todo lo inclusivo de presencial.
-un par de ejemplos de lo presencial.
-o adaptaciones que requiere la parte presencial cuando se vuelve híbrido (chat, q and a, streaming) opportunities of networking - que ser chair o keynote o presenter no cambie si estás online o presencial, no requiera que estés presente.
-%practical: choose the platforms and be aware that reach varies between countries
+%1.
+in 2021 an inclusive conference has to be an online or hybrid conference.
+this means a radical change in the functioning of all spaces, to include people who are online for any reason.
+we have elements to advocate for hybrid and lots of refs about online.
+cite joo and online refs.
+%practical: choose the platforms and be aware that reach varies between countries. check accessibility. understand bandwidth limitations. online does not make it automatically inclusive.
+%2.
+hybrid.
+we advocate for active hybrid and distributed hybrid conferences because it's the most inclusive formats (not passive hybrid or semi-passive hybrid).
+we want all the inclusiveness of the in-person component with the _practical_ adaptations for hybrid (chat platform, q and a, streaming and networking) being online or in-person should not affect the possibility to be a chair a keynote or a presenter.
+no need to be present.
+a potential problem: opportunities of networking.
 As stated in [the nature correspondence]: The COVID pandemic forced us to embrace online conferences (see C.

### 2021-11-29 plos_latex_template.tex
Minor changes to Intro & Pillars ~ @jananiravi

@@
 % Rocío: Main message of the paragraph: Conferences are great, but for some
-Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+Conferences, from the Latin \textit{conferre} `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
 The larger the conference, the larger the opportunities to network and learn from the cohort.
 However, opportunities for participating in conferences are not equal for all. 
-Just like in academia and tech in general, conferences can become exclusive spaces for privileged groups, i.e.
-with unearned advantages given by society to some but not all (e.g.
-white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+Although unintentionally, many academic and tech conferences have become exclusive spaces for privileged groups, \textit{i.e.}, with unearned advantages given by society to some but not all (\textit{e.g.}, white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
 % However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
 
-This article proposes 10 rules to host inclusive conferences.
-They are organized in three sections: the pillars (embracing diversity, creating a welcoming environment, and having a diverse organizing team), design rules (counteracting bias in the conference program, having a strong online component, achieving accessibility for people with disabilities, embracing language diversity, having inclusive communication strategies, and allocating financial resources to support inclusive practices), and a process rule: make the conference part of a long-term process for inclusion.
+This article proposes 10 simple rules to host inclusive conferences.
+They are organized in three sections: \textbf{the pillars} (embracing diversity, creating a welcoming environment, and having a diverse organizing team), \textbf{design rules} (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices), and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion.
 These practices stem from the experience of organizing useR!
@@
 
-The list of authors is composed of the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
-Therefore, the rules are a result of our lessons learned from diverse perspectives and expertise (i.e. our different backgrounds, profiles, and roles in the conference). 
-Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes.
-They are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
-The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
+The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other critical areas of the conference. 
+Therefore, the rules result from our lessons learned from diverse perspectives and expertise (\textit{i.e.}, our different backgrounds, profiles, and roles in the conference). 
+Even though these rules were designed based on our experience with a global event, they can be easily transferred or scaled to events with regional or local scopes.
+These rules will greatly benefit people who are part of meeting committees that oversee the site/location selection process, or that coordinates with the local organizers of conferences.
+The rules will also help local/virtual organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
+We have designed them for wide adoption to enable and ensure that the scientific and technical conferences post-pandemic are as accessible and inclusive as they can be.
 
@@
 
-\section*{The pillars}
+\section*{I. The pillars}
 
@@
 No form of oppression operates in isolation from the others, which means some groups are particularly vulnerable and systematically left behind.}
-While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
-Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against.
-These are the groups we need to make more effort to include.
-Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
+While no isolated initiative can drastically alter conference outcomes, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
+Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify the most marginalized and discriminated subgroups.
+These are the groups that warrant deliberate efforts that ensure fair inclusion.
+Investing a substantial effort towards the most excluded groups does not mean neglecting the other communities, but will instead ensure equity, guide the vision of diversity for your conference, and help design strategies to achieve it.
 % vision
@@
 % Rocío: more effort towards including the more excluded
-This may translate into different actions depending on your field, region, or community, e.g.
-a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
+This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people — especially Black people — among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
 
@@
 
-\section*{Design Rules}
+\section*{II. Design Rules}
 
@@
 
-\section*{The Process}
+\section*{III. The Process}
 

### 2021-11-30 plos_latex_template.tex
RJ: Accepted many of Janani's suggestions for the Intro

@@
 The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector.
-For 2021, we aimed to build a high-quality virtual and explicitly global conference providing a kind, inclusive, accessible, and welcoming environment for everyone.
+For 2021, we aimed to build a high-quality virtual and explicitly global conference creating a kind, inclusive, accessible, and welcoming environment for everyone.
 In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference.
 These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules also apply to regional events.
-These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events as we exit the pandemic.
+These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events even after the covid pandemic.
 
@@
 However, opportunities for participating in conferences are not equal for all. 
-Although unintentionally, many academic and tech conferences have become exclusive spaces for privileged groups, \textit{i.e.}, with unearned advantages given by society to some but not all (\textit{e.g.}, white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+Many academic and tech conferences have become exclusive spaces for privileged groups, \textit{i.e.}, with unearned advantages given by society to some but not all (\textit{e.g.}, white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
 % However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
@@
 This article proposes 10 simple rules to host inclusive conferences.
-They are organized in three sections: \textbf{the pillars} (embracing diversity, creating a welcoming environment, and having a diverse organizing team), \textbf{design rules} (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices), and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion.
+They are organized in three sections: \textbf{the pillars}, fundamental rules with key elements to conceive the work on diversity and inclusion in any conference (embracing diversity, creating a welcoming environment, and having a diverse organizing team); \textbf{design rules}, concerning diverse elements in the design of the conference and operational tasks (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices); and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion.
 These practices stem from the experience of organizing useR!
@@
 
-The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other critical areas of the conference. 
+The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
 Therefore, the rules result from our lessons learned from diverse perspectives and expertise (\textit{i.e.}, our different backgrounds, profiles, and roles in the conference). 
@@
 The rules will also help local/virtual organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
-We have designed them for wide adoption to enable and ensure that the scientific and technical conferences post-pandemic are as accessible and inclusive as they can be.
+We have designed them for wide adoption to enable and ensure that the scientific and technical conferences during and post-pandemic are as accessible and inclusive as they can be.
 
@@
 %sara: without success i'm trying to use pdfcomment package and pdftooltip to make the alt-text. any help?
-
+%ast: R/ Unfortunately, support for adding accessibility metadata to LaTeX documents is limited.
+You will probably need to use Adobe Acrobat to add missing accessibility metadata to your PDF file.https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#latex
 \pdftooltip{
@@
 % Rocío: more effort towards including the more excluded
-This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people — especially Black people — among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
+This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
 
@@
 
+\begin{figure}[!h]
+\centering
+\pdftooltip{
+\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
+}
+\caption{Margot the marmot was useR! 2021 mascot}
+\label{fig:marmots}
+\end{figure}
+
 \subsection*{Rule 9: Allocate adequate financial resources to support your conference goals}

### 2021-11-30 plos_latex_template.tex
RJ: edits to pillars and rule 1

@@
 \textbf{Rule 1} is about embracing all dimensions of diversity and prioritizing the participation of groups that are often marginalized.
-However, increasing diversity alone is not enough. \textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the conference attendees and the larger community. 
+However, increasing diversity alone is not enough. \textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the conference attendees. 
 Inclusion should start from the organizing team. \textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team, and provides tips on work dynamics.
@@
 Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
-\as{People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia.
-No form of oppression operates in isolation from the others, which means some groups are particularly vulnerable and systematically left behind.}
-While no isolated initiative can drastically alter conference outcomes, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.
-Recognizing this systematic exclusion, particularly in our field and in our scientific or professional community, will help identify the most marginalized and discriminated subgroups.
-These are the groups that warrant deliberate efforts that ensure fair inclusion.
-Investing a substantial effort towards the most excluded groups does not mean neglecting the other communities, but will instead ensure equity, guide the vision of diversity for your conference, and help design strategies to achieve it.
+People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; some people have to deal with several of these on a daily basis.
+% While no isolated initiative can drastically alter conference outcomes, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles.
+Recognizing this systematic exclusion, particularly in our fields and in our scientific or professional communities, will help identify the most marginalized and discriminated subgroups \cite{timperleyHeMoanaPukepuke2020}.
+% These are the groups that warrant deliberate efforts that ensure fair inclusion.
+As a starting point, identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (e.g. explicit discrimination, participation fees, etc.).
+This analysis should guide the vision of diversity and inclusion for your conference and help design strategies to achieve it. 
+Express this vision for your conference in a diversity statement. 
+Next, use this vision and the set strategies to define measurable indicators and goals of diversity and inclusion (\textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, or participation from key geographic regions).
+These numbers should guide you and help achieve your vision.
+
+% Investing a substantial effort towards the most excluded groups does not mean neglecting the other communities, but will instead ensure equity, guide the vision of diversity for your conference, and help design strategies to achieve it.
 % vision
@@
 %Express the diversity and inclusion vision for your conference by writing a diversity statement. 
+% As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors.
+Next, set measurable indicators of success when organizing a diverse conference.
+The numbers resulting from these indicators are not the primary goal, and this is not about performative allyship.
+Instead, these numbers should guide and help hold the organizing team accountable towards these diverse and inclusive goals.
 
 % Rocío: more effort towards including the more excluded
-This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
+% This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
 

### 2021-11-30 plos_latex_template.tex
RJ: Edits to rule 3

@@
 A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team in charge of the conference decision-making.
-Organizers should be a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
+Organizers should be a team with people from different regions, genders, ethnicities, socioeconomic statuses, and other aspects of diversity.
 Having diverse people in decision-making positions will positively affect all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
@@
 People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity.
-This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups \cite{costanzachockDesign2020}.
-People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
-However, they must have the freedom to choose in which areas of the conference they want to work and not be restricted to diversity and inclusion aspects. 
+The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions from people that have not been through the same experiences \cite{costanzachockDesign2020}.
+% People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
+However, do not require nor expect minoritized people to only work—or want to work—on diversity and inclusion aspects; 
+everybody must have the freedom to choose in which areas of the conference they want to work. 
 % Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
-When working with people from marginalized groups do not assume that self-nomination and voting will work as mechanisms to choose leadership positions.
+Furthermore, do not assume that self-nomination and voting will work as mechanisms to choose leadership positions.
 Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
 Build an environment in which every person can express their opinion and pay special attention to people from systematically excluded groups.
-Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.
-% Rocío: Add other types of recognition besides money (website, public recognition on social media, during the conference, etc.)
+Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (e.g.
+public recognition on social media, on the conference website, or during the conference).
 In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.

### 2021-12-01 plos_latex_template.tex
RJ: Edits to rule 4

@@
 % Rocío: Idea of this paragraph: non-transferability and do not restrict people
-People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity.
+People with disabilities often say: `Nothing about us without us' [ref]; the same holds for other dimensions of diversity.
 The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions from people that have not been through the same experiences \cite{costanzachockDesign2020}.
@@
 
-When choosing or inviting people for visible and valued roles in the conference––such as keynote speakers, program committee members, session chairs––, it is likely that there will not be much diversity in the first set of names.
+When choosing or inviting people for visible and valued roles in the conference––such as keynote speakers, program/scientific committee members, session chairs––, it is likely that there will not be much diversity in the first set of names.
 Our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
@@
 avoid `manels' and all-white panels, for instance).
-Make sure the composition of all these groups is balanced across topics, areas, and representation across regions, gender, and race.
+Make sure the composition of all these groups is balanced across topics, areas, and representation across regions, gender, race, career stage, among other aspects.
 %rita: For academic conferences consider also to select for a talk master and PhD students and not to give voice only to senior scientist or principal investigator.
 lo pongo acá porque sí faltan cosas prácticas.
+% RJ: I added career stage to the sentence above.
+That should be enough.
 
@@
 
-The final list of accepted talks can also be biased. 
-On the one hand, because people do not feel confident to submit their work. 
-To counteract bias in submission, promote your calls beyond the usual communities and keep in touch with local groups and communities of practice to encourage submissions by their members.
-Some of these communities have pre-submission mechanisms to help their members prepare and improve their submissions.%sin el ejemplo no se entiende
+The final list of accepted talks can also be biased in both submission and acceptance processes. 
+People who have never participated in conferences, and especially those regarded as big and prestigious, may not feel confident to submit their work.
+To counteract this self-selection bias, promote the submission call beyond the communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
+Some of these communities may have pre-submission mechanisms to help their members prepare and improve their submissions (e.g.
+the Latin-R community has a Slack channel with the purpose of helping community members improve their abstracts for conferences, and the R-Ladies community did the same for useR!
+2021).%sin el ejemplo no se entiende
 % 1. Clear rules on reviews: double blind, double open, kind of open, etc.
@@
 %see: https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
-On the other hand, proposals may be biased because of even diverse selection committees can be subjected to unconscious or implicit bias %explain!.
+% Unconscious (implicit) bias is the automatic impact of attitudes or stereotypes on our on our view of the world, actions, and decisions without our awareness (swartz 2019)
+% in this study, we defined implicit bias as associations, attitudes, or beliefs that exist and enact their influence outside of an individual’s conscious awareness.
+On the other hand, the review process may be biased because selection committees—even diverse ones—can be subjected to unconscious or implicit bias.
 Unconscious bias happens when we interpret information, such as names of the authors, we assume their origin or ethnicity, their institution of origin, and we attach a positive or negative value to their work according to negative or positive stereotypes.
-To counteract this...
-The review process should guarantee the fair evaluation of the submitted work, and this may  will be double blind, open, or partially open. 
-In some fields, it's not possible to guarantee anonymity of the presenters, and in some and external cues such as country of origin and links to repositories. 
-Create clear evaluation criteria for the submitted proposals, and share them with the authors and reviewers.
-Remind the reviewers that unconscious bias exists and encourage them to take unconscious 
-Selection committees need to be reminded of such biases, and be aware of them when evaluating submissions and presentations.
-\cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}
+Nobody is free from implicit bias [cite Ross2019].
+Selection committees need to be reminded of such biases (the implicit association test can help identifying them [cite Greenwald - Measuring individual differences in implicit cognition: The implicit association test]), and
+decide on the most appropriate strategy to counteract them, either with double blind, double open, or single blind reviews [cite discover cookbook].
+Double blind reviews can help counteract implicit bias only if it is possible to completely anonymize the identifying information from submission. 
+Double open, though more transparent, may still be subjected to implicit bias, and reviewers from minoritized groups might feel additional pressure because the reviews are signed.
+If this likely case, and in fields were it is not possible to guarantee anonymity of the authors during the selection process (e.g.
+non-anonymized github repositories need to be examined), single blind reviews are the only option left.
+In any case, create clear evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g.
+based on their number of publications or talks), and share them with the authors and reviewers.
+Advertising your practices to mitigate bias in proposal selection may also encourage people from diverse groups to submit their work, thus counteracting self selection bias.
+
+% \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}
 
 % Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
-Activities such as scientific publication and development of software are often regarded as higher value than other roles like community building or teaching. 
+Implicit bias also pertain the value that we give to certain activities; e.g.
+scientific publication and development of software are often regarded as higher value than other roles like community building or teaching.
 These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
-Research about these topics is equally undervalued compared to other areas of knowledge production.
-Counteracting bias also means changing the standard program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, // and giving visibility to the whole range of activities and practitioners that contribute to the overall topics of the conference.
+Research about these topics is undervalued relative to other areas of knowledge production.
+Counteracting bias also means innovating the program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, and giving visibility to the whole range of activities and practitioners that contribute to the overall topics of the conference.
 
 %Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
-You can also reframe the awards ceremony to acknowledge work beyond the traditional definition of merit/quality and include those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
+% You can also reframe the awards ceremony to acknowledge work beyond the traditional definition of merit/quality and include those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
 
@@
 (see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status).
-%In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
+
 Resources permitting, allow the possibility of a `pay what you can' option.

### 2021-12-01 plos_latex_template.tex
RJ: Edited rule 5 and cleaned comments on rule 4

@@
 Make sure the composition of all these groups is balanced across topics, areas, and representation across regions, gender, race, career stage, among other aspects.
-%rita: For academic conferences consider also to select for a talk master and PhD students and not to give voice only to senior scientist or principal investigator.
-lo pongo acá porque sí faltan cosas prácticas.
-% RJ: I added career stage to the sentence above.
-That should be enough.
 
@@
 the Latin-R community has a Slack channel with the purpose of helping community members improve their abstracts for conferences, and the R-Ladies community did the same for useR!
-2021).%sin el ejemplo no se entiende
-% 1. Clear rules on reviews: double blind, double open, kind of open, etc.
-% Be transparent on the evaluation criteria
-% 2. Diverse program committee (above)
-% 3. Unconscious bias (test)
-% 4. Curate the final list against 
-%see: https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
-% Unconscious (implicit) bias is the automatic impact of attitudes or stereotypes on our on our view of the world, actions, and decisions without our awareness (swartz 2019)
-% in this study, we defined implicit bias as associations, attitudes, or beliefs that exist and enact their influence outside of an individual’s conscious awareness.
+2021).
 On the other hand, the review process may be biased because selection committees—even diverse ones—can be subjected to unconscious or implicit bias.
@@
 
-%Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
-% You can also reframe the awards ceremony to acknowledge work beyond the traditional definition of merit/quality and include those who contributed to community building, presented teaching materials, and assumed other care-taking roles.
-
 \subsection*{Rule 5: Have a strong online component} 
@@
 % Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
-%1.
-in 2021 an inclusive conference has to be an online or hybrid conference.
-this means a radical change in the functioning of all spaces, to include people who are online for any reason.
-we have elements to advocate for hybrid and lots of refs about online.
-cite joo and online refs.
-%practical: choose the platforms and be aware that reach varies between countries. check accessibility. understand bandwidth limitations. online does not make it automatically inclusive.
-%2.
-hybrid.
-we advocate for active hybrid and distributed hybrid conferences because it's the most inclusive formats (not passive hybrid or semi-passive hybrid).
-we want all the inclusiveness of the in-person component with the _practical_ adaptations for hybrid (chat platform, q and a, streaming and networking) being online or in-person should not affect the possibility to be a chair a keynote or a presenter.
-no need to be present.
-a potential problem: opportunities of networking.
-As stated in [the nature correspondence]: The COVID pandemic forced us to embrace online conferences (see C.
-Woolston Nature 582, 135–136; 2020), which remove some of the barriers that disproportionately affect marginalized groups.
-These include the cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications (H.
-J.
-Niner \& S.
-N.
-Wassermann Front.
-Mar.
-Sci.
-8, 638025; 2021; Gewin, V.
-Nature 569, 297–299; 2019).
-Inclusive conferences must have an online component [cite all the other references].
-In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
-Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}.
-Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
-This means that online conferences have a greater reach, not only in terms of attendees but in terms of speakers  \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
-The online format may also make it easier to be inclusive of geographic regions by encompassing several time zones or deciding to rotate the favored time zone every year, without depending on a conference central location.
+Virtual conferences are more inclusive, removing barriers like costs of registration, transport, accommodation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}.
+After experiencing virtual conferences mostly due to the COVID pandemic, there is a widespread call to keep an online component in conferences, either maintaining conferences completely virtual or using hybrid formats combining in-person and online components \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
+Though more inclusive, an online component does not guarantee inclusion and accessibility for everyone.
+Key decisions must be made early on in the planning stages \cite{levitisCenteringInclusivityDesign2021}.
+For instance, when choosing the conference platforms (for submission, registration, chatting system, live streaming, etc.) account for accessibility to screen readers, geopolitical restrictions, and bandwidth requirements.
+You should also decide how to be inclusive of geographic regions with the schedule of the conference, asking for pre-recorded videos, recording sessions, and sharing them during the conference.
+Some possibilities for scheduling are: having different days dedicated to different time zones; having each day with the same conference hours, and scheduling time slots that are friendly--and unfriendly--to different time zones; or having time-zone hubs in each day and repeating the events in each (in some cases live and others a recorded version).
 
-%Saranjeet: About time zones, although we can rotate and have different days dedicated to different time zones, we could also try spreading the conference time to be longer.
-So that each day the conference runs live for a couple of hours or so and the recorded material is made available/uploaded on YouTube (say) soon (every day).
-This might help because, not everyone might have the luxury of attending 4-6 hours of the conference on a weekday, even though it is running in their local timezone.
-However, people might be able to attend a couple of hours even on weekdays and will not get the feeling that they have missed a huge chunk of the conference.
-%ast add something about online giving flexibility in formats and how that depends on to the context of the conference
 % Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
 Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
-%However, in-person settings might not be comfortable or appealing to everyone for socializing and some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
 However, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
 \cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}.
-% JPNG: "On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
-[29, 30]).
-Written media like chats create opportunities for long term interactions and networking by allowing more people to join the conversation even when it has started some time ago."
-Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community.
-It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference.
-%la: see https://www.trainingforchange.org/wp-content/uploads/2017/11/Mainstream-Margin-in-Groups.pdf
-Some ideas in this line are: offering the option of written chat in addition to voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}
-% we need to check if something here (newbies!) is not missing in the above paragraphs
-% As we mentioned previously, conference attendees should have ample opportunities to network with each other.
-However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants.
-If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}).
-Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
+Invest time in creating opportunities to meet and bond virtually that can appeal to people with different backgrounds, disabilities, preferences, and remembering that no single networking activity will ever serve the whole community.
+Try new and varied activities centered around different groups of people outside the mainstream of your conference; if you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive, successful, and inclusive.
+Some examples are the newbies sessions for first-timers, social mixers lead by specific subgroups or communities like LGBT+, or leisure activities that reunite subgroups with the same interests: virtual art exhibitions, yoga sessions, movie watch parties, trivia quizzes, etc.
+Offer some activities that require voice or video interactions, and others that involve only written chat.
 
-% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
-Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well and maximize the experience of everyone attending your conference.
-This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely.
-Conference organizers who have been holding hybrid events since 2011 [Wenger-Tayner, 2021] consider that inclusive hybrid events are the future.
-These authors emphasize the need that all participants of hybrid events must go through the same event quality to the point that in-person and online participants pay the same registration amount to participate in high-quality hybrid conferences.
-These authors recommend having a buddy system, where every person attending online has an in-person attendant that makes sure the online person can follow the event and participate in the event as if they were in the room [Wenger-Trayner, 2020].
-Wenger-Trayner (2020, 2021) also recommends having a chat function for back-channel conversations between all attendants, a virtual space for all online attendants to talk to each other, and allocating time to small group conversations, often mixing in-person and online groups.
-% Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
-The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{jooKeepOnlineOption2021,ninerBetterWhomLeveling2021}.
-\rita{Sessions of an hybrid conference could have both chairs online, as attendee or speaker.
-This mean that is important when organizing the conference to include a third person who is in the venue to help the in-person participants, take the questions that are both online and in-person and communicate with the remote session chairs.}
-
-\rita{The chat system of the conference can have a list of all the participants, with their registration name and if they are on-line or in-person, to facilitate the connections.}  The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space.
-\rita{The venue selection must respond to some criteria such as give online support both bot in-person and online participants, to have stable and secure connection for the in-person participant to join the online event.
-Moreover, offer the option to create virtual room to connect in-person and online participant for networking events, such as scientific interest group, or LGBTQIA+-friendly spaces.
-Exhibition stands in the venue need to have facilities to connect with online participants (virtual stand).
-Furthermore, the poster session need to be carefully thought to give the best experience to both in-person and online participants.
-This should have both components to allow online participants to see the poster and have the possibility to ask questions.
-However, it also need allow who want to present their poster virtually to be seen by in-person participants and respond to questions about their works.
-Lastly to better plan your hybrid conference ask in advance who wish to travel to the venue and who wish to participate online.
-This will save you time and money for looking for a suitable venue.}
-
-% [Wenger, 1999] https://doi.org/10.1017/CBO9780511803932
-% [Wenger-Trayner, 2020] https://wenger-trayner.com/all/in-person-and-online-events/
-% [Wenger-Trayner, 2021] https://wenger-trayner.com/reflections/in-person-online-hybrid-the-future/
+If you are organizing a hybrid event instead, you should decide on the relative importance that the online and in-person components will take.
+The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the format they choose to participate in.
+Under this format, the audience, chairs, and presenters--of talks, posters and workshops--could be either in-person or remote, and they should all be able to interact. 
+For that purpose, all questions and comments should pass through an online system. 
+The physical venue would have to include multiple ways to connect with online folks such as tablets on the meeting venue, cameras and microphones that allow online attendants to follow who is speaking in the in-person space.
+This requires the venue to ensure stable and secure internet connection for the in-person participants to join the online folks. 
+Having a hybrid format also means accounting for the fact that the venue does not need to host all the attendees.
+You should estimate in advance how many in-person attendees to expect, and ask them early on during registration or pre-registration.
+This could save you time and money when looking for a suitable venue.
+When planning for networking activities in hybrid events, use the same principles of the paragraph above and offer varied activities centered around different groups.
+Some of these activities could be only in-person, but to promote connections between all participants, there should be a fair proportion of activities connecting in-person and online attendees.
 

### 2021-12-01 plos_latex_template.tex
RJ: Edits to rule 8

@@
 As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference.
-If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute.
+If you do not consider accessibility from the conference inception, it is highly likely that you will be better off without trying to patch accessibility at the last minute.
 Key decisions in this respect are hard to correct.
@@
 Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly and keyboard accessible, and have alternative text.
-Any videos should have good-quality captions and a transcript.
-The organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
+Every video should have good-quality (not automated) captions and a transcript; live presentations should have good-quality captions too.
+Provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
 Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.
@@
 
-%Live captioning or interpretation!
-and surtitles (Rita) LA: provide live, high-quality (automated captions are often referred to as "craptions"), captioning or simultaneous translation in English for persons who only speak English.
+%LA: provide live, high-quality (automated captions are often referred to as "craptions"), captioning or simultaneous translation in English for persons who only speak English.
 After all, captions and simultaneous translations are common place in the audiovisual industry and your conference audience will likely be already familiar with this if they ever watched a movie from a non-English speaking country.
@@
 For example, useR!
-2021 created a mascot for the conference wearing different scarfs reflecting folks from different groups that the conference wanted to reach (Fig.
+2021 created a mascot for the conference wearing different scarves reflecting folks from different groups that the conference wanted to reach (Fig.
 2).
-The diverse organizing team should be able to help deciding which languages to emphasize and which social media platforms to utilize in the promotion of the conference (e.g.
+The diverse members of your team should be able to decide which languages to emphasize and which social media platforms to utilize in the promotion of the conference (e.g.
 Twitter, Facebook, LinkedIn, conference website, mailing lists).
+Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues.
+For instance, inform if the conference platform is not completely screen-reader friendly, but you are offering personalized help to navigate it, or if captions will be available for some talks but not all.
+%ast: responsible communication!
 
-%lA: in case it is not mentioned later: If your budget does not allow to accommodate everyone, at a bare minimum, you should inform potential attendant which accommodations you do provide and which you will not.
-This will help everyone to know up-front what to expect and save their time if your conference does not accomodate their needs adequately.%ast: responsible communication!
 %Rocío: Main idea: Use inclusive language
 Inclusive language--language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups--should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
-Make the effort to teach yourself the vocabulary about disabilities, racialized groups, and gender and sexual orientations. %
+Become familiar with the terminology used for disabilities, racialized groups, and gender and sexual orientations, and with the terms that should be avoided. %
 Do not expect minoritized people to teach you--it's not their role--, and accept feedback without being offended.
@@
 %practice: create a standards in inclusive language for your community, to build collective knowledge than can be passed to
-
-% Rocío: could add some ideas from here 
-% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
 

### 2021-12-01 plos_latex_template.tex
RJ: Minor edits until rule 10

@@
 Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms. 
-%
 % links, examples, this changes with the context
@@
 
-% Rocío: main idea: Budgets are limited and need to choose priorities.
+% Rocío: main idea: Need to budget for inclusive practices.
 Allocation of resources to support the goals of inclusion of the conference has to be intentional. 
@@
 
-% Conference budgets are limited and rely mostly on sponsorship and attendance fees.
-
 % Rocío: main idea: help take the burden out of participants for more inclusive participation
@@
 (see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
-
 Resources permitting, allow the possibility of a `pay what you can' option.
@@
 
-A conference can make a difference in the path towards inclusion/Conferences are powerful community-building spaces, that can foster change in future events and the community itself.
+A conference is a critical community-building space, and can foster change in future events and the community itself.
 The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encountered to future organizers.
@@
 
-% Rocío: main idea: no first attempt is guaranteed to succeed
+% Rocío: main idea: assess how inclusive was the conference
 Assess whether equity and inclusion goals were met during the conference. 
-Part of it can be done internally by evaluating the diversity indicators you defined at the beginning of the planning process. % examples
-
-% Ask for feedback to know the extent of your success
-Ask for feedback about the conference experience of the participants by posting a survey during and after the conference. 
-Make open answer options to let people self-identify and express their opinions freely. Let all responses be optional and anonymous and commit to confidentiality and data privacy. 
-The survey results will allow you to identify the aspects that worked towards your inclusion goals, and understand what went wrong and how people were excluded.
-Do not expect everything to be perfectly inclusive, or to encompass all the dimensions of diversity at once. %You're part of a process and contributing to structural change.
+You can rely on the diversity indicators defined at the beginning of the planning process (rule 1). 
+It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and others (e.g.
+countries of origin, or preferred language).
+Other indicators may need a special survey asking for additional information and feedback about the conference. 
+In the survey, providing options for open answers will allow people to express their opinions freely, and may give you more insights into how/why your inclusion practices worked or did not work. 
+% Do not expect everything to be perfectly inclusive, or to encompass all the dimensions of diversity at once. %You're part of a process and contributing to structural change.
 
 % Pass the information
-Document your processes, lessons, and guides in blog posts, reports, and presentations. 
-Facilitate contacts with the people you worked with along the process. 
-Make sure to share this information with future organizers,
-they will have a starting point and their work to improve inclusion in the following editions of the conference will be easier.
-If you are part of a stable meetings committee, encourage organizers to follow these rules and advocate for setting up new standards for inclusion. 
-%hablar del pasado - shoulder of people before, federico
-%sale? Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
-%\la{Due to these systemic problems, even the minimum changes to how conferences have been historically organized will find resistance}, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
+Document your processes, outcomes, and lessons learned in blog posts, reports, and presentations. 
+Make sure to share this information with future organizers, as well as the contact information of the people you worked with--if they agree to it, to give them a starting point to improve inclusion in the following editions of the conference.
+If you are part of a stable meetings committee, encourage organizers to follow these rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria.
+%something about the past - shoulder of people before, federico
 

### 2021-12-01 plos_latex_template.tex
Added :bulb and few changes to Abstract, Intro, Rules 1–2 ~ @jananiravi

@@
 \usepackage{pdfcomment}
+
+%Janani: for inserting sparingly few super relevant emojis
+\usepackage{emoji} %Rocio: https://www.overleaf.com/learn/latex/Questions/Inserting_emojis_in_LaTeX_documents_on_Overleaf#Emoji_characters_and_the_Overleaf_editor
 
@@
 The authors of this article recently organized useR!
-2021, the annual conference of the R Project for Statistical Computing.
-useR!
-conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation.
-The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector.
-For 2021, we aimed to build a high-quality virtual and explicitly global conference creating a kind, inclusive, accessible, and welcoming environment for everyone.
-In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference.
-These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules also apply to regional events.
-These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events even after the covid pandemic.
+2021, the annual global R conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector.
+To embrace the constraints posed by the pandemic, this year we hosted a 100\% virtual conference.
+We worked on building a high-quality virtual and explicitly global conference by creating a kind, inclusive, accessible, and welcoming environment for everyone.
+This article outlines our most important lessons learned abstracted as '10 simple rules to host an inclusive conference'.
+These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules seamlessly translate to regional events.
+These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events even after the COVID pandemic.
 
@@
 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
+useR!
+conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation.
 We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
@@
 Human diversity should be celebrated and respected in every way. 
-Nonetheless, we live in a world with implicit hierarchies along these axes. 
-Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built.
-People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; some people have to deal with several of these on a daily basis.
+Unfortunately, we live in a world with implicit hierarchies along these axes. 
+Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems -- including conferences -- are consciously and unconsciously built.
+People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; a few of them have to deal with several of these on a daily basis.
 % While no isolated initiative can drastically alter conference outcomes, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles.
-Recognizing this systematic exclusion, particularly in our fields and in our scientific or professional communities, will help identify the most marginalized and discriminated subgroups \cite{timperleyHeMoanaPukepuke2020}.
+Therefore, recognizing this systematic exclusion, particularly in our fields and in our scientific or professional communities, is paramount to inclusion, and will help identify the most marginalized and discriminated subgroups \cite{timperleyHeMoanaPukepuke2020}.
 % These are the groups that warrant deliberate efforts that ensure fair inclusion.
-As a starting point, identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (e.g. explicit discrimination, participation fees, etc.).
+
+\emoji{bulb} \textbf{Practical tips:} As a starting point, the meeting organizers could identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (\textit{e.g.}, explicit discrimination, participation fees, etc.).
 This analysis should guide the vision of diversity and inclusion for your conference and help design strategies to achieve it. 
@@
 % Rocío: main idea: Inclusion for everyone
-While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well.
+While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a genuinely inclusive environment means taking care of all aspects of diversity.
+Inclusivity should lie at the heart of setting the tone and meeting space for any meeting as we move forward in the 21st century!
 % Edit and make more specific (below)
-Promote respect of pronouns, schedule events exclusive for diverse groups (e.g., LGBTQIA+-friendly spaces), and if the conference has an in-person component, set spaces for religious practices, breastfeeding and child care, and include menu options that account for diverse dietary and religious requirements.
-These are just some examples of decisions that are relatively easy to implement and can have a large impact in making inclusion real (see \cite{noauthor_discover2021} for some great advice on `low-hanging fruits').
-Importantly, by being proactive in creating such a welcoming space, you can be respectful of everybody's privacy and avoid requiring anyone to disclose personal information–such as revealing a disability, the sexual orientation, gender identity, or mental health issues, just to a name a few.
+
+\emoji{bulb} \textbf{Practical tips:} Here are a few ways in which conference organizers can actively enforce inclusivity, such as i) Promote respect of pronouns, ii) schedule events exclusive for diverse groups (\textit{e.g.}, LGBTQIA+-friendly spaces), iii) if the conference has an in-person component, pre-assign spaces for religious practices, lactation, and child care, iv) include menu options that account for diverse dietary and religious requirements.
+These are just a few easy-to-implement examples with a significant impact on including key sub-groups that otherwise find it challenging (see \cite{noauthor_discover2021} for some great advice on 'low-hanging fruits').
+Importantly, by being proactive in creating such a welcoming space, the organizers can be respectful of attendees' privacy and avoid requiring anyone to disclose personal information, such as revealing a disability, sexual orientation, gender identity, or mental health issues, to a name a few.
 % Rocío: Could have a paragraph about active bystandership—Andrea%ö
@@
 % Rocío: main idea: Code of conduct for a safe place
-Adopting a Code of Conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
-The Code of Conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
-The Code of Conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly creating protocols on how to respond.
-This should be done during the early phases of the conference planning because Code of conduct violations can happen before the conference (e.g.
-social media, organization meetings).
-Since people from marginalized groups are more likely to be victims of Code of Conduct violations and to see their claims dismissed due to power dynamics, assembling a diverse Code of Conduct team will be essential to deal with discrimination and harassment issues and the power dynamics underneath them and deal with them in a sensitive manner.
-It may also be more understanding of different disciplinary cultures and geographical and cultural considerations.
-We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
+
+Adopting a Code of Conduct (CoC) and gathering a diverse team to enforce these ideas are vital in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
+The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
+The CoC team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize protocols on how to respond correspondingly.
+This should be done during the early phases of the conference planning because CoC violations can happen before the conference too (\textit{e.g.}, social media, organization meetings).
+Since people from marginalized groups are more likely to be victims of Code of Conduct violations and find their claims dismissed due to inadvertent power dynamics, assembling a diverse CoC team will be essential to appropriately deal with discrimination and harassment issues and the power dynamics underneath them with respect and sensitivity.
+A well-drafted CoC may also be more understanding and accommodating of different disciplinary cultures and geographical and cultural considerations.
+We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC team.
 

### 2021-12-02 plos_latex_template.tex
RJ: added author contributions (for now)

@@
 Sara Mortara\textsuperscript{3},
-Heather Turner\textsuperscript{4}, %\ddag},
-Dorothea Hug Peter\textsuperscript{5}, %\ddag},
-Matt Bannert\textsuperscript{6},
-Batool Almazrouq\textsuperscript{7},
-Liz Hare\textsuperscript{8},
-Janani Ravi\textsuperscript{9},
+Yanina Bellini Saibene\textsuperscript{4},
+Heather Turner\textsuperscript{5}, %\ddag},
+Dorothea Hug Peter\textsuperscript{6}, %\ddag},
+Matt Bannert\textsuperscript{7},
+Batool Almazrouq\textsuperscript{8},
+Liz Hare\textsuperscript{9},
 Laura Ación\textsuperscript{10},
@@
 Silvia Canelón\textsuperscript{19\Yinyang},
-Yanina Bellini Saibene\textsuperscript{20}%,3\textcurrency}
+Janani Ravi\textsuperscript{20}
+%,3\textcurrency}
 \\
@@
 
+\section*{Authors' contributions}
+
+To list author contributions, we used the high-level contributor roles in the CRediT framework relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and an additional category for contributions on their own ten simple rules for an inclusive conference without looking at the first draft (i.e.
+sharing rules).
+
+\begin{itemize}
+    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization
+    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules
+    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules
+    \item YBS: Conceptualization, Writing--major reviews and edits, Resources
+    \item HT: Conceptualization, Writing--major reviews and edits, Resources
+    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules
+    \item MB: Conceptualization
+    \item BA: Writing--major reviews and edits, Resources, Sharing rules
+    \item LH: Writing--major reviews and edits, Resources, Sharing rules
+    \item LA: Writing--major reviews and edits, Resources
+    \item JPN: Writing--major reviews and edits, Visualization
+    \item MAC: Writing--major reviews and edits
+    \item FM: Writing--major reviews and edits
+    \item RG: Writing--major reviews and edits
+    \item NM: Sharing rules
+    \item AU: Sharing rules
+    \item JC: Sharing rules
+    \item AE: Sharing rules
+    \item SC: Sharing rules
+    \item JR: Writing--major reviews and edits
+\end{itemize}
+
 \section*{Acknowledgments}

### 2021-12-02 plos_latex_template.tex
added a few changes to Rule 3 ~ @jananiravi

@@
 
-A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team in charge of the conference decision-making.
+A genuinely inclusive conference can only be organized by an inclusive and diverse team in charge of the conference decision-making.
 Organizers should be a team with people from different regions, genders, ethnicities, socioeconomic statuses, and other aspects of diversity.
-Having diverse people in decision-making positions will positively affect all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an important role in creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see Rule 2 about a safe and welcoming environment).
-If you already started assembling an organizing team, check for gaps in its composition.
-The regional and local communities, groups, or associations in your field are good sources to tap into.
+Having diverse people in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
+In addition, a diverse team plays an essential role in creating a welcoming space because representation -- seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes -- is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule 2} about creating a safe and welcoming environment).
 
 % Rocío: Idea of this paragraph: non-transferability and do not restrict people
-People with disabilities often say: `Nothing about us without us' [ref]; the same holds for other dimensions of diversity.
-The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions from people that have not been through the same experiences \cite{costanzachockDesign2020}.
+People with disabilities often say: 'Nothing about us without us' [ref]; the same holds for other dimensions of diversity.
+The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
 % People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
-However, do not require nor expect minoritized people to only work—or want to work—on diversity and inclusion aspects; 
-everybody must have the freedom to choose in which areas of the conference they want to work. 
+It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work—or want to work—on diversity and inclusion aspects; 
+every member should be granted the freedom to choose which areas of the conference they want to work on, in addition to contributing to decisions involving marginalized communities.
+Needless to say, they are respected members of the scientific community, and they should be treated as such.
+
 % Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
-Furthermore, do not assume that self-nomination and voting will work as mechanisms to choose leadership positions.
-Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
-Build an environment in which every person can express their opinion and pay special attention to people from systematically excluded groups.
-Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (e.g.
-public recognition on social media, on the conference website, or during the conference).
+
+\emoji{bulb} \textbf{Practical tips:} If you have already started assembling an organizing team, check for gaps in its composition and find representatives accordingly. 
+The regional and local communities, groups, or associations in your field are good sources to tap into. 
+Furthermore, self-nomination and voting do not always work as reliable mechanisms to fill leadership positions.
+Instead, it helps to nominate members directly and offer them leading positions in the organization team, especially those that would be typically occupied by people from privileged groups.
+Build an environment that encourages healthy conflict and constructive criticism, one in which every person can express their opinion and pay special attention to people from systematically excluded groups.
+Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (\textit{e.g.}, public recognition on social media, on the conference website, or during the conference, ideally whatever works best for the contributing member).
 In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.

### 2022-01-18 plos_latex_template.tex
Just in case, pushing latest changes from Janani

@@
 {\Large
-\textbf\newline{Ten simple rules towards hosting an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 
+% % Batool: I didn't change anything in the abstract - you can check the history - I copied the abstract and pasted, then the colour change for the abstract but it contains all text and suggestions/edits by Janani - my apologies (didn't commit the changes).
 The authors of this article recently organized useR!
@@
 % Rocío: main idea: Inclusion for everyone
-While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a genuinely inclusive environment means taking care of all aspects of diversity.
+While it is common to think about representation in terms of some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a genuinely inclusive environment means taking care of all aspects of diversity.
 Inclusivity should lie at the heart of setting the tone and meeting space for any meeting as we move forward in the 21st century!
@@
 Importantly, by being proactive in creating such a welcoming space, the organizers can be respectful of attendees' privacy and avoid requiring anyone to disclose personal information, such as revealing a disability, sexual orientation, gender identity, or mental health issues, to a name a few.
-% Rocío: Could have a paragraph about active bystandership—Andrea%ö
-%promote a supportive and respectful environment where attendants act as active bystanders. y definir '¬¬ 
+
+The organizers should promote a supportive and respectful environment, where all attendants can be themselves without fear of rejection or isolation.
+This includes promoting the idea that anyone can act in defense of other attendants in case anything unpleasant, incorrect, of offensive happens.
+Instead of looking away, act as "active bystanders" and intervene to keep the community safe.
+[ast: should we add pre-conference allyship workshops? i saw a conference that had one pre-conference events for the whole community, just brilliant]
 % Rocío: main idea: Code of conduct for a safe place
-
-Adopting a Code of Conduct (CoC) and gathering a diverse team to enforce these ideas are vital in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
+However, the most important step that organizers should take to ensure the safety of the community is
+Adopting a Code of Conduct (CoC) and gathering a diverse team to enforce it \cite{favaroYourScienceConference2016}.
 The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
@@
 
-These rules concern diverse elements in the design of the conference and operational tasks. They are all important pieces for inclusion.
-In rule 4, we introduce ways to counteract bias in the conference program, in keynotes, program committee, abstract selection, and other spaces. 
-Rule 5 describes the advantages of online conferences reducing barriers for participation, and provides advice for organizing online and hybrid conferences.
-Rule 6 focuses on practices to make all conference platforms and spaces accessible to people with disabilities. 
-In rule 7, we advocate for more language-inclusive conferences, and provide suggestions to encourage and ensure full participation of non-native English speakers. 
-Rule 8 provides tips for an inclusive media strategy. 
-In rule 9, we address budgeting for inclusive practices, and helping participants with affordable registration costs, scholarships, and other forms of financial support.
+The next set of rules concern diverse elements in the design of the conference and operational tasks. They focus on weaving inclusion into the conference design process.
+In \textbf{Rule 4}, we introduce ways to counteract bias in the conference program, keynotes, program committee, abstract selection, and other spaces. 
+\textbf{Rule 5} describes the advantages of online conferences by reducing barriers for participation and provides advice for organizing online and hybrid conferences.
+\textbf{Rule 6} focuses on best practices to make all conference platforms and spaces accessible to people with disabilities. 
+In \textbf{Rule 7}, we advocate for more language-inclusive conferences and provide suggestions to encourage and ensure full participation of non-native English speakers. 
+\textbf{Rule 8} provides tips for developing an inclusive media strategy. 
+In \textbf{Rule 9}, we address budgeting for inclusive practices, and helping participants with affordable registration costs, scholarships, and other forms of financial support.
 
@@
 If this likely case, and in fields were it is not possible to guarantee anonymity of the authors during the selection process (e.g.
-non-anonymized github repositories need to be examined), single blind reviews are the only option left.
+non-anonymized GitHub repositories need to be examined), single blind reviews are the only option left.
 In any case, create clear evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g.

### 2022-01-19 plos_latex_template.tex
edits to design rules, esp. Rules 4, 5 ~ @jananiravi

@@
 % Please keep the abstract below 300 words
-\section*{Abstract (optional from what I've seen)}
+\section*{Abstract}
 
-% % Batool: I didn't change anything in the abstract - you can check the history - I copied the abstract and pasted, then the colour change for the abstract but it contains all text and suggestions/edits by Janani - my apologies (didn't commit the changes).
 The authors of this article recently organized useR!
 2021, the annual global R conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector.
-To embrace the constraints posed by the pandemic, this year we hosted a 100\% virtual conference.
-We worked on building a high-quality virtual and explicitly global conference by creating a kind, inclusive, accessible, and welcoming environment for everyone.
-This article outlines our most important lessons learned abstracted as '10 simple rules to host an inclusive conference'.
+To embrace the constraints posed by the pandemic, we hosted a 100\% virtual conference.
+We worked on building a high-quality virtual and global conference by creating a kind, inclusive, accessible, and welcoming environment for everyone.
+This article outlines our most important lessons learned abstracted as `10 simple rules to host an inclusive conference'.
 These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules seamlessly translate to regional events.
-These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events even after the COVID pandemic.
+We have designed them for wide adoption to enable and ensure that the scientific and technical conferences are as accessible and inclusive as they can be.
 
@@
 % Rocío: Main message of the paragraph: Conferences are great, but for some
-Conferences, from the Latin \textit{conferre} `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
-The larger the conference, the larger the opportunities to network and learn from the cohort.
+Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+A good conference experience can be 
 However, opportunities for participating in conferences are not equal for all. 
-Many academic and tech conferences have become exclusive spaces for privileged groups, \textit{i.e.}, with unearned advantages given by society to some but not all (\textit{e.g.}, white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+Many academic and tech conferences have become spaces that reproduce systemic inequalities and privileges, i.e.
+advantages given by society to people from some groups, like being white, male, from a rich country, English-native speaker, with no physical disabilities \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
 % However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
@@
 They are organized in three sections: \textbf{the pillars}, fundamental rules with key elements to conceive the work on diversity and inclusion in any conference (embracing diversity, creating a welcoming environment, and having a diverse organizing team); \textbf{design rules}, concerning diverse elements in the design of the conference and operational tasks (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices); and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion.
-These practices stem from the experience of organizing useR!
-2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
-useR!
-conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation.
-We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
+These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
+We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
 
 The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference.
-Therefore, the rules result from our lessons learned from diverse perspectives and expertise (\textit{i.e.}, our different backgrounds, profiles, and roles in the conference). 
+We come from different backgrounds, career stages and even continents.
+The rules result from our lessons learned from diverse perspectives and expertise. 
 Even though these rules were designed based on our experience with a global event, they can be easily transferred or scaled to events with regional or local scopes.
-These rules will greatly benefit people who are part of meeting committees that oversee the site/location selection process, or that coordinates with the local organizers of conferences.
-The rules will also help local/virtual organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
-We have designed them for wide adoption to enable and ensure that the scientific and technical conferences during and post-pandemic are as accessible and inclusive as they can be.
+These rules will greatly benefit people who are part of meeting committees that oversee the site/location selection process, or that coordinate with the local organizers of conferences.
+The rules will also help organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
+We have designed them for wide adoption to enable and ensure that the scientific and technical conferences are as accessible and inclusive as they can be.
 
@@
 %ast: R/ Unfortunately, support for adding accessibility metadata to LaTeX documents is limited.
-You will probably need to use Adobe Acrobat to add missing accessibility metadata to your PDF file.https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#latex
+You will probably need to use Adobe Acrobat to add missing accessibility metadata to your PDF file.https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#latex %ast: let's try not to forget this when submitting - some edits to the PDF may be needed to add alt-text to the final version
 \pdftooltip{
@@
 
-To organize a more inclusive conference, it is necessary to identify upfront which groups of people are being excluded, the barriers that keep them from active partaking, and targeted strategies to overcome them.
+To organize a more inclusive conference, it is necessary to identify upfront which groups of people are being excluded, the barriers that keep them from active partaking, and strategies to overcome them.
 \textbf{Rule 1} is about embracing all dimensions of diversity and prioritizing the participation of groups that are often marginalized.
@@
 Unfortunately, we live in a world with implicit hierarchies along these axes. 
-Some statuses (\textit{e.g.}, cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems -- including conferences -- are consciously and unconsciously built.
-People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; a few of them have to deal with several of these on a daily basis.
-% While no isolated initiative can drastically alter conference outcomes, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles.
+Some statuses (e.g.
+cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built.
+People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; some of them have to deal with several of these on a daily basis.
+% While no isolated initiative can drastically alter xx, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles.
 Therefore, recognizing this systematic exclusion, particularly in our fields and in our scientific or professional communities, is paramount to inclusion, and will help identify the most marginalized and discriminated subgroups \cite{timperleyHeMoanaPukepuke2020}.
@@
 
-\emoji{bulb} \textbf{Practical tips:} As a starting point, the meeting organizers could identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (\textit{e.g.}, explicit discrimination, participation fees, etc.).
+As a starting point, the meeting organizers could identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (e.g.
+explicit discrimination, participation fees).
 This analysis should guide the vision of diversity and inclusion for your conference and help design strategies to achieve it. 
 Express this vision for your conference in a diversity statement. 
-Next, use this vision and the set strategies to define measurable indicators and goals of diversity and inclusion (\textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, or participation from key geographic regions).
+Next, use this vision and the set strategies to define measurable indicators and goals of diversity and inclusion (e.g.
+a gender distribution of your speakers that is representative of the general population, the presence of racialized people (e.g.
+Black people) among the organizers, speakers, and attendees, or participation from key geographic regions).
 These numbers should guide you and help achieve your vision.
 
-% Investing a substantial effort towards the most excluded groups does not mean neglecting the other communities, but will instead ensure equity, guide the vision of diversity for your conference, and help design strategies to achieve it.
-% vision
-% back to the indicators? 
 % realistic but bold (back?)
 % be aware and promote awareness
-%¡LAS BARRERAS!
-%Express the diversity and inclusion vision for your conference by writing a diversity statement. 
-% As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors.
-Next, set measurable indicators of success when organizing a diverse conference.
-The numbers resulting from these indicators are not the primary goal, and this is not about performative allyship.
-Instead, these numbers should guide and help hold the organizing team accountable towards these diverse and inclusive goals.
-
 % Rocío: more effort towards including the more excluded
-% This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.
+% This may translate into different actions depending on your field, region, or community
 
@@
 While it is common to think about representation in terms of some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a genuinely inclusive environment means taking care of all aspects of diversity.
-Inclusivity should lie at the heart of setting the tone and meeting space for any meeting as we move forward in the 21st century!
+%Inclusivity should lie at the heart of setting the tone and meeting space for any meeting as we move forward in the 21st century!
 % Edit and make more specific (below)
+%\emoji{bulb} \textbf{Practical tips:} 
+Here are a few ways in which conference organizers can actively enforce inclusiveness, such as i) Promote respect of pronouns, ii) schedule events exclusive for diverse groups (e.g.
+LGBTQIA+-friendly spaces), iii) if the conference has an in-person component, pre-assign spaces for religious practices, lactation, and child care, iv) include menu options that account for diverse dietary and religious requirements.
+These are just a few easy-to-implement examples with a significant impact on including key sub-groups that otherwise find it challenging (see \cite{noauthor_discover2021} for some great advice on `low-hanging fruits').
+Importantly, by being proactive in creating such a welcoming space, the organizers can be respectful of attendees' privacy and avoid requiring anyone to disclose personal information, such as revealing a disability, sexual orientation, gender identity, or mental health issues, to name a few.
 
-\emoji{bulb} \textbf{Practical tips:} Here are a few ways in which conference organizers can actively enforce inclusivity, such as i) Promote respect of pronouns, ii) schedule events exclusive for diverse groups (\textit{e.g.}, LGBTQIA+-friendly spaces), iii) if the conference has an in-person component, pre-assign spaces for religious practices, lactation, and child care, iv) include menu options that account for diverse dietary and religious requirements.
-These are just a few easy-to-implement examples with a significant impact on including key sub-groups that otherwise find it challenging (see \cite{noauthor_discover2021} for some great advice on 'low-hanging fruits').
-Importantly, by being proactive in creating such a welcoming space, the organizers can be respectful of attendees' privacy and avoid requiring anyone to disclose personal information, such as revealing a disability, sexual orientation, gender identity, or mental health issues, to a name a few.
-
-The organizers should promote a supportive and respectful environment, where all attendants can be themselves without fear of rejection or isolation.
+The organizers should promote a supportive and respectful environment, where all attendants can participate without fear of rejection or isolation.
 This includes promoting the idea that anyone can act in defense of other attendants in case anything unpleasant, incorrect, of offensive happens.
 Instead of looking away, act as "active bystanders" and intervene to keep the community safe.
-[ast: should we add pre-conference allyship workshops? i saw a conference that had one pre-conference events for the whole community, just brilliant]
-% Rocío: main idea: Code of conduct for a safe place
+% Rocío: main idea: Code of conduct for a safe place %i think the transition here should leave the two concepts closer. not however. 
 However, the most important step that organizers should take to ensure the safety of the community is
@@
 The CoC team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize protocols on how to respond correspondingly.
-This should be done during the early phases of the conference planning because CoC violations can happen before the conference too (\textit{e.g.}, social media, organization meetings).
+This should be done during the early phases of the conference planning because CoC violations can happen before the conference too (e.g.
+social media, organization meetings).
 Since people from marginalized groups are more likely to be victims of Code of Conduct violations and find their claims dismissed due to inadvertent power dynamics, assembling a diverse CoC team will be essential to appropriately deal with discrimination and harassment issues and the power dynamics underneath them with respect and sensitivity.
-A well-drafted CoC may also be more understanding and accommodating of different disciplinary cultures and geographical and cultural considerations.
+A well-drafted CoC should account for  different cultures and geographical and cultural considerations.
 We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC team.
@@
 Organizers should be a team with people from different regions, genders, ethnicities, socioeconomic statuses, and other aspects of diversity.
-Having diverse people in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an essential role in creating a welcoming space because representation -- seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes -- is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule 2} about creating a safe and welcoming environment).
+Having people from diverse backgrounds in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
+In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule 2} about creating a safe and welcoming environment).
 
 % Rocío: Idea of this paragraph: non-transferability and do not restrict people
-People with disabilities often say: 'Nothing about us without us' [ref]; the same holds for other dimensions of diversity.
+People with disabilities often say: `Nothing about us without us' [ref]; the same holds for other dimensions of diversity.
 The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
 % People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
-It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work—or want to work—on diversity and inclusion aspects; 
+It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
 every member should be granted the freedom to choose which areas of the conference they want to work on, in addition to contributing to decisions involving marginalized communities.
-Needless to say, they are respected members of the scientific community, and they should be treated as such.
 
@@
 
-\emoji{bulb} \textbf{Practical tips:} If you have already started assembling an organizing team, check for gaps in its composition and find representatives accordingly. 
+%\emoji{bulb} \textbf{Practical tips:} 
+If you have already started assembling an organizing team, check for gaps in its composition and find representatives accordingly. 
 The regional and local communities, groups, or associations in your field are good sources to tap into. 
@@
 Build an environment that encourages healthy conflict and constructive criticism, one in which every person can express their opinion and pay special attention to people from systematically excluded groups.
-Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (\textit{e.g.}, public recognition on social media, on the conference website, or during the conference, ideally whatever works best for the contributing member).
-In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.
+Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (e.g.
+public recognition on social media, on the conference website, or during the conference, ideally whatever works best for the contributing member).
+Furthermore, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.
 
@@
 The next set of rules concern diverse elements in the design of the conference and operational tasks. They focus on weaving inclusion into the conference design process.
-In \textbf{Rule 4}, we introduce ways to counteract bias in the conference program, keynotes, program committee, abstract selection, and other spaces. 
+In \textbf{Rule 4}, we introduce ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
 \textbf{Rule 5} describes the advantages of online conferences by reducing barriers for participation and provides advice for organizing online and hybrid conferences.
@@
 
-When choosing or inviting people for visible and valued roles in the conference––such as keynote speakers, program/scientific committee members, session chairs––, it is likely that there will not be much diversity in the first set of names.
-Our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
+When choosing or inviting people for visible and valued roles in the conference –– such as keynote speakers, program/scientific committee, session chairs –– it is likely that there will not be much diversity in the first pool of names.
+This is likely due to our inherent biases resulting from existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
 Rather than deter us, this should encourage us to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not stay with people in privileged groups (e.g.
-avoid `manels' and all-white panels, for instance).
-Make sure the composition of all these groups is balanced across topics, areas, and representation across regions, gender, race, career stage, among other aspects.
+avoid `manels' and all-white panels; \cite{else_how_2019}; see \textbf{Rule 8}).
+\emoji{bulb} It would help to ensure that the composition of all these groups is balanced across topics and areas with representation across regions, gender, race, and career stage, among other aspects.
 
@@
 
-The final list of accepted talks can also be biased in both submission and acceptance processes. 
-People who have never participated in conferences, and especially those regarded as big and prestigious, may not feel confident to submit their work.
-To counteract this self-selection bias, promote the submission call beyond the communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
-Some of these communities may have pre-submission mechanisms to help their members prepare and improve their submissions (e.g.
-the Latin-R community has a Slack channel with the purpose of helping community members improve their abstracts for conferences, and the R-Ladies community did the same for useR!
-2021).
-On the other hand, the review process may be biased because selection committees—even diverse ones—can be subjected to unconscious or implicit bias.
-Unconscious bias happens when we interpret information, such as names of the authors, we assume their origin or ethnicity, their institution of origin, and we attach a positive or negative value to their work according to negative or positive stereotypes.
-Nobody is free from implicit bias [cite Ross2019].
-Selection committees need to be reminded of such biases (the implicit association test can help identifying them [cite Greenwald - Measuring individual differences in implicit cognition: The implicit association test]), and
-decide on the most appropriate strategy to counteract them, either with double blind, double open, or single blind reviews [cite discover cookbook].
-Double blind reviews can help counteract implicit bias only if it is possible to completely anonymize the identifying information from submission. 
-Double open, though more transparent, may still be subjected to implicit bias, and reviewers from minoritized groups might feel additional pressure because the reviews are signed.
-If this likely case, and in fields were it is not possible to guarantee anonymity of the authors during the selection process (e.g.
-non-anonymized GitHub repositories need to be examined), single blind reviews are the only option left.
-In any case, create clear evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g.
+The final list of abstracts making it into the program could also be biased in both submission and acceptance processes. 
+For instance, first-time conference participants, especially for the flagship/prestigious meetings, may not feel confident to submit their work.
+\emoji{bulb} To counteract this self-selection bias, promote the submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
+Some of these communities may have pre-submission mechanisms to help their members receive feedback for their submissions (e.g.
+the Latin-R and R-Ladies communities have Slack channels to help their members improve their conference abstracts).
+On the other hand, the review process may be biased because selection committees — even diverse ones — can be subjected to unconscious or implicit bias.
+Unconscious bias happens when we interpret information such as names of the authors, assuming their origin, ethnicity, or their institution of origin, and we attach a positive or negative value to their work according to the stereotypes.
+Unfortunately, nobody is free from implicit bias \cite{ross_everyday_2020}.
+Selection committees need to be gently reminded of such biases with evaluation guidelines (the implicit association test can help identifying them \cite{greenwald_measuring_nodate}) and the organizers need to
+identify the most appropriate strategy to counteract them, either with double blind, double open, or single blind reviews \cite{numfocus_discover_2021}.
+
+Double blind reviews can help counteract implicit bias only if the identifying information is completely anonymized in the submission. 
+Double open reviews, though more transparent, may still be subjected to implicit bias, and reviewers from minoritized groups might feel additional pressure because the reviews are signed.
+In this likely case, and in fields where it is not possible to guarantee anonymity of the authors during the selection process (e.g.
+when non-anonymized GitHub repositories need to be examined), single blind reviews are the only remaining option.
+
+%\emoji{bulb} \textbf{Practical tips:} 
+In any case, design unambiguous evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g.
 based on their number of publications or talks), and share them with the authors and reviewers.
-Advertising your practices to mitigate bias in proposal selection may also encourage people from diverse groups to submit their work, thus counteracting self selection bias.
+Advertising your practices to mitigate bias in proposal selection may also encourage people from diverse groups to submit their work, thus counteracting self-selection bias.
 
@@
 % Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
-Implicit bias also pertain the value that we give to certain activities; e.g.
-scientific publication and development of software are often regarded as higher value than other roles like community building or teaching.
-These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
-Research about these topics is undervalued relative to other areas of knowledge production.
-Counteracting bias also means innovating the program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, and giving visibility to the whole range of activities and practitioners that contribute to the overall topics of the conference.
+Implicit bias also manifests in the weightage given to certain products/activities; e.g. scientific publication and development of software are often rated higher than community building or teaching. 
+The latter roles are equally, if not more, challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
+Likewise, research about these topics is undervalued relative to other areas of knowledge production. 
+Counteracting bias actively means innovating the program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, and giving visibility to the whole range of activities and practitioners that contribute to the overall range and quality of the conference.
 
@@
 Virtual conferences are more inclusive, removing barriers like costs of registration, transport, accommodation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}.
-After experiencing virtual conferences mostly due to the COVID pandemic, there is a widespread call to keep an online component in conferences, either maintaining conferences completely virtual or using hybrid formats combining in-person and online components \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
-Though more inclusive, an online component does not guarantee inclusion and accessibility for everyone.
-Key decisions must be made early on in the planning stages \cite{levitisCenteringInclusivityDesign2021}.
-For instance, when choosing the conference platforms (for submission, registration, chatting system, live streaming, etc.) account for accessibility to screen readers, geopolitical restrictions, and bandwidth requirements.
-You should also decide how to be inclusive of geographic regions with the schedule of the conference, asking for pre-recorded videos, recording sessions, and sharing them during the conference.
-Some possibilities for scheduling are: having different days dedicated to different time zones; having each day with the same conference hours, and scheduling time slots that are friendly--and unfriendly--to different time zones; or having time-zone hubs in each day and repeating the events in each (in some cases live and others a recorded version).
+After experiencing virtual conferences mostly due to the COVID pandemic, there is a widespread call to retain an online component in conferences, either hosting conferences completely virtual or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
+It is important to note that while being more inclusive, an online component does not guarantee inclusion and accessibility for everyone.
+Key decisions must be made early on during the planning phase \cite{levitisCenteringInclusivityDesign2021}.
+For instance, the chosen conference platforms (e.g.
+for submission, registration, chatting system, live streaming, etc.) must account for accessibility to accommodate screen reader users, geopolitical restrictions, and bandwidth requirements.
+\emoji{bulb} The organizers should also decide how to incorporate geographic regions within the schedule of the conference by asking for pre-recorded videos, recording sessions, and sharing them during the conference.
+Some possibilities for scheduling are: having different days dedicated to different time zones; having each day with the same conference hours, and scheduling time slots that are friendly to different time zones; or having time-zone hubs in each day and repeating the events in each (live for one/few and recorded versions for others).
+A flexible schedule and recorded material will allow many more attendees to partake in the conference, e.g.
+people with schedules conflicting with the live conference, those who cannot take off, or others who are caretakers.
 
 % Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
-Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
-However, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g.
-\cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}.
+Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to in-person interactions at social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
+However, there is evidence that virtual communication can make people from minoritized groups feel more included, thus participate and contribute more \cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}.
+
+%\emoji{bulb} Practical tips:
 Invest time in creating opportunities to meet and bond virtually that can appeal to people with different backgrounds, disabilities, preferences, and remembering that no single networking activity will ever serve the whole community.
-Try new and varied activities centered around different groups of people outside the mainstream of your conference; if you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive, successful, and inclusive.
-Some examples are the newbies sessions for first-timers, social mixers lead by specific subgroups or communities like LGBT+, or leisure activities that reunite subgroups with the same interests: virtual art exhibitions, yoga sessions, movie watch parties, trivia quizzes, etc.
-Offer some activities that require voice or video interactions, and others that involve only written chat.
+Try new activities centered on different groups of people outside the mainstream of your conference (e.g.
+\textbf{ABC}); if you think of activities that reach specific groups, preferably co-led by community leaders, these sessions can be more productive, successful, and inclusive.
+Some examples are newbies sessions for first-timers, social mixers led by specific subgroups or communities like LGBTQAI+, or leisure activities that reunite subgroups with the same interests: virtual art exhibitions, yoga sessions, movie watch parties, trivia quizzes, shared language, etc.
+Offer activities that require voice or video interactions, and others that involve only written chat, and make all participation optional and voluntary.
 
 If you are organizing a hybrid event instead, you should decide on the relative importance that the online and in-person components will take.
-The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the format they choose to participate in.
-Under this format, the audience, chairs, and presenters--of talks, posters and workshops--could be either in-person or remote, and they should all be able to interact. 
-For that purpose, all questions and comments should pass through an online system. 
-The physical venue would have to include multiple ways to connect with online folks such as tablets on the meeting venue, cameras and microphones that allow online attendants to follow who is speaking in the in-person space.
-This requires the venue to ensure stable and secure internet connection for the in-person participants to join the online folks. 
+The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the preferred format /cite{bajpai_towards_2021}.
+
+\emoji{bulb} For this format, the audience, chairs, and presenters -- of talks, posters, and workshops -- could be either in-person or remote, and they should all be able to interact. 
+To best include both kinds of attendees, all questions and comments should pass through an online system, or have an pre-designated point person who enables this cross-talk. 
+The physical venue would have to include multiple ways to connect with online folks such as tablets on the meeting venue, cameras and microphones that allow online attendants to follow who is speaking at the in-person space.
+This requires the venue to have stable and secure internet connection for the in-person participants to join the online folks. 
 Having a hybrid format also means accounting for the fact that the venue does not need to host all the attendees.
@@
 This could save you time and money when looking for a suitable venue.
-When planning for networking activities in hybrid events, use the same principles of the paragraph above and offer varied activities centered around different groups.
-Some of these activities could be only in-person, but to promote connections between all participants, there should be a fair proportion of activities connecting in-person and online attendees.
+When planning for networking activities in hybrid events, use the same principles outlined above and offer various activities engaging different groups.
+It is understandable that a few of these activities could be only in-person, but to promote connections between both groups of participants, there should be a fair proportion of activities connecting in-person and online attendees.
 
@@
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
-This is one key aspect where participation in the organizing team allows disabled people to take part in the decisions from the beginning (see rule 3 about the organizing team), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}.
-Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}.
+Including people with disabilities in the organizing team can have a large impact from the beginning (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making \cite{irishIncreasingParticipationUsing2020}.
 As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference.
@@
 
-For in-person components of conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, ADHD, autism and its spectrum).
+If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
+In addition to this, the organizers should take care proactively of invisible disabilities (e.g.
+dyslexia, anxiety, ADHD, autism).
 This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.
 
-Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly and keyboard accessible, and have alternative text.
+Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly, keyboard accessible, and have alternative text.
 Every video should have good-quality (not automated) captions and a transcript; live presentations should have good-quality captions too.
 Provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
+For instance, useR!
+2021 published accessibility guidelines for every format of the conference, a blog post with recommendations, and two post-conference blog posts regarding the whole accessibility experience.
+The participation of a blind person in the organizing team was key to several of these decisions.
 Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.
 
-Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability.
+Accessibility practices should cover social events and networking, which should include activities that do not restrict participation based on body type or ability.
+
 Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
 For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
-Accessibility efforts should be explicitly displayed in the official communication channels (see rule 8 about communication strategy).
+Accessibility efforts should be explicitly displayed in the official communication channels and inform potential attendees with disabilities a point of contact to solve any questions in this respect.
+Clear communication about the lack of accessibility of some spaces or activities is also highly recommended, to give a clear notion of the conference experience (see \textbf{Rule 8} about communication strategy).
 
-%LA: provide live, high-quality (automated captions are often referred to as "craptions"), captioning or simultaneous translation in English for persons who only speak English.
-After all, captions and simultaneous translations are common place in the audiovisual industry and your conference audience will likely be already familiar with this if they ever watched a movie from a non-English speaking country.
-
-\subsection*{Rule 7: Don't let language restrict high-quality participation}
+\subsection*{Rule 7: Include languages other than English}
 \label{rule_language}
@@
 In international meetings, the linguistic diversity of the participants is often overlooked. 
-English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
-While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.
+English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpages, and communications. 
+While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.%RJ: Could cite Amano's paper here.
 Non-native English speakers could miss opportunities to attend or actively participate in conferences (e.g. asking questions or participating in discussions),
@@
 % Rocío: main idea: make conferences more linguistically inclusive
-Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (rule 2 about a welcoming environment for all).
+Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (\textbf{Rule 2} about a welcoming environment for all).
 This may be done at different levels.
+Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.
+That way, the abstract will be judged primarily by the quality or relevance of the work instead of the English grammar.
 For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities) follow the presentations.
@@
 The following step would be having sessions and events in languages other than English, both regarding the technical content of the conference and the social and networking aspects.
-Networking sessions in languages other than English can include captions or interpretation or not, which has the benefit of not requiring major technical adjustments.
+Networking sessions in languages other than English can include captions, interpretation or none of them; the latter has the benefit of not requiring additional technical adjustments.
 Keynotes and regular talks should always be accompanied by captions or live interpretation to English.
-
 Promote the attendance to presentations in languages other than English as key components of your schedule and announce when captions are available. 
-Remind your native English speaking audience to be respectful of accents and mistakes in English from non-native English speakers.
+Remind your native English speaking audience to be respectful of accents and mistakes in English.
 
@@
 Actively reach out and promote the conference to people who have been systematically excluded. 
-Publish the diversity statement, the Code of Conduct, the accessibility guidelines, and  options for financial support (see rule 9 about financial resources), to emphasize that everybody is welcome in the event.
+Publish the diversity statement, the Code of Conduct, the accessibility guidelines, and  options for financial support (see \textbf{Rule 9} about financial resources), to emphasize that everybody is welcome in the event.
 Try to come up with creative ways to show that everyone is seen, respected, and welcome.
@@
 Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues.
-For instance, inform if the conference platform is not completely screen-reader friendly, but you are offering personalized help to navigate it, or if captions will be available for some talks but not all.
-%ast: responsible communication!
+For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering personalized help to navigate it, or if captions will be available for some talks but not all.
 
 %Rocío: Main idea: Use inclusive language
-Inclusive language--language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups--should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
-Become familiar with the terminology used for disabilities, racialized groups, and gender and sexual orientations, and with the terms that should be avoided. %
-Do not expect minoritized people to teach you--it's not their role--, and accept feedback without being offended.
+Inclusive language—language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups—should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
+Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, and with the terms that should be avoided. %
+Do not expect minoritized people to teach you—it's not their role—, and accept feedback without being offended.
 Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms. 
@@
 }
-\caption{Margot the marmot was useR! 2021 mascot}
+\caption{Margot the marmot was useR!
+2021 mascot.
+To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing diverse groups that are part of the large R community: (from left to right) MiR (Minorities in R), LGBTQAI+, R-Ladies, transgender people, and AfricaR.
+%RJ: Need to acknowledge Francisco's contribution.}
 \label{fig:marmots}
@@
 Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
-Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance).
+Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying because they think they don't stand a chance).
 Conferences usually ask for cover letters or applications to assign these scholarships, and this can be a time-consuming, emotionally-demanding task. 
 Simplify as much as you can the process of asking for financial support; 
-people who do not have the time (e.g., because they have family responsibilities) might decide not to apply if the process is overtly complicated. 
+people who do not have the time (e.g. because they have family responsibilities) might decide not to apply if the process is overtly complicated. 
 Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups.
-Whenever possible, facilitate ways around money transfer (e.g., book flights, hotel reservations, waive conference fees).
+Whenever possible, facilitate ways around money transfer (e.g.
+book flights, hotel reservations, waive conference fees).
 
@@
 A conference is a critical community-building space, and can foster change in future events and the community itself.
-The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encountered to future organizers.
+The organizers of a conference edition and members of selection committees should do their best to pass their knowledge on best practices and problems encountered to future organizers.
 
@@
 Assess whether equity and inclusion goals were met during the conference. 
-You can rely on the diversity indicators defined at the beginning of the planning process (rule 1). 
+You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
 It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and others (e.g.
@@
 Document your processes, outcomes, and lessons learned in blog posts, reports, and presentations. 
-Make sure to share this information with future organizers, as well as the contact information of the people you worked with--if they agree to it, to give them a starting point to improve inclusion in the following editions of the conference.
+Make sure to share this information with future organizers, as well as the contact information of the people you worked with—if they agree to it—, to give them a starting point to improve inclusion in the following editions of the conference.
 If you are part of a stable meetings committee, encourage organizers to follow these rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria.
@@
 
-This article suggests changes in conference planning towards diversity and inclusion to make them networking and learning opportunities for all. 
-Organizing a conference and implementing inclusive practices are both learning experiences;
-you may not achieve everything you aimed for, set priorities and remember rule 10: diversity and inclusion are processes. 
-Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that a single conference cannot change. 
-Due to these systemic problems, even the minimum changes to how conferences have been historically organized may find resistance, but your courage to make these changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of people who are often excluded.
-Moreover, if more conferences and domains apply these rules, there will be less resistance, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
-% Rocío: I don't like the ending.
-(Guidelines/documentation) may be useful not only for your conference series but also for other communities with the same questions that you have. % Give examples of how others have helped us.
+This article suggests changes in conference planning towards diversity and inclusion. 
+Organizing a conference and implementing inclusive practices are both learning experiences.
+As conference organizers, we started with different levels of clarity about the principles and practices described here, learned many of them together during the organization phase, and learned even more when translating them into ten rules for this paper.
+We hope that they can be useful guidelines to improve diversity and inclusion in your conference.  
+And that you can adapt them, improve them, and share your lessons and experience with as many people as possible.
+
+Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, personal issues affecting your organizing team.
+Try to account for these factors as best as you can to set realistic yet bold vision and goals of diversity and inclusion for the conference.
+At the end of the day, even if you do not achieve them at 100\%, striving for them will make you stretch what feels possible and help you make changes you would have otherwise not even attempted to make.
+And most importantly, you would have contributed to a great conference experience for the attendees. 
+They will notice the welcoming and inclusive spirit of the conference reflected in the pillar rules, and translated into design decisions—even if there are gaps in them due to the constraining factors mentioned above.
+Inclusive conferences can help to enhance career paths and create a sense of belonging in the community. 
+That is more than enough motivation to make the effort worth it.
 

### 2022-01-19 plos_latex_template.tex
edits to Rules 8–10 ~ @jananiravi

@@
 % These are the groups that warrant deliberate efforts that ensure fair inclusion.
-
 As a starting point, the meeting organizers could identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (e.g.
@@
 For instance, the chosen conference platforms (e.g.
-for submission, registration, chatting system, live streaming, etc.) must account for accessibility to accommodate screen reader users, geopolitical restrictions, and bandwidth requirements.
+for submission, registration, chatting system, live streaming, etc.) must account for accessibility to accommodate screen-reader users, geopolitical restrictions, and bandwidth requirements.
 \emoji{bulb} The organizers should also decide how to incorporate geographic regions within the schedule of the conference by asking for pre-recorded videos, recording sessions, and sharing them during the conference.
@@
 If you are organizing a hybrid event instead, you should decide on the relative importance that the online and in-person components will take.
-The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the preferred format /cite{bajpai_towards_2021}.
+The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the preferred format \cite{bajpai_towards_2021}.
 
@@
 
-\subsection*{Rule 8: Express the welcoming spirit in your communication strategy}
+\subsection*{Rule 8: Convey the welcoming spirit in your communication strategy}
 % media
@@
 The communication strategy of your conference should aim to increase the number and diversity of participants, and be inclusive by design.
-Actively reach out and promote the conference to people who have been systematically excluded. 
-Publish the diversity statement, the Code of Conduct, the accessibility guidelines, and  options for financial support (see \textbf{Rule 9} about financial resources), to emphasize that everybody is welcome in the event.
-Try to come up with creative ways to show that everyone is seen, respected, and welcome.
+Actively reach out and promote the conference to communities that have been systematically excluded. 
+Publish the diversity statement, Code of Conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9} about financial resources) to emphasize that everybody is welcome to join the event.
+Try to develop creative ways to show that everyone is seen, respected, and welcome.
 For example, useR!
-2021 created a mascot for the conference wearing different scarves reflecting folks from different groups that the conference wanted to reach (Fig.
-2).
-The diverse members of your team should be able to decide which languages to emphasize and which social media platforms to utilize in the promotion of the conference (e.g.
+2021 created a mascot for the conference wearing different scarves reflecting folks from various minoritized groups that the conference wanted to reach (\textbf{Fig.
+2}).
+The diverse members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g.
 Twitter, Facebook, LinkedIn, conference website, mailing lists).
@@
 %Rocío: Main idea: Use inclusive language
-Inclusive language—language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups—should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
-Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, and with the terms that should be avoided. %
+\emoji{bulb} Inclusive language — language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups — should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
+Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, and other terms that should be avoided. %
 Do not expect minoritized people to teach you—it's not their role—, and accept feedback without being offended.
@@
 2021 mascot.
-To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing diverse groups that are part of the large R community: (from left to right) MiR (Minorities in R), LGBTQAI+, R-Ladies, transgender people, and AfricaR.
-%RJ: Need to acknowledge Francisco's contribution.}
+To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing diverse groups that are part of the large R community: (from left to right) MiR (Minorities in R), LGBTQAI+, R-Ladies, transgender people, and AfricaR.} %RJ: Need to acknowledge Francisco's contribution.
 \label{fig:marmots}
@@
 (see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
-Resources permitting, allow the possibility of a `pay what you can' option.
-You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than non-free events \cite{eventbrite_ultimate_2017}.
+Resources permitting, allow the possibility of a 'pay what you can' option.
+You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}.
 %Other costs can be prohibitive-> scholarships
@@
 Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying because they think they don't stand a chance).
-Conferences usually ask for cover letters or applications to assign these scholarships, and this can be a time-consuming, emotionally-demanding task. 
-Simplify as much as you can the process of asking for financial support; 
-people who do not have the time (e.g. because they have family responsibilities) might decide not to apply if the process is overtly complicated. 
+Conferences usually ask for cover letters or applications to assign these scholarships, which can be time-consuming and emotionally demanding. 
+\emoji{bulb} Simplifying and deconstructing the process of requesting financial support, especially for 
+people who lack the time and resources (e.g. because they have family responsibilities) might greatly improve the probability of minoritized participants applying for the much-needed support. 
 Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups.
-Whenever possible, facilitate ways around money transfer (e.g.
+Whenever possible, facilitate alternate ways for money transfer (e.g.
 book flights, hotel reservations, waive conference fees).
@@
 A conference is a critical community-building space, and can foster change in future events and the community itself.
-The organizers of a conference edition and members of selection committees should do their best to pass their knowledge on best practices and problems encountered to future organizers.
+The organizers of a conference edition and members of selection committees should do their best to pass along their knowledge on best practices and problems encountered to future organizers.
 
@@
 % Rocío: main idea: assess how inclusive was the conference
-Assess whether equity and inclusion goals were met during the conference. 
-You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
-It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and others (e.g.
+As conference organizers, one of the key post-conference tasks is to analyze conference attendee, participant, and organizer data including their experiences.
+It is essential to assess whether equity and inclusion goals were met during the conference, as a result of the steps taken.
+\emoji{bulb} You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
+It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and participants (e.g.
 countries of origin, or preferred language).
 Other indicators may need a special survey asking for additional information and feedback about the conference. 
-In the survey, providing options for open answers will allow people to express their opinions freely, and may give you more insights into how/why your inclusion practices worked or did not work. 
+In the survey, providing options for open answers will allow people to express their opinions freely, and may give you more insights into how/why your inclusion practices worked or not. 
 % Do not expect everything to be perfectly inclusive, or to encompass all the dimensions of diversity at once. %You're part of a process and contributing to structural change.
@@
 % Pass the information
-Document your processes, outcomes, and lessons learned in blog posts, reports, and presentations. 
-Make sure to share this information with future organizers, as well as the contact information of the people you worked with—if they agree to it—, to give them a starting point to improve inclusion in the following editions of the conference.
-If you are part of a stable meetings committee, encourage organizers to follow these rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria.
+%\emoji{bulb} Practical tips
+Document your processes, outcomes, and lessons learned through blog posts, reports, and presentations. 
+Make sure you share this information with future organizers, as well as the contact information of the people you worked with (if they agree to it), to give them a good well-informed baseline to improve inclusion in the following editions of the conference.
+If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria.
 %something about the past - shoulder of people before, federico

### 2022-01-20 plos_latex_template.tex
edited Rules 6–7 ~ @jananiravi

@@
 % Be more direct
-Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}.
+Unfortunately, even today, conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}.
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
 Including people with disabilities in the organizing team can have a large impact from the beginning (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making \cite{irishIncreasingParticipationUsing2020}.
-As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference.
-If you do not consider accessibility from the conference inception, it is highly likely that you will be better off without trying to patch accessibility at the last minute.
-Key decisions in this respect are hard to correct.
+As with other aspects of inclusion, dealing with accessibility at the last minute is unlikely to yield much success.
+There are several faces of accessibility that organizers must address and accommodate.
+Therefore, accessibility must be considered right from the conference inception phase, since it's difficult to course correct at later stages.
 
-If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities (e.g.
+If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live translation, to name a few.
+In addition to this, the organizers should proactively take care of invisible disabilities (e.g.
 dyslexia, anxiety, ADHD, autism).
@@
 2021 published accessibility guidelines for every format of the conference, a blog post with recommendations, and two post-conference blog posts regarding the whole accessibility experience.
-The participation of a blind person in the organizing team was key to several of these decisions.
+The participation of a blind person in the organizing team was key to the successful implementation of several of these decisions.
 Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.
@@
 Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
-For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
-Accessibility efforts should be explicitly displayed in the official communication channels and inform potential attendees with disabilities a point of contact to solve any questions in this respect.
-Clear communication about the lack of accessibility of some spaces or activities is also highly recommended, to give a clear notion of the conference experience (see \textbf{Rule 8} about communication strategy).
+For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection.
+Accessibility efforts should be explicitly displayed in the official communication channels and potential attendees with disabilities must be provided with a point of contact to help clarify any questions.
+Clear communication about the lack of accessibility for some spaces or activities is also highly recommended, to give a clear notion of the conference experience and accommodation (see \textbf{Rule 8} about communication strategy).
 
@@
 While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.%RJ: Could cite Amano's paper here.
-Non-native English speakers could miss opportunities to attend or actively participate in conferences (e.g. asking questions or participating in discussions),
+Non-native English speakers may miss opportunities to attend or actively participate in conferences (e.g. asking questions or partaking in discussions),
 and conferences may in turn miss innovative contributions.
@@
 This may be done at different levels.
-Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.
-That way, the abstract will be judged primarily by the quality or relevance of the work instead of the English grammar.
+\emoji{bulb} Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.
+That way, the abstract will be judged primarily by the quality or relevance of the work instead of English proficiency.
 For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities) follow the presentations.
-Whenever possible, identify other key languages for the conference and provide translated captions or live interpretation into these key languages, including multilingual Q \& A sessions.
-The following step would be having sessions and events in languages other than English, both regarding the technical content of the conference and the social and networking aspects.
-Networking sessions in languages other than English can include captions, interpretation or none of them; the latter has the benefit of not requiring additional technical adjustments.
-Keynotes and regular talks should always be accompanied by captions or live interpretation to English.
-Promote the attendance to presentations in languages other than English as key components of your schedule and announce when captions are available. 
-Remind your native English speaking audience to be respectful of accents and mistakes in English.
+Whenever possible, identify other prominent languages for the conference and provide translated captions or live interpretation into these key languages, including multilingual Q \& A sessions.
+Another way to engage non-native English participants is to host sessions and events in languages other than English, both technical and social/networking aspects.
+Networking sessions in other languages can include captions, interpretation or none of them; the latter has the benefit of not requiring additional technical adjustments.
+Keynotes and regular talks should always be accompanied by captions or live translation to English.
+Promote the attendance to presentations in languages other than English as key components of your schedule and announce when captions are made available. 
+To ensure inclusion without language barriers, remind your native English-speaking audience to be respectful of accents and possible grammatical mistakes, and embrace the backgrounds of non-English speakers while focusing on the core content.
 

### 2022-01-24 plos_latex_template.tex
Version sent to coauthors (w changes by @jananiravi, @andreasancheztapia, @rociojoo)

--- 2022-01-19 plos_latex_template.tex
@@
 
-\newcommand{\fede}[1]{\textcolor{ForestGreen}{Fede: #1}}
-\newcommand{\rocio}[1]{\textcolor{MidnightBlue}{Rocio: #1}}
-\newcommand{\doro}[1]{\textcolor{RedViolet}{Dorothea: #1}}
-\newcommand{\jani}[1]{\textcolor{YellowOrange}{Janina: #1}}
-\newcommand{\la}[1]{\textcolor{RawSienna}{Laura: #1}}
-\newcommand{\as}[1]{\textcolor{Violet}{Andrea: #1}}
-\newcommand{\rita}[1]{\textcolor{CarnationPink}{Rita: #1}}
-
 \vspace*{0.2in}
@@
 \\
-Rocío Joo\textsuperscript{1*}, %,2\Yinyang},
-Andrea Sánchez-Tapia\textsuperscript{2},
+Rocío Joo\textsuperscript{1*\Yinyang}, 
+Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
 Sara Mortara\textsuperscript{3},
@@
 Laura Ación\textsuperscript{10},
-Juan Pablo Narváez\textsuperscript{11},
+Juan Pablo Narváez-Gómez\textsuperscript{11},
 Marcela Alfaro Córdoba\textsuperscript{12},
@@
 Adithi Upadhya\textsuperscript{16},
-Joselyn Chávez\textsuperscript{17\Yinyang},
-Anicet Ebou\textsuperscript{18\Yinyang},
-Silvia Canelón\textsuperscript{19\Yinyang},
+Joselyn Chávez\textsuperscript{17},
+Anicet Ebou\textsuperscript{18},
+Silvia Canelón\textsuperscript{19},
 Janani Ravi\textsuperscript{20}
@@
 \\
-\textbf{5} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{5} Department of Statistics, University of Warwick, Coventry, UK
 \\
@@
 \\
-\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{16} ILK Labs/Bengaluru, India - 560046
 \\
@@
 \\
-\textbf{18} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{18} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
 \\
@@
 \\
-\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{20} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
 \\
@@
 % Primary Equal Contribution Note
-\Yinyang These authors contributed equally to this work.
+\Yinyang These authors share first authorship of this work.
 
@@
 
-The authors of this article recently organized useR!
-2021, the annual global R conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector.
-To embrace the constraints posed by the pandemic, we hosted a 100\% virtual conference.
-We worked on building a high-quality virtual and global conference by creating a kind, inclusive, accessible, and welcoming environment for everyone.
-This article outlines our most important lessons learned abstracted as `10 simple rules to host an inclusive conference'.
-These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules seamlessly translate to regional events.
-We have designed them for wide adoption to enable and ensure that the scientific and technical conferences are as accessible and inclusive as they can be.
+Conferences are spaces to meet and network within an academic or technical field, to learn about new advances and share our work.
+They can help define career paths and create long-lasting collaborations and opportunities.
+However, these opportunities are not equal for all. 
+This article introduces 10 simple rules to host an inclusive conference. 
+The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
+Coming from different backgrounds, career stages and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
+The rules result from our lessons learned before, during, and after the organization of the conference. 
+They have been mainly written for potential organizers or selection committees of conferences, and contain multiple practical tips so that they can help diverse types of scientific and technological conferences to become as accessible and inclusive as they can and should be.
 
@@
 
-% Rocío: Main message of the paragraph: Conferences are great, but for some
+% Conferences are great, but only for some
 Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
-A good conference experience can be 
+A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
 However, opportunities for participating in conferences are not equal for all. 
-Many academic and tech conferences have become spaces that reproduce systemic inequalities and privileges, i.e.
-advantages given by society to people from some groups, like being white, male, from a rich country, English-native speaker, with no physical disabilities \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
-% However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
+Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from rich countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
+We had the opportunity to organize a virtual and global statistical computing conference, useR!
+2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
+Coming from different backgrounds, career stages and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
 
-This article proposes 10 simple rules to host inclusive conferences.
-They are organized in three sections: \textbf{the pillars}, fundamental rules with key elements to conceive the work on diversity and inclusion in any conference (embracing diversity, creating a welcoming environment, and having a diverse organizing team); \textbf{design rules}, concerning diverse elements in the design of the conference and operational tasks (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices); and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion.
-These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
-We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
+% Roadmap of the paper
+The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
+They are organized in three sections: the pillars, design rules, and the continuity rule (\textbf{Fig. \ref{fig:diagram}}).
+\textbf{The pillars} comprise three fundamental rules with key elements to conceive the work on diversity and inclusion in any conference. 
+\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
+\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
+\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
+Six \textbf{design rules} focus on weaving inclusion into the conference design process.
+In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
+\textbf{Rule 5} provides advice for organizing inclusive online and hybrid conferences.
+\textbf{Rule 6} focuses on accessibility practices to include people with disabilities. 
+In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
+\textbf{Rule 8} provides tips for developing an inclusive communication strategy. 
+In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
+Finally, 
+a \textbf{continuity Rule} to make the conference part of a long-term commitment to inclusion.
 
-The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference.
-We come from different backgrounds, career stages and even continents.
-The rules result from our lessons learned from diverse perspectives and expertise. 
-Even though these rules were designed based on our experience with a global event, they can be easily transferred or scaled to events with regional or local scopes.
-These rules will greatly benefit people who are part of meeting committees that oversee the site/location selection process, or that coordinate with the local organizers of conferences.
-The rules will also help organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
-We have designed them for wide adoption to enable and ensure that the scientific and technical conferences are as accessible and inclusive as they can be.
+We believe that these rules will greatly benefit organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
+The rules will also help people who are part of meeting committees that oversee the site/location selection process, or that coordinate with the local organizers of conferences. 
+We have designed them for wide adoption to help scientific and technical conferences to become as accessible and inclusive as they can and should be.
 
-% This message is not new. 
-% Barriers such as unattainable registration costs, sexism, prejudice against disable people (ableism), among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020},
+% refs for past works in case we add them in line 311.
 % %timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018, marks2021meeting
-% and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021},
-% with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/
-% %By inclusion, we mean creating conferences organized by and dedicated to a group of attendees that is representative of the whole population.
-This means creating a welcoming, respectful, and safe conference that tries to counteract systemic oppression.
-
-% % Rocío: Main message of the paragraph: Roadmap of the paper
-% This article focuses on practices for hosting inclusive conferences.
-% The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
-% The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
-% These tips \rita{tips looks to me too informal for an article, maybe used the word `suggestions` or `advice`?} stem from the authors' experience of organizing useR!
-2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}.
-% We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.
-% Here, we share the lessons learned within the time organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
-% The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
-% %sara: still need to better format figure size and get the alt text right :P
-
-% Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
-% The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
-% Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of counteracting implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs.
-% The third group includes rules 5 to 9.
-These rules are about components of the conference that should be carefully planned for: an online component, accessibility for people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion.
-% These rules apply equally to conferences where the expected audience is academic, from the industry, or a mixed group.
-While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
-% \fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}
+% some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, numfocus_discover_2021},
 
@@
 \centering
-%sara: without success i'm trying to use pdfcomment package and pdftooltip to make the alt-text. any help?
-%ast: R/ Unfortunately, support for adding accessibility metadata to LaTeX documents is limited.
-You will probably need to use Adobe Acrobat to add missing accessibility metadata to your PDF file.https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#latex %ast: let's try not to forget this when submitting - some edits to the PDF may be needed to add alt-text to the final version
+%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file.
+let's try not to forget this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
 \pdftooltip{
 \includegraphics[width=\textwidth]{figs/rules.png}}{
-%sara: suggestion for alt text:
-%Diagram of how the 10 rules are organized into three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9).
-The diagram has five rows.
-Each line contains rectangular boxes with the rule number and a short title for each rule.
-Each rule box is colored based on their group (grounding rules are grey, people-related rules are green and design rules are pink).
-First row is for the grounding rules 1.
-Multidimensional diversity and 2.
-Safe and inclusive environment in grey boxes.
-Second row is for people-related rules 3.
-Organizing team and 4.
-Unbias in green boxes.
-Third row is for the design rules 5.
-Online component, 6.
-Accessibility for disabilities, 7.
-Language inclusivity in pink boxes.
-The fourth row is for the design rules 8.
-Communication and 9.
-Financial support and budget in pink boxes.
-The last row is for the grounding rule 10.
-Diversity and inclusion as a process in grey boxes.
+%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: pillar rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10).
+The overall diagram depicts an arrow from left to right which contains all the rules.
+From left to right, "the pillars" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team".
+Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial".
+To the right, and forming the head of the arrow, the Tenth Rule is under the label "Continuity" and consists of a blue bubble "10 Assess and pass along".
 }
-\caption{Diagram of the rules organized in three groups: the pillars (rules 1, 2, and 3), design rules (rules 4 to 9), and the process (rule 10).}
+\caption{Schematic diagram of the rules organized in three groups: the pillars (Rules 1, 2, and 3), design rules (Rules 4 to 9), and continuity (Rule 10).}
 \label{fig:diagram}
@@
 
-\section*{I. The pillars}
-
-To organize a more inclusive conference, it is necessary to identify upfront which groups of people are being excluded, the barriers that keep them from active partaking, and strategies to overcome them.
-\textbf{Rule 1} is about embracing all dimensions of diversity and prioritizing the participation of groups that are often marginalized.
-However, increasing diversity alone is not enough. \textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the conference attendees. 
-Inclusion should start from the organizing team. \textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team, and provides tips on work dynamics.
-
-\subsection*{Rule 1: Embrace all dimensions of diversity}
+\subsection*{Rule 1: Set a vision for diversity and inclusion}
 \label{rule_diversity}
 
-% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
 Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
 Human diversity should be celebrated and respected in every way. 
-Unfortunately, we live in a world with implicit hierarchies along these axes. 
+Nonetheless, we live in a world with implicit hierarchies along these axes. 
 Some statuses (e.g.
-cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built.
-People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; some of them have to deal with several of these on a daily basis.
-% While no isolated initiative can drastically alter xx, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles.
-Therefore, recognizing this systematic exclusion, particularly in our fields and in our scientific or professional communities, is paramount to inclusion, and will help identify the most marginalized and discriminated subgroups \cite{timperleyHeMoanaPukepuke2020}.
-% These are the groups that warrant deliberate efforts that ensure fair inclusion.
-As a starting point, the meeting organizers could identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (e.g.
-explicit discrimination, participation fees).
-This analysis should guide the vision of diversity and inclusion for your conference and help design strategies to achieve it. 
-Express this vision for your conference in a diversity statement. 
-Next, use this vision and the set strategies to define measurable indicators and goals of diversity and inclusion (e.g.
-a gender distribution of your speakers that is representative of the general population, the presence of racialized people (e.g.
-Black people) among the organizers, speakers, and attendees, or participation from key geographic regions).
-These numbers should guide you and help achieve your vision.
-
-% realistic but bold (back?)
-% be aware and promote awareness
-% Rocío: more effort towards including the more excluded
-% This may translate into different actions depending on your field, region, or community
+being cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built.
+People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways.
+As a result, some groups of people have been systematically excluded from or only partially included in academia and scientific and professional circles.
+Recognizing this systematic exclusion is paramount to inclusion.
+Assess how it translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
+Define measurable indicators for these goals (e.g.
+a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.
+The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.
 
@@
 
-% Rocío: main idea: Inclusion for everyone
-While it is common to think about representation in terms of some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a genuinely inclusive environment means taking care of all aspects of diversity.
-%Inclusivity should lie at the heart of setting the tone and meeting space for any meeting as we move forward in the 21st century!
-% Edit and make more specific (below)
-%\emoji{bulb} \textbf{Practical tips:} 
-Here are a few ways in which conference organizers can actively enforce inclusiveness, such as i) Promote respect of pronouns, ii) schedule events exclusive for diverse groups (e.g.
-LGBTQIA+-friendly spaces), iii) if the conference has an in-person component, pre-assign spaces for religious practices, lactation, and child care, iv) include menu options that account for diverse dietary and religious requirements.
-These are just a few easy-to-implement examples with a significant impact on including key sub-groups that otherwise find it challenging (see \cite{noauthor_discover2021} for some great advice on `low-hanging fruits').
-Importantly, by being proactive in creating such a welcoming space, the organizers can be respectful of attendees' privacy and avoid requiring anyone to disclose personal information, such as revealing a disability, sexual orientation, gender identity, or mental health issues, to name a few.
+% Inclusion for everyone
+% Explain why we need this
+Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community.
+Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees.
+This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}).
+For example, promote the use and respect of preferred pronouns and commit to respect all genders and sexual orientations.
+If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care.
+Include menu options that account for diverse dietary and religious requirements.
+Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and  great advice on `low-hanging fruits').
+Open the schedule for events organized by thematic groups and communities (e.g.
+taskforce meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.
 
-The organizers should promote a supportive and respectful environment, where all attendants can participate without fear of rejection or isolation.
-This includes promoting the idea that anyone can act in defense of other attendants in case anything unpleasant, incorrect, of offensive happens.
-Instead of looking away, act as "active bystanders" and intervene to keep the community safe.
-% Rocío: main idea: Code of conduct for a safe place %i think the transition here should leave the two concepts closer. not however. 
-However, the most important step that organizers should take to ensure the safety of the community is
-Adopting a Code of Conduct (CoC) and gathering a diverse team to enforce it \cite{favaroYourScienceConference2016}.
-The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}.
-The CoC team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize protocols on how to respond correspondingly.
-This should be done during the early phases of the conference planning because CoC violations can happen before the conference too (e.g.
+% A culture of support (active bystandership) and code of conduct for a safe place 
+Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, incorrect, or offensive happens.
+This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
+However, the most important step that you should take to ensure the safety of the community is
+adopting a code of conduct (coc) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
+The coc is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
+The coc response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community.
+The team should be prepared early during conference planning because coc violations can happen before the conference too (e.g.
 social media, organization meetings).
-Since people from marginalized groups are more likely to be victims of Code of Conduct violations and find their claims dismissed due to inadvertent power dynamics, assembling a diverse CoC team will be essential to appropriately deal with discrimination and harassment issues and the power dynamics underneath them with respect and sensitivity.
-A well-drafted CoC should account for  different cultures and geographical and cultural considerations.
-We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC team.
+We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong coc team.
 
@@
 
-% Rocío: Idea of this paragraph: build a diverse team, and representation
-
-A genuinely inclusive conference can only be organized by an inclusive and diverse team in charge of the conference decision-making.
-Organizers should be a team with people from different regions, genders, ethnicities, socioeconomic statuses, and other aspects of diversity.
+% Build a diverse team, and representation
+A genuinely inclusive conference can only be organized by an inclusive and diverse team.
+Organizers should comprise a team with people from as many different backgrounds as possible regarding regions, genders, ethnicities, career stages, and other aspects of diversity.
+People with disabilities often say: `Nothing about us without us' \cite{charlton_nothing_1998, werner_nothing_1998}; the same holds for other dimensions of diversity. 
 Having people from diverse backgrounds in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}.
-In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule 2} about creating a safe and welcoming environment).
-
-% Rocío: Idea of this paragraph: non-transferability and do not restrict people
-People with disabilities often say: `Nothing about us without us' [ref]; the same holds for other dimensions of diversity.
-The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
-% People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
+The experience of people in marginalized groups
+is especially important; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
+In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
 It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
-every member should be granted the freedom to choose which areas of the conference they want to work on, in addition to contributing to decisions involving marginalized communities.
+every member should have the freedom to choose which areas of the conference they want to contribute to.
 
 % Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
-
-%\emoji{bulb} \textbf{Practical tips:} 
-If you have already started assembling an organizing team, check for gaps in its composition and find representatives accordingly. 
+If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
 The regional and local communities, groups, or associations in your field are good sources to tap into. 
-Furthermore, self-nomination and voting do not always work as reliable mechanisms to fill leadership positions.
-Instead, it helps to nominate members directly and offer them leading positions in the organization team, especially those that would be typically occupied by people from privileged groups.
-Build an environment that encourages healthy conflict and constructive criticism, one in which every person can express their opinion and pay special attention to people from systematically excluded groups.
-Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (e.g.
-public recognition on social media, on the conference website, or during the conference, ideally whatever works best for the contributing member).
-Furthermore, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.
-
-% % Rocío: might go to the next rule
-% Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).
-
-\section*{II. Design Rules}
-
-The next set of rules concern diverse elements in the design of the conference and operational tasks. They focus on weaving inclusion into the conference design process.
-In \textbf{Rule 4}, we introduce ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
-\textbf{Rule 5} describes the advantages of online conferences by reducing barriers for participation and provides advice for organizing online and hybrid conferences.
-\textbf{Rule 6} focuses on best practices to make all conference platforms and spaces accessible to people with disabilities. 
-In \textbf{Rule 7}, we advocate for more language-inclusive conferences and provide suggestions to encourage and ensure full participation of non-native English speakers. 
-\textbf{Rule 8} provides tips for developing an inclusive media strategy. 
-In \textbf{Rule 9}, we address budgeting for inclusive practices, and helping participants with affordable registration costs, scholarships, and other forms of financial support.
+When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
+People from underrepresented groups sometimes refrain to nominate themselves while people in dominant groups tend to do it more easily. 
+Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
+Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
+Furthermore, tasks such as receiving and responding to code of conduct reports can be emotionally intense work and should be additionally rewarded.
+And most importantly, take care of your team and their well-being!
+Check on them regularly and make sure that everyone is comfortable.
+This represents a great amount of work and you will need to support each other in the long run.
 
@@
 \label{rule_unbias}
-%idea of paragraph 1: our lists are biased
 
-When choosing or inviting people for visible and valued roles in the conference –– such as keynote speakers, program/scientific committee, session chairs –– it is likely that there will not be much diversity in the first pool of names.
-This is likely due to our inherent biases resulting from existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
-Rather than deter us, this should encourage us to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not stay with people in privileged groups (e.g.
-avoid `manels' and all-white panels; \cite{else_how_2019}; see \textbf{Rule 8}).
-\emoji{bulb} It would help to ensure that the composition of all these groups is balanced across topics and areas with representation across regions, gender, race, and career stage, among other aspects.
+The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered.
+It is important to counteract biases in all of these aspects.
+When choosing or inviting people for visible and valued roles in the conference it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
+Rather than deter you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges (e.g.
+`manels' and all-white panels \cite{else_how_2019}).
 
-% our committees are biased and unconscious bias exists
-%Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field.
+Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
+To mitigate this self-selection bias, 
+some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g.
+the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
+If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
 
-The final list of abstracts making it into the program could also be biased in both submission and acceptance processes. 
-For instance, first-time conference participants, especially for the flagship/prestigious meetings, may not feel confident to submit their work.
-\emoji{bulb} To counteract this self-selection bias, promote the submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
-Some of these communities may have pre-submission mechanisms to help their members receive feedback for their submissions (e.g.
-the Latin-R and R-Ladies communities have Slack channels to help their members improve their conference abstracts).
-On the other hand, the review process may be biased because selection committees — even diverse ones — can be subjected to unconscious or implicit bias.
-Unconscious bias happens when we interpret information such as names of the authors, assuming their origin, ethnicity, or their institution of origin, and we attach a positive or negative value to their work according to the stereotypes.
-Unfortunately, nobody is free from implicit bias \cite{ross_everyday_2020}.
-Selection committees need to be gently reminded of such biases with evaluation guidelines (the implicit association test can help identifying them \cite{greenwald_measuring_nodate}) and the organizers need to
-identify the most appropriate strategy to counteract them, either with double blind, double open, or single blind reviews \cite{numfocus_discover_2021}.
+The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.
+Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019}.
+Evaluate the plausibility of anonymizing all of the submission information to
+define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
+Define clear evaluation criteria and share them with the reviewers and authors.
+Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe; even well-meaning and carefully crafted processes can be prone to reproduce the biases observed in academia and society.
 
-Double blind reviews can help counteract implicit bias only if the identifying information is completely anonymized in the submission. 
-Double open reviews, though more transparent, may still be subjected to implicit bias, and reviewers from minoritized groups might feel additional pressure because the reviews are signed.
-In this likely case, and in fields where it is not possible to guarantee anonymity of the authors during the selection process (e.g.
-when non-anonymized GitHub repositories need to be examined), single blind reviews are the only remaining option.
+In many academic or technological fields, typically competitive activities or areas—such as scientific publications or software development—are more valued than those focused on sharing and cooperation, like community building, teaching or mentoring.
+The latter activities are equally, if not more, relevant and challenging, and are usually performed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020, light_gender_2022}.
+When planning the program for the conference, consider giving visibility to the whole range of activities and practitioners that contribute to the field by proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials.
 
-%\emoji{bulb} \textbf{Practical tips:} 
-In any case, design unambiguous evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g.
-based on their number of publications or talks), and share them with the authors and reviewers.
-Advertising your practices to mitigate bias in proposal selection may also encourage people from diverse groups to submit their work, thus counteracting self-selection bias.
-
-% \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}
-
-% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
-Implicit bias also manifests in the weightage given to certain products/activities; e.g. scientific publication and development of software are often rated higher than community building or teaching. 
-The latter roles are equally, if not more, challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
-Likewise, research about these topics is undervalued relative to other areas of knowledge production. 
-Counteracting bias actively means innovating the program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, and giving visibility to the whole range of activities and practitioners that contribute to the overall range and quality of the conference.
-
-\subsection*{Rule 5: Have a strong online component} 
+\subsection*{Rule 5: Design a strong online component} 
 \label{rule_online}
 
-% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
-Virtual conferences are more inclusive, removing barriers like costs of registration, transport, accommodation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}.
-After experiencing virtual conferences mostly due to the COVID pandemic, there is a widespread call to retain an online component in conferences, either hosting conferences completely virtual or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
-It is important to note that while being more inclusive, an online component does not guarantee inclusion and accessibility for everyone.
-Key decisions must be made early on during the planning phase \cite{levitisCenteringInclusivityDesign2021}.
-For instance, the chosen conference platforms (e.g.
-for submission, registration, chatting system, live streaming, etc.) must account for accessibility to accommodate screen-reader users, geopolitical restrictions, and bandwidth requirements.
-\emoji{bulb} The organizers should also decide how to incorporate geographic regions within the schedule of the conference by asking for pre-recorded videos, recording sessions, and sharing them during the conference.
-Some possibilities for scheduling are: having different days dedicated to different time zones; having each day with the same conference hours, and scheduling time slots that are friendly to different time zones; or having time-zone hubs in each day and repeating the events in each (live for one/few and recorded versions for others).
-A flexible schedule and recorded material will allow many more attendees to partake in the conference, e.g.
-people with schedules conflicting with the live conference, those who cannot take off, or others who are caretakers.
+After experiencing virtual conferences mostly due to the COVID-19 pandemic, there have been multiple calls to retain an online component in conferences, either as completely virtual events or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
+Virtual conferences remove barriers for inclusion like costs of participation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}.
+Make sure that you are not creating additional barriers when choosing conference platforms, schedule, and activities. 
+Check every platform's screen reader accessibility (see \textbf{Rule 6}), bandwidth requirements (a high-speed internet prerequisite could exclude attendees from many regions), and geopolitical restrictions (some services are not available in all countries).
+Be mindful of the different time zones and time constraints of the potential attendees and create flexible schedules.
+Avoid creating a strictly synchronous live program, ask for pre-recorded talks, record sessions, and make them available during the conference.
 
-% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
-Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to in-person interactions at social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}.
-However, there is evidence that virtual communication can make people from minoritized groups feel more included, thus participate and contribute more \cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}.
+If you are organizing a hybrid event, decide on the relative importance that the online and in-person components will take \cite{bajpai_towards_2021}.
+The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending the conference.
+The audience, chairs, and presenters should all be able to interact regardless of their in-person or remote status. 
+To best integrate in-person and remote attendees, all questions and comments should pass through a unique online system. 
+The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with them.
 
-%\emoji{bulb} Practical tips:
-Invest time in creating opportunities to meet and bond virtually that can appeal to people with different backgrounds, disabilities, preferences, and remembering that no single networking activity will ever serve the whole community.
-Try new activities centered on different groups of people outside the mainstream of your conference (e.g.
-\textbf{ABC}); if you think of activities that reach specific groups, preferably co-led by community leaders, these sessions can be more productive, successful, and inclusive.
-Some examples are newbies sessions for first-timers, social mixers led by specific subgroups or communities like LGBTQAI+, or leisure activities that reunite subgroups with the same interests: virtual art exhibitions, yoga sessions, movie watch parties, trivia quizzes, shared language, etc.
-Offer activities that require voice or video interactions, and others that involve only written chat, and make all participation optional and voluntary.
-
-If you are organizing a hybrid event instead, you should decide on the relative importance that the online and in-person components will take.
-The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the preferred format \cite{bajpai_towards_2021}.
-
-\emoji{bulb} For this format, the audience, chairs, and presenters -- of talks, posters, and workshops -- could be either in-person or remote, and they should all be able to interact. 
-To best include both kinds of attendees, all questions and comments should pass through an online system, or have an pre-designated point person who enables this cross-talk. 
-The physical venue would have to include multiple ways to connect with online folks such as tablets on the meeting venue, cameras and microphones that allow online attendants to follow who is speaking at the in-person space.
-This requires the venue to have stable and secure internet connection for the in-person participants to join the online folks. 
-Having a hybrid format also means accounting for the fact that the venue does not need to host all the attendees.
-You should estimate in advance how many in-person attendees to expect, and ask them early on during registration or pre-registration.
-This could save you time and money when looking for a suitable venue.
-When planning for networking activities in hybrid events, use the same principles outlined above and offer various activities engaging different groups.
-It is understandable that a few of these activities could be only in-person, but to promote connections between both groups of participants, there should be a fair proportion of activities connecting in-person and online attendees.
+Online networking and socializing can make people from minoritized groups feel more included, thus participate and contribute more, than when meeting face-to-face for the first time \cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}.
+Take advantage of the online component to have a broader range of social activities that can appeal to people with different backgrounds and preferences.
+Offer some activities that require voice or video interactions, and others that involve only written chat.
+You can adapt in-person activities for online and hybrid settings or create new activities. 
+Virtual art exhibitions, yoga sessions, movie watch parties, trivia quizzes, and virtual city tours are just some examples.
+When planning for networking activities in hybrid events, try to
+promote connections between both groups of participants.
+It is understandable that a few of these activities would only be in-person, but 
+there should be a fair proportion of activities connecting in-person and online attendees.
 
@@
 
-% Be more direct
 Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}.
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
-Including people with disabilities in the organizing team can have a large impact from the beginning (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making \cite{irishIncreasingParticipationUsing2020}.
-As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference.
-If you do not consider accessibility from the conference inception, it is highly likely that you will be better off without trying to patch accessibility at the last minute.
-Key decisions in this respect are hard to correct.
+Including people with disabilities in the organizing team from the start can have a large impact (see \textbf{Rule 3}), since planning for accessibility requires time, experience, early decision-making, and inaccessible features are difficult to correct at later stages \cite{irishIncreasingParticipationUsing2020}.
 
-If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features.
-In addition to this, the organizers should take care proactively of invisible disabilities (e.g.
-dyslexia, anxiety, ADHD, autism).
-This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.
+If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and language interpretation, to name a few.
+Invisible disabilities should be accounted for proactively (e.g.
+dyslexia, anxiety, ADHD, autism), for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.
 
-Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly, keyboard accessible, and have alternative text.
-Every video should have good-quality (not automated) captions and a transcript; live presentations should have good-quality captions too.
-Provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have.
-For instance, useR!
-2021 published accessibility guidelines for every format of the conference, a blog post with recommendations, and two post-conference blog posts regarding the whole accessibility experience.
-The participation of a blind person in the organizing team was key to several of these decisions.
-Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.
+Regardless of the conference format, all platforms (website, registration, abstract submission, chat system, conference administration tools, live streaming) should be screen-reader friendly and keyboard accessible.
+All images used in digital conference spaces should have alternative text and use color-blind safe palettes.
+If there is video streaming, it should have good-quality (not automated) captions and a transcript; live presentations should have good-quality live captions too.
+If the conference has a regional scope, sign-language interpreters can be a better option than captions.
 
-Accessibility practices should cover social events and networking, which should include activities that do not restrict participation based on body type or ability.
+Encourage the preparation of accessible slides and presentations by providing accessibility guidelines and be available for any questions that presenters and attendees may have; see \cite{sanchez-tapia_user_2021} for an example of accessibility guidelines and \cite{chavez_preparing_2021, sanchez-tapia_making_2021, joo_how_2021} for further accessibility recommendations.
+Social events and networking should also have accessibility features like captioning or sign language interpretation, and include activities that do not restrict participation based on body type or ability.
 
-Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
-For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc.
-Accessibility efforts should be explicitly displayed in the official communication channels and inform potential attendees with disabilities a point of contact to solve any questions in this respect.
-Clear communication about the lack of accessibility of some spaces or activities is also highly recommended, to give a clear notion of the conference experience (see \textbf{Rule 8} about communication strategy).
+Importantly, accessibility practices are inclusive not only for people with disabilities but for everyone.
+For instance, captions are helpful for non-native speakers and having slides available for download helps attendees with low bandwidth connection.
 
-\subsection*{Rule 7: Include languages other than English}
+\subsection*{Rule 7: Make room for the linguistic diversity of your community}
 \label{rule_language}
@@
 English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpages, and communications. 
-While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.%RJ: Could cite Amano's paper here.
-Non-native English speakers could miss opportunities to attend or actively participate in conferences (e.g. asking questions or participating in discussions),
+While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege \cite{amanoTenTipsOvercoming2021}.
+Non-native English speakers may miss opportunities to attend or to actively participate in conferences (e.g. presenting, asking questions, or partaking in discussions),
 and conferences may in turn miss innovative contributions.
@@
 % Rocío: main idea: make conferences more linguistically inclusive
-Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (\textbf{Rule 2} about a welcoming environment for all).
-This may be done at different levels.
+Create an inclusive environment by encouraging the full participation of non-native English speakers. This may be done in multiple ways.
 Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.
-That way, the abstract will be judged primarily by the quality or relevance of the work instead of the English grammar.
-For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities) follow the presentations.
-Whenever possible, identify other key languages for the conference and provide translated captions or live interpretation into these key languages, including multilingual Q \& A sessions.
-The following step would be having sessions and events in languages other than English, both regarding the technical content of the conference and the social and networking aspects.
-Networking sessions in languages other than English can include captions, interpretation or none of them; the latter has the benefit of not requiring additional technical adjustments.
-Keynotes and regular talks should always be accompanied by captions or live interpretation to English.
-Promote the attendance to presentations in languages other than English as key components of your schedule and announce when captions are available. 
-Remind your native English speaking audience to be respectful of accents and mistakes in English.
+That way, the abstract will be judged primarily by the quality or relevance of the work instead of English proficiency.
+For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities, see \textbf{Rule 6}) follow the presentations.
+Whenever possible, identify other prominent languages for the conference and provide translated captions or live language interpretation into these key languages, including multilingual Q\&A sessions.
+Another way to engage non-native English participants is to host sessions and events in languages other than English. 
+Promote these sessions among all conference attendees and announce if captions will be available; all sessions, regardless of the language, should be given the same relevance.
+Importantly, consideration and respect of accents or linguistic mistakes can make a significant difference in the conference experience of non-native English speakers.
+Kindly remind your native English-speaking audience to be mindful of that.
+
+This rule also applies to cases when English is not the official language of the conference (e.g.
+a Latin American conference with Spanish as the official or most popular language).
+No language should be a barrier for inclusion.
 
@@
 
-\subsection*{Rule 8: Convey the welcoming spirit in your communication strategy}
+\subsection*{Rule 8: Build an inclusive communication strategy}
 % media
@@
 
-%Rocío: Main idea: Include people in your communication 
-The communication strategy of your conference should aim to increase the number and diversity of participants, and be inclusive by design.
+The communication strategy of your conference should aim to reach broader audiences, express the commitment with diversity and inclusion, and the welcoming spirit of the conference.
 Actively reach out and promote the conference to communities that have been systematically excluded. 
-Publish the diversity statement, Code of Conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9} about financial resources) to emphasize that everybody is welcome to join the event.
-Try to develop creative ways to show that everyone is seen, respected, and welcome.
+Promote the abstract submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
+The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g.
+Twitter, Facebook, LinkedIn, conference website, mailing lists).%#promotion2
+
+Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}) to communicate your practices and to hold yourself accountable for them.
+Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. 
+For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering help to navigate it, or if captions will be available for some talks but not all.
+Provide a point of contact to help clarify any questions regarding accessibility, financial support, and others.
+
+Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups–in all communications \cite{hallDesigningDiversityInclusion2019}.
+Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
+Do not expect minoritized people to teach you and accept feedback without being offended.
+Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms.
+
+Communication should be fun! Try to develop creative ways to show that everyone is seen, respected, and welcome. 
 For example, useR!
-2021 created a mascot for the conference wearing different scarves reflecting folks from various minoritized groups that the conference wanted to reach (\textbf{Fig.
-2}).
-The diverse members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g.
-Twitter, Facebook, LinkedIn, conference website, mailing lists).
-Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues.
-For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering personalized help to navigate it, or if captions will be available for some talks but not all.
-
-%Rocío: Main idea: Use inclusive language
-\emoji{bulb} Inclusive language — language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups — should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
-Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, and other terms that should be avoided. %
-Do not expect minoritized people to teach you—it's not their role—, and accept feedback without being offended.
-Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms. 
-% links, examples, this changes with the context
-%practice: create a standards in inclusive language for your community, to build collective knowledge than can be passed to
+2021 created a mascot for the conference wearing different scarves reflecting folks from communities and various minoritized groups that were part of the potential attendees (\textbf{Fig.
+\ref{fig:marmots}}).
 
@@
 }
-\caption{Margot the marmot was useR!
+\caption{Margot the marmot was the useR!
 2021 mascot.
-To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing diverse groups that are part of the large R community: (from left to right) MiR (Minorities in R), LGBTQAI+, R-Ladies, transgender people, and AfricaR.} %RJ: Need to acknowledge Francisco's contribution.
+To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly: (from left to right) scarves with MiR (Minorities in R https://mircommunity.com/), R-Ladies https://rladies.org/, AfricaR https://africa-r.org/, rainbow LGBTQIA+ and transgender colors.
+Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license}
 \label{fig:marmots}
 \end{figure}
-
-\subsection*{Rule 9: Allocate adequate financial resources to support your conference goals}
+%alt-text for the figure, to be included in the final PDF: Five marmots with colorful scarves.
+From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand.
+\subsection*{Rule 9: Allocate adequate financial resources to support inclusion goals}
 \label{rule_financial}
@@
 % Rocío: main idea: Need to budget for inclusive practices.
-Allocation of resources to support the goals of inclusion of the conference has to be intentional. 
-Estimate the costs for these practices and define your priorities in advance (e.g. paying the organizing team, Code of Conduct training, captioning).
-Additional support for attendees could also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component).
-Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
-When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion than making generic demands for funding.
+Allocation of resources to support the diversity and inclusion vision and goals has to be intentional. 
+Estimate the costs of these practices and define your priorities in advance (e.g. paying the organizing team, code of conduct training, captioning, accessible software, scholarships).
+Consider that an online conference might reduce overall costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards inclusive practices. 
+When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion than making generic requests for funding.
 
@@
 (see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
-Resources permitting, allow the possibility of a 'pay what you can' option.
+Resources permitting, enable a `pay what you can' option.
 You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}.
-%Other costs can be prohibitive-> scholarships
 Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
-Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying because they think they don't stand a chance).
+Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
 Conferences usually ask for cover letters or applications to assign these scholarships, which can be time-consuming and emotionally demanding. 
-\emoji{bulb} Simplifying and deconstructing the process of requesting financial support, especially for 
-people who lack the time and resources (e.g. because they have family responsibilities) might greatly improve the probability of minoritized participants applying for the much-needed support. 
-Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups.
+Simplifying the process of requesting financial support, especially for 
+people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
+Additional aid for attendees could also be offered—e.g. child care support or internet connection services for the virtual component. 
+Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups.
 Whenever possible, facilitate alternate ways for money transfer (e.g.
 book flights, hotel reservations, waive conference fees).
-
-\section*{III. The Process}
-
-A conference is a critical community-building space, and can foster change in future events and the community itself.
-The organizers of a conference edition and members of selection committees should do their best to pass along their knowledge on best practices and problems encountered to future organizers.
 
@@
 % Rocío: main idea: assess how inclusive was the conference
-As conference organizers, one of the key post-conference tasks is to analyze conference attendee, participant, and organizer data including their experiences.
-It is essential to assess whether equity and inclusion goals were met during the conference, as a result of the steps taken.
-\emoji{bulb} You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
+As conference organizers, one of the key post-conference tasks is to make a final assessment of the achievement of the diversity and inclusion goals. 
+You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
 It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and participants (e.g.
 countries of origin, or preferred language).
-Other indicators may need a special survey asking for additional information and feedback about the conference. 
-In the survey, providing options for open answers will allow people to express their opinions freely, and may give you more insights into how/why your inclusion practices worked or not. 
-% Do not expect everything to be perfectly inclusive, or to encompass all the dimensions of diversity at once. %You're part of a process and contributing to structural change.
+You may have goals concerning how welcomed and included participants felt during the conference, and wonder if your practices translated into real inclusion.
+Surveys, focal groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
 
 % Pass the information
-%\emoji{bulb} Practical tips
-Document your processes, outcomes, and lessons learned through blog posts, reports, and presentations. 
-Make sure you share this information with future organizers, as well as the contact information of the people you worked with (if they agree to it), to give them a good well-informed baseline to improve inclusion in the following editions of the conference.
+The feedback received from the participants and your whole assessment of the conference should be documented and shared with future organizers. 
+Gather and do your best to organize all the documentation produced during the organization process—or even after.
+Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors.
+Create a list with contact information of the people of the organizing team, as well as previous and potential speakers—if they agree to share their contact information—, and pass it along.
+For conference series, it could be a good idea to create an official document (e.g.
+a wiki, knowledge base or similar guide, see for instance \cite{sanchez-tapia_user_2021-2}) that can be used and kept up to date by future organizers.
 If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria.
-%something about the past - shoulder of people before, federico
 
@@
 Organizing a conference and implementing inclusive practices are both learning experiences.
-As conference organizers, we started with different levels of clarity about the principles and practices described here, learned many of them together during the organization phase, and learned even more when translating them into ten rules for this paper.
-We hope that they can be useful guidelines to improve diversity and inclusion in your conference.  
-And that you can adapt them, improve them, and share your lessons and experience with as many people as possible.
+As conference organizers ourselves, we started with different levels of clarity about the principles and practices described here, learned many of them together during the organization phase, and learned even more when translating them into ten rules for this paper.
+We hope that they can be useful guidelines to improve diversity and inclusion in your conference, and that you can adapt them, improve them, and share your lessons and experience with as many people as possible.
 
-Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, personal issues affecting your organizing team.
+Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, or personal issues affecting your organizing team.
 Try to account for these factors as best as you can to set realistic yet bold vision and goals of diversity and inclusion for the conference.
@@
 
-To list author contributions, we used the high-level contributor roles in the CRediT framework relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and an additional category for contributions on their own ten simple rules for an inclusive conference without looking at the first draft (i.e.
+To list author contributions, we used the high-level contributor roles in the CRediT framework that we considered relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and an additional category to acknowledge those who shared their own ten simple rules for an inclusive conference without looking at the first draft (i.e.
 sharing rules).
@@
 The authors of this piece would like to thank every single member of the organizing team of useR!
-2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for trusting us with the organization of useR!
+2021 (\url{https://user2021.r-project.org/about/global-team/}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR!
 2021 and supporting us through the process.
+Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.
+
+Heather Turner was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR!
+2021.
 

### 2022-01-25 plos_latex_template.tex
Tiny edits by @andreasancheztapia, @rociojoo + author affiliations

@@
 Matt Bannert\textsuperscript{7},
-Batool Almazrouq\textsuperscript{8},
-Liz Hare\textsuperscript{9},
-Laura Ación\textsuperscript{10},
-Juan Pablo Narváez-Gómez\textsuperscript{11},
-Marcela Alfaro Córdoba\textsuperscript{12},
-Federico Marini\textsuperscript{13},
-Rita Giordano\textsuperscript{14},
-Natalia Morandeira\textsuperscript{15},
-Adithi Upadhya\textsuperscript{16},
-Joselyn Chávez\textsuperscript{17},
-Anicet Ebou\textsuperscript{18},
-Silvia Canelón\textsuperscript{19},
-Janani Ravi\textsuperscript{20}
+Batool Almazrouq\textsuperscript{8,9},
+Liz Hare\textsuperscript{10},
+Laura Ación\textsuperscript{11},
+Juan Pablo Narváez-Gómez\textsuperscript{12},
+Marcela Alfaro Córdoba\textsuperscript{13},
+Federico Marini\textsuperscript{14},
+Rita Giordano\textsuperscript{15},
+Natalia Morandeira\textsuperscript{16},
+Adithi Upadhya\textsuperscript{17},
+Joselyn Chávez\textsuperscript{18},
+Anicet Ebou\textsuperscript{19},
+Silvia Canelón\textsuperscript{20},
+Janani Ravi\textsuperscript{21}
 %,3\textcurrency}
@@
 \\
-\textbf{8} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{8} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
 \\
-\textbf{9} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{9} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
 \\
@@
 \\
-\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{13} Department of Statistics, University of California, Santa Cruz, CA, USA
 \\
@@
 \\
-\textbf{16} ILK Labs/Bengaluru, India - 560046
+\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{17} ILK Labs/Bengaluru, India - 560046
 \\
@@
 \\
-\textbf{18} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\textbf{19} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
 \\
-\textbf{19} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
 \\
-\textbf{20} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
+\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
 \\
@@
 
-Conferences are spaces to meet and network within an academic or technical field, to learn about new advances and share our work.
+Conferences are spaces to meet and network within an academic or technical field, to learn about new advances, and share our work.
 They can help define career paths and create long-lasting collaborations and opportunities.
@@
 The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
-Coming from different backgrounds, career stages and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
+Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
 The rules result from our lessons learned before, during, and after the organization of the conference. 
-They have been mainly written for potential organizers or selection committees of conferences, and contain multiple practical tips so that they can help diverse types of scientific and technological conferences to become as accessible and inclusive as they can and should be.
+They have been mainly written for potential organizers or selection committees of conferences, and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.
 
@@
 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
-Coming from different backgrounds, career stages and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
+Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
 
@@
 The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
-They are organized in three sections: the pillars, design rules, and the continuity rule (\textbf{Fig. \ref{fig:diagram}}).
-\textbf{The pillars} comprise three fundamental rules with key elements to conceive the work on diversity and inclusion in any conference. 
+They are organized in three sections: three pillars, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
+\textbf{The pillars} comprise fundamental rules with key elements to conceive the work on diversity and inclusion in any conference. 
 \textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
@@
 \textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
-Six \textbf{design rules} focus on weaving inclusion into the conference design process.
+The \textbf{design rules} focus on weaving inclusion into the conference design process.
 In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
-\textbf{Rule 5} provides advice for organizing inclusive online and hybrid conferences.
+\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
 \textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
@@
 Finally, 
-a \textbf{continuity Rule} to make the conference part of a long-term commitment to inclusion.
+a \textbf{continuity Rule} emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.
 
@@
 However, the most important step that you should take to ensure the safety of the community is
-adopting a code of conduct (coc) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
-The coc is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
-The coc response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community.
-The team should be prepared early during conference planning because coc violations can happen before the conference too (e.g.
+adopting a code of conduct (CoC) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
+The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
+The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community.
+The team should be prepared early during conference planning because CoC violations can happen before the conference too (e.g.
 social media, organization meetings).
-We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong coc team.
+We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.
 
@@
 Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
-Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
+Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
 Furthermore, tasks such as receiving and responding to code of conduct reports can be emotionally intense work and should be additionally rewarded.
@@
 Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
-To mitigate this self-selection bias, 
+To mitigate this, 
 some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g.
@@
 
-In many academic or technological fields, typically competitive activities or areas—such as scientific publications or software development—are more valued than those focused on sharing and cooperation, like community building, teaching or mentoring.
-The latter activities are equally, if not more, relevant and challenging, and are usually performed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020, light_gender_2022}.
+In many academic or technological fields, typically competitive activities or areas—such as scientific publications or software development—are more valued than those focused on sharing and cooperation, like community building, teaching, or mentoring.
+The latter activities are equally, if not more, relevant and challenging, and are usually performed by women, racialized people, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020, light_gender_2022}.
 When planning the program for the conference, consider giving visibility to the whole range of activities and practitioners that contribute to the field by proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials.
@@
 If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and language interpretation, to name a few.
-Invisible disabilities should be accounted for proactively (e.g.
-dyslexia, anxiety, ADHD, autism), for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.
+Invisible disabilities  (e.g.
+dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.
 
 Regardless of the conference format, all platforms (website, registration, abstract submission, chat system, conference administration tools, live streaming) should be screen-reader friendly and keyboard accessible.
-All images used in digital conference spaces should have alternative text and use color-blind safe palettes.
+All images used in digital conference spaces should have alternative text and use colorblind-safe palettes.
 If there is video streaming, it should have good-quality (not automated) captions and a transcript; live presentations should have good-quality live captions too.
-If the conference has a regional scope, sign-language interpreters can be a better option than captions.
+If the conference has a regional scope, sign language interpreters can be a better option than captions.
 
@@
 
-The communication strategy of your conference should aim to reach broader audiences, express the commitment with diversity and inclusion, and the welcoming spirit of the conference.
+The communication strategy of your conference should aim to reach broader audiences and express the commitment with diversity and inclusion and the welcoming spirit of the conference.
 Actively reach out and promote the conference to communities that have been systematically excluded. 
@@
 2021 mascot.
-To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly: (from left to right) scarves with MiR (Minorities in R https://mircommunity.com/), R-Ladies https://rladies.org/, AfricaR https://africa-r.org/, rainbow LGBTQIA+ and transgender colors.
+To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly.
+From left to right: Margot wearing scarves with the colors of MiR (Minorities in R https://mircommunity.com/), R-Ladies https://rladies.org/, AfricaR https://africa-r.org/, and the LGBTQIA+ and transgender communities.
 Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license}
@@
 From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand.
+
 \subsection*{Rule 9: Allocate adequate financial resources to support inclusion goals}
@@
 Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
-Conferences usually ask for cover letters or applications to assign these scholarships, which can be time-consuming and emotionally demanding. 
+To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
 Simplifying the process of requesting financial support, especially for 
 people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
-Additional aid for attendees could also be offered—e.g. child care support or internet connection services for the virtual component. 
+Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the virtual component. 
 Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups.
 Whenever possible, facilitate alternate ways for money transfer (e.g.
-book flights, hotel reservations, waive conference fees).
+book flights, hotel reservations, waive conference fees, give gift cards).
+Depending on the country, services like OpenCollective (https://opencollective.com/) can be very useful to do these transfers.
 
@@
 Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, or personal issues affecting your organizing team.
-Try to account for these factors as best as you can to set realistic yet bold vision and goals of diversity and inclusion for the conference.
+Try to account for these factors as best as you can to set a realistic yet bold vision and goals of diversity and inclusion for the conference.
 At the end of the day, even if you do not achieve them at 100\%, striving for them will make you stretch what feels possible and help you make changes you would have otherwise not even attempted to make.
 And most importantly, you would have contributed to a great conference experience for the attendees. 
-They will notice the welcoming and inclusive spirit of the conference reflected in the pillar rules, and translated into design decisions—even if there are gaps in them due to the constraining factors mentioned above.
-Inclusive conferences can help to enhance career paths and create a sense of belonging in the community. 
+They will notice the welcoming and inclusive spirit of the conference reflected in the pillar rules and the effort put in every aspect of the conference design. 
+Inclusive conferences can have lasting impacts in career paths and create a sense of belonging in the community. 
 That is more than enough motivation to make the effort worth it.
@@
     \item SC: Sharing rules
-    \item JR: Writing--major reviews and edits
+    \item JR: Conceptualization, Writing--major reviews and edits
 \end{itemize}

### 2022-01-28 plos_latex_template.tex
Recent edits

@@
 Sara Mortara\textsuperscript{3},
-Yanina Bellini Saibene\textsuperscript{4},
-Heather Turner\textsuperscript{5}, %\ddag},
-Dorothea Hug Peter\textsuperscript{6}, %\ddag},
-Matt Bannert\textsuperscript{7},
-Batool Almazrouq\textsuperscript{8,9},
-Liz Hare\textsuperscript{10},
-Laura Ación\textsuperscript{11},
-Juan Pablo Narváez-Gómez\textsuperscript{12},
-Marcela Alfaro Córdoba\textsuperscript{13},
-Federico Marini\textsuperscript{14},
-Rita Giordano\textsuperscript{15},
-Natalia Morandeira\textsuperscript{16},
-Adithi Upadhya\textsuperscript{17},
-Joselyn Chávez\textsuperscript{18},
-Anicet Ebou\textsuperscript{19},
-Silvia Canelón\textsuperscript{20},
-Janani Ravi\textsuperscript{21}
+Yanina Bellini Saibene\textsuperscript{4,5,6},
+Heather Turner\textsuperscript{7}, %\ddag},
+Dorothea Hug Peter\textsuperscript{8}, %\ddag},
+Natalia Soledad Morandeira\textsuperscript{9},
+Matt Bannert\textsuperscript{10},
+Batool Almazrouq\textsuperscript{11,12},
+Liz Hare\textsuperscript{13},
+Laura Ación\textsuperscript{5,14},
+Juan Pablo Narváez-Gómez\textsuperscript{15},
+Marcela Alfaro Córdoba\textsuperscript{16},
+Federico Marini\textsuperscript{17},
+Rita Giordano\textsuperscript{18},
+Adithi Upadhya\textsuperscript{19},
+Joselyn Chávez\textsuperscript{20},
+Anicet Ebou\textsuperscript{21},
+Silvia Canelón\textsuperscript{22},
+Janani Ravi\textsuperscript{23}
 %,3\textcurrency}
@@
 \\
-\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
 \\
-\textbf{4} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina
 \\
-\textbf{5} Department of Statistics, University of Warwick, Coventry, UK
+\textbf{5} MetaDocencia, Argentina
 \\
-\textbf{6} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{6} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
 \\
-\textbf{7} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{7} Department of Statistics, University of Warwick, Coventry, UK
 \\
-\textbf{8} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
+\textbf{8} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
 \\
-\textbf{9} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
+\textbf{9} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
 \\
@@
 \\
-\textbf{11} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{11} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
 \\
-\textbf{12} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{12} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
 \\
-\textbf{13} Department of Statistics, University of California, Santa Cruz, CA, USA
+\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
 \\
-\textbf{14} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{14} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
 \\
-\textbf{15} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{15} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
 \\
-\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{16} Department of Statistics, University of California, Santa Cruz, CA, USA
 \\
-\textbf{17} ILK Labs/Bengaluru, India - 560046
+\textbf{17} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
 \\
-\textbf{17} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{18} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
 \\
-\textbf{19} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\textbf{19} ILK Labs/Bengaluru, Karnataka 560046, India
 \\
-\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
 \\
-\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
+\textbf{21} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\\
+\textbf{22} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
+\\
+\textbf{23} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
 \\
@@
 
-Conferences are spaces to meet and network within an academic or technical field, to learn about new advances, and share our work.
+Conferences are spaces to meet and network within and across academic and technical fields, to learn about new advances, and share our work.
 They can help define career paths and create long-lasting collaborations and opportunities.
 However, these opportunities are not equal for all. 
-This article introduces 10 simple rules to host an inclusive conference. 
+This article introduces ten simple rules to host an inclusive conference. 
 The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
@@
 The rules result from our lessons learned before, during, and after the organization of the conference. 
-They have been mainly written for potential organizers or selection committees of conferences, and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.
+They have been mainly written for potential organizers or selection committees of conferences and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.
 
@@
 However, opportunities for participating in conferences are not equal for all. 
-Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from rich countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
 The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
@@
 %ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file.
-let's try not to forget this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
+ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
 \pdftooltip{
@@
 Open the schedule for events organized by thematic groups and communities (e.g.
-taskforce meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.
+task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.
 
@@
 Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
-Furthermore, tasks such as receiving and responding to code of conduct reports can be emotionally intense work and should be additionally rewarded.
+Furthermore, tasks such as receiving and responding to CoC reports can be emotionally intense work and should be additionally rewarded.
 And most importantly, take care of your team and their well-being!
@@
 When choosing or inviting people for visible and valued roles in the conference it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
-Rather than deter you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges (e.g.
-`manels' and all-white panels \cite{else_how_2019}).
+Rather than deter you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.
 
@@
 In international meetings, the linguistic diversity of the participants is often overlooked. 
-English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpages, and communications. 
+English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, websites, and communications. 
 While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege \cite{amanoTenTipsOvercoming2021}.
@@
 
-Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups–in all communications \cite{hallDesigningDiversityInclusion2019}.
+Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups—in all communications \cite{hallDesigningDiversityInclusion2019}.
 Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
@@
 To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly.
-From left to right: Margot wearing scarves with the colors of MiR (Minorities in R https://mircommunity.com/), R-Ladies https://rladies.org/, AfricaR https://africa-r.org/, and the LGBTQIA+ and transgender communities.
-Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license}
+From left to right: Margot wearing scarves with the colors of MiR (Minorities in R \url{https://mircommunity.com}), R-Ladies \url{https://rladies.org}, AfricaR \url{https://africa-r.org}, and the LGBTQIA+ and transgender communities.
+Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license (\url{https://github.com/useRconf/visuals-2021}).}
 \label{fig:marmots}
@@
 book flights, hotel reservations, waive conference fees, give gift cards).
-Depending on the country, services like OpenCollective (https://opencollective.com/) can be very useful to do these transfers.
+Depending on the country, services like OpenCollective (\url{https://opencollective.com}) can be very useful to do these transfers.
 
@@
     \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules
+    \item NSM: Writing--major reviews and edits, Resources, Visualization, Sharing rules
     \item MB: Conceptualization
@@
     \item RG: Writing--major reviews and edits
-    \item NM: Sharing rules
     \item AU: Sharing rules
@@
 The authors of this piece would like to thank every single member of the organizing team of useR!
-2021 (\url{https://user2021.r-project.org/about/global-team/}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR!
+2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR!
 2021 and supporting us through the process.
@@
 
-Heather Turner was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR!
+\section*{Funding}
+HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR!
 2021.
+YBS was supported by the Instituto Nacional de Tecnología Agropecuaria, Argentina during the organization of useR! 2021.
+... ...
 

### 2022-02-03 plos_latex_template.tex
co-author edits incorporated by @AndreaSanchezTapia

@@
 \\
-\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
 \\
@@
 \\
-\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{10} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
 \\
@@
 \\
-\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{13} Dog Genetics LLC, Astoria, NY, USA
 \\
@@
 \\
-\textbf{21} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\textbf{21} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
 \\
@@
 However, opportunities for participating in conferences are not equal for all. 
-Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
 The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
@@
 The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
-They are organized in three sections: three pillars, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
-\textbf{The pillars} comprise fundamental rules with key elements to conceive the work on diversity and inclusion in any conference. 
+They are organized in three sections: three foundational rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
+The \textbf{foundational rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
 \textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
@@
 Finally, 
-a \textbf{continuity Rule} emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.
+\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.
 
@@
 \includegraphics[width=\textwidth]{figs/rules.png}}{
-%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: pillar rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10).
+%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundational rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10).
 The overall diagram depicts an arrow from left to right which contains all the rules.
-From left to right, "the pillars" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team".
+From left to right, "the foundations" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team".
 Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial".
@@
 }
-\caption{Schematic diagram of the rules organized in three groups: the pillars (Rules 1, 2, and 3), design rules (Rules 4 to 9), and continuity (Rule 10).}
+\caption{Schematic diagram of the rules organized in three groups: the foundations (Rules 1, 2, and 3), conference design (Rules 4 to 9), and continuity (Rule 10).}
 \label{fig:diagram}
@@
 
-Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
+Diversity encompasses multiple dimensions: age, disability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
 Human diversity should be celebrated and respected in every way. 
@@
 being cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built.
-People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways.
+People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways \cite{crenshawDemarginalizingIntersectionRace1989}.
 As a result, some groups of people have been systematically excluded from or only partially included in academia and scientific and professional circles.
@@
 This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}).
-For example, promote the use and respect of preferred pronouns and commit to respect all genders and sexual orientations.
+For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations.
 If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care.
@@
 Make sure that you are not creating additional barriers when choosing conference platforms, schedule, and activities. 
-Check every platform's screen reader accessibility (see \textbf{Rule 6}), bandwidth requirements (a high-speed internet prerequisite could exclude attendees from many regions), and geopolitical restrictions (some services are not available in all countries).
+Check every platform's screen reader accessibility (see \textbf{Rule 6}), technological requirements (a high-speed internet or a latest generation computer prerequisite could exclude attendees from many regions), and geopolitical restrictions (some services are not available in all countries).
 Be mindful of the different time zones and time constraints of the potential attendees and create flexible schedules.
@@
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
-Including people with disabilities in the organizing team from the start can have a large impact (see \textbf{Rule 3}), since planning for accessibility requires time, experience, early decision-making, and inaccessible features are difficult to correct at later stages \cite{irishIncreasingParticipationUsing2020}.
+Including people with disabilities in the organizing team from the start can have a large impact (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making, because inaccessible features are difficult to correct at later stages \cite{irishIncreasingParticipationUsing2020}.
+To ensure that disabled organizers can contribute substantially, the tools selected for behind-the-scenes organizing, communication, and planning the conference must be compatible with adaptive technology like screen readers, as well as providing captions for spoken conversations for deaf people.
+Many products currently available to organizers come with barriers to full participation.
+The people who use these technologies have expertise on which products work best with them.
 
-If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and language interpretation, to name a few.
+If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, having signs in Braille, and using a sound system compatible with hearing devices and language interpretation, to name a few.
 Invisible disabilities  (e.g.
@@
 
-Encourage the preparation of accessible slides and presentations by providing accessibility guidelines and be available for any questions that presenters and attendees may have; see \cite{sanchez-tapia_user_2021} for an example of accessibility guidelines and \cite{chavez_preparing_2021, sanchez-tapia_making_2021, joo_how_2021} for further accessibility recommendations.
+Encourage the preparation of accessible slides and presentations by providing accessibility guidelines and presentation templates and be available for any questions that presenters and attendees may have; see \cite{sanchez-tapia_user_2021} for an example of accessibility guidelines and \cite{chavez_preparing_2021, sanchez-tapia_making_2021, joo_how_2021} for further accessibility recommendations.
 Social events and networking should also have accessibility features like captioning or sign language interpretation, and include activities that do not restrict participation based on body type or ability.
@@
 The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g.
-Twitter, Facebook, LinkedIn, conference website, mailing lists).%#promotion2
+Twitter, Facebook, LinkedIn, conference website, mailing lists).
 
@@
 Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups—in all communications \cite{hallDesigningDiversityInclusion2019}.
+Avoid as well all offensive or sexualized imagery in slides, conference materials, social events, and.
 Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
@@
 You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}.
-Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
+Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from minoritized groups, by offering support for travel and lodging expenses.
 Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
@@
 people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
-Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the virtual component. 
+Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the online component. 
 Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups.
@@
 And most importantly, you would have contributed to a great conference experience for the attendees. 
-They will notice the welcoming and inclusive spirit of the conference reflected in the pillar rules and the effort put in every aspect of the conference design. 
+They will notice the welcoming and inclusive spirit of the conference. 
 Inclusive conferences can have lasting impacts in career paths and create a sense of belonging in the community. 
@@
 2021 and supporting us through the process.
-Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.
+Special thanks to Saranjeet Kaur and Gwynn Sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.
 

### 2022-02-04 plos_latex_template.tex
Submitted version

@@
 %Janani: for inserting sparingly few super relevant emojis
-\usepackage{emoji} %Rocio: https://www.overleaf.com/learn/latex/Questions/Inserting_emojis_in_LaTeX_documents_on_Overleaf#Emoji_characters_and_the_Overleaf_editor
+% \usepackage{emoji} %Rocio: https://www.overleaf.com/learn/latex/Questions/Inserting_emojis_in_LaTeX_documents_on_Overleaf#Emoji_characters_and_the_Overleaf_editor
 
@@
 Rita Giordano\textsuperscript{18},
-Adithi Upadhya\textsuperscript{19},
-Joselyn Chávez\textsuperscript{20},
-Anicet Ebou\textsuperscript{21},
-Silvia Canelón\textsuperscript{22},
+Silvia Canelón\textsuperscript{19},
+Anicet Ebou\textsuperscript{20},
+Adithi Upadhya\textsuperscript{21},
+Joselyn Chávez\textsuperscript{22},
 Janani Ravi\textsuperscript{23}
@@
 \\
-\textbf{19} ILK Labs/Bengaluru, Karnataka 560046, India
+\textbf{19} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
 \\
-\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
+\textbf{20} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
 \\
-\textbf{21} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\textbf{21} ILK Labs/Bengaluru, Karnataka 560046, India
 \\
-\textbf{22} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
+\textbf{22} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
 \\
@@
 However, these opportunities are not equal for all. 
-This article introduces ten simple rules to host an inclusive conference. 
-The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
+This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR!
+statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector.
 Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
 The rules result from our lessons learned before, during, and after the organization of the conference. 
-They have been mainly written for potential organizers or selection committees of conferences and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.
+They have been mainly written for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive.
+We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
 
@@
 % Conferences are great, but only for some
-Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
+Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share recent contributions.
 A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
-However, opportunities for participating in conferences are not equal for all. 
-Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
-The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
+However, opportunities for participating in conferences are not equally available for all. 
+Many academic and tech conferences have been spaces that reproduce systemic inequalities, by failing to overcome the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income backgrounds, prestigious institutions, native English speakers with no disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+Like many others, the authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
 We had the opportunity to organize a virtual and global statistical computing conference, useR!
@@
 % Roadmap of the paper
-The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
+Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference. 
 They are organized in three sections: three foundational rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
@@
 In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
-\textbf{Rule 8} provides tips for developing an inclusive communication strategy. 
+\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
 In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
 Finally, 
-\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.
+\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion, and passing the torch to future organizers.
 
-We believe that these rules will greatly benefit organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
+We believe that these rules can guide organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
 The rules will also help people who are part of meeting committees that oversee the site/location selection process, or that coordinate with the local organizers of conferences. 
-We have designed them for wide adoption to help scientific and technical conferences to become as accessible and inclusive as they can and should be.
+We see the rules as a starting point for conversations and efforts towards building more inclusive conferences across the world.
 
@@
 \pdftooltip{
-\includegraphics[width=\textwidth]{figs/rules.png}}{
-%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundational rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10).
+\includegraphics[width=\textwidth]{figs/fig1_rules.png}}{
+%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundation rules (Rules 1-3), design rules (Rules 4 to 9), and a continuity rule (Rule 10).
 The overall diagram depicts an arrow from left to right which contains all the rules.
-From left to right, "the foundations" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team".
-Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial".
-To the right, and forming the head of the arrow, the Tenth Rule is under the label "Continuity" and consists of a blue bubble "10 Assess and pass along".
+In the left column, under the first label" Foundation", there are three text boxes in light blue: "1.
+Vision", "2.
+Safe environment", and "3.
+Organizing Team".
+In the middle column, under the label "Design Rules" there are six light green boxes with Rules "4.
+Counter bias", "5.
+Virtual / hybrid", "6.
+Accessibility", "7.
+Language", "8.
+Communication",  "9.
+Finance".
+To the right, below the arrowhead labeled Continuity, is the Tenth Rule in a light pink box "10.
+Pass the torch".
 }
-\caption{Schematic diagram of the rules organized in three groups: the foundations (Rules 1, 2, and 3), conference design (Rules 4 to 9), and continuity (Rule 10).}
+\caption{Schematic diagram of the rules organized in three groups: foundation (Rules 1, 2, and 3), design (Rules 4 to 9), and continuity (Rule 10).}
 \label{fig:diagram}
@@
 
-\subsection*{Rule 1: Set a vision for diversity and inclusion}
+\section*{Rule 1: Set a vision for diversity and inclusion}
 \label{rule_diversity}
@@
 being cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built.
-People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways \cite{crenshawDemarginalizingIntersectionRace1989}.
-As a result, some groups of people have been systematically excluded from or only partially included in academia and scientific and professional circles.
+People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways (a concept known as `intersectionality' \cite{crenshawDemarginalizingIntersectionRace1989}).
+As a result, some groups of people have been systematically excluded from or only partially included in academic, scientific, and professional circles.
 Recognizing this systematic exclusion is paramount to inclusion.
-Assess how it translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
+Assess how lack of diversity translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
 Define measurable indicators for these goals (e.g.
@@
 
-\subsection*{Rule 2: Create a safe and welcoming environment}
+\section*{Rule 2: Create a safe and welcoming environment}
 \label{rule_inclusion}
@@
 Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees.
-This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}).
+This means respecting all the aspects included in your diversity vision (see \textbf{Rule 1}).
 For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations.
@@
 Include menu options that account for diverse dietary and religious requirements.
-Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and  great advice on `low-hanging fruits').
+Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it's good to have them in mind from the beginning.
 Open the schedule for events organized by thematic groups and communities (e.g.
@@
 social media, organization meetings).
+Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding their work to your budget (see \textbf{Rule 9}).
 We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.
 
-\subsection*{Rule 3: Have an inclusive and diverse organizing team}
+\section*{Rule 3: Gather an inclusive and diverse organizing team}
 \label{rule_organizing_team}
@@
 A genuinely inclusive conference can only be organized by an inclusive and diverse team.
-Organizers should comprise a team with people from as many different backgrounds as possible regarding regions, genders, ethnicities, career stages, and other aspects of diversity.
+Organizers should comprise a team with people from as many different backgrounds as possible in terms of regions, genders, ethnicities, career stages, and other aspects of diversity.
 People with disabilities often say: `Nothing about us without us' \cite{charlton_nothing_1998, werner_nothing_1998}; the same holds for other dimensions of diversity. 
@@
 When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably.
-People from underrepresented groups sometimes refrain to nominate themselves while people in dominant groups tend to do it more easily. 
-Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
-Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
-Furthermore, tasks such as receiving and responding to CoC reports can be emotionally intense work and should be additionally rewarded.
+Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
+Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
+In any case, try to find meaningful ways to acknowledge everybody's work.
+Use the conference spaces (e.g.
+webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
 And most importantly, take care of your team and their well-being!
@@
 
-\subsection*{Rule 4: Consciously counteract bias in the conference program}
+\section*{Rule 4: Consciously counteract bias in the conference program}
 \label{rule_unbias}
@@
 It is important to counteract biases in all of these aspects.
-When choosing or inviting people for visible and valued roles in the conference it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
-Rather than deter you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.
+When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
+Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.
 
 Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
-To mitigate this, 
-some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g.
+To mitigate such self-selection, some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g.
 the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
@@
 The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.
-Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019}.
+Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
 Evaluate the plausibility of anonymizing all of the submission information to
@@
 Define clear evaluation criteria and share them with the reviewers and authors.
-Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe; even well-meaning and carefully crafted processes can be prone to reproduce the biases observed in academia and society.
+Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
 
@@
 
-\subsection*{Rule 5: Design a strong online component} 
+\section*{Rule 5: Design a strong online component} 
 \label{rule_online}
@@
 After experiencing virtual conferences mostly due to the COVID-19 pandemic, there have been multiple calls to retain an online component in conferences, either as completely virtual events or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
-Virtual conferences remove barriers for inclusion like costs of participation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}.
+Virtual conferences remove barriers for inclusion like costs of participation, the logistics of long-distance and international travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}.
 Make sure that you are not creating additional barriers when choosing conference platforms, schedule, and activities. 
@@
 Be mindful of the different time zones and time constraints of the potential attendees and create flexible schedules.
-Avoid creating a strictly synchronous live program, ask for pre-recorded talks, record sessions, and make them available during the conference.
+Avoid creating a strictly synchronous live program; ask for pre-recorded talks, record sessions, and make them available during the conference.
 
-If you are organizing a hybrid event, decide on the relative importance that the online and in-person components will take \cite{bajpai_towards_2021}.
+If you are organizing a hybrid event, decide on the relative importance that online and in-person components will take \cite{bajpai_towards_2021}.
 The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending the conference.
@@
 To best integrate in-person and remote attendees, all questions and comments should pass through a unique online system. 
-The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with them.
+The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.
 
-Online networking and socializing can make people from minoritized groups feel more included, thus participate and contribute more, than when meeting face-to-face for the first time \cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}.
+Online networking and socializing can make people from minoritized groups feel more included, thus participate and contribute more, than when meeting face-to-face for the first time (e.g.
+\cite{trianaDoesOrderFacetoFace2012,blackEngenderingBelongingThoughtful2020}).
 Take advantage of the online component to have a broader range of social activities that can appeal to people with different backgrounds and preferences.
@@
 
-\subsection*{Rule 6: Make the conference accessible to people with disabilities}
+\section*{Rule 6: Make the conference accessible to people with disabilities}
 \label{rule_accessibility}
@@
 Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}.
-Including people with disabilities in the organizing team from the start can have a large impact (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making, because inaccessible features are difficult to correct at later stages \cite{irishIncreasingParticipationUsing2020}.
-To ensure that disabled organizers can contribute substantially, the tools selected for behind-the-scenes organizing, communication, and planning the conference must be compatible with adaptive technology like screen readers, as well as providing captions for spoken conversations for deaf people.
-Many products currently available to organizers come with barriers to full participation.
-The people who use these technologies have expertise on which products work best with them.
+Including people with disabilities in the organizing team from the start can have a large impact (see \textbf{Rule 3}).
+Planning for accessibility requires time, experience, and early decision-making, because inaccessible features are extremely difficult to correct at later stages \cite{irishIncreasingParticipationUsing2020}.
+To ensure that disabled organizers can contribute substantially, the tools selected for behind-the-scenes organizing, communication, and planning the conference must be compatible with adaptive technology like screen readers.
+When working with deaf and hard-of-hearing people, spoken conversations should have captions.
 
@@
 Invisible disabilities (e.g.
-dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.
+dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces (but see \cite{pun_dos_2016} for other examples).
 
@@
 
-\subsection*{Rule 7: Make room for the linguistic diversity of your community}
+\section*{Rule 7: Make room for the linguistic diversity of your community}
 \label{rule_language}
@@
 That way, the abstract will be judged primarily by the quality or relevance of the work instead of English proficiency.
-For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities, see \textbf{Rule 6}) follow the presentations.
 Whenever possible, identify other prominent languages for the conference and provide translated captions or live language interpretation into these key languages, including multilingual Q\&A sessions.
 Another way to engage non-native English participants is to host sessions and events in languages other than English. 
-Promote these sessions among all conference attendees and announce if captions will be available; all sessions, regardless of the language, should be given the same relevance.
+Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same relevance as the rest of the program.
 Importantly, consideration and respect of accents or linguistic mistakes can make a significant difference in the conference experience of non-native English speakers.
@@
 
-\subsection*{Rule 8: Build an inclusive communication strategy}
+\section*{Rule 8: Build an inclusive communication strategy}
 % media
@@
 
-Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}) to communicate your practices and to hold yourself accountable for them.
+Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}) to communicate your values and practices, and to hold yourself accountable for them.
 Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. 
@@
 Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups—in all communications \cite{hallDesigningDiversityInclusion2019}.
-Avoid as well all offensive or sexualized imagery in slides, conference materials, social events, and.
-Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
+Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group, and the terms that should be avoided.
 Do not expect minoritized people to teach you and accept feedback without being offended.
@@
 
-\subsection*{Rule 9: Allocate adequate financial resources to support inclusion goals}
+\section*{Rule 9: Allocate adequate financial resources to support inclusion goals}
 \label{rule_financial}
@@
 % Rocío: main idea: Need to budget for inclusive practices.
-Allocation of resources to support the diversity and inclusion vision and goals has to be intentional. 
+Allocation of resources to support the diversity and inclusion vision and goals has to be intentional and demonstrate true commitment. 
 Estimate the costs of these practices and define your priorities in advance (e.g. paying the organizing team, code of conduct training, captioning, accessible software, scholarships).
-Consider that an online conference might reduce overall costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards inclusive practices. 
+An online conference might reduce overall costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards inclusive practices. 
 When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion than making generic requests for funding.
@@
 Resources permitting, enable a `pay what you can' option.
-You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}.
+You could also aim to have a conference with no registration costs; this could especially apply for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}.
 Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from minoritized groups, by offering support for travel and lodging expenses.
@@
 book flights, hotel reservations, waive conference fees, give gift cards).
-Depending on the country, services like OpenCollective (\url{https://opencollective.com}) can be very useful to do these transfers.
 
-\subsection*{Rule 10: Make the conference part of a long-term process for inclusion}
+\section*{Rule 10: Make the conference part of a long-term process for inclusion}
 \label{rule_process}
@@
 
-This article suggests changes in conference planning towards diversity and inclusion. 
+This article recommends changes in conference planning towards diversity and inclusion. 
 Organizing a conference and implementing inclusive practices are both learning experiences.
@@
 Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, or personal issues affecting your organizing team.
-Try to account for these factors as best as you can to set a realistic yet bold vision and goals of diversity and inclusion for the conference.
-At the end of the day, even if you do not achieve them at 100\%, striving for them will make you stretch what feels possible and help you make changes you would have otherwise not even attempted to make.
-And most importantly, you would have contributed to a great conference experience for the attendees. 
+Try to account for these factors as best as you can to set bold yet realistic goals of diversity and inclusion for the conference.
+And most importantly, your efforts will enhance the conference experience for your attendees. 
 They will notice the welcoming and inclusive spirit of the conference. 
-Inclusive conferences can have lasting impacts in career paths and create a sense of belonging in the community. 
+Remember that inclusive conferences can have lasting impacts in career paths and create a sense of belonging in the community.
 That is more than enough motivation to make the effort worth it.
@@
 
-To list author contributions, we used the high-level contributor roles in the CRediT framework that we considered relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and an additional category to acknowledge those who shared their own ten simple rules for an inclusive conference without looking at the first draft (i.e.
-sharing rules).
+To list author contributions, we used the high-level contributor roles in the CRediT framework that we considered relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and additional categories to acknowledge those who shared their own ten simple rules for an inclusive conference without looking at the first draft (i.e.
+sharing rules), and those who translated the extended abstract into different languages.
 
 \begin{itemize}
-    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization
-    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules
-    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules
-    \item YBS: Conceptualization, Writing--major reviews and edits, Resources
+    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization, Extended abstract translation
+    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
+    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
+    \item YBS: Conceptualization, Writing--major reviews and edits, Resources, Extended abstract translation
     \item HT: Conceptualization, Writing--major reviews and edits, Resources
-    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules
-    \item NSM: Writing--major reviews and edits, Resources, Visualization, Sharing rules
-    \item MB: Conceptualization
-    \item BA: Writing--major reviews and edits, Resources, Sharing rules
+    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules, Extended abstract translation
+    \item NSM: Writing--major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
+    \item MB: Conceptualization, Extended abstract translation
+    \item BA: Writing--major reviews and edits, Resources, Sharing rules, Extended abstract translation
     \item LH: Writing--major reviews and edits, Resources, Sharing rules
@@
     \item MAC: Writing--major reviews and edits
-    \item FM: Writing--major reviews and edits
-    \item RG: Writing--major reviews and edits
+    \item FM: Writing--major reviews and edits, Extended abstract translation
+    \item RG: Writing--major reviews and edits, Resources, Extended abstract translation
+    \item SC: Resources, Sharing rules
+    \item AE: Sharing rules, Extended abstract translation
     \item AU: Sharing rules
     \item JC: Sharing rules
-    \item AE: Sharing rules
-    \item SC: Sharing rules
-    \item JR: Conceptualization, Writing--major reviews and edits
+    \item JR: Conceptualization, Writing--major reviews and edits, Visualization, Extended abstract translation
 \end{itemize}
@@
 2021 and supporting us through the process.
-Special thanks to Saranjeet Kaur and Gwynn Sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.
+Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.
+We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript.
+In addition to some of the authors, we would like to acknowledge the following contributors for translating our extended abstract to several languages: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], and Kewalin Samart [Thai].
 
 \section*{Funding}
+
 HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR!
 2021.
-YBS was supported by the Instituto Nacional de Tecnología Agropecuaria, Argentina during the organization of useR! 2021.
-... ...
+HT was a board member of the The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
+conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups.
+The R Foundation is currently funding YBS to support the useR!
+2022 organization team.
+HT was administrator and volunteer for the R Project Google Season of Docs 2021 project, that funded work by AST (writer) and YBS (steering committee).
+The organization of useR!
+2021 was largely a volunteer effort, but stipends were available from the conference funds for specific roles.
+RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively.
 
@@
 % % Compile your BiBTeX database using our plos2015.bst
-\bibliography{community-science}
+%\bibliography{community-science}
 % % style file and paste the contents of your .bbl file
@@
 % % % 
-% \begin{thebibliography}{10}
+\begin{thebibliography}{10}
 
-% \bibitem{bib1}
-% Conant GC, Wolfe KH.
-% \newblock {{T}urning a hobby into a job: how duplicated genes find new
-%   functions}.
-% \newblock Nat Rev Genet. 2008 Dec;9(12):938--950.
+\bibitem{arendDisparityConferenceRegistration2019}
+Arend ME, Bruijns SR.
+\newblock Disparity in Conference Registration Cost for Delegates from Low- and
+  Middle-Income Backgrounds.
+\newblock African Journal of Emergency Medicine. 2019;9(3):156--161.
+\newblock doi:{10.1016/j.afjem.2019.01.016}.
 
-% \bibitem{bib2}
-% Ohno S.
-% \newblock Evolution by gene duplication.
-% \newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
-%   Springer-Verlag.; 1970.
+\bibitem{biggsAcademicConferenceChilly2018}
+Biggs J, Hawley PH, Biernat M.
+\newblock The {{Academic Conference}} as a {{Chilly Climate}} for {{Women}}:
+  Effects of {{Gender Representation}} on {{Experiences}} of {{Sexism}},
+  {{Coping Responses}}, and {{Career Intentions}}.
+\newblock Sex Roles. 2018;78(5):394--408.
+\newblock doi:{10.1007/s11199-017-0800-9}.
 
-% \bibitem{bib3}
-% Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
-% \newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
-%   infection through a transposon insertion followed by a {D}uplication}.
-% \newblock PLoS Genet. 2011 Oct;7(10):e1002337.
+\bibitem{depickerRethinkingInclusionDisability2020a}
+De~Picker M.
+\newblock Rethinking Inclusion and Disability Activism at Academic Conferences:
+  Strategies Proposed by a {{PhD}} Student with a Physical Disability.
+\newblock Disability \& Society. 2020;35(1):163--167.
+\newblock doi:{10.1080/09687599.2019.1619234}.
 
-% \end{thebibliography}
+\bibitem{irishIncreasingParticipationUsing2020}
+Irish JEN.
+\newblock Increasing Participation: Using the Principles of Universal Design to
+  Create Accessible Conferences.
+\newblock Journal of Convention \& Event Tourism. 2020;21(4):308--330.
+\newblock doi:{10.1080/15470148.2020.1814469}.
+
+\bibitem{sanchez-tapia_user_2021-2}
+{S{\'a}nchez-Tapia} A, Tamir N, Moy~Das S, Bellini~Saibene Y, Joo R, Morandeira
+  N, et~al.. {{useR}}! {{Knowledgebase}}; 2021.
+\newblock https://rconf.gitlab.io/userknowledgebase/.
+
+\bibitem{r_core_team_2021}
+{R Core Team}. R: A Language and Environment for Statistical Computing; 2021.
+\newblock Available from: \url{https://www.R-project.org/}.
+
+\bibitem{crenshawDemarginalizingIntersectionRace1989}
+Crenshaw K.
+\newblock Demarginalizing the {{Intersection}} of {{Race}} and {{Sex}}: A
+  {{Black Feminist Critique}} of {{Antidiscrimination Doctrine}}, {{Feminist
+  Theory}} and {{Antiracist Politics}}.
+\newblock University of Chicago Legal Forum. 1989;1989:139.
+
+\bibitem{numfocus_discover_2021}
+{NumFOCUS Diversity \& Inclusion in Scientific Computing (DISC) Program}.
+  {{DISCOVER Cookbook}} {$\cdot$} {{Diverse}} \& {{Inclusive Spaces}} and
+  {{Conferences}}; 2021.
+\newblock https://discover-cookbook.numfocus.org/.
+
+\bibitem{favaroYourScienceConference2016}
+Favaro B, Oester S, Cigliano JA, Cornick LA, Hind EJ, Parsons ECM, et~al.
+\newblock Your {{Science Conference Should Have}} a {{Code}} of {{Conduct}}.
+\newblock Frontiers in Marine Science. 2016;0.
+\newblock doi:{10.3389/fmars.2016.00103}.
+
+\bibitem{auroraHowRespondCode2019}
+Aurora V, Gardiner M.
+\newblock How to {{Respond}} to {{Code}} of {{Conduct Reports}}. {{A}}
+  Practical Step-by-Step Guide to Handling Code of Conduct Issues.
+\newblock {Frame Shift Consulting LLC}; 2019.
+\newblock Available from:
+  \url{https://frameshiftconsulting.com/resources/code-of-conduct-book/}.
+
+\bibitem{charlton_nothing_1998}
+Charlton JI.
+\newblock Nothing about us without us.
+\newblock University of California Press; 1998.
+
+\bibitem{werner_nothing_1998}
+Werner D, Thuman C, Maxwell J.
+\newblock Nothing about us without us.
+\newblock Developing innovative technologies for, by and with disabled persons
+  Palo Alto: Healthwrights. 1998;.
+
+\bibitem{hongGroupsDiverseProblem2004}
+Hong L, Page SE.
+\newblock Groups of Diverse Problem Solvers Can Outperform Groups of
+  High-Ability Problem Solvers.
+\newblock Proceedings of the National Academy of Sciences.
+  2004;101(46):16385--16389.
+\newblock doi:{10.1073/pnas.0403723101}.
+
+\bibitem{costanzachockDesign2020}
+{Costanza-Chock} S.
+\newblock Design Justice: Community-Led Practices to Build the Worlds We Need.
+\newblock Braman S, editor. Information {{Policy}}. {Cambridge, MA, USA}: {MIT
+  Press}; 2020.
+
+\bibitem{dwyerNoticeWhoScience2021}
+Dwyer JR.
+\newblock Notice Who the Science System Honours, and How.
+\newblock Nature. 2021;595(7865):30--30.
+\newblock doi:{10.1038/d41586-021-01785-3}.
+
+\bibitem{swartzScienceValueDiversity2019}
+Swartz TH, Palermo AGS, Masur SK, Aberg JA.
+\newblock The {{Science}} and {{Value}} of {{Diversity}}: Closing the {{Gaps}}
+  in {{Our Understanding}} of {{Inclusion}} and {{Diversity}}.
+\newblock The Journal of Infectious Diseases. 2019;220(Supplement\_2):S33--S41.
+\newblock doi:{10.1093/infdis/jiz174}.
+
+\bibitem{wongBuildDiversityScience2020}
+Wong VNL, Shaw JD.
+\newblock Build Diversity among Science Prize Winners.
+\newblock Nature. 2020;580(7802):185--185.
+\newblock doi:{10.1038/d41586-020-01033-0}.
+
+\bibitem{dignazioUnicornsJanitorsNinjas2020}
+D'Ignazio C, Klein L.
+\newblock 5. {{Unicorns}}, {{Janitors}}, {{Ninjas}}, {{Wizards}}, and {{Rock
+  Stars}}.
+\newblock In: Data {{Feminism}}. {PubPub}; 2020.Available from:
+  \url{https://data-feminism.mitpress.mit.edu/pub/2wu7aft8/release/2}.
+
+\bibitem{ross_everyday_2020}
+Ross HJ.
+\newblock Everyday {{Bias}}: {{Identifying}} and {{Navigating Unconscious
+  Judgments}} in {{Our Daily Lives}}.
+\newblock {Rowman \& Littlefield}; 2020.
+
+\bibitem{andersson_implicit_2019}
+Andersson ER, Hagberg C, H{\"a}gg S. Implicit Bias Is Strongest When Assessing
+  Top Candidates; 2019.
+
+\bibitem{cheng2020x+}
+Cheng E.
+\newblock X+ Y: A Mathematician's Manifesto for Rethinking Gender.
+\newblock Basic Books; 2020.
+
+\bibitem{burfordHomelinessMeantHaving2020}
+Burford J, Bosanquet A, Smith J.
+\newblock `{{Homeliness}} Meant Having the Fucking Vacuum Cleaner out': The
+  Gendered Labour of Maintaining Conference Communities.
+\newblock Gender and Education. 2020;32(1):86--100.
+\newblock doi:{10.1080/09540253.2019.1680809}.
+
+\bibitem{light_gender_2022}
+Light AE, {Benson-Greenwald} TM, Diekman AB.
+\newblock Gender Representation Cues Labels of Hard and Soft Sciences.
+\newblock Journal of Experimental Social Psychology. 2022;98:104234.
+
+\bibitem{jooKeepOnlineOption2021}
+Joo R.
+\newblock Keep Online Option at Conferences \textemdash{} It Makes Them More
+  Inclusive.
+\newblock Nature. 2021;598(7880):257--257.
+\newblock doi:{10.1038/d41586-021-02752-8}.
+
+\bibitem{woolstonLearningLoveVirtual2020}
+Woolston C.
+\newblock Learning to Love Virtual Conferences in the Coronavirus Era.
+\newblock Nature. 2020;582(7810):135--136.
+\newblock doi:{10.1038/d41586-020-01489-0}.
+
+\bibitem{ninerBetterWhomLeveling2021}
+Niner HJ, Wassermann SN.
+\newblock Better for {{Whom}}? Leveling the {{Injustices}} of {{International
+  Conferences}} by {{Moving Online}}.
+\newblock Frontiers in Marine Science. 2021;0.
+\newblock doi:{10.3389/fmars.2021.638025}.
+
+\bibitem{roosOnlineConferencesNew2020}
+Roos G, Ol{\'a}h J, Ingle R, Kobayashi R, Feldt M.
+\newblock Online Conferences \textendash{} {{Towards}} a New (Virtual) Reality.
+\newblock Computational and Theoretical Chemistry. 2020;1189:112975.
+\newblock doi:{10.1016/j.comptc.2020.112975}.
+
+\bibitem{levitisCenteringInclusivityDesign2021}
+Levitis E, {van Praag} CDG, Gau R, Heunis S, DuPre E, Kiar G, et~al.
+\newblock Centering Inclusivity in the Design of Online
+  Conferences\textemdash{{An OHBM}}\textendash{{Open Science}} Perspective.
+\newblock GigaScience. 2021;10(8).
+\newblock doi:{10.1093/gigascience/giab051}.
+
+\bibitem{sarabipourChangingScientificMeetings2021}
+Sarabipour S, Khan A, Seah YFS, Mwakilili AD, Mumoki FN, S{\'a}ez PJ, et~al.
+\newblock Changing Scientific Meetings for the Better.
+\newblock Nature Human Behaviour. 2021;5(3):296--300.
+\newblock doi:{10.1038/s41562-021-01067-y}.
+
+\bibitem{salibaGettingGripsOnline2020}
+Saliba M.
+\newblock Getting to Grips with Online Conferences.
+\newblock Nature Energy. 2020;5(7):488--490.
+\newblock doi:{10.1038/s41560-020-0656-z}.
+
+\bibitem{gichoraTenSimpleRules2010a}
+Gichora NN, Fatumo SA, Ngara MV, Chelbat N, Ramdayal K, Opap KB, et~al.
+\newblock Ten {{Simple Rules}} for {{Organizing}} a {{Virtual
+  Conference}}\textemdash{{Anywhere}}.
+\newblock PLOS Computational Biology. 2010;6(2):e1000650.
+\newblock doi:{10.1371/journal.pcbi.1000650}.
+
+\bibitem{bajpai_towards_2021}
+Bajpai V, Hohlfeld O, Crowcroft J, Keshav S.
+\newblock Towards Climate-Friendly Internet Research (Dagstuhl Seminar 21272).
+\newblock In: Dagstuhl Reports. vol.~11. Schloss Dagstuhl-Leibniz-Zentrum
+  f{\"u}r Informatik; 2021.
+
+\bibitem{trianaDoesOrderFacetoFace2012}
+Triana MdC, Kirkman BL, Wagstaff MF.
+\newblock Does the {{Order}} of {{Face}}-to-{{Face}} and
+  {{Computer}}-{{Mediated Communication Matter}} in {{Diverse Project Teams}}?
+  An {{Investigation}} of {{Communication Order Effects}} on {{Minority
+  Inclusion}} and {{Participation}}.
+\newblock Journal of Business and Psychology. 2012;27(1):57--70.
+
+\bibitem{blackEngenderingBelongingThoughtful2020}
+Black AL, Crimmins G, Dwyer R, Lister V.
+\newblock Engendering Belonging: Thoughtful Gatherings with/in Online and
+  Virtual Spaces.
+\newblock Gender and Education. 2020;32(1):115--129.
+\newblock doi:{10.1080/09540253.2019.1680808}.
+
+\bibitem{priceAccessImaginedConstruction2009}
+Price M.
+\newblock Access {{Imagined}}: The {{Construction}} of {{Disability}} in
+  {{Conference Policy Documents}}.
+\newblock Disability Studies Quarterly. 2009;29(1).
+\newblock doi:{10.18061/dsq.v29i1.174}.
+
+\bibitem{marks2021meeting}
+Marks GS, Solomon C, Whitney KS.
+\newblock Meeting frameworks must be even more inclusive.
+\newblock Nature Ecology \& Evolution. 2021;5(5):552--552.
+
+\bibitem{pun_dos_2016}
+Pun K. Dos and Don'ts on Designing for Accessibility - {{Accessibility}} in
+  Government; 2016.
+\newblock
+  https://accessibility.blog.gov.uk/2016/09/02/dos-and-donts-on-designing-for-accessibility/.
+
+\bibitem{sanchez-tapia_user_2021}
+{S{\'a}nchez-Tapia} A, Hare L, Ch{\'a}vez J, Canel{\'o}n S, Hug~Peter D.
+  {{useR}}! 2021 {{Accessibility}} Guidelines; 2021.
+
+\bibitem{chavez_preparing_2021}
+Ch{\'a}vez J. Preparing for an {{Accessible Online Conference}}; 2021.
+\newblock
+  https://user2021.r-project.org/blog/2021/02/17/preparing-for-an-accessible-conference//.
+
+\bibitem{sanchez-tapia_making_2021}
+{S{\'a}nchez-Tapia} A, Hare L, Ch{\'a}vez J, Mortara S, Joo R. Making
+  Accessible Presentations at {{useR}}! 2021: The Story behind the Scenes;
+  2021.
+\newblock
+  https://user2021.r-project.org/blog/2021/12/07/accessibility\_awards\_interview//.
+
+\bibitem{joo_how_2021}
+Joo R, Canel{\'o}n S, Ch{\'a}vez J, {S{\'a}nchez-Tapia} A. How to Improve
+  Conference Accessibility for Screen-Reader Users\textemdash an Interview with
+  {{Liz Hare}}; 2021.
+\newblock
+  https://user2021.r-project.org/blog/2021/11/04/accessibility\_interview\_liz\_hare//.
+
+\bibitem{amanoTenTipsOvercoming2021}
+Amano T, Rios~Rojas C, Boum~II Y, Calvo M, Misra BB.
+\newblock Ten tips for overcoming language barriers in science.
+\newblock Nature Human Behaviour. 2021;5(9):1119--1122.
+\newblock doi:{10.1038/s41562-021-01137-1}.
+
+\bibitem{hallDesigningDiversityInclusion2019}
+Hall B. Designing a Diversity and Inclusion ({{D}}\&{{I}}) Communications
+  Strategy; 2019.
+\newblock Available from:
+  \url{https://www.interactsoftware.com/blog/diversity-inclusion-communications-strategy/}.
+
+\bibitem{andalibPostdocQueueLabour2018}
+Andalib MA, Ghaffarzadegan N, Larson RC.
+\newblock The {{Postdoc Queue}}: A {{Labour Force}} in {{Waiting}}.
+\newblock Systems Research and Behavioral Science. 2018;35(6):675--686.
+\newblock doi:{10.1002/sres.2510}.
+
+\bibitem{kaplanPostdocNot2012}
+Kaplan K.
+\newblock Postdoc or Not?
+\newblock Nature. 2012;483(7390):499--500.
+\newblock doi:{10.1038/nj7390-499a}.
+
+\bibitem{canelon2021cost}
+Canelón S. useR!2021 Cost Conversion Tool: Cost Conversion Tool; 2021.
+\newblock Available from:
+  \url{https://spcanelon.github.io/useR2021-cost-conversion-tool}.
+
+\bibitem{eventbrite_ultimate_2017}
+{Eventbrite}. The Ultimate Way to Reduce No-Shows at Free Events;.
+\newblock Available from:
+  \url{https://www.eventbrite.com/blog/asset/ultimate-way-reduce-no-shows-free-events/}.
+
+\end{thebibliography}
 

### 2022-03-09 response_reviewers.tex
Adding response to reviewers + comments by @yabellini

\documentclass{article}
\usepackage[utf8]{inputenc}
%\usepackage{standalone}
% \usepackage[breaklinks=true,colorlinks=true,linkcolor=blue,citecolor=black,urlcolor=blue]{hyperref} % url
\usepackage[top=0.85in,left=1in,footskip=0.75in]{geometry}
\usepackage[dvipsnames]{xcolor} % colors
% \definecolor{blue-pigment}{rgb}{0.2, 0.2, 0.6}
\usepackage[colorinlistoftodos,textwidth=4.3cm]{todonotes} % for comments
\usepackage[normalem]{ulem} % for sout (strikeout)
\usepackage{url}

\usepackage[colorinlistoftodos]{todonotes} % for comments

\usepackage{lineno}
%\usepackage{xr-hyper} 
\usepackage{xr} 
%\usepackage{hyperref} 
\newenvironment{Reply}{\noindent\color{BlueViolet}\textbf{Reply.}}{\vspace{1em}}

%\usepackage{xr} % for getting lines with modifs from the other doc
%\externaldocument[ms-]{plos_latex_template}
% In your preamble

\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{plos_latex_template}
%%%% defining colors for coauthors %%%%

\setlength{\parskip}{0.3cm plus 4mm minus 3mm} % space between paragraphs

\title{Ten simple rules to host an inclusive conference\\ \vspace{0.5cm}
\textbf{Response to reviewers}}
\author{}
\date{}

\begin{document}

\maketitle

When you are ready to resubmit, please upload the following:

[1] A letter containing a detailed list of your responses to all review comments, and a description of the changes you have made in the manuscript. Please note while forming your response, if your article is accepted, you may have the opportunity to make the peer review history publicly available. The record will include editor decision letters (with reviews) and your responses to reviewer comments. If eligible, we will contact you to opt in or out

[2] Two versions of the revised manuscript: one with either highlights or tracked changes denoting where the text has been changed; the other a clean version (uploaded as the manuscript file).

Important additional instructions are given below your reviewer comments.

 Thank you again for your submission to our journal. We hope that our editorial process has been constructive so far, and we welcome your feedback at any time. Please don't hesitate to contact us if you have any questions or comments.

Reviewer's Responses to Questions

Comments to the Authors:

\textbf{Reviewer \#1:}
I greatly enjoyed reading the paper "Ten simple rules to host an inclusive conference", and think it provides an easy to follow set of rules that organizers can use to ensure an inclusive conference. I found many tips for inclusiveness in the article to be especially helpful, such as multilingual accommodations and ensuring screen-reader accessibility.

I just have a few aspects I wasn't 100\% clear on.

(1) In Rule 4, the manuscript states "Define measurable indicators for these goals e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions", but how do organizers choose the correct demographics to monitor and the expected background rates? Should it match global, the local, or field-specific rates? Or should it be remediative, where it specifically selects for communities historically ignored in the field beyond the regional population rate?

\begin{Reply}
   We are going to refer to line numbers in the following way: In the plos\_latex\_template.tex document, add a line number label where you want to edit like this: \\ linelabel\{rev1-com1-a\}
   To refer to these lines in this document, add a reference here, like this: Lines   \ref{rev1-com1-a} to \ref{rev1-com1-b}
\end{Reply}

(2) Adding on to the first point, how are the indicators measured? Should the conference organizers quantify the demographics of the attendees/participants? If so, should these details be published to the participants after the conference? How do organizers hold themselves accountable? And to whom are they held accountable to; the organizing committee themselves or the greater community?

\begin{Reply}
   % Yani: I think we can mention and link to the info board and to the blog post post conference analyzing this numbers like the Latin American one and the accessibility one. The surveys are also another source of data for this accountability. I think this is related with the other review asking about governance.

\end{Reply}

(3) In Rule 4, it states "Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;" Does this pertain to invited speakers only, or speakers selected from submitted abstracts? If this pertains to scored abstracts, how does one ensure that this is done in a transparent and reproducible way? Or maybe I'm not fully grasping what curate means in this context?

\begin{Reply}

\end{Reply}

(4) In Rule 5, it states " The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants." How do the tablets help in-person attendees interact with remote participants? Especially participants that may be in a different time-zone. Or are the tablets used to help the online activities mentioned in the next paragraph?

(5) In Rule 7 it states " Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language." It is not trivial to submit an abstract in two languages, and if there is no chance that the non-English abstract won't be reviewed I feel like this would be wasting their time. Is it possible to denote the accepted languages ahead of time? Or for the submitter to put in a language request before the abstract deadline?

Overall, I think these rules are great. My biggest concern is on the concept of organizing committee accountability. They are defining what the diversity goals are and how they are measured, but it isn't clear to me what input the community has on the conference. I feel like the greater community should be able to have input into the organizing process as well as the ability to see if their concerns are being addressed.

\textbf{Reviewer \#2:}
L78-L80 As a female scientist I would ask myself "where are we" (because we still have parts of the World where female scientist are rare on leading positions), but I consider science apolitical in every segment, and I certainly suggest that it concentrate here on the scientific youth, the segment of research in industry, etc., and not on the above.
The rest of the paperit should certainly be praised because it enters into the essence of the exchange of knowledge today.

\textbf{Reviewer \#3:} 
This is a timely, well-written commentary on organising an inclusive conference and should act as a reference compendium for researchers, practitioners as well as conference organisers.

I have some comments for authors to consider, and possibly expand upon.

1. As much as I enjoyed reading the ten rules, I missed finding mention of how this is to be implemented and governed. Therefore, it is poignant that authors should brief discuss having an inclusion oversight or governance committee that tracks these metrics and their implementation.

2. Secondly, I also missed detailed methodology on how these ten rules were developed or consensus were reached at? Were stakeholders from diverse backgrounds consulted? Was the consultation process iterative? Were the recommendations thus made by those from diverse backgrounds incorporated? How were the final draft reached at - through consensus or voting or a process to that effect? This is important as for such a set of indications to be indicative, it is important that opinions from all backgrounds are taken into account and co-developed.

\begin{Reply}
   %Yani: I think this is not on the scope of the paper, but we can give an explanation of the process, which I think was fantastic and inclusive and have a diverse background on the construction of the rules. 
\end{Reply}

3. In Rule 1, though gender or ethnic diversity has been highlighted, geographical diversity needs to be discussed in detail. There is significant "north-south" divide in scientific research - with several of institutions as well as researchers from LMICs or global south not being given appropriate representation on conference committees as well as on speakers diversity.

4. In Rule 2, safe and inclusive environments, I find little attention to "racist" aberrations or incidents - and how they should be mitigated and when reported be appropriately addressed with fairness and justice.
Several speakers from ethnically diverse backgrounds face racism, exclusion or apathy at such conferences. This need to be discussed.

5. In Rule 4, via a vis bias, bias against people with disability and ethnic minorities should also be highlighted.

6. In Rule 8, inclusive communication strategy - it is important that the speakers from diverse backgrounds, especially from ethnic minorities etc are highlighted or given prominence in social media strategy or communication. Often, these speakers face relatively less focus. This may be discussed.

\begin{Reply}
   %Yani: I think all keynote deserve the _same_ communication, if we give prominence to minoritized group could be perceived as tokenism. So I would take the suggestion on NOT give more prominence to one group of keynote and make sure that all of them are promote in the same spaces and in the same way.  We can mention here that communities partners can help with the higlight of the minoritazed groups (tricky).
\end{Reply}

7. In Conclusion, suggest also adding that issues of diversity and inclusion have come to fore even more during COVID-19 as we witness exacerbation of inequities disadvantaging those from vulnerable backgrounds (see https://pubmed.ncbi.nlm.nih.gov/33343410/
PMID: 33343410)

\end{document}

### 2022-03-09 plos_latex_template.tex
Adding response to reviewers + comments by @yabellini

@@
 
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
-
 % line numbers
 \usepackage[right]{lineno}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{xr-hyper}%for cross ref in overleaf
+\usepackage{nameref,hyperref}
+%always load hyperref last
 
@@
 Batool Almazrouq\textsuperscript{11,12},
-Liz Hare\textsuperscript{13},
+Elizabeth Hare\textsuperscript{13},
 Laura Ación\textsuperscript{5,14},
@@
 Assess how lack of diversity translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
-Define measurable indicators for these goals (e.g.
-a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.
-The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.
+\linelabel{rev1-com1-a} \hl{Define measurable indicators for these goals (e.g.
+a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.}
+The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization. \linelabel{rev1-com1-b}
 
@@
 Define clear evaluation criteria and share them with the reviewers and authors.
-Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
+\hl{Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe}; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
 
@@
 To best integrate in-person and remote attendees, all questions and comments should pass through a unique online system. 
-The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.
+\hl{The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.}
 
@@
 Create an inclusive environment by encouraging the full participation of non-native English speakers. This may be done in multiple ways.
-Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.
-That way, the abstract will be judged primarily by the quality or relevance of the work instead of English proficiency.
+\hl{Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.} That way, the abstract will be judged primarily by the quality or relevance of the work instead of English proficiency.
 Whenever possible, identify other prominent languages for the conference and provide translated captions or live language interpretation into these key languages, including multilingual Q\&A sessions.

### 2022-03-14 response_reviewers.tex
comments to reviewers by @malfaro2 and @andreasancheztapia

@@
    To refer to these lines in this document, add a reference here, like this: Lines   \ref{rev1-com1-a} to \ref{rev1-com1-b}
+
+   %ast: We do not have a correct answer for this, unfortunately, and we believe there is a reflexion work that every team has to do in their context.
+We had extensive discussions about the indicators and our conclusion is that the "correct" demographic to monitor will vary for every conference, field, and local and regional context.
+   % Depending on the field, greater achievements can be made, and in some areas, this exercise will be more difficult and yet extremely important even if the apparent advancements are less evident. 
+   % Our team is inclined to propose what the reviewer calls "remediative measures", or affirnative actions, and be explicitly more than inclusive, but we understand that not every team can set such targets.
+And even with bold intentions, there is no guarantee that the final result is going to be as expected.
+   % A support from the community may be a great way to strenghen bold goals.
+However, in some communities the common sense shared by members may represent a wish to preserve the status quo, so not always a broad support from the community can be expected -but this may indicate also a need to broaden the definition of the community.
+   % We would like to encourage the readers of any field and regional context to make this exercise independently of the status of the field and the initial impression that these efforts won't be successful or sufficient.
+   % to edit: We have made this explicit in the text.
+We also included the fact that the community 1.
+should be defined and 2.
+can be consulted before the conference to make sure that this construction is collective and participative.
+
+%Marce: I agree with reviewer 1 that this should be remediative, quoting: "where it specifically selects for communities historically ignored in the field beyond the regional population rate".
+To respond to the same reviewer, I think we can mention how  UseR presented the demographics of the attendees/participants, and how this was shared with the community as a way to be held accountable.
+%Marce: About rule 10: both reviewers 1 and 3 suggested to discuss further how to implement these rules. I think rule 10 is already discussing that in a clear manner.
 \end{Reply}
@@
 I think this is related with the other review asking about governance.
+   %ast: The tenth rule is directly connected to the first, because these indicators can be measured and evaluated via surveys, direct impressions from the atendance, and exercises such as focus groups.
+Of all these options, surveys are the most frequently used but we wanted to highlight that there are other options.
+This link to rule 10 was not explicit in the text (edit!)
+   % About publishing the details after the conference, the privacy and anonymity of the participants should be respected, so some kind of binning/grouping of the data and care with categories that have a low number of respondents should be taken.
+%make sure this in the manuscript!
+we had this in previous versions.
+   % Accountability of the team should be posible with or without public indicators.
+Open channels for communication and open fields and the possibility to make anonymous comments should be provided.
+Especially if there is a claim for diversity and inclusion, the general community should be able to give feedback about the success of these efforts.
+   %to edit: cite the tenth rule in the first, telling there will be accountability. but also say that the indicators should not become a goal themselves.
 
 \end{Reply}
-
 (3) In Rule 4, it states "Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;" Does this pertain to invited speakers only, or speakers selected from submitted abstracts?
@@
 \begin{Reply}
-
+   %ast: Although curation is easier and definitely encouraged for invited speakers, we would like to reinforce that even transparent scored processes can be influenced by implicit bias, once the reviewers start evaluating there might be "shifting criteria" and other considerations entering mid-process.
+   %We have added some extra explanation about this curation, that include at least: 
+   %1.
+determining clear criteria before seeing the list of submissions, 2.
+examining the final list of selected work for imbalances in gender, geographical representation and other visible aspects of diversity.
+3.
+check if close competitors (similar quality work) were excluded for some apparent bias.
+   % This process is hard because there are many aspects of diversity that will be unreachable for reviewers and are prone to perceived cues rather than real information (so not every dicersity dimension can be considered here).
+The invitation is to examine similar quality work and try to understand why some work was included rather than other and to consider whether bias could have influenced the outcome.
+   %criteria should be clear before the abstract evaluation. Some adaptation of this can be made: https://honors.agu.org/files/2014/11/AWARDS-SUGGESTED-BEST-PRACTICES-MAY2011.pdf
 \end{Reply}
 
-(4) In Rule 5, it states " The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants." How do the tablets help in-person attendees interact with remote participants?
+(4) In Rule 5, it states "The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants." How do the tablets help in-person attendees interact with remote participants?
 Especially participants that may be in a different time-zone.
 Or are the tablets used to help the online activities mentioned in the next paragraph?
+
+\begin{Reply}
+   %ast: Although one can assume that an ever growing proportion of people have their own handheld device to interact online with the remote participants, we cannot assume that everyone does, or that the models are compatible with the options of the conference.
+Some devices (and support for those and the participants') should be provided by the organization.
+With everyone having a device, the idea is to unify the networking activities and Q & A sessions to be online, this more integrated with the remote component.
+\end{Reply}
 
@@
 Or for the submitter to put in a language request before the abstract deadline?
+
+\begin{Reply}
+%ast: "and" should be "or".
+Yes, definitely, rather than two abstracts, one in English and the other in a second language, the best way would be to be able to have abstracts in languages different than English without the need for a duplication.
+That depends on the conference capabilities but we would recommend both submission and presentation be open to other languages.
+The idea for a previous request/screening about the languages that will be needed is good.
+Another option is to have a language editing help for English, via the communities or via the organizing team, that helps ensure that the submitted abstracts are conveying the meaning the authors are trying to give.
+This could include also, abstracts written in English by students and early career researchers.
+\end{Reply}
 
@@
 
+\begin{Reply}
+%ast: We thank the reviewer's feedback.
+We have added some pre-conference participation of the community, also inspired by examples we got from colleagues at pycon Germany.
+The accountability is very hard to define because no formal structure requires that in the present, it has to be the team's will to be held accountable to their participant community.
+Clarity in the communication strategy and open spaces for feedback are key here.
+\end{Reply}
+
 \textbf{Reviewer \#2:}
 L78-L80 As a female scientist I would ask myself "where are we" (because we still have parts of the World where female scientist are rare on leading positions), but I consider science apolitical in every segment, and I certainly suggest that it concentrate here on the scientific youth, the segment of research in industry, etc., and not on the above.
-The rest of the paperit should certainly be praised because it enters into the essence of the exchange of knowledge today.
+The rest of the paper it should certainly be praised because it enters into the essence of the exchange of knowledge today.
+
+\begin{Reply}
+
+\end{Reply}
 
@@
 Therefore, it is poignant that authors should brief discuss having an inclusion oversight or governance committee that tracks these metrics and their implementation.
+
+\begin{Reply}
+%ast: my first hunch in general: the metrics part is one of the hardest part to implement and reviewer1 talked about it, but it is also one of the "goals" that stick with some readers and we were trying to avoid putting them at the center of decision-making because they can become a end in itself and not a way to understand if your efforts were successful
+%The tenth rule should have accounted for this but we may need to be more specific in the messages of not letting yourself be prey of the indicators. 
+%Team accountability and governance were part of rule 10 so maybe we could strengthen this.
+\end{Reply}
 
@@
    %Yani: I think this is not on the scope of the paper, but we can give an explanation of the process, which I think was fantastic and inclusive and have a diverse background on the construction of the rules.
+   %ast: i agree, this may break the format for a ten simple rules piece so maybe a very short mention in the introduction or as SupMat?
 \end{Reply}
@@
 
+\begin{Reply}
+
+%ast: All specific examples of offenses should be included in the code of conduct construction process, and that's why we didn't make a list here, but maybe we can make explicit mentions of what type of offenses can/have to be in the code of conduct.
+%suggestion for edits: add this, talk a little bit more about the content of the CoC text.
+
+\end{Reply}
+
 5. In Rule 4, via a vis bias, bias against people with disability and ethnic minorities should also be highlighted.
+
+\begin{Reply}
+%ast: We have tried to make a distinction between visible and invisible aspects of human diversity, and we have to respect the right of every individual to occupy the space and assert/disclaim who they are and the right to remain silent about some aspects of their lives.
+Rule 2 is about being proactive regarding all dimensions of diversity, and Rule 4 focuses more on the visible aspects of diversity, especifically those that can be grasped via written communication.
+% We mention this because counteracting bias actions during the selection of speakers and submitted work, we--unfortunately--only have *perceived* cues of these dimensions.
+We can cite Laura's piece on video abstracts being exclusionary and be more explicit about the fact that perceived cues and implicit and explicit racism go hand in hand.
+\end{Reply}
 
@@
 So I would take the suggestion on NOT give more prominence to one group of keynote and make sure that all of them are promote in the same spaces and in the same way.
-We can mention here that communities partners can help with the higlight of the minoritazed groups (tricky).
+We can mention here that communities partners can help with the highlight of the minoritazed groups (tricky).
+   %ast: I agree with Yani, and we had it partially covered in the rule 7, "Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same relevance as the rest of the program." but we can add something here.
 \end{Reply}

### 2022-03-17 response_reviewers.tex
Comments by @jupanago and @rociojoo, start editing by @andreasancheztapia

@@
 
+Dear Dr. Joo,
+
+Thank you very much for submitting your manuscript "Ten simple rules to host an inclusive conference" for consideration at PLOS Computational Biology.
+As with all papers reviewed by the journal, your manuscript was reviewed by members of the editorial board and by several independent reviewers.
+The reviewers appreciated the attention to an important topic.
+Based on the reviews, we are likely to accept this manuscript for publication, providing that you modify the manuscript according to the review recommendations.
+
+\begin{Reply}
+    Thank you blabla.
+\end{Reply}
+
 When you are ready to resubmit, please upload the following:
@@
 
+\begin{Reply}
+    Thank you blabla.
+\end{Reply}
+
 I just have a few aspects I wasn't 100\% clear on.
+
+% RJ: General comment about the responses to this reviewer.
+They tend to write several comments in one.
+I'd prefer to break them down in several bits and reply to each one, to make our responses clearer.
 
@@
    % Depending on the field, greater achievements can be made, and in some areas, this exercise will be more difficult and yet extremely important even if the apparent advancements are less evident. 
-   % Our team is inclined to propose what the reviewer calls "remediative measures", or affirnative actions, and be explicitly more than inclusive, but we understand that not every team can set such targets.
+   % Our team is inclined to propose what the reviewer calls "remediative measures", or affirmative actions, and be explicitly more than inclusive, but we understand that not every team can set such targets.
 And even with bold intentions, there is no guarantee that the final result is going to be as expected.
-   % A support from the community may be a great way to strenghen bold goals.
+   % A support from the community may be a great way to strenghten bold goals.
 However, in some communities the common sense shared by members may represent a wish to preserve the status quo, so not always a broad support from the community can be expected -but this may indicate also a need to broaden the definition of the community.
@@
 %Marce: About rule 10: both reviewers 1 and 3 suggested to discuss further how to implement these rules. I think rule 10 is already discussing that in a clear manner.
+
+%JPNG: Assessing the correct demographics depends on what information is available when defining inclusiveness goals.
+When DEI studies in the particular field of the conference are available, these should be used.
+Otherwise, global metrics in STEM, social sciences, and humanities should be used.
+I think a "remediative" selection of historically ignored communities must be the standard regardless of the availability of DEI studies in the field.
+
+%RJ: I agree with everyone. Marce has a good point. We could add to the text an example about useR! but I'm wondering if we should be "useR!-explicit" in every rule.
 \end{Reply}
 
-(2) Adding on to the first point, how are the indicators measured?
-Should the conference organizers quantify the demographics of the attendees/participants?
+(2) Adding on to the first point, how are the indicators measured? Should the conference organizers quantify the demographics of the attendees/participants?
+
+\begin{Reply}
+\end{Reply}
+
 If so, should these details be published to the participants after the conference? 
-How do organizers hold themselves accountable?
-And to whom are they held accountable to; the organizing committee themselves or the greater community?
+\begin{Reply}
+% About publishing the details after the conference, the privacy and anonymity of the participants should be respected, so some kind of binning/grouping of the data and care with categories that have a low number of respondents should be taken.
+%make sure this in the manuscript!
+we had this in previous versions.
+\end{Reply}
+
+How do organizers hold themselves accountable? And to whom are they held accountable to; the organizing committee themselves or the greater community?
 
@@
 Of all these options, surveys are the most frequently used but we wanted to highlight that there are other options.
-This link to rule 10 was not explicit in the text (edit!)
-   % About publishing the details after the conference, the privacy and anonymity of the participants should be respected, so some kind of binning/grouping of the data and care with categories that have a low number of respondents should be taken.
-%make sure this in the manuscript!
-we had this in previous versions.
-   % Accountability of the team should be posible with or without public indicators.
+This link to rule 10 was not explicit in the text (edit!) - RJ: Yes, but let's try to give as much practical advice as possible.
+
+   % Accountability of the team should be possible with or without public indicators.
 Open channels for communication and open fields and the possibility to make anonymous comments should be provided.
 Especially if there is a claim for diversity and inclusion, the general community should be able to give feedback about the success of these efforts.
-   %to edit: cite the tenth rule in the first, telling there will be accountability. but also say that the indicators should not become a goal themselves.
+   %to edit: cite the tenth rule in the first, telling there will be accountability.
+but also say that the indicators should not become a goal themselves.
+% RJ: sounds like a blog post we didn't write.
 
+% JPNG: 
+% This is a good question.
+I think this was put into practice during UseR2021.
+I think that accountability should be self-referenced by looking at previous successes of the same conference and by comparison with the DEI local/global measures available in the field.
+By publishing these results to the participants, the organising committee becomes accountable both to the greater community and the committee itself.
+To increase community participation, a final survey about the community's perception of DEI goals of the meeting could be done.
+We would need to think about how to compare these perceptions with the metrics taken by the organising committee.
+Questions about "nice to haves" for the next conference can be a good way to include the community in the medium/long term.
 \end{Reply}
+
 (3) In Rule 4, it states "Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;" Does this pertain to invited speakers only, or speakers selected from submitted abstracts?
@@
 \begin{Reply}
-   %ast: Although curation is easier and definitely encouraged for invited speakers, we would like to reinforce that even transparent scored processes can be influenced by implicit bias, once the reviewers start evaluating there might be "shifting criteria" and other considerations entering mid-process.
-   %We have added some extra explanation about this curation, that include at least: 
+This specific section refers to the final selection of submitted works.
+The invited roles were mentioned previously; counteracting bias in invited roles was mentioned earlier in the text in Line \ref{invited}.
+We agree with the reviewer that `curation' may be a misleading term.
+We would like to reinforce that even transparent scored processes can be influenced by implicit bias, once the reviewers start evaluating there might be "shifting criteria" and other considerations entering mid-process.
+We edited that section to be explicit that the information should be as anonymous as possible (Line \ref{anonymous}), that reviewers should receive and understand the criteria before seeing the submission list (Line \ref{criteria}), and that the final list should be examined (Line \ref{not-curation}):
+`Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias', especially to be sure that some close contenders were not judged differently.
    %1.
@@
 check if close competitors (similar quality work) were excluded for some apparent bias.
-   % This process is hard because there are many aspects of diversity that will be unreachable for reviewers and are prone to perceived cues rather than real information (so not every dicersity dimension can be considered here).
-The invitation is to examine similar quality work and try to understand why some work was included rather than other and to consider whether bias could have influenced the outcome.
-   %criteria should be clear before the abstract evaluation. Some adaptation of this can be made: https://honors.agu.org/files/2014/11/AWARDS-SUGGESTED-BEST-PRACTICES-MAY2011.pdf
+   %  https://honors.agu.org/files/2014/11/AWARDS-SUGGESTED-BEST-PRACTICES-MAY2011.pdf
+
+   % JPNG: This point is important.
+I think this "curation" is more relevant to the more visible keynotes who are usually those that become role models and attract participants to join the conference.
+However, increasing representation and diversity at the level of the submitted abstracts is desirable.
+I think this second point is more difficult and can be approached only by proactively inviting researchers and students from less privileged backgrounds to participate in the conference.
+Fee weavers and networking events for these specific communities seem key to me.
+To clarify what "curation" means, we could say "review the list of keynote speakers and make changes to maximise DEI" or something like that.
+Instead of using a word we could paraphrase the concept.
+
+   %RJ: I agree with Juan Pablo. Curation may sound odd. Let's be more explicit.
+   %ast. ok, made some changes but this section was referring only to the final selection of abstracts, not invited roles, and i said so
 \end{Reply}
@@
 Some devices (and support for those and the participants') should be provided by the organization.
-With everyone having a device, the idea is to unify the networking activities and Q & A sessions to be online, this more integrated with the remote component.
+With everyone having a device, the idea is to unify the networking activities and Q & A sessions to be online, thus more integrated with the remote component.
+
+   %JPNG: Some ideas: 1) tablets/laptops available at the physical venue can be used to allow physical and digital participants to interact through chat rooms and video meetings both during main and satellite conference events.
+2) We should highlight somehow that this encounter between the physical and digital dimensions of a conference is a source of innovation.
+Thinking creatively about how to use technology to design engaging interactions can open the way to new forms of interaction we have not thought of before.
+
+   %RJ: Agreed with both. I like Juan Pablo's tone. 
 \end{Reply}
@@
 This could include also, abstracts written in English by students and early career researchers.
+
+%JPNG: I disagree.
+Writing a non-English language abstract is not a waste of time even if there is a slight chance of not having someone to review it.
+This comment forgets that we write for the community, which is multilingual, and not only for ourselves.
+I think we could denote main languages instead of accepted languages only to stress those languages that have a greater chance of being reviewed.
+
+% RJ: I agree with both. I'm just wondering if it's necessary to add something to the text in case we think many people with linguistic privileges would think the same way
 \end{Reply}
@@
 Clarity in the communication strategy and open spaces for feedback are key here.
+%RJ: Could you paste here some of those things you're thinking about?
+
+%JPNG: I agree.
+The input the community has on the conference is key.
+I suggested above that surveys and "nice to haves" can be among the possible options.
+DEI processes are community learning endeavours.
+We should stress that new ways of community participation will emerge in the process of understanding the needs of the community and better ways to increase diversity and representation.
 \end{Reply}
@@
 \begin{Reply}
+Thanks for reading and reviewing. The list given was meant to be an example of extant thematic groups and communities
+%i'm tempted to add this but feel free to say if we do:
+and we aimed to go beyond gender equality. 
+We have added "women in science" as an example of such communities inside the parenthesis in Line  \ref{wis}.
 
+% RJ: If that doesn't hurt, we can do what Juan Pablo suggests and ignore everything else.
+Unless we want to suggest her some reading and training.
+We can start with "thanks for reading and reviewing" to be polite.
 \end{Reply}
@@
 This is a timely, well-written commentary on organising an inclusive conference and should act as a reference compendium for researchers, practitioners as well as conference organisers.
+
+\begin{Reply}
+    We thank reviewer 3 for their comments and careful revision 
+\end{Reply}
 
@@
 \begin{Reply}
-%ast: my first hunch in general: the metrics part is one of the hardest part to implement and reviewer1 talked about it, but it is also one of the "goals" that stick with some readers and we were trying to avoid putting them at the center of decision-making because they can become a end in itself and not a way to understand if your efforts were successful
+Many teams do work by having an DEI committee--and we did too, led by the two first authors of this manuscript and composed by seven other coauthors \url{https://user2021.r-project.org/about/global-team/}.
+However, the whole organizing team should be involved in the goal of broadening participation and the circumstances will vary from team to team, so we haven't made specific recommendations regarding team organization and governance.
+Line \ref{team}
+%ast: i changed my mind because i think i misunderstood this comment.
+a brief edit about subteams can be added (although we had hints regarding this) in the team rule.
+but i think we agree that the existence of a DEI team is not 1.
+necessary 2.
+sufficient if the whole team is onboard accross conference areas.
 %The tenth rule should have accounted for this but we may need to be more specific in the messages of not letting yourself be prey of the indicators. 
 %Team accountability and governance were part of rule 10 so maybe we could strengthen this.
+
+%JPNG: Please, correct me if I am wrong, but I think it is clear that this paper is addressed to the organising committees.
+
+% RJ: Agreed with Andrea. We need to address accountability a bit more. Briefly though.
+
 \end{Reply}
@@
    %ast: i agree, this may break the format for a ten simple rules piece so maybe a very short mention in the introduction or as SupMat?
+
+   % JPNG: Not sure that the "Ten rules" format requires a detailed methodology.
+However, we could include a few lines in the second paragraph of the introduction to briefly describe how we wrote the paper.
+Rocío and Andrea, do you think this is feasible and desirable?
+
+   % RJ: Yep. We can do this and add it to the Intro.
 \end{Reply}
@@
 There is significant "north-south" divide in scientific research - with several of institutions as well as researchers from LMICs or global south not being given appropriate representation on conference committees as well as on speakers diversity.
+
+\begin{Reply}
+   % JPNG: I think this reviewer is missing the point of this paper.
+We are not diagnosing a situation, we are showing how things could be done if DEI is our goal based on a first experience of making a conference as inclusive as possible.
+Global North and the Global South are problematic terms.
+It is not the geography but the power relationships which is the key to the discussion.
+I would suggest discussing better if it is worth it to include these terms in the paper.
+
+   % RJ: Talking about regions should be enough.
+100% agreed with Juan Pablo about global north and global south.
+So I would be in favor of not adding anything to the paper and just reply to the reviewer here.
+It'd be good if we can back our response with some refs.
+\end{Reply}
 
@@
 
+% JPNG: This is what the code of conduct is for.
+This reviewer is asking for a detailed account of some aspects that are beyond the scope of the paper.
+What the reviewer is asking us to refer to is already writtem in the text but uses different words and phrasing.
+
+% RJ: Agreed. Out of the scope. We could, eventually, cite examples of code of conducts enumerating unacceptable behavior (some that mention racism), without having to mention racism ourselves.
+
 \end{Reply}
@@
 We can cite Laura's piece on video abstracts being exclusionary and be more explicit about the fact that perceived cues and implicit and explicit racism go hand in hand.
+
+% JPNG: I think this is already there.
+
+% RJ: Agreed with Andrea but we need to carefully write a nice yet simple and convincing response.
 \end{Reply}
@@
    %ast: I agree with Yani, and we had it partially covered in the rule 7, "Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same relevance as the rest of the program." but we can add something here.
+% RJ: Agreed.
+
+   % JPNG: I think this is already there.   
 \end{Reply}
@@
 
+\begin{Reply}
+    % RJ: We could try to add a line about this in the intro.
+\end{Reply}
+
 \end{document}

### 2022-03-17 plos_latex_template.tex
Comments by @jupanago and @rociojoo, start editing by @andreasancheztapia

@@
 Open the schedule for events organized by thematic groups and communities (e.g.
-task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.
+task force meetings, LGBTQIA+-friendly spaces, Black in STEM\hl{, women in science}\linelabel{wis}), and sessions aimed to welcome first-timers to the conference.
 
@@
 The regional and local communities, groups, or associations in your field are good sources to tap into. 
-When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably.
+When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. \linelabel{team}
 Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
@@
 When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
+\linelabel{invited}
 Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.
@@
 Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
-Evaluate the plausibility of anonymizing all of the submission information to
+Evaluate the plausibility of anonymizing \hl{as much of the submission information as possible} \linelabel{anonymous}, and
 define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
-Define clear evaluation criteria and share them with the reviewers and authors.
-\hl{Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe}; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
+Define clear evaluation criteria and share them with the reviewers and authors \hl{before they read any of the submissions} \linelabel{criteria}.
+\hl{Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias} \linelabel{not-curation}; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
 

### 2022-03-19 response_reviewers.tex
RJ: wrote some responses and edited/commented Andrea's

@@
 
+\section*{Editorial Summary}
+
 Dear Dr. Joo,
 
-Thank you very much for submitting your manuscript "Ten simple rules to host an inclusive conference" for consideration at PLOS Computational Biology.
+Thank you very much for submitting your manuscript `Ten simple rules to host an inclusive conference' for consideration at PLOS Computational Biology.
 As with all papers reviewed by the journal, your manuscript was reviewed by members of the editorial board and by several independent reviewers.
@@
 \begin{Reply}
-    Thank you blabla.
+    Thank you very much for these good news, and special gratitude to the members of the editorial board and the independent reviewers for the positive reviews.
 \end{Reply}
@@
 
-Reviewer's Responses to Questions
-
-Comments to the Authors:
-
-\textbf{Reviewer \#1:}
+\section*{Reviewer \#1:}
 I greatly enjoyed reading the paper "Ten simple rules to host an inclusive conference", and think it provides an easy to follow set of rules that organizers can use to ensure an inclusive conference.
@@
 \begin{Reply}
-    Thank you blabla.
+    Thank you for the thorough and positive review.
 \end{Reply}
@@
 
-(1) In Rule 4, the manuscript states "Define measurable indicators for these goals e.g.
-a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions", but how do organizers choose the correct demographics to monitor and the expected background rates?
+(1) In Rule 4, the manuscript states `Define measurable indicators for these goals e.g.
+a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions', but how do organizers choose the correct demographics to monitor and the expected background rates?
 Should it match global, the local, or field-specific rates?
@@
 \begin{Reply}
-   We are going to refer to line numbers in the following way: In the plos\_latex\_template.tex document, add a line number label where you want to edit like this: \\ linelabel\{rev1-com1-a\}
-   To refer to these lines in this document, add a reference here, like this: Lines   \ref{rev1-com1-a} to \ref{rev1-com1-b}
+    % RJ: This is actually rule 1.
+   The reviewer asked very important questions, for which there are no correct answers or formulas. 
+   The `correct' demographic will vary for every conference, field, and local and regional contexts. 
+   We believe that every meeting committee or organizing team will have to ask themselves these questions.
+   We now explicitly mention the need of this reflection and a suggestion to involve the community (the subset of the general public that would ideally attend the conference) early on as a way to understand its current state and make the goal setting phase a participative process (lines \ref{indicators-1} to \ref{indicators-2}).
 
-   %ast: We do not have a correct answer for this, unfortunately, and we believe there is a reflexion work that every team has to do in their context.
-We had extensive discussions about the indicators and our conclusion is that the "correct" demographic to monitor will vary for every conference, field, and local and regional context.
-   % Depending on the field, greater achievements can be made, and in some areas, this exercise will be more difficult and yet extremely important even if the apparent advancements are less evident. 
-   % Our team is inclined to propose what the reviewer calls "remediative measures", or affirmative actions, and be explicitly more than inclusive, but we understand that not every team can set such targets.
-And even with bold intentions, there is no guarantee that the final result is going to be as expected.
+   Our team agrees with implementing "remediative measures", or affirmative actions, and believe they should be the standard, but we understand that not every team can set such targets from the beginning.
+% RJ: Explain more what you mean about understading that not every team can set such targets from the beginning
+
    % A support from the community may be a great way to strenghten bold goals.
@@
 
-%Marce: I agree with reviewer 1 that this should be remediative, quoting: "where it specifically selects for communities historically ignored in the field beyond the regional population rate".
-To respond to the same reviewer, I think we can mention how  UseR presented the demographics of the attendees/participants, and how this was shared with the community as a way to be held accountable.
+%Marce:To respond to the same reviewer, I think we can mention how  UseR presented the demographics of the attendees/participants, and how this was shared with the community as a way to be held accountable.
 %Marce: About rule 10: both reviewers 1 and 3 suggested to discuss further how to implement these rules. I think rule 10 is already discussing that in a clear manner.
@@
 Otherwise, global metrics in STEM, social sciences, and humanities should be used.
-I think a "remediative" selection of historically ignored communities must be the standard regardless of the availability of DEI studies in the field.
-
 %RJ: I agree with everyone. Marce has a good point. We could add to the text an example about useR! but I'm wondering if we should be "useR!-explicit" in every rule.
@@
 \begin{Reply}
+The evaluation of the indicators is part of Rule 10, and can be done via: 
+`the information that you have already collected from your organizing team, speakers, and participants' (line \ref{info}) and `Surveys, focal groups, and informal conversations during and after the conference' (line \ref{info2}).
+To avoid repetition of this evaluation methods, we have added in Rule 1 an explicit link to Rule 10 in line \ref{rule10}, to let the reader know that these indicators will be used as part of the conference wrap up and continuity.
+%This mirrors the link to Rule 1 in Rule 10. % RJ: I would remove the last sentence.
 \end{Reply}
@@
 If so, should these details be published to the participants after the conference?
+
 \begin{Reply}
-% About publishing the details after the conference, the privacy and anonymity of the participants should be respected, so some kind of binning/grouping of the data and care with categories that have a low number of respondents should be taken.
+Yes, the final assessment of the success of the inclusion actions should be shared with everyone that attended and helped create these goals, as well as with people who would use the data in future conferences.
+We added more details about this in Rule 10, line \ref{data-share}:
+`The feedback received from the participants and your whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendants, and future organizers.'
+However, before publishing this information after the conference, the privacy and anonymity of the participants should be respected, so some kind of binning/grouping of the data should be performed to take care with categories that have a low number of respondents.
 %make sure this in the manuscript!
 we had this in previous versions.
+i left a note close to  \linelabel{data-share} % RJ: I agree.
+We thank the reviewer for this question because it allows us to clarify details around the definition of the indicators and their evaluation. 
 \end{Reply}
@@
 \begin{Reply}
+% Accountability in the context of conferences is a difficult concept, because historically there are some situations in which no real consequences whenever something goes wrong. 
+% For most aspects, we can only trust that organizing teams that decide to build more inclusive conferences have a commitment to their community.
+% RJ: The first two sentences are a bit negative, assuming that the reviewer is thinking about bad organizing committees.
+Let's not go that route.
+I also think that the reviewer is referring to the use of accountable in Rule 1, so don't start by Rule 8 or 10.
+%We have used this word sparingly, regarding Rule 10 and the transparency needed for responsible communication but 
+There are many accountability actions that we have included throughout the text: 
+As we said, the main tool for accountability (taking responsibility) is the Code of Conduct and having a team to enforce it.
+Other actions include making public the diversity statement and the accessibility guidelines (Rule 8), to make clear to the attendants how the conference experience has been designed and who they can contact.
+For example in Line \ref{accountability-1} `Be transparent with potential attendees and communicate limitations [...] to let them know what to expect, and ways in which you will try to mitigate these issues.
+[...] Provide a point of contact to help clarify any questions regarding accessibility, financial support, and others.'
+Rule 10 is also about evaluating internally and with the community whether the goals you set were met. 
+% RJ: Andrea, I'm assuming you haven't finished with this reply so I will stop there for now.
    % Yani: I think we can mention and link to the info board and to the blog post post conference analyzing this numbers like the Latin American one and the accessibility one.
@@
 I think this is related with the other review asking about governance.
-   %ast: The tenth rule is directly connected to the first, because these indicators can be measured and evaluated via surveys, direct impressions from the atendance, and exercises such as focus groups.
-Of all these options, surveys are the most frequently used but we wanted to highlight that there are other options.
-This link to rule 10 was not explicit in the text (edit!) - RJ: Yes, but let's try to give as much practical advice as possible.
 
@@
 
-(3) In Rule 4, it states "Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;" Does this pertain to invited speakers only, or speakers selected from submitted abstracts?
+(3) In Rule 4, it states `Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;' Does this pertain to invited speakers only, or speakers selected from submitted abstracts?
 If this pertains to scored abstracts, how does one ensure that this is done in a transparent and reproducible way?
@@
 This specific section refers to the final selection of submitted works.
-The invited roles were mentioned previously; counteracting bias in invited roles was mentioned earlier in the text in Line \ref{invited}.
+The invited roles were mentioned previously; counteracting bias in invited roles was mentioned earlier in the text in line \ref{invited}.
 We agree with the reviewer that `curation' may be a misleading term.
-We would like to reinforce that even transparent scored processes can be influenced by implicit bias, once the reviewers start evaluating there might be "shifting criteria" and other considerations entering mid-process.
-We edited that section to be explicit that the information should be as anonymous as possible (Line \ref{anonymous}), that reviewers should receive and understand the criteria before seeing the submission list (Line \ref{criteria}), and that the final list should be examined (Line \ref{not-curation}):
-`Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias', especially to be sure that some close contenders were not judged differently.
+We rephrased it to `Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias' (lines \ref{criteria}-\ref{not-curation}).
+This manual examination at a final stage is unlikely to be reproducible.
+Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts (lines \ref{criteria-1}-\ref{criteria}).
+% RJ: I'm not against the idea of adding more information, but the text below does not help to make things clearer, in my opinion.
+
+% We would like to reinforce that even transparent scored processes can be influenced by implicit bias, once the reviewers start evaluating there might be "shifting criteria" and other considerations entering mid-process.
+% RJ: This last sentence is a bit confusing.
+What do you want to communicate exactly?
+% We edited that section to be explicit that the information should be as anonymous as possible (line \ref{anonymous}), that reviewers should receive and understand the criteria before seeing the submission list (line \ref{criteria}), and that the final list should be examined (line \ref{not-curation}):
+% `Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias', especially to be sure that some close contenders were not judged differently.
    %1.
@@
 
-(4) In Rule 5, it states "The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants." How do the tablets help in-person attendees interact with remote participants?
+(4) In Rule 5, it states `The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.' How do the tablets help in-person attendees interact with remote participants?
 Especially participants that may be in a different time-zone.
@@
 \begin{Reply}
-   %ast: Although one can assume that an ever growing proportion of people have their own handheld device to interact online with the remote participants, we cannot assume that everyone does, or that the models are compatible with the options of the conference.
-Some devices (and support for those and the participants') should be provided by the organization.
-With everyone having a device, the idea is to unify the networking activities and Q & A sessions to be online, thus more integrated with the remote component.
 
-   %JPNG: Some ideas: 1) tablets/laptops available at the physical venue can be used to allow physical and digital participants to interact through chat rooms and video meetings both during main and satellite conference events.
-2) We should highlight somehow that this encounter between the physical and digital dimensions of a conference is a source of innovation.
-Thinking creatively about how to use technology to design engaging interactions can open the way to new forms of interaction we have not thought of before.
-
-   %RJ: Agreed with both. I like Juan Pablo's tone. 
+Although one can assume that an ever growing proportion of people have their own handheld device to interact online with the remote participants, we cannot assume that everyone does, or that the models are compatible with the software platforms used in the conference.
+If the conference organizers provides such devices, Q\&A sessions and networking activities can be online and allow for more integration between in-person and online participants. 
+Thinking creatively about how to use technology to design engaging interactions can open the way to new forms of interaction we have not thought of before.  %RJ: Should we keep this last sentence?
+%RJ: I'm not convinced that we need to add anything to the text in the manuscript, but if you see a way to do that, great.
 \end{Reply}
@@
 
-\textbf{Reviewer \#2:}
+\section*{Reviewer \#2:}
 L78-L80 As a female scientist I would ask myself "where are we" (because we still have parts of the World where female scientist are rare on leading positions), but I consider science apolitical in every segment, and I certainly suggest that it concentrate here on the scientific youth, the segment of research in industry, etc., and not on the above.
@@
 \begin{Reply}
-Thanks for reading and reviewing. The list given was meant to be an example of extant thematic groups and communities
-%i'm tempted to add this but feel free to say if we do:
-and we aimed to go beyond gender equality. 
-We have added "women in science" as an example of such communities inside the parenthesis in Line  \ref{wis}.
-
-% RJ: If that doesn't hurt, we can do what Juan Pablo suggests and ignore everything else.
-Unless we want to suggest her some reading and training.
-We can start with "thanks for reading and reviewing" to be polite.
+    Thanks for reading and reviewing. The list given was meant to be an example of extant thematic groups and communities, aiming to go beyond gender equality. 
+    We agree about adding women as an example of such groups in line \ref{wis}:
+    `Open the schedule for events organized by thematic groups and communities (e.g.
+task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.'
 \end{Reply}
 
-\textbf{Reviewer \#3:} 
+\section*{Reviewer \#3:} 
 This is a timely, well-written commentary on organising an inclusive conference and should act as a reference compendium for researchers, practitioners as well as conference organisers.
@@
 \begin{Reply}
-    We thank reviewer 3 for their comments and careful revision 
+    We thank the reviewer for their comments and careful revision.
 \end{Reply}
@@
 \begin{Reply}
-Many teams do work by having an DEI committee--and we did too, led by the two first authors of this manuscript and composed by seven other coauthors \url{https://user2021.r-project.org/about/global-team/}.
+Organizing teams often have DEI committee--and we did too, led by the two first authors of this manuscript and composed by seven other coauthors \url{https://user2021.r-project.org/about/global-team/}.
 However, the whole organizing team should be involved in the goal of broadening participation and the circumstances will vary from team to team, so we haven't made specific recommendations regarding team organization and governance.
@@
 sufficient if the whole team is onboard accross conference areas.
+% RJ: I agree that it is not necessary and certainly not sufficient.
+But it can really help.
+And honestly I would've not done so much work on diversity nor learned as much as I did if I had not been part of this team.
+So we could suggest having one if there's not one already.
+But it should not be such a strong recommendation, because the most important thing is that the whole organizing team should be committed to DEI (please don't argue that we had a DAI instead).
+We can connect the latter idea with accountability (briefly), also mentioned in the responses to reviewer 1.
 %The tenth rule should have accounted for this but we may need to be more specific in the messages of not letting yourself be prey of the indicators. 
@@
 \begin{Reply}
-   %Yani: I think this is not on the scope of the paper, but we can give an explanation of the process, which I think was fantastic and inclusive and have a diverse background on the construction of the rules.
-   %ast: i agree, this may break the format for a ten simple rules piece so maybe a very short mention in the introduction or as SupMat?
-
-   % JPNG: Not sure that the "Ten rules" format requires a detailed methodology.
-However, we could include a few lines in the second paragraph of the introduction to briefly describe how we wrote the paper.
-Rocío and Andrea, do you think this is feasible and desirable?
-
-   % RJ: Yep. We can do this and add it to the Intro.
+    The rules were first drafted by the first author; then the leading group and the members of the diversity, accessibility, and inclusion team were invited to write their own ten simple rules—everyone's rules are in XXX.
+    The rules in this manuscript are a result of all of these rules and further discussions with and contributions from all the coauthors, fueled by personal experiences and literature review.
+    This is now stated in lines \ref{methods-a}-\ref{methods-b} of the current version of the manuscript.
+    The rules got to their final state by consensus, not popularity. 
+    The coauthors are from a diversity of backgrounds, career stages, and even continents, and have experienced inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege, as stated in lines \ref{background-a}-\ref{background-b}.
+    This diversity, as well as the awareness and experience of organizing a conference that aimed to be a kind, inclusive, and accessible experience for as many people as possible, made the manuscript as diverse and inclusive as it is now.
+    We kept the methodology as brief as possible to keep the focus on the rules and not our team or conference. 
 \end{Reply}
@@
 \begin{Reply}
-    % RJ: We could try to add a line about this in the intro.
+% RJ: I'm leaning towards not adding this now.
+    We appreciate the reviewers' suggestion and agree that the COVID-19 pandemic contributed to the exacerbation of inequity. 
+    However, we would like the manuscript to help inspire and guide the organization of more inclusive conferences regardless of the state of the pandemic in the different places in the world, and even when the pandemic subsides.
+For that reason, we have tried to mention the pandemic very few times and only when deemed necessary: to give context to the conference that we organized, and to talk about virtual conferences which became the prevailing format during the last years.
 \end{Reply}

### 2022-03-19 plos_latex_template.tex
RJ: wrote some responses and edited/commented Andrea's

@@
 Many academic and tech conferences have been spaces that reproduce systemic inequalities, by failing to overcome the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income backgrounds, prestigious institutions, native English speakers with no disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
-Like many others, the authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
+\linelabel{background-a}Like many others, the authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
+\linelabel{background-b}
 We had the opportunity to organize a virtual and global statistical computing conference, useR!
@@
 Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
+\linelabel{methods-a} \hl{The rules were first drafted by the first author; then the leading group and the members of the diversity, accessibility, and inclusion team were invited to write their own ten simple rules---everyone's rules are in XXX.
+The rules in this manuscript are a result of all of these rules and further discussions with and contributions from all the coauthors, fueled by personal experiences and literature review.} \linelabel{methods-b}
 They are organized in three sections: three foundational rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
@@
 Assess how lack of diversity translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
-\linelabel{rev1-com1-a} \hl{Define measurable indicators for these goals (e.g.
-a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.}
-The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization. \linelabel{rev1-com1-b}
+%depends on what information is available when defining inclusiveness goals.
+When DEI studies in the particular field of the conference are available, these should be used.
+Otherwise, global metrics in STEM, social sciences, and humanities should be used.
+Other indicators, like gender balance, can be set without the need for concrete studies in the area.
+\linelabel{indicators-1} \hl{The result of this assessment will vary between areas of knowledge and depending on the geographic context of the conference. 
+Involving the community---that is, the subset of the general public that would ideally attend the conference---%and for whom you are organizing the conference
+early in this stage can help you set these goals in a collective and participative way.%be very revealing of their perception about the people who usually get to attend the conferences in the area.
+not this.
+judgemental
+} \linelabel{indicators-2}
+
+Define measurable indicators for these goals (e.g.
+a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions)
+that will help you be accountable along the way
+\linelabel{rule10} \hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
+The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.
 
@@
 Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
-Evaluate the plausibility of anonymizing \hl{as much of the submission information as possible} \linelabel{anonymous}, and
+Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
 define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
-Define clear evaluation criteria and share them with the reviewers and authors \hl{before they read any of the submissions} \linelabel{criteria}.
+\linelabel{criteria-1}Define clear evaluation criteria and share them with the reviewers and authors \hl{preferably during the call for abstracts} \linelabel{criteria}.
 \hl{Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias} \linelabel{not-curation}; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
@@
 Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}) to communicate your values and practices, and to hold yourself accountable for them.
-Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. 
+\linelabel{accountability-1}Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues.
 For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering help to navigate it, or if captions will be available for some talks but not all.
@@
 You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
-It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and participants (e.g.
+It is likely that you will be able to calculate some of your indicators directly from \linelabel{info}the information that you have already collected from your organizing team, speakers, and participants (e.g.
 countries of origin, or preferred language).
 You may have goals concerning how welcomed and included participants felt during the conference, and wonder if your practices translated into real inclusion. 
-Surveys, focal groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
+\linelabel{info2}Surveys, focal groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
 
 % Pass the information
-The feedback received from the participants and your whole assessment of the conference should be documented and shared with future organizers. 
+The feedback received from the participants and your whole assessment of the conference should be documented and shared with \linelabel{data-share}\hl{those who helped define the inclusion goals, the attendants, and} future organizers.
+%ast: Rocío, we had in previous versions a part talking about data privacy and anonymization, maybe we could add a quick mention to this here, share with whom (and add the community) with some precaution.
 Gather and do your best to organize all the documentation produced during the organization process—or even after.
@@
 \begin{itemize}
-    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization, Extended abstract translation
-    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
-    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
-    \item YBS: Conceptualization, Writing--major reviews and edits, Resources, Extended abstract translation
-    \item HT: Conceptualization, Writing--major reviews and edits, Resources
-    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules, Extended abstract translation
-    \item NSM: Writing--major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
+    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing---major reviews and edits, Resources, Visualization, Extended abstract translation
+    \item AST: Project administration, Supervision, Conceptualization,  Writing---major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
+    \item SM: Conceptualization, Writing---major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
+    \item YBS: Conceptualization, Writing---major reviews and edits, Resources, Extended abstract translation
+    \item HT: Conceptualization, Writing---major reviews and edits, Resources
+    \item DHP: Conceptualization, Writing---major reviews and edits, Sharing rules, Extended abstract translation
+    \item NSM: Writing---major reviews and edits, Resources, Visualization, Sharing rules, Extended abstract translation
     \item MB: Conceptualization, Extended abstract translation
-    \item BA: Writing--major reviews and edits, Resources, Sharing rules, Extended abstract translation
-    \item LH: Writing--major reviews and edits, Resources, Sharing rules
-    \item LA: Writing--major reviews and edits, Resources
-    \item JPN: Writing--major reviews and edits, Visualization
-    \item MAC: Writing--major reviews and edits
-    \item FM: Writing--major reviews and edits, Extended abstract translation
-    \item RG: Writing--major reviews and edits, Resources, Extended abstract translation
+    \item BA: Writing---major reviews and edits, Resources, Sharing rules, Extended abstract translation
+    \item LH: Writing---major reviews and edits, Resources, Sharing rules
+    \item LA: Writing---major reviews and edits, Resources
+    \item JPN: Writing---major reviews and edits, Visualization
+    \item MAC: Writing---major reviews and edits
+    \item FM: Writing---major reviews and edits, Extended abstract translation
+    \item RG: Writing---major reviews and edits, Resources, Extended abstract translation
     \item SC: Resources, Sharing rules
@@
     \item JC: Sharing rules
-    \item JR: Conceptualization, Writing--major reviews and edits, Visualization, Extended abstract translation
+    \item JR: Conceptualization, Writing---major reviews and edits, Visualization, Extended abstract translation
 \end{itemize}

### 2022-04-14 response_reviewers.tex
Finished response to reviewers. This version was sent to coauthors :tada:

--- 2022-03-14 response_reviewers.tex
@@
 \usepackage{url}
-
+\usepackage{soul}
 \usepackage[colorinlistoftodos]{todonotes} % for comments
@@
 
+\section*{Note from the authors}
+
+Dear Dr. Markel,
+
+Please find the responses to all the reviewers comments and suggestions below in this document.
+We also highlighted our changes in the manuscript and added line numbers and quotes to describe our changes.
+We would like to thank the reviewers for their detailed and useful comments and questions.
+Comments from Reviewer 1 led to more explicit mentions of the involvement of the community when defining diversity and inclusion priorities (in Rule 1) and ways to ensure accountability during the organizing process (e.g.
+in Rule 2 with a DEI team and Rule 8 with publications like the diversity statement and the code of conduct) and after the conference (in Rule 10).
+Reviewer 2 brought to our attention that women had not been mentioned in Rule 2 and we corrected that. 
+Reviewer 3 raised several issues that led to changes in the text such as the reference to a DEI team, the mention of video abstracts as exclusionary, and the connection between the code of conduct and the diversity statement.
+We appreciate the interest in the process that we followed to put together this list of simple rules.
+The process is described in detail in the response to Reviewer 3 and summarized in the current version of the manuscript.
+We included the lists of rules gathered by the team as Supplementary Information.
+
+In the process of writing and revising the manuscript, we tried to minimize referring to the useR!
+experience.
+Though the conference was the main source of inspiration and lessons for this manuscript, we preferred to write the rules in a way that they could be applied to conferences in all regions and areas of knowledge, and only self-reference when necessary.
+
+To be consistent with our own rule of making room for the linguistic diversity of our community, and more broadly, the scientific community that would potentially read this manuscript, we elaborated versions of the abstract and list of rules in ten languages other than English.
+We now mention them at the end of the abstract in English.
+The translated abstracts are now included as Supplementary Information but we would like to make a special request for them to appear in the main page of the manuscript online.
+
+We sincerely hope that you will now find our manuscript suitable for publication. Please let us know if you require any additional information.
+
+All my best,
+on behalf of the authors,
+
+Roc\'io Joo
+
+\section*{Editorial Summary}
+
+Dear Dr. Joo,
+
+Thank you very much for submitting your manuscript `Ten simple rules to host an inclusive conference' for consideration at PLOS Computational Biology.
+As with all papers reviewed by the journal, your manuscript was reviewed by members of the editorial board and by several independent reviewers.
+The reviewers appreciated the attention to an important topic.
+Based on the reviews, we are likely to accept this manuscript for publication, providing that you modify the manuscript according to the review recommendations.
+
+\begin{Reply}
+    Thank you very much for these good news, and special gratitude to the members of the editorial board and the independent reviewers for the positive reviews.
+\end{Reply}
+
 When you are ready to resubmit, please upload the following:
@@
 \begin{Reply}
-   We are going to refer to line numbers in the following way: In the plos\_latex\_template.tex document, add a line number label where you want to edit like this: \\ linelabel\{rev1-com1-a\}
-   To refer to these lines in this document, add a reference here, like this: Lines   \ref{rev1-com1-a} to \ref{rev1-com1-b}
+    % RJ: This is actually rule 1.
+   The reviewer asked very important questions, for which there are no correct answers or formulas. 
+   The `correct' demographic will vary for every conference, field, and local and regional contexts. 
+   Our team agrees with implementing remediative measures but we believe that every meeting committee or organizing team will have to ask themselves these questions.
+   We now explicitly mention the need of this reflection and suggest to involve the community as well as using available information on diversity in the field (lines \ref{indicators-1} to \ref{indicators-2}):
 
-   %ast: We do not have a correct answer for this, unfortunately, and we believe there is a reflexion work that every team has to do in their context.
-We had extensive discussions about the indicators and our conclusion is that the "correct" demographic to monitor will vary for every conference, field, and local and regional context.
-   % Depending on the field, greater achievements can be made, and in some areas, this exercise will be more difficult and yet extremely important even if the apparent advancements are less evident. 
-   % Our team is inclined to propose what the reviewer calls "remediative measures", or affirnative actions, and be explicitly more than inclusive, but we understand that not every team can set such targets.
-And even with bold intentions, there is no guarantee that the final result is going to be as expected.
-   % A support from the community may be a great way to strenghen bold goals.
-However, in some communities the common sense shared by members may represent a wish to preserve the status quo, so not always a broad support from the community can be expected -but this may indicate also a need to broaden the definition of the community.
-   % We would like to encourage the readers of any field and regional context to make this exercise independently of the status of the field and the initial impression that these efforts won't be successful or sufficient.
-   % to edit: We have made this explicit in the text.
-We also included the fact that the community 1.
-should be defined and 2.
-can be consulted before the conference to make sure that this construction is collective and participative.
+   \textit{
+    Lack of diversity will vary between areas of knowledge and depend on the geographic context of the conference.
+    Assess how it translates to your particular field and scientific or professional community.
+    Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM) if they exist.
+    Most importantly, involve the community---that is, the subset of the general public that would ideally attend the conference---to help make this assessment and to set the vision and goals of diversity and inclusion for your conference.
+   }
 
-%Marce: I agree with reviewer 1 that this should be remediative, quoting: "where it specifically selects for communities historically ignored in the field beyond the regional population rate".
-To respond to the same reviewer, I think we can mention how  UseR presented the demographics of the attendees/participants, and how this was shared with the community as a way to be held accountable.
-%Marce: About rule 10: both reviewers 1 and 3 suggested to discuss further how to implement these rules. I think rule 10 is already discussing that in a clear manner.
 \end{Reply}
 
-(2) Adding on to the first point, how are the indicators measured?
-Should the conference organizers quantify the demographics of the attendees/participants?
-If so, should these details be published to the participants after the conference?
-How do organizers hold themselves accountable?
-And to whom are they held accountable to; the organizing committee themselves or the greater community?
+(2) Adding on to the first point, how are the indicators measured? Should the conference organizers quantify the demographics of the attendees/participants?
 
 \begin{Reply}
-   % Yani: I think we can mention and link to the info board and to the blog post post conference analyzing this numbers like the Latin American one and the accessibility one.
-The surveys are also another source of data for this accountability.
-I think this is related with the other review asking about governance.
-   %ast: The tenth rule is directly connected to the first, because these indicators can be measured and evaluated via surveys, direct impressions from the atendance, and exercises such as focus groups.
-Of all these options, surveys are the most frequently used but we wanted to highlight that there are other options.
-This link to rule 10 was not explicit in the text (edit!)
-   % About publishing the details after the conference, the privacy and anonymity of the participants should be respected, so some kind of binning/grouping of the data and care with categories that have a low number of respondents should be taken.
-%make sure this in the manuscript!
-we had this in previous versions.
-   % Accountability of the team should be posible with or without public indicators.
-Open channels for communication and open fields and the possibility to make anonymous comments should be provided.
-Especially if there is a claim for diversity and inclusion, the general community should be able to give feedback about the success of these efforts.
-   %to edit: cite the tenth rule in the first, telling there will be accountability. but also say that the indicators should not become a goal themselves.
+The evaluation of goals and outcomes is part of Rule 10, and was addressed in the manuscript, now lines \ref{info} to \ref{info2}:
+\textit{
+(...) It is likely that you will be able to evaluate the outcomes associated to these goals directly from the information that you have already collected from your organizing team, speakers, and participants (e.g.
+city/state/country of origin, or preferred language).
+You may have goals concerning how welcomed and included participants felt during the conference, and wonder if your practices translated into real inclusion. 
+Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
+}
+We also added in Rule 1 an explicit link to Rule 10 in line \ref{rule10}, to let the reader know that the evaluation results will be used as part of the conference wrap up and continuity.
 
 \end{Reply}
-(3) In Rule 4, it states "Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;" Does this pertain to invited speakers only, or speakers selected from submitted abstracts?
+
+If so, should these details be published to the participants after the conference?
+
+\begin{Reply}
+Yes, the final assessment of the success of the inclusion actions should be shared with everyone that attended and helped create these goals, as well as with people who would use the data in future conferences.
+We added more details about this in Rule 10, lines \ref{data-share} to \ref{data-share2}:
+\textit{The feedback received from the participants and your whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendees, and future organizers.
+Make sure that the privacy and anonymity of the participants is respected; ask for consent and consider binning/grouping the data for categories that have a low number of respondents.}
+\end{Reply}
+
+How do organizers hold themselves accountable? And to whom are they held accountable to; the organizing committee themselves or the greater community?
+
+\begin{Reply}
+Conferences are organized for the community, thus the organizers should hold themselves accountable to the whole community and not just the organizing committee themselves.
+As we mentioned in the manuscript, this should be done along the way (line \ref{rule10}), during the organization process, by being transparent and communicating inclusion efforts and limitations (lines \ref{accountability-1} to \ref{accountability-2}):
+\textit{Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
+This is not only to communicate your values and practices, but as a way to hold yourself accountable for them.
+Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. 
+For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering help to navigate it, or if captions will be available for some talks but not all.
+Provide a point of contact to help clarify any questions regarding accessibility, financial support, and others.}
+
+As stated in Rule 10 (lines \ref{account-3} to \ref{data-share2}), there should be a final assessment of the achievement of the diversity and inclusion goals at the end of the conference.
+Many tools can be used for that (e.g.
+demographic information, surveys, focus groups, information from informal conversations).
+Publishing these final assessments is an important way to hold the organizers accountable both to the greater community and the committee itself.
+\end{Reply}
+
+(3) In Rule 4, it states `Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;' Does this pertain to invited speakers only, or speakers selected from submitted abstracts?
 If this pertains to scored abstracts, how does one ensure that this is done in a transparent and reproducible way?
@@
 \begin{Reply}
-   %ast: Although curation is easier and definitely encouraged for invited speakers, we would like to reinforce that even transparent scored processes can be influenced by implicit bias, once the reviewers start evaluating there might be "shifting criteria" and other considerations entering mid-process.
-   %We have added some extra explanation about this curation, that include at least: 
-   %1.
-determining clear criteria before seeing the list of submissions, 2.
-examining the final list of selected work for imbalances in gender, geographical representation and other visible aspects of diversity.
-3.
-check if close competitors (similar quality work) were excluded for some apparent bias.
-   % This process is hard because there are many aspects of diversity that will be unreachable for reviewers and are prone to perceived cues rather than real information (so not every dicersity dimension can be considered here).
-The invitation is to examine similar quality work and try to understand why some work was included rather than other and to consider whether bias could have influenced the outcome.
-   %criteria should be clear before the abstract evaluation. Some adaptation of this can be made: https://honors.agu.org/files/2014/11/AWARDS-SUGGESTED-BEST-PRACTICES-MAY2011.pdf
+This specific section refers to the final selection of submitted proposals. Counteracting bias in invited roles was mentioned earlier in the text in line \ref{invited}.
+We agree with the reviewer that `curation' was a misleading term.
+We rephrased it to \textit{
+Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
+Examine the final list of selected abstracts to verify if there is lack of diversity.
+If proposals with similar quality were still judged differently because of apparent bias, consider giving preference to the work from people in minoritized groups.
+This manual examination at a final stage is unlikely to be reproducible.
+Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} (lines \ref{criteria-1}-\ref{criteria-2}).
+
 \end{Reply}
 
-(4) In Rule 5, it states "The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants." How do the tablets help in-person attendees interact with remote participants?
+(4) In Rule 5, it states `The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.' How do the tablets help in-person attendees interact with remote participants?
 Especially participants that may be in a different time-zone.
@@
 \begin{Reply}
-   %ast: Although one can assume that an ever growing proportion of people have their own handheld device to interact online with the remote participants, we cannot assume that everyone does, or that the models are compatible with the options of the conference.
-Some devices (and support for those and the participants') should be provided by the organization.
-With everyone having a device, the idea is to unify the networking activities and Q & A sessions to be online, this more integrated with the remote component.
+Although one can assume that an ever growing proportion of people have their own handheld device to interact virtually with the remote participants, we cannot assume that everyone does, or that the models are compatible with the software platforms used in the conference.
+If the conference organizers provides such devices, Q\&A sessions and networking activities can be online and allow for more integration between in-person and online participants. 
 \end{Reply}
 
-(5) In Rule 7 it states " Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language." It is not trivial to submit an abstract in two languages, and if there is no chance that the non-English abstract won't be reviewed I feel like this would be wasting their time.
+(5) In Rule 7 it states `Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.' It is not trivial to submit an abstract in two languages, and if there is no chance that the non-English abstract won't be reviewed I feel like this would be wasting their time.
 Is it possible to denote the accepted languages ahead of time?
@@
 \begin{Reply}
-%ast: "and" should be "or".
-Yes, definitely, rather than two abstracts, one in English and the other in a second language, the best way would be to be able to have abstracts in languages different than English without the need for a duplication.
-That depends on the conference capabilities but we would recommend both submission and presentation be open to other languages.
-The idea for a previous request/screening about the languages that will be needed is good.
-Another option is to have a language editing help for English, via the communities or via the organizing team, that helps ensure that the submitted abstracts are conveying the meaning the authors are trying to give.
-This could include also, abstracts written in English by students and early career researchers.
+    Thank you for this suggestion. 
+    The organizing team should decide whether to ask for two languages or just one depending on their capabilities. 
+    It is true that it is not trivial to submit abstracts in two languages, but for many people, writing in English may be cumbersome and, if given the choice, they would opt for also adding a version in their native language if that increases the chances to communicate their ideas accurately.
+    The text now states: 
+    \textit{Allow for abstract submission in both English and the language the person feels more comfortable with, and whenever possible, assign a reviewer who is fluent in that language. 
+    Alternatively, the organizing team could define a set of languages that are accepted for abstract submission, and allow applicants to choose from one of those languages. 
+    The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency} (lines \ref{abstractlanguage1} to \ref{abstractlanguage2}).
+
+    We also added a mention to accepting multilingual submissions when counteracting self-selection during the call for abstracts in Rule 4, linking to rule 7 (line \ref{multilingual-submission}).
 \end{Reply}
@@
 \begin{Reply}
-%ast: We thank the reviewer's feedback.
-We have added some pre-conference participation of the community, also inspired by examples we got from colleagues at pycon Germany.
-The accountability is very hard to define because no formal structure requires that in the present, it has to be the team's will to be held accountable to their participant community.
-Clarity in the communication strategy and open spaces for feedback are key here.
+    Thank you for the thorough and positive review.
+We hope to have addressed all of your concerns and particularly accountability.
+We agree with involving the community in every part of the process, in an active and participative manner, using different channels of open communication and integrating members of the community in the organizing team.
 \end{Reply}
 
-\textbf{Reviewer \#2:}
+\section*{Reviewer \#2:}
 L78-L80 As a female scientist I would ask myself "where are we" (because we still have parts of the World where female scientist are rare on leading positions), but I consider science apolitical in every segment, and I certainly suggest that it concentrate here on the scientific youth, the segment of research in industry, etc., and not on the above.
@@
 \begin{Reply}
-
+    Thanks for reading and reviewing. The list given was meant to be an example of extant thematic groups and communities, aiming to go beyond gender representation.
+    We agree about adding women as an example of such groups in line \ref{wis}:
+    `Open the schedule for events organized by thematic groups and communities (e.g.
+task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.'
 \end{Reply}
 
-\textbf{Reviewer \#3:} 
+\section*{Reviewer \#3:} 
 This is a timely, well-written commentary on organising an inclusive conference and should act as a reference compendium for researchers, practitioners as well as conference organisers.
@@
 \begin{Reply}
-%ast: my first hunch in general: the metrics part is one of the hardest part to implement and reviewer1 talked about it, but it is also one of the "goals" that stick with some readers and we were trying to avoid putting them at the center of decision-making because they can become a end in itself and not a way to understand if your efforts were successful
-%The tenth rule should have accounted for this but we may need to be more specific in the messages of not letting yourself be prey of the indicators. 
-%Team accountability and governance were part of rule 10 so maybe we could strengthen this.
+    A Diversity, Equity and Inclusion (DEI) committee would be the best suited to make sure that the inclusion vision and goals are achieved. 
+    Establishing a DEI team is now suggested in Rule 3 in the manuscript (lines \ref{subteam-1} to \ref{subteam-2}). Thank you for that. 
+    \textit{ 
+    Having a dedicated DEI team may be a good way to give the attention and energy the topic requires.
+    The leading team should provide support to the DEI initiatives and advocate for inclusion across all conference areas.}
+    On the other hand, the existence of a DEI team is neither absolutely necessary nor sufficient, and we believe the most important thing is the commitment of the 
+    whole organizing team towards inclusion, and the willingness to hold themselves accountable, with concrete actions mentioned in Rule 8 (during the conference, lines \ref{accountability-1} to \ref{accountability-2})  and in Rule 10 (after the conference, lines \ref{account-3} to \ref{data-share2}) .
+
 \end{Reply}
@@
 \begin{Reply}
-   %Yani: I think this is not on the scope of the paper, but we can give an explanation of the process, which I think was fantastic and inclusive and have a diverse background on the construction of the rules.
-   %ast: i agree, this may break the format for a ten simple rules piece so maybe a very short mention in the introduction or as SupMat?
+    The rules were first drafted by the first author, based on the collective experience of organizing useR!
+2021 and discussions with other organizers; then the leading group and the members of the diversity, accessibility, and inclusion team were invited to write their own ten simple rules—the rules proposed at that stage are now in the Supplementary Information.
+    We reached a final list by consensus, not popularity.
+All the other members of the organizing team and the people who acted as helpers in the conference were invited to give their feedback to the resulting list, in order to check for omissions and errors and suggest editions to the text.
+    We also sent the manuscript to friendly reviewers who have the experience of attending or organizing conferences. 
+    These rules, or preliminary versions of them, were shared with over 100 people, though not all of them provided feedback.
+    The rules in this manuscript are a result of this process and further discussions with and contributions from all the coauthors, fueled by personal experiences and literature review.
+    A brief description of this process is now stated in lines \ref{methods-a}-\ref{methods-b} of the current version of the manuscript.
+    We kept the methodology as brief as possible to keep the focus on the rules and not our team or conference.
+
+    The coauthors are from a diversity of backgrounds, career stages, and even continents, and have experienced inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege, as stated in lines \ref{background-a}-\ref{background-b}.
+    The manuscript is the result our work together, as well as the awareness and experience of organizing a conference that aimed to be a kind, inclusive, and accessible experience for as many people as possible.
 \end{Reply}
@@
 
+\begin{Reply}
+   In a global scale, there is an imbalance associated to what could be considered a `north-south' divide, though the definition of north and south in this sense may not be clear, and we agree that being from a lower income country hinders participation in conferences.
+For that reason, we offer an example of registration rates based on country of origin among other aspects (\ref{example-countries}).
+   On the other hand, while the manuscript is heavily based on our experience organizing a global conference—which is why there are some global examples—, the intention is to make the rules general and flexible enough so that they can be useful for local or regional conferences around the world.
+   For that reason, we explicitly mention in Rule 1 that \textit{Lack of diversity will vary between areas of knowledge and depend on the geographic context of the conference.} (line \ref{indicators-1}).
+
+\end{Reply}
+
 4.
@@
 \begin{Reply}
-
-%ast: All specific examples of offenses should be included in the code of conduct construction process, and that's why we didn't make a list here, but maybe we can make explicit mentions of what type of offenses can/have to be in the code of conduct.
-%suggestion for edits: add this, talk a little bit more about the content of the CoC text.
-
+    Racist acts should be unacceptable behavior in any conference, but it is not our intention to prepare a comprehensive list of unacceptable behaviors in Rule 2.
+    All specific examples of offenses should be included in the code of conduct of each conference, and listing them should the task of the organizers, as well as establishing mechanisms for reporting and addressing those reports.
+    We also made explicit the link between the diversity statement and the code of conduct, to highlight that all dimensions of diversity should be respected (lines \ref{unacceptable-1} to \ref{unacceptable-2}):
+    \textit{The CoC is a living document meant to keep the community safe and should state clearly the behaviors that are deemed unacceptable by the community, the consequences for engaging in such behaviors, and the way to report violations.
+Bear in mind that every dimension of diversity can be a target of unacceptable behavior.
+The CoC should help honor the vision expressed in the diversity statement.}
 \end{Reply}
 
-5. In Rule 4, via a vis bias, bias against people with disability and ethnic minorities should also be highlighted.
+5. In Rule 4, vis a vis bias, bias against people with disability and ethnic minorities should also be highlighted.
 
 \begin{Reply}
-%ast: We have tried to make a distinction between visible and invisible aspects of human diversity, and we have to respect the right of every individual to occupy the space and assert/disclaim who they are and the right to remain silent about some aspects of their lives.
-Rule 2 is about being proactive regarding all dimensions of diversity, and Rule 4 focuses more on the visible aspects of diversity, especifically those that can be grasped via written communication.
-% We mention this because counteracting bias actions during the selection of speakers and submitted work, we--unfortunately--only have *perceived* cues of these dimensions.
-We can cite Laura's piece on video abstracts being exclusionary and be more explicit about the fact that perceived cues and implicit and explicit racism go hand in hand.
+    In the selection of invited speakers, biases could be related to multiple dimensions of diversity identified in the assessment mentioned in Rule 1. 
+    All of these should be equally addressed.
+
+    In addition, special attention should be given to unconscious bias in the revision of submitted work, which is usually related to aspects of human diversity that can be inferred via written communication.
+    We had mentioned [perceived] ethnicity as one of these aspects  (line \ref{bias}, \textit{`their perceived origin, ethnicity, gender, or native language'}).
+
+    Conscious and unconscious bias can also operate when the submitted work is not written, but rather in video form.
+We added a short mention and citation about how video abstracts can be exclusionary (Spitschan et al 2021) by adding barriers to people with visible disabilities, diverse body types, non-native language speakers, or less resources to create videos (lines \ref{video-1} to \ref{video-2}).
+    We thank the reviewer for this comment that led us to the current addition.
 \end{Reply}
@@
 \begin{Reply}
-   %Yani: I think all keynote deserve the _same_ communication, if we give prominence to minoritized group could be perceived as tokenism.
-So I would take the suggestion on NOT give more prominence to one group of keynote and make sure that all of them are promote in the same spaces and in the same way.
-We can mention here that communities partners can help with the highlight of the minoritazed groups (tricky).
-   %ast: I agree with Yani, and we had it partially covered in the rule 7, "Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same relevance as the rest of the program." but we can add something here.
+    We added the following text to the manuscript (lines \ref{promotion-1} to \ref{promotion-2}): 
+    \textit{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility than all other sessions.}
 \end{Reply}
@@
 7.
-In Conclusion, suggest also adding that issues of diversity and inclusion have come to fore even more during COVID-19 as we witness exacerbation of inequities disadvantaging those from vulnerable backgrounds (see https://pubmed.ncbi.nlm.nih.gov/33343410/
-PMID: 33343410)
+In Conclusion, suggest also adding that issues of diversity and inclusion have come to fore even more during COVID-19 as we witness exacerbation of inequities disadvantaging those from vulnerable backgrounds (see https://pubmed.ncbi.nlm.nih.gov/33343410/PMID: 33343410)
+
+\begin{Reply}
+% RJ: I'm leaning towards not adding this now.
+    We appreciate the suggestion and agree that the COVID-19 pandemic contributed to the exacerbation of inequity. 
+    However, we would like the manuscript to help inspire and guide the organization of more inclusive conferences regardless of the state of the pandemic in the different places in the world, and even when the pandemic subsides.
+For that reason, we tried to mention the pandemic very few times and only when deemed necessary: to give context to the conference that we organized, and to talk about virtual conferences which became the prevailing format during the last two years.
+\end{Reply}
+
+\begin{Reply}
+    We thank the reviewer for their comments and careful revision.
+\end{Reply}
 

### 2022-04-14 plos_latex_template.tex
Finished response to reviewers. This version was sent to coauthors :tada:

@@
 We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
+\hl{* Versions of this abstract in Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai are available in Supplementary Information.}
 
@@
 However, opportunities for participating in conferences are not equally available for all. 
-Many academic and tech conferences have been spaces that reproduce systemic inequalities, by failing to overcome the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income backgrounds, prestigious institutions, native English speakers with no disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
+Many academic and tech conferences have been spaces that reproduce systemic inequalities, by failing to overcome the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income backgrounds, prestigious institutions, \linelabel{power} native English speakers with no disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
 \linelabel{background-a}Like many others, the authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
@@
 Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
-\linelabel{methods-a} \hl{The rules were first drafted by the first author; then the leading group and the members of the diversity, accessibility, and inclusion team were invited to write their own ten simple rules---everyone's rules are in XXX.
-The rules in this manuscript are a result of all of these rules and further discussions with and contributions from all the coauthors, fueled by personal experiences and literature review.} \linelabel{methods-b}
-They are organized in three sections: three foundational rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
-The \textbf{foundational rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
+\linelabel{methods-a} \hl{The rules were first drafted by the leading group and  members of the diversity, accessibility, and inclusion team of useR!
+2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
+A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
+The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
+The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
 \textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
@@
 The \textbf{design rules} focus on weaving inclusion into the conference design process.
-In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
+In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
 \textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
 \textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
-In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
+In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
 \textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
@@
 Finally, 
-\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion, and passing the torch to future organizers.
+\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and passing the torch to future organizers.
 
@@
 Recognizing this systematic exclusion is paramount to inclusion.
-Assess how lack of diversity translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
-%depends on what information is available when defining inclusiveness goals.
-When DEI studies in the particular field of the conference are available, these should be used.
-Otherwise, global metrics in STEM, social sciences, and humanities should be used.
-Other indicators, like gender balance, can be set without the need for concrete studies in the area.
-\linelabel{indicators-1} \hl{The result of this assessment will vary between areas of knowledge and depending on the geographic context of the conference. 
-Involving the community---that is, the subset of the general public that would ideally attend the conference---%and for whom you are organizing the conference
-early in this stage can help you set these goals in a collective and participative way.%be very revealing of their perception about the people who usually get to attend the conferences in the area.
-not this.
-judgemental
-} \linelabel{indicators-2}
-
-Define measurable indicators for these goals (e.g.
-a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions)
-that will help you be accountable along the way
-\linelabel{rule10} \hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
+
+\linelabel{indicators-1}\hl{
+Lack of diversity will vary between areas of knowledge and depend on the geographic context of the conference.
+Assess how it translates to your particular field and scientific or professional community.
+Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM) if they exist. 
+Most importantly, involve the community---that is, the subset of the general public that would ideally attend the conference---to help make this assessment and to set the vision and goals of diversity and inclusion for your conference.}
+\linelabel{indicators-2}
+Define measurable outcomes for these goals (e.g.
+a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions).
+They will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
 The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.
@@
 Include menu options that account for diverse dietary and religious requirements.
-Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it's good to have them in mind from the beginning.
+Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it is good to have them in mind from the beginning.
 Open the schedule for events organized by thematic groups and communities (e.g.
@@
 adopting a code of conduct (CoC) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
-The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
+\linelabel{unacceptable-1}The CoC is a living document meant to keep the community safe and should state clearly the behaviors that are deemed unacceptable by the community, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
+\hl{Bear in mind that every dimension of diversity can be a target of unacceptable behavior. The CoC should help honor the vision expressed in the diversity statement.}\linelabel{unacceptable-2}
+
 The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community.
@@
 In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
+\hl{If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
+The regional and local communities, groups, or associations in your field are good sources to tap into.}
+
+\linelabel{subteam-1} \hl{Create subteams to distribute fairly the large amount of work that organizing a conference represents. 
+Having a dedicated DEI team may be a good way to give the attention and energy the topic requires.
+The leading team should provide support to the DEI initiatives and advocate for inclusion across all conference areas.} \linelabel{subteam-2} 
 It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
 every member should have the freedom to choose which areas of the conference they want to contribute to.
-
-% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
-If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
-The regional and local communities, groups, or associations in your field are good sources to tap into. 
-When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. \linelabel{team}
+When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
 Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
@@
 
-The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered.
+The conference program may lack diversity in the invited roles––-such as keynote speakers, program/scientific committee, session chairs, panelists––-, the accepted abstracts, and even the scope of the topics covered.
 It is important to counteract biases in all of these aspects.
@@
 If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
-
-The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.
+\linelabel{multilingual-submission}\hl{To mitigate self-selection from non-native speakers, consider accepting submissions in several languages} (see \textbf{Rule 7}).
+
+The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias}
 Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
@@
 define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
-\linelabel{criteria-1}Define clear evaluation criteria and share them with the reviewers and authors \hl{preferably during the call for abstracts} \linelabel{criteria}.
-\hl{Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias} \linelabel{not-curation}; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
-
-In many academic or technological fields, typically competitive activities or areas—such as scientific publications or software development—are more valued than those focused on sharing and cooperation, like community building, teaching, or mentoring.
+\linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
+Examine the final list of selected abstracts to verify if there is lack of diversity.
+If proposals with similar quality were still judged differently because of apparent bias, consider giving preference to the work from people in minoritized groups.
+This manual examination at a final stage is unlikely to be reproducible.
+Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
+\linelabel{video-1}\hl{It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.} \cite{spitschanVideoGrantProposals2021}\linelabel{video-2}
+
+Bias can also be expressed in the level of visibility given to the diverse topics in the conference program.
+In many academic or technological fields, typically competitive activities or areas---such as scientific publications or software development---are more valued than those focused on sharing and cooperation, like community building, teaching, or mentoring.
 The latter activities are equally, if not more, relevant and challenging, and are usually performed by women, racialized people, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020, light_gender_2022}.
@@
 Make sure that you are not creating additional barriers when choosing conference platforms, schedule, and activities. 
-Check every platform's screen reader accessibility (see \textbf{Rule 6}), technological requirements (a high-speed internet or a latest generation computer prerequisite could exclude attendees from many regions), and geopolitical restrictions (some services are not available in all countries).
+Check every platform's screen reader accessibility (see \textbf{Rule 6}), technological requirements (a high-speed internet or a latest generation computer prerequisite could exclude attendees), and geopolitical restrictions depending on the scale of the conference (some services are not available in all countries).
 Be mindful of the different time zones and time constraints of the potential attendees and create flexible schedules.
@@
 To best integrate in-person and remote attendees, all questions and comments should pass through a unique online system. 
-\hl{The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.}
+The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.
 
@@
 % Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
-In international meetings, the linguistic diversity of the participants is often overlooked. 
+In academic and technical events, the linguistic diversity of the participants is often overlooked. 
 English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, websites, and communications. 
@@
 Create an inclusive environment by encouraging the full participation of non-native English speakers. This may be done in multiple ways.
-\hl{Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.} That way, the abstract will be judged primarily by the quality or relevance of the work instead of English proficiency.
+\linelabel{abstractlanguage1}\hl{Allow for abstract submission in both English and the language the person feels more comfortable with, and whenever possible, assign a reviewer who is fluent in that language.
+Alternatively, the organizing team could define a set of languages that are accepted for abstract submission, and allow applicants to choose from one of those languages. 
+The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency}.\linelabel{abstractlanguage2}
 Whenever possible, identify other prominent languages for the conference and provide translated captions or live language interpretation into these key languages, including multilingual Q\&A sessions.
@@
 Twitter, Facebook, LinkedIn, conference website, mailing lists).
-
-Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}) to communicate your values and practices, and to hold yourself accountable for them.
-\linelabel{accountability-1}Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues.
+\linelabel{promotion-1}\hl{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility than all other sessions.}\linelabel{promotion-2}
+
+\linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
+\hl{This is not only to communicate your values and practices, but as a way to hold yourself accountable for them.}
+Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. 
 For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering help to navigate it, or if captions will be available for some talks but not all.
 Provide a point of contact to help clarify any questions regarding accessibility, financial support, and others.
-
-Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups—in all communications \cite{hallDesigningDiversityInclusion2019}.
+\linelabel{accountability-2}
+
+Use inclusive language---language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups---in all communications \cite{hallDesigningDiversityInclusion2019}.
 Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group, and the terms that should be avoided.
@@
 % Rocío: main idea: help take the burden out of participants for more inclusive participation
-When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
-(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
+When determining the registration rates, the socioeconomic context of participants, their career status, and their country of origin—if applicable—should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
+(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status).\linelabel{example-countries}
 Resources permitting, enable a `pay what you can' option.
@@
 % Rocío: main idea: assess how inclusive was the conference
-As conference organizers, one of the key post-conference tasks is to make a final assessment of the achievement of the diversity and inclusion goals. 
-You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
-It is likely that you will be able to calculate some of your indicators directly from \linelabel{info}the information that you have already collected from your organizing team, speakers, and participants (e.g.
-countries of origin, or preferred language).
+\linelabel{account-3}As conference organizers, one of the key post-conference tasks is to make a final assessment of the achievement of the diversity and inclusion goals. 
+You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
+\linelabel{info}It is likely that you will be able to evaluate the outcomes associated to these goals directly from the information that you have already collected from your organizing team, speakers, and participants (e.g.
+\hl{city/state/country of origin}, or preferred language).
 You may have goals concerning how welcomed and included participants felt during the conference, and wonder if your practices translated into real inclusion. 
-\linelabel{info2}Surveys, focal groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
+Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
+\linelabel{info2}
 
 % Pass the information
-The feedback received from the participants and your whole assessment of the conference should be documented and shared with \linelabel{data-share}\hl{those who helped define the inclusion goals, the attendants, and} future organizers.
-%ast: Rocío, we had in previous versions a part talking about data privacy and anonymization, maybe we could add a quick mention to this here, share with whom (and add the community) with some precaution.
-Gather and do your best to organize all the documentation produced during the organization process—or even after.
+The feedback received from the participants and your whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers.
+\hl{Make sure that the privacy and anonymity of the participants is respected; ask for consent and consider binning/grouping the data for categories that have a low number of respondents}.\linelabel{data-share2}
+Gather and do your best to organize all the documentation produced during the organization process---or even after.
 Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors.
-Create a list with contact information of the people of the organizing team, as well as previous and potential speakers—if they agree to share their contact information—, and pass it along.
+Create a list with contact information of the people of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
 For conference series, it could be a good idea to create an official document (e.g.
@@
 We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript.
-In addition to some of the authors, we would like to acknowledge the following contributors for translating our extended abstract to several languages: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], and Kewalin Samart [Thai].
+In addition to some of the authors, we would like to acknowledge the following contributors for translating our extended abstract to several languages: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae [Thai].
+The comments of the anonymous reviewers contributed to the improvement of the manuscript and we thank them for that.
 
@@
 
+\bibitem{spitschanVideoGrantProposals2021}
+Spitschan M, Yehudi Y, Acion L.
+\newblock Video Grant Proposals Could Be Exclusionary.
+\newblock Nature. 2021;592(7852):26--26.
+\newblock doi:{10.1038/d41586-021-00862-x}.
+
 \bibitem{ross_everyday_2020}

### 2022-04-20 response_reviewers.tex
RJ: accepted changes to responses to the reviewers

@@
 %\usepackage{hyperref} 
-\newenvironment{Reply}{\noindent\color{BlueViolet}\textbf{Reply.}}{\vspace{1em}}
+\newenvironment{Reply}{\noindent\color{BlueViolet}\textbf{Reponse.}}{\vspace{1em}}
 
@@
 
-Please find the responses to all the reviewers comments and suggestions below in this document.
-We also highlighted our changes in the manuscript and added line numbers and quotes to describe our changes.
-We would like to thank the reviewers for their detailed and useful comments and questions.
-Comments from Reviewer 1 led to more explicit mentions of the involvement of the community when defining diversity and inclusion priorities (in Rule 1) and ways to ensure accountability during the organizing process (e.g.
-in Rule 2 with a DEI team and Rule 8 with publications like the diversity statement and the code of conduct) and after the conference (in Rule 10).
-Reviewer 2 brought to our attention that women had not been mentioned in Rule 2 and we corrected that. 
-Reviewer 3 raised several issues that led to changes in the text such as the reference to a DEI team, the mention of video abstracts as exclusionary, and the connection between the code of conduct and the diversity statement.
-We appreciate the interest in the process that we followed to put together this list of simple rules.
-The process is described in detail in the response to Reviewer 3 and summarized in the current version of the manuscript.
-We included the lists of rules gathered by the team as Supplementary Information.
+Thank you for considering our article for publication. We would like to thank the reviewers for their detailed feedback and thoughtful comments and suggestions.
 
-In the process of writing and revising the manuscript, we tried to minimize referring to the useR!
+Please find the detailed responses to all the reviewers' comments below. We added line numbers and quotes and highlighted our changes in the manuscript. 
+Comments from the Reviewers led to editing and paraphrasing of several parts of the manuscript, including
+i) direct mentions of the involvement of the community when defining diversity and inclusion priorities (in Rule 1)
+ii) ways to ensure accountability during the organizing process and after the conference (e.g.\ in Rule 2 with a Diversity, Equity, and Inclusion (DEI) team, Rule 8 with guidelines like the diversity statement and the code of conduct, and Rule 10 with post-conference continuity)
+iii) mention of women in science in Rule 2
+iv) a reference to the DEI team, biases due to video abstracts, and the connection between the code of conduct and the diversity statement
+v) the process we followed to put together this list of simple rules; now described in detail below and summarized in the current manuscript version.
+We also included the sets of rules gathered by the team as Supplementary Information.
+
+While writing and revising the manuscript, we tried to minimize references to the useR!
 experience.
-Though the conference was the main source of inspiration and lessons for this manuscript, we preferred to write the rules in a way that they could be applied to conferences in all regions and areas of knowledge, and only self-reference when necessary.
+Though the conference was the primary source of inspiration and lessons learned and implemented in this manuscript, we wrote the rules in a way that could be applied to conferences in all regions and areas of knowledge.
 
-To be consistent with our own rule of making room for the linguistic diversity of our community, and more broadly, the scientific community that would potentially read this manuscript, we elaborated versions of the abstract and list of rules in ten languages other than English.
-We now mention them at the end of the abstract in English.
-The translated abstracts are now included as Supplementary Information but we would like to make a special request for them to appear in the main page of the manuscript online.
+To be consistent with our own rule of making room for linguistic diversity and to ensure accessibility across the scientific community, we translated the extended version of the abstract and list of rules into ten other languages.
+We now mention them at the end of the English abstract.
+The translated abstracts are also included as Supplementary Information.
+To ensure broader reach, we would like to make a special request for them to appear on the main page of the manuscript online.
 
-We sincerely hope that you will now find our manuscript suitable for publication. Please let us know if you require any additional information.
+We sincerely hope that you will now find our revised manuscript suitable for publication. Please let us know if you require any additional information.
 
-All my best,
+Sincerely,
 on behalf of the authors,
 
-Roc\'io Joo
+Roc\'io Joo and Janani
 
@@
 
-Thank you very much for submitting your manuscript `Ten simple rules to host an inclusive conference' for consideration at PLOS Computational Biology.
+Thank you very much for submitting your manuscript `Ten simple rules to host an inclusive conference' for consideration at PLoS Computational Biology.
 As with all papers reviewed by the journal, your manuscript was reviewed by members of the editorial board and by several independent reviewers.
@@
 \begin{Reply}
-    Thank you very much for these good news, and special gratitude to the members of the editorial board and the independent reviewers for the positive reviews.
+
+    Thank you very much for this good news, and special gratitude to the members of the editorial board and the independent reviewers for the positive feedback.
 \end{Reply}
 
-When you are ready to resubmit, please upload the following:
+% When you are ready to resubmit, please upload the following:
 
-[1] A letter containing a detailed list of your responses to all review comments, and a description of the changes you have made in the manuscript.
+% [1] A letter containing a detailed list of your responses to all review comments, and a description of the changes you have made in the manuscript.
 Please note while forming your response, if your article is accepted, you may have the opportunity to make the peer review history publicly available.
@@
 
-[2] Two versions of the revised manuscript: one with either highlights or tracked changes denoting where the text has been changed; the other a clean version (uploaded as the manuscript file).
+% [2] Two versions of the revised manuscript: one with either highlights or tracked changes denoting where the text has been changed; the other a clean version (uploaded as the manuscript file).
 
-Important additional instructions are given below your reviewer comments.
+% Important additional instructions are given below your reviewer comments.
 
- Thank you again for your submission to our journal.
+%  Thank you again for your submission to our journal.
 We hope that our editorial process has been constructive so far, and we welcome your feedback at any time.
@@
 \section*{Reviewer \#1:}
-I greatly enjoyed reading the paper "Ten simple rules to host an inclusive conference", and think it provides an easy to follow set of rules that organizers can use to ensure an inclusive conference.
+I greatly enjoyed reading the paper `Ten simple rules to host an inclusive conference', and think it provides an easy to follow set of rules that organizers can use to ensure an inclusive conference.
 I found many tips for inclusiveness in the article to be especially helpful, such as multilingual accommodations and ensuring screen-reader accessibility.
@@
 
-(1) In Rule 4, the manuscript states `Define measurable indicators for these goals e.g.
-a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions', but how do organizers choose the correct demographics to monitor and the expected background rates?
+(1) In Rule 4, the manuscript states `Define measurable indicators for these goals e.g.\ a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions', but how do organizers choose the correct demographics to monitor and the expected background rates?
 Should it match global, the local, or field-specific rates?
@@
 \begin{Reply}
+
     % RJ: This is actually rule 1.
-   The reviewer asked very important questions, for which there are no correct answers or formulas. 
-   The `correct' demographic will vary for every conference, field, and local and regional contexts. 
-   Our team agrees with implementing remediative measures but we believe that every meeting committee or organizing team will have to ask themselves these questions.
-   We now explicitly mention the need of this reflection and suggest to involve the community as well as using available information on diversity in the field (lines \ref{indicators-1} to \ref{indicators-2}):
+   The reviewer asked very important questions, for which there are no straightforward answers or formulae. 
+   The `correct' demographic will vary for every conference, field, and local/regional contexts. 
+   We believe that while remediation measures should be implemented in all conferences, every meeting committee or organizing team will have to ask themselves these questions and formulate their own inclusive plan.
+   We now explicitly mention the need for this reflection and suggest involving the community as well as using available information on diversity in the field (lines \ref{indicators-1} to \ref{indicators-2}):
 
    \textit{
-    Lack of diversity will vary between areas of knowledge and depend on the geographic context of the conference.
-    Assess how it translates to your particular field and scientific or professional community.
-    Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM) if they exist.
-    Most importantly, involve the community---that is, the subset of the general public that would ideally attend the conference---to help make this assessment and to set the vision and goals of diversity and inclusion for your conference.
+    Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
+    Assess how it translates to the conference's particular field and scientific or professional community.
+    Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if it exists. 
+    Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.
     }
@@
 \begin{Reply}
-The evaluation of goals and outcomes is part of Rule 10, and was addressed in the manuscript, now lines \ref{info} to \ref{info2}:
+
+The evaluation of goals and outcomes is part of Rule 10, and has now been addressed in the manuscript in lines \ref{info} to \ref{info2}:
 \textit{
-(...) It is likely that you will be able to evaluate the outcomes associated to these goals directly from the information that you have already collected from your organizing team, speakers, and participants (e.g.
-city/state/country of origin, or preferred language).
-You may have goals concerning how welcomed and included participants felt during the conference, and wonder if your practices translated into real inclusion. 
-Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
+(...) The organizers will likely be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ city/state/country of origin or preferred language).
+The planning committee may include goals concerning how welcome and included participants felt during the conference and consider if the implemented practices translated into true inclusion. 
+Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
 }
-We also added in Rule 1 an explicit link to Rule 10 in line \ref{rule10}, to let the reader know that the evaluation results will be used as part of the conference wrap up and continuity.
+We also added in Rule 1 an explicit link to Rule 10 in line \ref{rule10} to let the reader know that the evaluation results will be used as part of the conference wrap-up and continuity.
 
@@
 \begin{Reply}
-Yes, the final assessment of the success of the inclusion actions should be shared with everyone that attended and helped create these goals, as well as with people who would use the data in future conferences.
+
+Yes, the organizing team should share the final assessment of the success of their inclusive actions with everyone who attended, helped create these goals, or could be part of future conference committees.
 We added more details about this in Rule 10, lines \ref{data-share} to \ref{data-share2}:
-\textit{The feedback received from the participants and your whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendees, and future organizers.
-Make sure that the privacy and anonymity of the participants is respected; ask for consent and consider binning/grouping the data for categories that have a low number of respondents.}
+\textit{The feedback received from the participants, and the whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendees, and future organizers.
+Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents.}
 \end{Reply}
@@
 \begin{Reply}
-Conferences are organized for the community, thus the organizers should hold themselves accountable to the whole community and not just the organizing committee themselves.
+
+Conferences are organized for the community; thus, the organizers should hold themselves accountable to the whole community and not just the organizing committee themselves.
 As we mentioned in the manuscript, this should be done along the way (line \ref{rule10}), during the organization process, by being transparent and communicating inclusion efforts and limitations (lines \ref{accountability-1} to \ref{accountability-2}):
 \textit{Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
-This is not only to communicate your values and practices, but as a way to hold yourself accountable for them.
-Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. 
-For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering help to navigate it, or if captions will be available for some talks but not all.
-Provide a point of contact to help clarify any questions regarding accessibility, financial support, and others.}
+These documents will not only communicate the conference values and practices but will act as a built-in mechanism to hold the organizers accountable.
+Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues.
+For instance, inform participants if the conference platform is not completely screen-reader friendly and let people know if you are offering help to navigate it, or highlight on the conference website if captions will be available for some talks but not all.
+Provide a point of contact to help clarify any questions regarding accessibility, financial support, and other issues affecting participation.}
 
 As stated in Rule 10 (lines \ref{account-3} to \ref{data-share2}), there should be a final assessment of the achievement of the diversity and inclusion goals at the end of the conference.
-Many tools can be used for that (e.g.
-demographic information, surveys, focus groups, information from informal conversations).
+Many tools can be used for that (e.g.\ demographic information, surveys, focus groups, information from informal conversations).
 Publishing these final assessments is an important way to hold the organizers accountable both to the greater community and the committee itself.
@@
 \begin{Reply}
+
 This specific section refers to the final selection of submitted proposals. Counteracting bias in invited roles was mentioned earlier in the text in line \ref{invited}.
-We agree with the reviewer that `curation' was a misleading term.
-We rephrased it to \textit{
+We agree with the reviewer that `curation' was an ambiguous term.
+We rephrased the section as follows: \textit{
 Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
-Examine the final list of selected abstracts to verify if there is lack of diversity.
-If proposals with similar quality were still judged differently because of apparent bias, consider giving preference to the work from people in minoritized groups.
-This manual examination at a final stage is unlikely to be reproducible.
+It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity.
+If proposals with similar quality were judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
+This manual examination at the final stage is unlikely to be reproducible.
 Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} (lines \ref{criteria-1}-\ref{criteria-2}).
@@
 \begin{Reply}
-Although one can assume that an ever growing proportion of people have their own handheld device to interact virtually with the remote participants, we cannot assume that everyone does, or that the models are compatible with the software platforms used in the conference.
-If the conference organizers provides such devices, Q\&A sessions and networking activities can be online and allow for more integration between in-person and online participants. 
+
+Although one can assume that an ever growing proportion of people have their own handheld device to interact virtually with remote participants, we cannot assume that everyone does, or that the models are compatible with the software platforms used in the conference.
+If the conference organizers provide such devices, Q\&A sessions and networking activities can be held virtually and allow for more integration between in-person and online participants. 
 \end{Reply}
@@
 \begin{Reply}
+
     Thank you for this suggestion. 
-    The organizing team should decide whether to ask for two languages or just one depending on their capabilities. 
-    It is true that it is not trivial to submit abstracts in two languages, but for many people, writing in English may be cumbersome and, if given the choice, they would opt for also adding a version in their native language if that increases the chances to communicate their ideas accurately.
+    The organizing team should decide whether to ask for two languages or just one, depending on their capabilities. 
+    Indeed, it is not trivial to submit abstracts in two languages.
+However, from personal experience, we know that many people find English so cumbersome that, if given the choice, they would opt for adding a version in their native language to increase the chances of communicating their ideas accurately.
     The text now states: 
-    \textit{Allow for abstract submission in both English and the language the person feels more comfortable with, and whenever possible, assign a reviewer who is fluent in that language. 
-    Alternatively, the organizing team could define a set of languages that are accepted for abstract submission, and allow applicants to choose from one of those languages. 
+    \textit{Allow for abstract submission in both English and the applicant's preferred native language, and whenever possible, assign a reviewer who is fluent in that language. 
+    Alternatively, the organizing team could define a set of accepted languages for abstract submission, and allow applicants to choose from this list. 
     The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency} (lines \ref{abstractlanguage1} to \ref{abstractlanguage2}).
 
-    We also added a mention to accepting multilingual submissions when counteracting self-selection during the call for abstracts in Rule 4, linking to rule 7 (line \ref{multilingual-submission}).
+    We also added a mention to accepting multilingual submissions when counteracting self-selection during the call for abstracts in Rule 4, linking to Rule 7 (line \ref{multilingual-submission}).
 \end{Reply}
@@
 \begin{Reply}
+
     Thank you for the thorough and positive review.
 We hope to have addressed all of your concerns and particularly accountability.
-We agree with involving the community in every part of the process, in an active and participative manner, using different channels of open communication and integrating members of the community in the organizing team.
+We agree with involving the community in every part of the process in an active and participative manner, using different channels of open communication and integrating members of the community into the organizing team.
 \end{Reply}
@@
 \begin{Reply}
-    Thanks for reading and reviewing. The list given was meant to be an example of extant thematic groups and communities, aiming to go beyond gender representation.
-    We agree about adding women as an example of such groups in line \ref{wis}:
-    `Open the schedule for events organized by thematic groups and communities (e.g.
-task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.'
+
+    We thank the reviewer for this comment. The list given was meant to be an example of existing thematic groups and communities, aiming to go beyond gender representation.
+    We have now added women as an example in line \ref{wis}:
+    `Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.'
 \end{Reply}
@@
 This is a timely, well-written commentary on organising an inclusive conference and should act as a reference compendium for researchers, practitioners as well as conference organisers.
+
+\begin{Reply}
+
+    We thank the reviewer for the positive comment and the detailed suggestions.
+\end{Reply}
 
@@
 \begin{Reply}
-    A Diversity, Equity and Inclusion (DEI) committee would be the best suited to make sure that the inclusion vision and goals are achieved. 
-    Establishing a DEI team is now suggested in Rule 3 in the manuscript (lines \ref{subteam-1} to \ref{subteam-2}). Thank you for that. 
+
+    A Diversity, Equity, and Inclusion (DEI) committee would be best suited to ensure that the inclusion vision and goals are achieved. 
+    Establishing a DEI team is now suggested in Rule 3 in the manuscript (lines \ref{subteam-1} to \ref{subteam-2}). Thank you for pointing this out. 
     \textit{ 
-    Having a dedicated DEI team may be a good way to give the attention and energy the topic requires.
-    The leading team should provide support to the DEI initiatives and advocate for inclusion across all conference areas.}
-    On the other hand, the existence of a DEI team is neither absolutely necessary nor sufficient, and we believe the most important thing is the commitment of the 
-    whole organizing team towards inclusion, and the willingness to hold themselves accountable, with concrete actions mentioned in Rule 8 (during the conference, lines \ref{accountability-1} to \ref{accountability-2})  and in Rule 10 (after the conference, lines \ref{account-3} to \ref{data-share2}) .
+    Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
+    The leadership team should support the DEI initiatives and advocate for inclusion across all conference areas.}
+    On the other hand, the existence of a DEI team is neither necessary nor sufficient, and we believe the most important thing is the commitment of the 
+    whole organizing team towards inclusion, and the willingness to hold themselves accountable, with concrete actions mentioned in Rule 8 (during the conference, lines \ref{accountability-1} to \ref{accountability-2}) and Rule 10 (after the conference, lines \ref{account-3} to \ref{data-share2}) .
 
@@
 \begin{Reply}
+
     The rules were first drafted by the first author, based on the collective experience of organizing useR!
-2021 and discussions with other organizers; then the leading group and the members of the diversity, accessibility, and inclusion team were invited to write their own ten simple rules—the rules proposed at that stage are now in the Supplementary Information.
+2021 and discussions with other organizers; then, the core group and the members of the diversity, accessibility, and inclusion team were invited to write their own ten simple rules—the rules proposed at that stage are now included in the Supplementary Information.
     We reached a final list by consensus, not popularity.
-All the other members of the organizing team and the people who acted as helpers in the conference were invited to give their feedback to the resulting list, in order to check for omissions and errors and suggest editions to the text.
+All the other members of the organizing team and the people who helped during the conference (designated helpers) were invited to give their feedback to the resulting list to check for omissions and errors and suggest editions to the text.
     We also sent the manuscript to friendly reviewers who have the experience of attending or organizing conferences. 
-    These rules, or preliminary versions of them, were shared with over 100 people, though not all of them provided feedback.
-    The rules in this manuscript are a result of this process and further discussions with and contributions from all the coauthors, fueled by personal experiences and literature review.
+    These rules, or their preliminary versions, were shared with 54 people; we received feedback from $\sim$40\%.
+    The rules in this manuscript result from this iterative process, discussions with and contributions from all the coauthors, fueled by personal experiences, and extensive literature review.
     A brief description of this process is now stated in lines \ref{methods-a}-\ref{methods-b} of the current version of the manuscript.
-    We kept the methodology as brief as possible to keep the focus on the rules and not our team or conference.
+    We kept the methodology as brief as possible to focus on the rules instead of our team or conference.
 
-    The coauthors are from a diversity of backgrounds, career stages, and even continents, and have experienced inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege, as stated in lines \ref{background-a}-\ref{background-b}.
-    The manuscript is the result our work together, as well as the awareness and experience of organizing a conference that aimed to be a kind, inclusive, and accessible experience for as many people as possible.
+    The coauthors are from diverse backgrounds, career stages, and different continents.
+They have experienced inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege, as stated in lines \ref{background-a}-\ref{background-b}.
+    The manuscript summarizes our work together and is strongly supported by the awareness and experience of organizing a global conference that aims to be a kind, inclusive, and accessible experience for as many people and intersectionalities as possible.
 \end{Reply}
@@
 \begin{Reply}
-   In a global scale, there is an imbalance associated to what could be considered a `north-south' divide, though the definition of north and south in this sense may not be clear, and we agree that being from a lower income country hinders participation in conferences.
-For that reason, we offer an example of registration rates based on country of origin among other aspects (\ref{example-countries}).
-   On the other hand, while the manuscript is heavily based on our experience organizing a global conference—which is why there are some global examples—, the intention is to make the rules general and flexible enough so that they can be useful for local or regional conferences around the world.
-   For that reason, we explicitly mention in Rule 1 that \textit{Lack of diversity will vary between areas of knowledge and depend on the geographic context of the conference.} (line \ref{indicators-1}).
+
+   On a global scale, there is an imbalance associated with what could be considered a `north-south' divide, though the definition of north and south may not be clear.
+We agree that coming from a lower income country hinders participation in conferences.
+For that reason, we offer an example of registration rates based on country of origin and career status (https://spcanelon.github.io/useR2021-cost-conversion-tool/).
+   On the other hand, while the manuscript is heavily based on our experience organizing a global conference—which is why there are some global examples—the intention is to make the rules general and flexible enough to be useful for local or regional conferences around the world.
+   For that reason, we explicitly mention in Rule 1 that \textit{Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.} (line \ref{indicators-1}).
 
@@
 \begin{Reply}
-    Racist acts should be unacceptable behavior in any conference, but it is not our intention to prepare a comprehensive list of unacceptable behaviors in Rule 2.
-    All specific examples of offenses should be included in the code of conduct of each conference, and listing them should the task of the organizers, as well as establishing mechanisms for reporting and addressing those reports.
-    We also made explicit the link between the diversity statement and the code of conduct, to highlight that all dimensions of diversity should be respected (lines \ref{unacceptable-1} to \ref{unacceptable-2}):
+
+    Racist acts should be unacceptable behavior in any conference, but we did not mention any example of offenses (including expressions of racism, sexism, ableism, and others).
+They should be included explicitly in the code of conduct of each conference, as well as the mechanisms for reporting and addressing those reports.
+    To further address this comment, we made explicit the link between the diversity statement and the code of conduct, to highlight that all dimensions of diversity should be respected (lines \ref{unacceptable-1} to \ref{unacceptable-2}):
     \textit{The CoC is a living document meant to keep the community safe and should state clearly the behaviors that are deemed unacceptable by the community, the consequences for engaging in such behaviors, and the way to report violations.
@@
 \begin{Reply}
-    In the selection of invited speakers, biases could be related to multiple dimensions of diversity identified in the assessment mentioned in Rule 1. 
-    All of these should be equally addressed.
 
-    In addition, special attention should be given to unconscious bias in the revision of submitted work, which is usually related to aspects of human diversity that can be inferred via written communication.
-    We had mentioned [perceived] ethnicity as one of these aspects  (line \ref{bias}, \textit{`their perceived origin, ethnicity, gender, or native language'}).
+    We had mentioned [perceived] ethnicity as one of the aspects of diversity that may be inferred (correctly or incorrectly) via written communication and be a source of unconscious bias  (line \ref{bias}, \textit{`their perceived origin, ethnicity, gender, or native language'}).
 
@@
 We added a short mention and citation about how video abstracts can be exclusionary (Spitschan et al 2021) by adding barriers to people with visible disabilities, diverse body types, non-native language speakers, or less resources to create videos (lines \ref{video-1} to \ref{video-2}).
+
+    Spitschan M, Yehudi Y, Acion L. Video Grant Proposals Could Be Exclusionary. 
+    Nature. 2021; 592(7852):26–26. doi:10.1038/d41586-021-00862-x.
+
     We thank the reviewer for this comment that led us to the current addition.
@@
 \begin{Reply}
-    We added the following text to the manuscript (lines \ref{promotion-1} to \ref{promotion-2}): 
-    \textit{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility than all other sessions.}
+
+    We added the following text to the manuscript (lines \ref{promotion-1} to \ref{promotion-2}) to address this comment: 
+    \textit{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility as all other sessions.}
 \end{Reply}
@@
 \begin{Reply}
-% RJ: I'm leaning towards not adding this now.
+
     We appreciate the suggestion and agree that the COVID-19 pandemic contributed to the exacerbation of inequity. 
-    However, we would like the manuscript to help inspire and guide the organization of more inclusive conferences regardless of the state of the pandemic in the different places in the world, and even when the pandemic subsides.
-For that reason, we tried to mention the pandemic very few times and only when deemed necessary: to give context to the conference that we organized, and to talk about virtual conferences which became the prevailing format during the last two years.
-\end{Reply}
+    We would like our manuscript to help inspire and guide the organization of more inclusive conferences across the world, well beyond the pandemic.
+To ensure this broad applicability, we minimized mentions of the pandemic to only when deemed necessary: to give context to the conference we organized, and to talk about virtual conference formats that prevailed during the last two years.
 
-\begin{Reply}
-    We thank the reviewer for their comments and careful revision.
+    Once again, we thank all the reviewers for their comments and careful consideration.  The feedback has helped clarify the language and the message.
 \end{Reply}

### 2022-04-20 plos_latex_template.tex
RJ: accepted changes to responses to the reviewers

@@
 Sara Mortara\textsuperscript{3},
-Yanina Bellini Saibene\textsuperscript{4,5,6},
-Heather Turner\textsuperscript{7}, %\ddag},
-Dorothea Hug Peter\textsuperscript{8}, %\ddag},
-Natalia Soledad Morandeira\textsuperscript{9},
-Matt Bannert\textsuperscript{10},
-Batool Almazrouq\textsuperscript{11,12},
-Elizabeth Hare\textsuperscript{13},
-Laura Ación\textsuperscript{5,14},
-Juan Pablo Narváez-Gómez\textsuperscript{15},
-Marcela Alfaro Córdoba\textsuperscript{16},
-Federico Marini\textsuperscript{17},
-Rita Giordano\textsuperscript{18},
-Silvia Canelón\textsuperscript{19},
-Anicet Ebou\textsuperscript{20},
-Adithi Upadhya\textsuperscript{21},
-Joselyn Chávez\textsuperscript{22},
-Janani Ravi\textsuperscript{23}
+Yanina Bellini Saibene\textsuperscript{4,5},
+Heather Turner\textsuperscript{6}, %\ddag},
+Dorothea Hug Peter\textsuperscript{7}, %\ddag},
+Natalia Soledad Morandeira\textsuperscript{8},
+Matt Bannert\textsuperscript{9},
+Batool Almazrouq\textsuperscript{10},
+Elizabeth Hare\textsuperscript{11},
+Laura Ación\textsuperscript{5,12},
+Juan Pablo Narváez-Gómez\textsuperscript{13},
+Marcela Alfaro Córdoba\textsuperscript{14},
+Federico Marini\textsuperscript{15},
+Rita Giordano\textsuperscript{16},
+Silvia Canelón\textsuperscript{17},
+Anicet Ebou\textsuperscript{18},
+Adithi Upadhya\textsuperscript{19},
+Joselyn Chávez\textsuperscript{20},
+Janani Ravi\textsuperscript{21*}
 %,3\textcurrency}
@@
 \\
-\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina
+\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina; Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
 \\
@@
 \\
-\textbf{6} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
-\\
-\textbf{7} Department of Statistics, University of Warwick, Coventry, UK
-\\
-\textbf{8} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
-\\
-\textbf{9} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
-\\
-\textbf{10} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
-\\
-\textbf{11} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
-\\
-\textbf{12} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
-\\
-\textbf{13} Dog Genetics LLC, Astoria, NY, USA
-\\
-\textbf{14} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
-\\
-\textbf{15} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
-\\
-\textbf{16} Department of Statistics, University of California, Santa Cruz, CA, USA
-\\
-\textbf{17} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
-\\
-\textbf{18} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
-\\
-\textbf{19} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
-\\
-\textbf{20} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
-\\
-\textbf{21} ILK Labs/Bengaluru, Karnataka 560046, India
-\\
-\textbf{22} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
-\\
-\textbf{23} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
+\textbf{6} Department of Statistics, University of Warwick, Coventry, UK
+\\
+\textbf{7} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
+\\
+\textbf{8} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
+\\
+\textbf{9} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
+\\
+\textbf{10} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK; King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
+\\
+\textbf{11} Dog Genetics LLC, Astoria, NY, USA
+\\
+\textbf{12} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
+\\
+\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
+\\
+\textbf{14} Department of Statistics, University of California, Santa Cruz, CA, USA
+\\
+\textbf{15} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
+\\
+\textbf{16} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
+\\
+\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
+\\
+\textbf{18} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\\
+\textbf{19} ILK Labs, Bengaluru, Karnataka, 560046, India
+\\
+\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
+\\
+\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
 \\
@@
 % Primary Equal Contribution Note
-\Yinyang These authors share first authorship of this work.
+\Yinyang Co-primary authors with equal contribution.
 
@@
 % Use the asterisk to denote corresponding authorship and provide email address in note below.
-* rocio.joo@globalfishingwatch.org
+*Co-corresponding authors: rocio.joo@globalfishingwatch.org, janani@msu.edu
 
@@
 
-Conferences are spaces to meet and network within and across academic and technical fields, to learn about new advances, and share our work.
+Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work.
 They can help define career paths and create long-lasting collaborations and opportunities.
@@
 The rules result from our lessons learned before, during, and after the organization of the conference. 
-They have been mainly written for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive.
+They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive.
 We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
-\hl{* Versions of this abstract in Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai are available in Supplementary Information.}
+\hl{* Translated versions of the extended abstract are available in 10 languages in the Supplementary Information--Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}
 
@@
 Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
-\linelabel{methods-a} \hl{The rules were first drafted by the leading group and  members of the diversity, accessibility, and inclusion team of useR!
+\linelabel{methods-a} \hl{The rules were first drafted by the core team and members of the diversity, accessibility, and inclusion team of useR!
 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
@@
 Finally, 
-\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and passing the torch to future organizers.
+\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.
 
@@
 \linelabel{indicators-1}\hl{
-Lack of diversity will vary between areas of knowledge and depend on the geographic context of the conference.
-Assess how it translates to your particular field and scientific or professional community.
-Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM) if they exist. 
-Most importantly, involve the community---that is, the subset of the general public that would ideally attend the conference---to help make this assessment and to set the vision and goals of diversity and inclusion for your conference.}
+Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
+Assess how it translates to the conference's particular field and scientific or professional community.
+Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering and mathematics) if they exist. 
+Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.}
 \linelabel{indicators-2}
@@
 a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions).
-They will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
+These goals and measures will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
 The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.
@@
 % Explain why we need this
-Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community.
+Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part in the conference activities and community.
 Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees.
@@
 Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it is good to have them in mind from the beginning.
-Open the schedule for events organized by thematic groups and communities (e.g.
-task force meetings, LGBTQIA+-friendly spaces, Black in STEM\hl{, women in science}\linelabel{wis}), and sessions aimed to welcome first-timers to the conference.
+Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM\hl{, women in science}\linelabel{wis}), and sessions aimed to welcome first-timers to the conference.
 
 % A culture of support (active bystandership) and code of conduct for a safe place 
-Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, incorrect, or offensive happens.
+Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, inappropriate, or offensive happens.
 This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
@@
 The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community.
-The team should be prepared early during conference planning because CoC violations can happen before the conference too (e.g.
-social media, organization meetings).
-Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding their work to your budget (see \textbf{Rule 9}).
-We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.
+The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media, or in organization meetings).
+Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding funds for their work to your budget (see \textbf{Rule 9}).
+We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.
 
@@
 is especially important; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
-In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
+In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupying critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
 \hl{If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
@@
 
-\linelabel{subteam-1} \hl{Create subteams to distribute fairly the large amount of work that organizing a conference represents. 
-Having a dedicated DEI team may be a good way to give the attention and energy the topic requires.
-The leading team should provide support to the DEI initiatives and advocate for inclusion across all conference areas.} \linelabel{subteam-2} 
+\linelabel{subteam-1} \hl{Create sub-teams to distribute fairly the large amount of work that organizing a conference represents. 
+Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
+The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} \linelabel{subteam-2} 
 It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
@@
 Check on them regularly and make sure that everyone is comfortable.
-This represents a great amount of work and you will need to support each other in the long run.
+This represents a great amount of work and you will need to support each other for the long haul.
 
@@
 \linelabel{invited}
-Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.
+Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and on-board great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.
 
@@
 The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias}
-Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
+Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
 Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
@@
 \linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
-Examine the final list of selected abstracts to verify if there is lack of diversity.
-If proposals with similar quality were still judged differently because of apparent bias, consider giving preference to the work from people in minoritized groups.
-This manual examination at a final stage is unlikely to be reproducible.
+It is, therefore, essential to examine the final list of selected abstracts to check if there is lack of diversity.
+If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
+This manual examination at the final stage is unlikely to be reproducible.
 Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
@@
 The audience, chairs, and presenters should all be able to interact regardless of their in-person or remote status. 
-To best integrate in-person and remote attendees, all questions and comments should pass through a unique online system. 
+To best integrate in-person and remote attendees, all questions and comments should pass through a single online system. 
 The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.
@@
 Invisible disabilities (e.g.
-dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces (but see \cite{pun_dos_2016} for other examples).
+dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces (see \cite{pun_dos_2016} for other examples).
 
@@
 While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege \cite{amanoTenTipsOvercoming2021}.
-Non-native English speakers may miss opportunities to attend or to actively participate in conferences (e.g. presenting, asking questions, or partaking in discussions),
+Non-native English speakers may miss opportunities to attend or to actively participate in conferences (e.g. presenting, asking questions, or taking part in discussions),
 and conferences may in turn miss innovative contributions.
@@
 Create an inclusive environment by encouraging the full participation of non-native English speakers. This may be done in multiple ways.
-\linelabel{abstractlanguage1}\hl{Allow for abstract submission in both English and the language the person feels more comfortable with, and whenever possible, assign a reviewer who is fluent in that language.
-Alternatively, the organizing team could define a set of languages that are accepted for abstract submission, and allow applicants to choose from one of those languages. 
+\linelabel{abstractlanguage1}\hl{Allow for abstract submission in both English and the applicant's preferred native language, and whenever possible, assign a reviewer who is fluent in that language. 
+Alternatively, the organizing team could define a set of accepted languages for abstract submission, and allow applicants to choose from this list. 
 The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency}.\linelabel{abstractlanguage2}
@@
 Another way to engage non-native English participants is to host sessions and events in languages other than English. 
-Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same relevance as the rest of the program.
+Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same importance as the rest of the program.
 Importantly, consideration and respect of accents or linguistic mistakes can make a significant difference in the conference experience of non-native English speakers.
@@
 
-The communication strategy of your conference should aim to reach broader audiences and express the commitment with diversity and inclusion and the welcoming spirit of the conference.
+The communication strategy of your conference should aim to reach broader audiences and express the commitment to diversity and inclusion and the welcoming spirit of the conference.
 Actively reach out and promote the conference to communities that have been systematically excluded. 
@@
 Twitter, Facebook, LinkedIn, conference website, mailing lists).
-\linelabel{promotion-1}\hl{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility than all other sessions.}\linelabel{promotion-2}
+\linelabel{promotion-1}\hl{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility as all other sessions.}\linelabel{promotion-2}
 
 \linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
-\hl{This is not only to communicate your values and practices, but as a way to hold yourself accountable for them.}
-Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. 
-For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering help to navigate it, or if captions will be available for some talks but not all.
-Provide a point of contact to help clarify any questions regarding accessibility, financial support, and others.
+\hl{These documents will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.}
+Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues.
+For instance, inform participants if the conference platform is not completely screen-reader friendly and let people know if you are offering help to navigate it, or highlight on the conference website if captions will be available for some talks but not all.
+Provide a point of contact to help clarify any questions regarding accessibility, financial support, and other issues affecting participation.
 \linelabel{accountability-2}
@@
 Do not expect minoritized people to teach you and accept feedback without being offended.
-Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms.
+Inclusive language also encompasses avoidance of excessive and over-specific technical jargon and acronyms.
 
@@
 You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
-\linelabel{info}It is likely that you will be able to evaluate the outcomes associated to these goals directly from the information that you have already collected from your organizing team, speakers, and participants (e.g.
-\hl{city/state/country of origin}, or preferred language).
-You may have goals concerning how welcomed and included participants felt during the conference, and wonder if your practices translated into real inclusion. 
-Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this type of information and great opportunities to receive feedback from the attendees.
+\linelabel{info}It is likely that the organizers will be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin}, or preferred language).
+The planning committee may include goals concerning how welcomed and included participants felt during the conference, and consider if the implemented practices translated into true inclusion. 
+Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
 \linelabel{info2}
@@
 % Pass the information
-The feedback received from the participants and your whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers.
-\hl{Make sure that the privacy and anonymity of the participants is respected; ask for consent and consider binning/grouping the data for categories that have a low number of respondents}.\linelabel{data-share2}
+The feedback received from the participants and the whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers.
+\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data for categories when there is a low number of respondents}.\linelabel{data-share2}
 Gather and do your best to organize all the documentation produced during the organization process---or even after.
@@
 a wiki, knowledge base or similar guide, see for instance \cite{sanchez-tapia_user_2021-2}) that can be used and kept up to date by future organizers.
-If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria.
+If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit about the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organizers and defining the selection criteria.
 
@@
 
-This article recommends changes in conference planning towards diversity and inclusion. 
+This article recommends changes in conference planning to support greater diversity and inclusion. 
 Organizing a conference and implementing inclusive practices are both learning experiences.
@@
 Try to account for these factors as best as you can to set bold yet realistic goals of diversity and inclusion for the conference.
-And most importantly, your efforts will enhance the conference experience for your attendees. 
+Most importantly, your efforts will enhance the conference experience for all your attendees. 
 They will notice the welcoming and inclusive spirit of the conference. 
@@
     \item BA: Writing---major reviews and edits, Resources, Sharing rules, Extended abstract translation
-    \item LH: Writing---major reviews and edits, Resources, Sharing rules
+    \item EH: Writing---major reviews and edits, Resources, Sharing rules
     \item LA: Writing---major reviews and edits, Resources
@@
 2021.
-HT was a board member of the The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
+HT is a board member of the The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
 conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups.
@@
 RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively.
+
+For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission.
 

### 2022-04-20 response_reviewers.tex
manuscript and responses match

@@
 Comments from the Reviewers led to editing and paraphrasing of several parts of the manuscript, including
-i) direct mentions of the involvement of the community when defining diversity and inclusion priorities (in Rule 1)
-ii) ways to ensure accountability during the organizing process and after the conference (e.g.\ in Rule 2 with a Diversity, Equity, and Inclusion (DEI) team, Rule 8 with guidelines like the diversity statement and the code of conduct, and Rule 10 with post-conference continuity)
-iii) mention of women in science in Rule 2
-iv) a reference to the DEI team, biases due to video abstracts, and the connection between the code of conduct and the diversity statement
-v) the process we followed to put together this list of simple rules; now described in detail below and summarized in the current manuscript version.
+i) direct mentions of the involvement of the community when defining diversity and inclusion priorities (in Rule 1),
+ii) ways to ensure accountability during the organizing process and after the conference (e.g.\ in Rule 2 with a Diversity, Equity, and Inclusion (DEI) team, Rule 8 with guidelines like the diversity statement and the code of conduct, and Rule 10 with post-conference continuity),
+iii) mention of women in science in Rule 2,
+iv) biases due to video abstracts, 
+v) the connection between the code of conduct and the diversity statement, and
+v) the process we followed to put together this list of simple rules---now described in detail below and summarized in the current manuscript version.
 We also included the sets of rules gathered by the team as Supplementary Information.
@@
 
-Roc\'io Joo and Janani
+Roc\'io Joo and Janani Ravi
 
@@
     Assess how it translates to the conference's particular field and scientific or professional community.
-    Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if it exists. 
+    Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if they exist. 
     Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.
@@
 (...) The organizers will likely be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ city/state/country of origin or preferred language).
-The planning committee may include goals concerning how welcome and included participants felt during the conference and consider if the implemented practices translated into true inclusion. 
+The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
 Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
@@
 We added more details about this in Rule 10, lines \ref{data-share} to \ref{data-share2}:
-\textit{The feedback received from the participants, and the whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendees, and future organizers.
+\textit{The feedback received from the participants and the whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendees, and future organizers.
 Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents.}
@@
 \textit{Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
-These documents will not only communicate the conference values and practices but will act as a built-in mechanism to hold the organizers accountable.
+These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.
 Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues.
@@
 It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity.
-If proposals with similar quality were judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
+If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
 This manual examination at the final stage is unlikely to be reproducible.
@@
     A Diversity, Equity, and Inclusion (DEI) committee would be best suited to ensure that the inclusion vision and goals are achieved. 
-    Establishing a DEI team is now suggested in Rule 3 in the manuscript (lines \ref{subteam-1} to \ref{subteam-2}). Thank you for pointing this out. 
+    Establishing a DEI team is now suggested in Rule 3 in the manuscript (lines \ref{subteam-1} to \ref{subteam-2}):
     \textit{ 
     Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
-    The leadership team should support the DEI initiatives and advocate for inclusion across all conference areas.}
+    The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} 
+    Thank you for pointing this out. 
     On the other hand, the existence of a DEI team is neither necessary nor sufficient, and we believe the most important thing is the commitment of the 
@@
 They have experienced inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege, as stated in lines \ref{background-a}-\ref{background-b}.
-    The manuscript summarizes our work together and is strongly supported by the awareness and experience of organizing a global conference that aims to be a kind, inclusive, and accessible experience for as many people and intersectionalities as possible.
+    The manuscript summarizes our work together and is strongly supported by the awareness and experience of organizing a global conference that aims to be a kind, inclusive, and accessible experience for as many people as possible.
 \end{Reply}

### 2022-04-20 plos_latex_template.tex
manuscript and responses match

@@
 Assess how it translates to the conference's particular field and scientific or professional community.
-Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering and mathematics) if they exist. 
+Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if they exist. 
 Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.}
@@
 a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions).
-These goals and measures will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
+They will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
 The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.
@@
 \linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
-It is, therefore, essential to examine the final list of selected abstracts to check if there is lack of diversity.
+It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity.
 If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
@@
 \linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
-\hl{These documents will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.}
+\hl{These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.}
 Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues.
@@
 You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
-\linelabel{info}It is likely that the organizers will be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin}, or preferred language).
-The planning committee may include goals concerning how welcomed and included participants felt during the conference, and consider if the implemented practices translated into true inclusion. 
+\linelabel{info}\hl{The organizers will likely be able} to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin} or preferred language).
+The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
 Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
@@
 The feedback received from the participants and the whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers.
-\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data for categories when there is a low number of respondents}.\linelabel{data-share2}
+\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents}.\linelabel{data-share2}
 Gather and do your best to organize all the documentation produced during the organization process---or even after.

### 2022-04-20 response_reviewers.tex
Response submitted :sunglasses:

@@
 \textit{
-(...) The organizers will likely be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ city/state/country of origin or preferred language).
-The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
+(...) You will likely be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ city/state/country of origin or preferred language).
+You may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
 Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
@@
 
-This specific section refers to the final selection of submitted proposals. Counteracting bias in invited roles was mentioned earlier in the text in line \ref{invited}.
+This specific section refers to the final selection of submitted proposals. Counteracting bias in invited roles was mentioned earlier in the text in lines \ref{invited} to \ref{invited-2}.
 We agree with the reviewer that `curation' was an ambiguous term.
@@
 If the conference organizers provide such devices, Q\&A sessions and networking activities can be held virtually and allow for more integration between in-person and online participants.
+We edited the manuscript for clarity: \textit{The physical venue would have to include multiple ways to connect with online folks [...], and tablets to help in-person attendees without suitable personal devices to interact with remote participants during Q\&A or networking sessions} (line \ref{tablets}).
+
 \end{Reply}
@@
 
-    We also added a mention to accepting multilingual submissions when counteracting self-selection during the call for abstracts in Rule 4, linking to Rule 7 (line \ref{multilingual-submission}).
+    We also added a mention to accepting multilingual submissions when counteracting self-selection during the call for abstracts in Rule 4, linking to Rule 7 (lines \ref{multilingual-submission} to \ref{multilingual-submission2}).
 \end{Reply}
@@
     We have now added women as an example in line \ref{wis}:
-    `Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.'
+    \textit{Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.}
 \end{Reply}
@@
     On the other hand, the existence of a DEI team is neither necessary nor sufficient, and we believe the most important thing is the commitment of the 
-    whole organizing team towards inclusion, and the willingness to hold themselves accountable, with concrete actions mentioned in Rule 8 (during the conference, lines \ref{accountability-1} to \ref{accountability-2}) and Rule 10 (after the conference, lines \ref{account-3} to \ref{data-share2}) .
+    whole organizing team towards inclusion, and the willingness to hold themselves accountable, with concrete actions mentioned in Rule 8 (during the conference, lines \ref{accountability-1} to \ref{accountability-2}) and Rule 10 (after the conference, lines \ref{account-3} to \ref{info2}) .
 
@@
    On the other hand, while the manuscript is heavily based on our experience organizing a global conference—which is why there are some global examples—the intention is to make the rules general and flexible enough to be useful for local or regional conferences around the world.
-   For that reason, we explicitly mention in Rule 1 that \textit{Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.} (line \ref{indicators-1}).
+   For that reason, we explicitly mention in Rule 1 that \textit{Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference} (line \ref{indicators-1}).
 

### 2022-04-20 clean.tex
Response submitted :sunglasses:

--- 2022-04-20 plos_latex_template.tex
@@
 We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
-\hl{* Translated versions of the extended abstract are available in 10 languages in the Supplementary Information--Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}
+* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.
 
@@
 Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
-\linelabel{methods-a} \hl{The rules were first drafted by the core team and members of the diversity, accessibility, and inclusion team of useR!
+\linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR!
 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
-A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
+A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
 The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
@@
 
-\linelabel{indicators-1}\hl{
+\linelabel{indicators-1}
 Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
@@
 Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if they exist. 
-Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.}
+Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.
 \linelabel{indicators-2}
@@
 a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions).
-They will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
+They will help you be accountable along the way (see \textbf{Rule 10}).
 The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.
@@
 Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it is good to have them in mind from the beginning.
-Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM\hl{, women in science}\linelabel{wis}), and sessions aimed to welcome first-timers to the conference.
+Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.
 
@@
 \linelabel{unacceptable-1}The CoC is a living document meant to keep the community safe and should state clearly the behaviors that are deemed unacceptable by the community, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
-\hl{Bear in mind that every dimension of diversity can be a target of unacceptable behavior. The CoC should help honor the vision expressed in the diversity statement.}\linelabel{unacceptable-2}
+Bear in mind that every dimension of diversity can be a target of unacceptable behavior. The CoC should help honor the vision expressed in the diversity statement.
 
 The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community.
-The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media, or in organization meetings).
+The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media or in organization meetings).
 Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding funds for their work to your budget (see \textbf{Rule 9}).
@@
 In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupying critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
-\hl{If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
-The regional and local communities, groups, or associations in your field are good sources to tap into.}
-
-\linelabel{subteam-1} \hl{Create sub-teams to distribute fairly the large amount of work that organizing a conference represents. 
+If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
+The regional and local communities, groups, or associations in your field are good sources to tap into.
+
+Create sub-teams to distribute fairly the large amount of work that organizing a conference represents. 
 Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
-The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} \linelabel{subteam-2} 
+The core team should support the DEI initiatives and advocate for inclusion across all conference areas.
 It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
-every member should have the freedom to choose which areas of the conference they want to contribute to.
+every person should have the freedom to choose which areas of the conference they want to contribute to.
 When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
-Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
-Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
+Instead, nominate folks directly and offer them leadership positions in the organizing team, especially those positions that would be typically occupied by people from privileged groups.
+Moreover, some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
 In any case, try to find meaningful ways to acknowledge everybody's work.
@@
 If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
-\linelabel{multilingual-submission}\hl{To mitigate self-selection from non-native speakers, consider accepting submissions in several languages} (see \textbf{Rule 7}).
+To mitigate self-selection from non-native speakers, consider accepting submissions in several languages (see \textbf{Rule 7}).
 
@@
 Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
-Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
+Evaluate the plausibility of anonymizing the submission information as much as possible and
 define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
-\linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
+Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
 It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity.
@@
 This manual examination at the final stage is unlikely to be reproducible.
-Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
-\linelabel{video-1}\hl{It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.} \cite{spitschanVideoGrantProposals2021}\linelabel{video-2}
+Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.
+It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.
+\cite{spitschanVideoGrantProposals2021}
 
@@
 To best integrate in-person and remote attendees, all questions and comments should pass through a single online system. 
-The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.
+The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees without suitable personal devices to interact with remote participants during Q\&A or networking sessions.
 
@@
 Create an inclusive environment by encouraging the full participation of non-native English speakers. This may be done in multiple ways.
-\linelabel{abstractlanguage1}\hl{Allow for abstract submission in both English and the applicant's preferred native language, and whenever possible, assign a reviewer who is fluent in that language. 
+Allow for abstract submission in both English and the applicant's preferred native language, and whenever possible, assign a reviewer who is fluent in that language. 
 Alternatively, the organizing team could define a set of accepted languages for abstract submission, and allow applicants to choose from this list. 
-The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency}.\linelabel{abstractlanguage2}
+The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency.
 Whenever possible, identify other prominent languages for the conference and provide translated captions or live language interpretation into these key languages, including multilingual Q\&A sessions.
@@
 Twitter, Facebook, LinkedIn, conference website, mailing lists).
-\linelabel{promotion-1}\hl{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility as all other sessions.}\linelabel{promotion-2}
+During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility as all other sessions.
 
 \linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
-\hl{These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.}
+These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.
 Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues.
@@
 You could also aim to have a conference with no registration costs; this could especially apply for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}.
-Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from minoritized groups, by offering support for travel and lodging expenses.
-Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
+Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from minoritized groups by offering support for travel and lodging expenses.
+Design the diversity-related scholarship to be explicit about the groups you want to support and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
 To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
@@
 You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
-\linelabel{info}\hl{The organizers will likely be able} to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin} or preferred language).
-The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
+You will likely be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ city/state/country of origin or preferred language).
+You may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
 Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
-\linelabel{info2}
 
 % Pass the information
-The feedback received from the participants and the whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers.
-\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents}.\linelabel{data-share2}
+The feedback received from the participants and the whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendees, and future organizers.
+Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents.
 Gather and do your best to organize all the documentation produced during the organization process---or even after.
 Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors.
-Create a list with contact information of the people of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
+Create a list with contact information of the members of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
 For conference series, it could be a good idea to create an official document (e.g.
@@
 2021.
-HT is a board member of the The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
+HT is a board member of The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
 conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups.

### 2022-04-20 plos_latex_template.tex
Updates from Overleaf

@@
 \\
-\textbf{6} Department of Statistics, University of Warwick, Coventry, UK
+\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation
 \\
@@
 We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
-\hl{* Translated versions of the extended abstract are available in 10 languages in the Supplementary Information--Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}
+\hl{* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}
 
@@
 Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
-\linelabel{methods-a} \hl{The rules were first drafted by the core team and members of the diversity, accessibility, and inclusion team of useR!
+\linelabel{methods-a} \hl{The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR!
 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
@@
 The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community.
-The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media, or in organization meetings).
+The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media or in organization meetings).
 Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding funds for their work to your budget (see \textbf{Rule 9}).
@@
 It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
-every member should have the freedom to choose which areas of the conference they want to contribute to.
+every \hl{person} should have the freedom to choose which areas of the conference they want to contribute to.
 When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
-Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
-Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
+Instead, nominate folks directly and offer them leadership positions in the organizing team, especially those positions that would be typically occupied by people from privileged groups.
+\hl{Moreover,} some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
 In any case, try to find meaningful ways to acknowledge everybody's work.
@@
 
-The conference program may lack diversity in the invited roles––-such as keynote speakers, program/scientific committee, session chairs, panelists––-, the accepted abstracts, and even the scope of the topics covered.
+\linelabel{invited}The conference program may lack diversity in the invited roles––-such as keynote speakers, program/scientific committee, session chairs, panelists––-, the accepted abstracts, and even the scope of the topics covered.
 It is important to counteract biases in all of these aspects.
 When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
-\linelabel{invited}
-Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and on-board great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.
+Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and on-board great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.\linelabel{invited-2}
 
@@
 If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
-\linelabel{multilingual-submission}\hl{To mitigate self-selection from non-native speakers, consider accepting submissions in several languages} (see \textbf{Rule 7}).
-
-The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias}
+\linelabel{multilingual-submission}\hl{To mitigate self-selection from non-native speakers, consider accepting submissions in several languages} (see \textbf{Rule 7}).\linelabel{multilingual-submission2}
+
+The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and \linelabel{bias} their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.
 Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
-Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
+Evaluate the plausibility of anonymizing \hl{the submission information as much as possible}\linelabel{anonymous} and
 define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
@@
 To best integrate in-person and remote attendees, all questions and comments should pass through a single online system. 
-The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.
+The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees \hl{without suitable personal devices to interact with remote participants during Q\&A or networking sessions}.\linelabel{tablets}
 
@@
 You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
-\linelabel{info}\hl{The organizers will likely be able} to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin} or preferred language).
-The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
+\linelabel{info}\hl{You will likely be able} to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin} or preferred language).
+You may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
 Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
@@
 % Pass the information
-The feedback received from the participants and the whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers.
+\linelabel{data-share}The feedback received from the participants and the whole assessment of the conference should be documented and shared with \hl{those who helped define the inclusion goals, the attendees, and} future organizers.
 \hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents}.\linelabel{data-share2}
@@
 Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors.
-Create a list with contact information of the people of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
+Create a list with contact information of the members of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
 For conference series, it could be a good idea to create an official document (e.g.
@@
 
-HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR!
+HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing for the organization of useR!
 2021.
-HT is a board member of the The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
+HT is a board member of The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
 conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups.
@@
 RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively.
-
+JR is funded by start-up award from Michigan State University; publication cost will waived due to PLoS-MSU partnership.
+
+\section*{Data availability}
 For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission.
+The manuscript and the supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.
 

### 2022-04-21 manuscript.tex
fig name update
Updates from Overleaf

--- 2022-04-20 plos_latex_template.tex
@@
 We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
-\hl{* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}
+\hl{* Translated versions of the English abstract and the list of rules are available in ten languages in the Supporting Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}
 
@@
 \linelabel{methods-a} \hl{The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR!
-2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
+2021 who wrote their own ten simple rules (see Supporting Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
 A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
@@
 \pdftooltip{
-\includegraphics[width=\textwidth]{figs/fig1_rules.png}}{
+\includegraphics[width=\textwidth]{figs/fig1.png}}{
 %Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundation rules (Rules 1-3), design rules (Rules 4 to 9), and a continuity rule (Rule 10).
@@
 \pdftooltip{
-\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
+\includegraphics[width=\textwidth]{figs/fig2.pdf}}{
 }
@@
 To list author contributions, we used the high-level contributor roles in the CRediT framework that we considered relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and additional categories to acknowledge those who shared their own ten simple rules for an inclusive conference without looking at the first draft (i.e.
-sharing rules), and those who translated the extended abstract into different languages.
+sharing rules), and those who translated the English abstract and the list of rules into different languages.
 
@@
 2021 and supporting us through the process.
-Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.
+Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the Marmot figures.
 We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript.
-In addition to some of the authors, we would like to acknowledge the following contributors for translating our extended abstract to several languages: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae [Thai].
+In addition to some of the authors (listed in the Translated Abstracts section of Supporting Information), we would like to acknowledge the following contributors for translating our extended abstract: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae [Thai].
 The comments of the anonymous reviewers contributed to the improvement of the manuscript and we thank them for that.
@@
 
-HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing for the organization of useR!
-2021.
-HT is a board member of The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
-conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups.
-The R Foundation is currently funding YBS to support the useR!
-2022 organization team.
-HT was administrator and volunteer for the R Project Google Season of Docs 2021 project, that funded work by AST (writer) and YBS (steering committee).
+JR is funded by start-up award from Michigan State University (MSU); publication costs were waived because of MSU’s institutional partnership with PLoS.
+HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.
+
 The organization of useR!
-2021 was largely a volunteer effort, but stipends were available from the conference funds for specific roles.
-RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively.
-JR is funded by start-up award from Michigan State University; publication cost will waived due to PLoS-MSU partnership.
+2021---which inspired this work---was largely a volunteer effort (from many people including all the authors), but stipends were given from the conference funds for some specific roles.
+AE, AU, AST, BA, JC, JPNG, MAC, NM, RG, RJ, SM, and YBS received stipends for roles that include code of conduct team, zoom host, and diversity and social program leads.
 
 \section*{Data availability}
-For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission.
-The manuscript and the supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.
+For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any author accepted manuscript version arising from this submission.
+The manuscript, figures, supplementary material, and supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.
 
@@
 
+\section*{Supporting Information}
+
+Supporting Information 1. Abstract and list of rules in multiple languages: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.
+
+Supporting Information 2. Rules proposed in the process of shaping these ten simple rules.
+
 \end{document}

### 2022-04-21 final_manuscript_no_markup.tex

--- 2022-04-20 clean.tex
@@
 We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
-* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.
+* Translated versions of the English abstract and the list of rules are available in ten languages in the Supporting Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.
 
@@
 \linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR!
-2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
+2021 who wrote their own ten simple rules (see Supporting Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
 A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
@@
 
-HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR!
-2021.
-HT is a board member of The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR!
-conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups.
-The R Foundation is currently funding YBS to support the useR!
-2022 organization team.
-HT was administrator and volunteer for the R Project Google Season of Docs 2021 project, that funded work by AST (writer) and YBS (steering committee).
+JR is funded by start-up award from Michigan State University (MSU); publication costs were waived because of MSU’s institutional partnership with PLoS.
+HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.
+
 The organization of useR!
-2021 was largely a volunteer effort, but stipends were available from the conference funds for specific roles.
-RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively.
-
-For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission.
+2021---which inspired this work---was largely a volunteer effort (from many people including all the authors), but stipends were given from the conference funds for some specific roles.
+AE, AU, AST, BA, JC, JPNG, MAC, NM, RG, RJ, SM, and YBS received stipends for roles that include code of conduct team, zoom host, and diversity and social program leads.
+
+\section*{Data availability}
+For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any author accepted manuscript version arising from this submission.
+The manuscript, figures, supplementary material, and supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.
 

### 2022-05-07 final_manuscript_no_markup.tex
the real submitted version :tada: :rocket:
some editing for final version

@@
 \bigskip
-\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
+\textbf{1} Global Fishing Watch, Washington,  District of Columbia, United States of America
 \\
@@
 \\
-\textbf{11} Dog Genetics LLC, Astoria, NY, USA
+\textbf{11} Dog Genetics LLC, Astoria, New York, United States of America
 \\
@@
 \\
-\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
-\\
-\textbf{14} Department of Statistics, University of California, Santa Cruz, CA, USA
+\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brazil
+\\
+\textbf{14} Department of Statistics, University of California, Santa Cruz, California, United States of America
 \\
@@
 \\
-\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
+\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, Pennsylvania, United States of America.
 \\
@@
 \\
-\textbf{19} ILK Labs, Bengaluru, Karnataka, 560046, India
-\\
-\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
-\\
-\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
+\textbf{19} ILK Labs, Bengaluru, Karnataka, India
+\\
+\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, Mexico
+\\
+\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, Michigan, United States of America
 \\
@@
 % Use the asterisk to denote corresponding authorship and provide email address in note below.
-*Co-corresponding authors: rocio.joo@globalfishingwatch.org, janani@msu.edu
+*rocio.joo@globalfishingwatch.org, janani@msu.edu
 
@@
 We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
-* Translated versions of the English abstract and the list of rules are available in ten languages in the Supporting Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.
+* Translated versions of the English abstract and the list of rules are available in ten languages in S1 Text: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.
 
@@
 \linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR!
-2021 who wrote their own ten simple rules (see Supporting Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
+2021 who wrote their own ten simple rules (see S1 Text) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
 A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
@@
 2021 and supporting us through the process.
-Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the Marmot figures.
+Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, Francisco Etchart for creating the Marmot figures, and Koki Tsuyuzaki, Kozo Nishida, KwangChun Lee, Arjun Krishnan, Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae for their extended abstract translations.
 We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript.
-In addition to some of the authors (listed in the Translated Abstracts section of Supporting Information), we would like to acknowledge the following contributors for translating our extended abstract: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae [Thai].
-The comments of the anonymous reviewers contributed to the improvement of the manuscript and we thank them for that.
 
@@
 
-JR is funded by start-up award from Michigan State University (MSU); publication costs were waived because of MSU’s institutional partnership with PLoS.
+JR is funded by start-up award from Michigan State University (MSU).
 HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.
 
-The organization of useR!
-2021---which inspired this work---was largely a volunteer effort (from many people including all the authors), but stipends were given from the conference funds for some specific roles.
-AE, AU, AST, BA, JC, JPNG, MAC, NM, RG, RJ, SM, and YBS received stipends for roles that include code of conduct team, zoom host, and diversity and social program leads.
+The funders had no role in study design, decision to publish, or preparation of the manuscript.
 
@@
 
+\bibitem{ross_everyday_2020}
+Ross HJ.
+\newblock Everyday {{Bias}}: {{Identifying}} and {{Navigating Unconscious
+  Judgments}} in {{Our Daily Lives}}.
+\newblock {Rowman \& Littlefield}; 2020.
+
 \bibitem{spitschanVideoGrantProposals2021}
@@
 \newblock doi:{10.1038/d41586-021-00862-x}.
-
-\bibitem{ross_everyday_2020}
-Ross HJ.
-\newblock Everyday {{Bias}}: {{Identifying}} and {{Navigating Unconscious
-  Judgments}} in {{Our Daily Lives}}.
-\newblock {Rowman \& Littlefield}; 2020.
 
@@
 
-\section*{Supporting Information}
-
-Supporting Information 1. Abstract and list of rules in multiple languages: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.
-
-Supporting Information 2. Rules proposed in the process of shaping these ten simple rules.
+\section*{Supporting Information Legends}
+
+S1 Text - Abstract and list of rules in multiple languages
+and rules proposed in the process of shaping these ten simple rules.
 

### 2022-05-10 manuscript.tex
the real submitted version :tada: :rocket:
affiliation corrections and file renaming + reference final PDF

@@
 \\
-\textbf{18} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\textbf{18} Bioinformatics team, Département de Formation et de Recherche Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, Yamoussoukro, Côte d’Ivoire.
 \\
@@
 
+The funders had no role in study design, decision to publish, or preparation of the manuscript.
+
 \section*{Data availability}

### 2022-05-10 reference_PDF.tex
affiliation corrections and file renaming + reference final PDF

--- 2022-05-07 final_manuscript_no_markup.tex
@@
 \\
-\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina; Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
-\\
-\textbf{5} MetaDocencia, Argentina
-\\
-\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation
-\\
-\textbf{7} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
+\textbf{4} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina; Instituto Nacional de Tecnología Agropecuaria, Anguil, La Pampa, Argentina; R-Ladies, California, United States of America  
+\\
+\textbf{5} MetaDocencia, Ciudad Autónoma de Buenos Aires, Argentina
+\\
+\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation for Statistical Computing, Vienna, Austria
+\\
+\textbf{7} Mountain Hydrology and Mass Movements, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
 \\
@@
 \\
-\textbf{10} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK; King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
+\textbf{10} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, UK; King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
 \\
@@
 \\
-\textbf{12} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
+\textbf{12} Instituto de Cálculo, Facultad de Ciencias Exactas y Naturales, Universidad de Buenos Aires - Consejo Nacional de Investigaciones Científicas y Técnicas, Ciudad Autónoma de Buenos Aires, Argentina
 \\

### 2022-05-14 final_manuscript_no_markup.tex
:tada: final version for production

@@
 \\
-\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
-\\
-\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
-\\
-\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina; Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
-\\
-\textbf{5} MetaDocencia, Argentina
-\\
-\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation
-\\
-\textbf{7} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
+\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Brazil
+\\
+\textbf{3} International Institute for Sustainability, Rio de Janeiro, Brazil
+\\
+\textbf{4} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina; Instituto Nacional de Tecnología Agropecuaria, Anguil, La Pampa, Argentina; R-Ladies, California, United States of America  
+\\
+\textbf{5} MetaDocencia, Ciudad Autónoma de Buenos Aires, Argentina
+\\
+\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation for Statistical Computing, Vienna, Austria
+\\
+\textbf{7} Mountain Hydrology and Mass Movements, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
 \\
@@
 \\
-\textbf{12} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
-\\
-\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brazil
+\textbf{12} Instituto de Cálculo, Facultad de Ciencias Exactas y Naturales, Universidad de Buenos Aires - Consejo Nacional de Investigaciones Científicas y Técnicas, Ciudad Autónoma de Buenos Aires, Argentina
+\\
+\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, Brazil
 \\
@@
 \\
-\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, Pennsylvania, United States of America.
-\\
-\textbf{18} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
+\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, Pennsylvania, United States of America
+\\
+\textbf{18} Bioinformatics team, Département de Formation et de Recherche Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, Yamoussoukro, Côte d’Ivoire.
 \\
@@
 \\
-\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, Mexico
+\textbf{20} Departamento de Microbiología Molecular, Instituto de Biotecnología UNAM, Cuernavaca, Morelos, Mexico
 \\
@@
 
-Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work.
+Conferences are spaces to meet and network within and across academic and technical fields, learn about new advances, and share our work.
 They can help define career paths and create long-lasting collaborations and opportunities.
@@
 2021 and supporting us through the process.
-Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, Francisco Etchart for creating the Marmot figures, and Koki Tsuyuzaki, Kozo Nishida, KwangChun Lee, Arjun Krishnan, Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae for their extended abstract translations.
+Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, Francisco Etchart for creating the Marmot figures, and Koki Tsuyuzaki, Kozo Nishida, KwangChun Lee, Arjun Krishnan, Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae for their extended abstract translations (S1 Text).
 We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript.
 
-\section*{Funding}
-
+\section*{Financial disclosure}
 JR is funded by start-up award from Michigan State University (MSU).
 HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.
-
 The funders had no role in study design, decision to publish, or preparation of the manuscript.
+
+\section*{Competing interests}
+I have read the journal's policy and the authors of this manuscript have the following competing interests: HT is a board member of The R Foundation, serving on the R Foundation conference committee that arranges hosts for the useR!
+conference series.
 

### 2022-05-23 final_manuscript_no_markup.tex
adding alternative text section to the main file

@@
 
+\section*{Alternative text for the figures}
+
+\textbf{Fig 1.} Diagram of the ten rules organized into three groups: foundation rules (Rules 1-3), design rules (Rules 4 to 9), and a continuity rule (Rule 10).
+The overall diagram depicts an arrow from left to right which contains all the rules.
+In the left column, under the first label "Foundation", there are three text boxes in light blue: "1.
+Vision", "2.
+Safe environment", and "3.
+Organizing Team".
+In the middle column, under the label "Design Rules" there are six light green boxes with Rules "4.
+Counter bias", "5.
+Virtual / hybrid", "6.
+Accessibility", "7.
+Language", "8.
+Communication”, “9.
+Finance".
+To the right, below the arrowhead labeled Continuity, is the tenth Rule in a light pink box "10.
+Pass the torch".
+
+\textbf{Fig 2.}  Five Margot the marmot with colorful scarves.
+From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride
+flag waves her hand.
+
 \section*{Financial disclosure}

### 2022-06-17 final_manuscript_no_markup.tex
add recent PDF

--- 2022-05-14 final_manuscript_no_markup.tex
@@
 }
-\caption{Schematic diagram of the rules organized in three groups: foundation (Rules 1, 2, and 3), design (Rules 4 to 9), and continuity (Rule 10).}
+\caption{Schematic diagram of the rules organized in three groups: foundation (Rules 1, 2, and 3), design (Rules 4 to 9), and continuity (Rule 10).
+See alt-text in the `Alternative text for figures' section.}
 \label{fig:diagram}
@@
 From left to right: Margot wearing scarves with the colors of MiR (Minorities in R \url{https://mircommunity.com}), R-Ladies \url{https://rladies.org}, AfricaR \url{https://africa-r.org}, and the LGBTQIA+ and transgender communities.
-Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license (\url{https://github.com/useRconf/visuals-2021}).}
+Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license (\url{https://github.com/useRconf/visuals-2021}).
+See alt-text in the `Alternative text for figures' section.}
 \label{fig:marmots}
@@
 
+\section*{Alternative text for the figures}
+
+\textbf{Fig 1.} Diagram of the ten rules organized into three groups: foundation rules (Rules 1-3), design rules (Rules 4 to 9), and a continuity rule (Rule 10).
+The overall diagram depicts an arrow from left to right containing all the rules.
+In the left column, under the first label `Foundation', there are three text boxes in light blue: `1.
+Vision', `2.
+Safe environment', and `3.
+Organizing Team'.
+In the middle column, under the label `Design Rules' there are six light green boxes with Rules `4.
+Counter bias', `5.
+Virtual/hybrid', `6.
+Accessibility', `7.
+Language', `8.
+Communication', `9.
+Finance'.
+To the right, below the arrowhead labeled `Continuity', is the tenth Rule in a light pink box `10.
+Pass the torch'.
+
+\textbf{Fig 2.}  Five `Margot the marmot' with colorful scarves.
+From left to right: a marmot with a dark and light green scarf of the MiR community is looking at her tablet, a marmot with the purple and gray scarf of R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride
+flag waves her hand.
+
 \section*{Financial disclosure}
